### 2017-09-20 paper/paper.tex
Start writing a bit

% vim: set spell spelllang=en tw=100 et sw=4 sts=4 :

\documentclass{article}
\usepackage{ijcai17}
\usepackage{times}
\usepackage{microtype}

\usepackage{showframe}

\title{Value-Ordering, Discrepancies, and Restarts in Subgraph Solvers}
\author{Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences
    Research Council [grant number EP/026842/1]}\\ University of Glasgow, Glasgow, Scotland \\
    ciaran.mccreesh@glasgow.ac.uk}

\begin{document}

\maketitle

\begin{abstract}
    Modern subgraph isomorphism solvers use reasonably good value-ordering heuristics to direct
    search, but they are not perfect, particularly for early choices. We investigate various forms
    of discrepancy search, but get best results by combining restarts with biased randomness in
    value-ordering heuristics.
\end{abstract}

\section{Introduction}

\section{The Basic Algorithm}

?? Started with the Glasgow algorithm, removed parts which did not contribute hugely to speedups,
and ended up with a simpler algorithm that performs as well in practice. In particular, large
graphs, no ILF, and not doing supplemental graphs of length 3.

\subsection{Ordering Heuristics}

\begin{description}
    \item[Random]
    \item[Tailored]
    \item[Anti-Tailored]
    \item[Tailor-Weighted Random]
\end{description}

\subsection{Discrepancy Searches}

?? Not dealing with binary search. Make it binary. Other versions where we increase the discrepancy
each time the degree changes, and where we count every ``against'' branch as one discrepancy.

\subsection{Restarts}

?? Aimed primarily at value-ordering, not value-ordering.

?? Value-ordering options

?? Propagation

?? We use the Luby scheme, because everyone else does. We count the number of backtracks (that is,
when we reach the end of the main for loop) to decide when to restart. Following established wisdom,
we multiply each item in the sequence by a magic constant. Preliminary experiments demonstrated that
this is useful, but we failed to determine a principled way of selecting the constant's value, so we
fell back on divine revelation and set it to 666. As with everyone else, we are not entirely clear
why we are doing this: it could be to allow the solver to spend more time deep in search, it could
be to reduce the overheads from repeatedly visiting inner nodes, or it could be to keep the nogoods
small.

?? Increasing nogoods etc. We are not confident enough in our programming abilities to implement
these correctly. Cannot propagate on negative decisions.

\section{Empirical Evaluation}

\subsection{Our Solver is Good}

?? Non-induced, compare to Glasgow, LAD, PathLAD, SND?, VF2

?? Induced, compare to Glasgow, LAD, PathLAD, VF2, VF3, McSplit.

\subsection{Discrepancies Don't Help}

?? DDS counting bias both ways, LDS

\subsection{Restarts Make it Better}

?? Random, tailored, antitailored, random with restarts, tailored biased, tailored biased with
restarts. Satisfiable and unsatisfiable split.

?? Breakdown by family

\subsection{Near the Phase Transition}

?? Closer look at SAT instances near the phase transition

\section{Other Subgraph Solvers}

?? Also do this in McSplit and kdown?

\section{Conclusion}

?? Parallel

\bibliographystyle{named}
\bibliography{dblp}

\end{document}

### 2017-09-24 paper/paper.tex
Start scribbling

@@
 \usepackage{microtype}
+\usepackage{amsmath}
 
@@
 
-\title{Value-Ordering, Discrepancies, and Restarts in Subgraph Solvers}
+\title{Some Half-Baked Mucking Around with Value-Ordering, Discrepancies, and Restarts in Subgraph Solvers}
 \author{Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences
@@
     Modern subgraph isomorphism solvers use reasonably good value-ordering heuristics to direct
-    search, but they are not perfect, particularly for early choices. We investigate various forms
-    of discrepancy search, but get best results by combining restarts with biased randomness in
-    value-ordering heuristics.
+    search, but they are not perfect, particularly for early choices. We investigate two mechanisms
+    for offsetting this weakness: discrepancy searches, and restarts. Our results suggest that
+    discrepancy search is too expensive to be effective, but that restarts with a random
+    value-ordering heuristic beat tailored heuristics. We then design a slightly-random
+    value-ordering heuristic which, when combined with restarts, performs ?? times better on
+    satisfiable instances without substantially weakening performance on unsatisfiable instances.
 \end{abstract}
@@
 \section{Introduction}
+
+?? Subgraph isomorphism is useful.
+
+The current strongest subgraph isomorphism solver uses degree as a value-ordering heuristic. This is
+much better than not having a value-ordering heuristic, but it is far from perfect, particularly for
+early branching choices. There are two reasons we suspect we can do better. Firstly, . Secondly, even if degree
+were a perfect source of information, many graphs do not have a large degree distribution, and so we
+still need to determine \emph{which} vertex of highest degree would be best.
+
+We could try to introduce tiebreaking mechanisms. However, given the extreme cost of making an
+incorrect branching choice early on in search, we believe it is more fruitful to investigate
+alternatives to simple backtracking (depth-first) search.
 
@@
 
+?? Variable ordering is smallest domain first with degree tiebreaking. This is not the focus of this
+paper.
+
+In our experiments, we compare five value-ordering heuristics.
+
 \begin{description}
-    \item[Random]
-    \item[Tailored]
-    \item[Anti-Tailored]
-    \item[Tailor-Weighted Random]
+    \item[Tailored] Where we select vertices from highest degree to lowest degree (which is done by
+        permuting the graph at the top of search).
+    \item[Anti-Tailored] Where we select vertices from lowest degree to highest degree.
+    \item[Extra Tiebreaking] Where we select vertices from highest degree to lowest degree,
+        tiebreaking on highest sum of neighbour degrees.
+    \item[Random] Where we select a branching vertex at random by (dynamically, that is, we shuffle
+        the vertices before entering the for loop, rather than by randomly permuting the graph once at the
+        top of search).
+    \item[Tailor-Weighted Random] Where we shuffle the vertices before each for loop, but instead of
+        picking a vertex randomly, we weight the chances of a vertex $v$ being chosen next by \[
+            2^{1 + \left|\operatorname{N}(v)\right|} \textnormal{.} \]
+        This will heavily prefer vertices of high degree, but will give the same chance of being
+        selected next to two vertices of equal degree. This also introduces an element of
+        randomness, which will be important for restarts.
 \end{description}

### 2017-09-25 code/unit.cc
Betterererer

@@ -167,12 +167,17 @@ namespace
                     if (pattern_graphs.at(0).adjacent(i, j))
                         pattern_degree_tiebreak.at(j).second += pattern_degree_tiebreak.at(i).first;
 
+            int max_degree = 0;
+            for (unsigned j = 0 ; j < target_size ; ++j)
+                max_degree = std::max(max_degree, target_graphs.at(0).degree(j));
+
             if (params.biased_shuffle) {
                 for (unsigned j = 0 ; j < target_size ; ++j) {
-                    unsigned long long bias = 1ull << (1 + target_graphs.at(0).degree(j));
-                    if (0 == bias)
-                        bias = std::numeric_limits<unsigned long long>::max();
-                    target_vertex_biases.push_back(bias);
+                    int degree = target_graphs.at(0).degree(j);
+                    if (max_degree - degree >= 50)
+                        target_vertex_biases.push_back(1);
+                    else
+                        target_vertex_biases.push_back(1ull << (50 - (max_degree - degree)));
                 }
             }
         }
@@ -418,23 +423,25 @@ namespace
             // pull out the remaining values in this domain for branching
             auto remaining = branch_domain->values;
 
-            std::vector<unsigned> branch_v;
+            std::array<unsigned, n_words_ * bits_per_word + 1> branch_v;
+            unsigned branch_v_end = 0;
             for (int f_v = remaining.first_set_bit() ; f_v != -1 ; f_v = remaining.first_set_bit()) {
                 remaining.unset(f_v);
-                branch_v.push_back(f_v);
+                branch_v[branch_v_end++] = f_v;
             }
 
-            if (params.shuffle)
-                std::shuffle(branch_v.begin(), branch_v.end(), global_rand);
+            if (params.shuffle) {
+                std::shuffle(branch_v.begin(), branch_v.begin() + branch_v_end, global_rand);
+            }
             else if (params.biased_shuffle) {
                 // sum up the bias scores of every branch vertex
                 unsigned long long remaining_score = 0;
-                for (auto & v : branch_v)
-                    remaining_score += target_vertex_biases.at(v);
+                for (unsigned v = 0 ; v < branch_v_end ; ++v)
+                    remaining_score += target_vertex_biases[branch_v[v]];
 
                 // now repeatedly pick a biased-random vertex, move it to the front of branch_v,
                 // and then only consider items further to the right in the next iteration.
[…]

### 2017-09-25 paper/paper.tex
I should go to sleep

@@
 
-\usepackage{showframe}
+% \usepackage{showframe}
 
-\title{Some Half-Baked Mucking Around with Value-Ordering, Discrepancies, and Restarts in Subgraph Solvers}
+\title{Value-Ordering, Discrepancies, and Restarts in Subgraph Solvers: \\ Some Half-Baked Mucking
+Around}
 \author{Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences
@@
     \item[Tailor-Weighted Random] Where we shuffle the vertices before each for loop, but instead of
-        picking a vertex randomly, we weight the chances of a vertex $v$ being chosen next by \[
-            2^{1 + \left|\operatorname{N}(v)\right|} \textnormal{.} \]
-        This will heavily prefer vertices of high degree, but will give the same chance of being
-        selected next to two vertices of equal degree. This also introduces an element of
-        randomness, which will be important for restarts.
+        picking a vertex uniformly at random, we use a weight of
+        $2^{\left|\operatorname{N}(v)\right|}$ when deciding whether to branch on vertex $v$
+        next.\footnote{Actually, we cheat a bit to avoid range issues with random number generation:
+        we give the vertex with highest degree in the entire graph weight $2^{50}$, and then give
+        lower degree vertices smaller powers, giving every vertex weight at least 1. This is a bit
+        of hack but getting it perfectly right is hard.} This scheme heavily prefer vertices of high
+        degree, but will give the same chance of being selected next to two vertices of equal
+        degree. This also introduces an element of randomness, which will be important for restarts.
 \end{description}
@@
 \section{Empirical Evaluation}
+
+\subsection{Instances}
+
+these, partly because some earlier solvers find many of these instances extremely hard, and partly
+to avoid skewing a portfolio. We want larger numbers of satisfiable instances, so we include all
+pattern / target pairs.
 

### 2017-09-25 paper/paper.tex
Al Gore rythmn

@@
 \usepackage{amsmath}
+\usepackage[ruled,vlined]{algorithm2e}
+\usepackage{tikz}
 
-% \usepackage{showframe}
+\usepackage{showframe}
+
+\newcommand{\neighbourhood}{\operatorname{N}}
+\newcommand{\vertexset}{\operatorname{V}}
+\newcommand{\nds}{\operatorname{S}}
+
+\usetikzlibrary{decorations, decorations.pathreplacing, decorations.pathmorphing,
+    calc, backgrounds, positioning, tikzmark, patterns, fit}
+
+\definecolor{chromaygb1}{rgb}{1.0, 1.0, 0.878431}
+\definecolor{chromaygb1b}{rgb}{0.8, 0.972549, 0.72549}
+\definecolor{chromaygb2}{rgb}{0.560784,0.929412,0.564706}
+\definecolor{chromaygb3}{rgb}{0.407843,0.776471,0.560784}
+\definecolor{chromaygb4}{rgb}{0.227451,0.631373,0.552941}
+\definecolor{chromaygb5}{rgb}{0.301961,0.431373,0.670588}
+\definecolor{chromaygb6}{rgb}{0.317647,0.254902,0.588235}
+\definecolor{chromaygb7}{rgb}{0.294118,0.0,0.509804}
 
@@
 
+\newcommand{\siplabel}[1]{\label{line:sip:#1}}
+\begin{algorithm}[p]\DontPrintSemicolon\small
+    \begin{tikzpicture}[remember picture,overlay]
+        \coordinate (restart1sc) at ($(pic cs:restart1s) + (0, 0.15)$);
+        \coordinate (restart1ec) at ($(pic cs:restart1e) + (0, 0.00)$);
+        \node [fill=chromaygb3, rounded corners, fit=(restart1sc) (restart1ec)] { };
+        \coordinate (restart2sc) at ($(pic cs:restart2s) + (0, 0.15)$);
+        \coordinate (restart2ec) at ($(pic cs:restart2e) + (0, 0.00)$);
+        \node [fill=chromaygb3, rounded corners, fit=(restart2sc) (restart2ec)] { };
+        \coordinate (restart3sc) at ($(pic cs:restart3s) + (0, 0.15)$);
+        \coordinate (restart3ec) at ($(pic cs:restart3e) + (0, 0.00)$);
+        \node [fill=chromaygb3, rounded corners, fit=(restart3sc) (restart3ec)] { };
+        \coordinate (restart4sc) at ($(pic cs:restart4s) + (0, 0.15)$);
+        \coordinate (restart4ec) at ($(pic cs:restart4e) + (0, 0.00)$);
+        \node [fill=chromaygb3, rounded corners, fit=(restart4sc) (restart4ec)] { };
+        \coordinate (restart5sc) at ($(pic cs:restart5s) + (0, 0.15)$);
+        \coordinate (restart5ec) at ($(pic cs:restart5e) - (0, 0.10)$);
+        \coordinate (restart5rc) at ($(pic cs:restart5r) + (0, 0.00)$);
+        \node [fill=chromaygb3, rounded corners, fit=(restart5sc) (restart5ec) (restart5rc)] { };
+        \coordinate (restart6sc) at ($(pic cs:restart6s) + (0, 0.15)$);
+        \coordinate (restart6ec) at ($(pic cs:restart6e) + (0, 0.00)$);
+        \node [fill=chromaygb3, rounded corners, fit=(restart6sc) (restart6ec)] { };
+    \end{tikzpicture}
+
+    \nl $\FuncSty{subgraphIsomorphism}$ (Graph $\mathcal{P}$, Graph $\mathcal{T}$) $\rightarrow$ Bool \;
+\nl \Begin{
+    \nl \lIf{$\left|\vertexset(\mathcal{P})\right| >
+    \left|\vertexset(\mathcal{T})\right|$}{$\KwSty{return}~\KwSty{false}$\siplabel{enough}}
+    \nl Discard isolated vertices in $\mathcal{P}$\siplabel{isolated} \;
+    \nl \KwSty{global} $L \gets \big[ (\mathcal{P},\ \mathcal{T}), \siplabel{supplemental}
+        (\mathcal{P}^{1,2}, \mathcal{T}^{1,2}),
+        (\mathcal{P}^{2,2},\ \mathcal{T}^{2,2}), $\\\hspace*{6em}$ (\mathcal{P}^{3,2},\ \mathcal{T}^{3,2}),
+        (\mathcal{P}^{4,2},\ \mathcal{T}^{4,2}) \big]$ \;
+    \nl \ForEach{$v \in \vertexset(\mathcal{P})$}{
+        \nl $D_v \gets \vertexset(\mathcal{T})$\;
+        \nl \ForEach{$(P,\,T) \in L$}{
+            \nl $D_v \gets \{ w \in D_v : $ \siplabel{degreefiltering} \\\hspace*{4em}$v \sim_P v \Rightarrow w \sim_T w~\wedge$
+            \\\hspace*{4em}$\nds_P(v) \preceq \nds_T(w)\}$ \;
+        }
+    }
+    \nl \lIf{$\KwSty{not}~\FuncSty{propagate}(D)$}{$\KwSty{return}~\KwSty{false}$}
+    \nl \While{$\KwSty{true}$}{
+        \nl \tikzmark{restart1s}$\KwSty{global}~(R, R') \gets (0,\,\textnormal{the next Luby value})$\tikzmark{restart1e} \;
+        \nl $S \gets \FuncSty{search}(\{ E \in D : \left|E\right| > 1 \}, \top)$ \;
+        \nl \lIf{$S = \KwSty{true}$}{$\KwSty{return}~\KwSty{true}$}
+        \nl \lElseIf{$S = \KwSty{false}$}{$\KwSty{return}~\KwSty{false}$}
+        \nl \tikzmark{restart2s}Trigger nogoods containing only one clause\tikzmark{restart2e} \;
+    }
+}
+\medskip
+    \nl $\FuncSty{search}$ (Domains $D$, Decisions $B$) $\rightarrow$ Bool
+    \KwSty{or} \KwSty{restart} \;
+\nl \Begin{
+    \nl \lIf{$D = \emptyset$}{$\KwSty{return}~\KwSty{true}$}
+    \nl $D_v \gets \textnormal{the smallest domain in}~D$ \siplabel{pickd} \;
+    \nl \tikzmark{restart3s}$C \gets \emptyset$\tikzmark{restart3e} \;
+    \nl \ForEach {$v' \in D_v~\textnormal{ordered by chosen heuristic}$\siplabel{parforeach}}{
+        \nl \tikzmark{restart4s}$C \gets C \cup \{ v' \}$\tikzmark{restart4e} \;
+        \nl $D' \gets \FuncSty{clone}(D)$ \;
+        \nl $D'_v \gets \{ v' \} $ \;
+        \nl \If{\FuncSty{propagate}(D')}{
+            \nl $S \gets \FuncSty{search}(\{ E \in D' : \left|E\right| > 1 \},$ \\\hspace*{3em} $B
+            \wedge (v \mapsto v'))$ \;
+            \nl \lIf{$S = \KwSty{true}$}{$\KwSty{return}~\KwSty{true}$}
+            \nl \tikzmark{restart5s}\ElseIf{$S = \KwSty{restart}$}{
+                \nl Post nogoods $\{B \wedge (v \mapsto c) \rightarrow \bot : c \in C \}$\tikzmark{restart5r}\;
+                \nl $\KwSty{return}~\KwSty{restart}$\tikzmark{restart5e} \;
+            }
+        }
+    }
+    \nl \tikzmark{restart6s}\lIf{$(R \gets R + 1) = R'$}{$\KwSty{return}~\KwSty{restart}\tikzmark{restart6e}$}
+    \nl $\KwSty{return}~\KwSty{false}$
+}
+\medskip
+\nl $\FuncSty{propagate}$ (Domains $D$) $\rightarrow$ Bool \;
+\nl \Begin{
+    \nl \While{$D_v \gets$~\textnormal{a unit domain from}~$D$}{
+        \nl $v' \gets $ the single value in $D_v$ \;
+        \nl Trigger nogoods containing $(v \mapsto v')$ \;
+        \nl \ForEach{$D_w \in D - D_v$}{
+            \nl $D_w \gets D_w - v'$ \siplabel{removev} \;
+            \nl \ForEach{$(P,\,T) \in L$}{
+                \nl \lIf{$v \sim_P w$}{$D_w \gets D_w \cap \neighbourhood_T(v')$
+                }
+            }
+            \nl \lIf{$D_w = \emptyset$}{$\KwSty{return}~\KwSty{false}$}
+        }
+    }
+\nl $(H,\,A,\,n) \gets (\emptyset,\,\emptyset,\,0)$ \;
+\nl \ForEach{$D_v \in D$ \textnormal{from smallest cardinality upwards\siplabel{eachdomain}}}{
+    \nl $D_v \gets D_v \setminus H$ \siplabel{elimhall} \;
+    \nl $(A,\,n) \gets (A \cup D_v,\,n + 1)$ \siplabel{acc} \;
+    \nl \lIf{$D_v = \emptyset~\vee~|A| < n$}{$\KwSty{return}~\KwSty{false}$\siplabel{failhall}}
+    \nl \lIf{$|A| = n$}{$(H,\,A,\,n) \gets (H \cup A,\,\emptyset,\,0)$\siplabel{hall}}
+}
+\nl $\KwSty{return}~\KwSty{true}$ \;
+}
+\caption{Subgraph isomorphism. Highlighted lines are ignored for the non-restarting algorithm.}
+\end{algorithm}
+
 \subsection{Ordering Heuristics}
@@
         picking a vertex uniformly at random, we use a weight of
-        $2^{\left|\operatorname{N}(v)\right|}$ when deciding whether to branch on vertex $v$
+        $2^{\left|\neighbourhood(v)\right|}$ when deciding whether to branch on vertex $v$
         next.\footnote{Actually, we cheat a bit to avoid range issues with random number generation:

### 2017-09-25 paper/paper.tex
Heuristics in text
Hail Santa!
Put words back in

@@
     \nl \While{$\KwSty{true}$}{
-        \nl \tikzmark{restart1s}$\KwSty{global}~(R, R') \gets (0,\,\textnormal{the next Luby value})$\tikzmark{restart1e} \;
+        \nl \tikzmark{restart1s}$\KwSty{global}~(R, R') \gets (0, 666 \times \textnormal{the next Luby value})$\tikzmark{restart1e} \;
         \nl $S \gets \FuncSty{search}(\{ E \in D : \left|E\right| > 1 \}, \top)$ \;
@@
     \nl \lIf{$D = \emptyset$}{$\KwSty{return}~\KwSty{true}$}
-    \nl $D_v \gets \textnormal{the smallest domain in}~D$ \siplabel{pickd} \;
+    \nl $D_v \gets \textnormal{a domain from}~D~\textnormal{chosen by variable heuristic}$ \siplabel{pickd} \;
     \nl \tikzmark{restart3s}$C \gets \emptyset$\tikzmark{restart3e} \;
-    \nl \ForEach {$v' \in D_v~\textnormal{ordered by chosen heuristic}$\siplabel{parforeach}}{
+    \nl \ForEach {$v' \in D_v~\textnormal{ordered by chosen value heuristic}$\siplabel{parforeach}}{
         \nl \tikzmark{restart4s}$C \gets C \cup \{ v' \}$\tikzmark{restart4e} \;
@@
 
-these, partly because some earlier solvers find many of these instances extremely hard, and partly
-to avoid skewing a portfolio. We want larger numbers of satisfiable instances, so we include all
-pattern / target pairs.
+?? Other studies use a random selection of 200 of the mesh instances, partly because some earlier
+solvers find many of these instances extremely hard, and partly to avoid skewing a portfolio. We
+want larger numbers of satisfiable instances, so we include all pattern / target pairs.
 

### 2017-09-25 paper/paper.tex
Small skip to fit
Notational consistency
I like green boxes
Plotting

@@
 \usepackage{tikz}
+\usepackage{cleveref}
 
-\usepackage{showframe}
+% \usepackage{showframe}
 
@@
         \node [fill=chromaygb3, rounded corners, fit=(restart6sc) (restart6ec)] { };
+        \coordinate (restart7sc) at ($(pic cs:restart7s) + (0, 0.15)$);
+        \coordinate (restart7ec) at ($(pic cs:restart7e) + (0, 0.00)$);
+        \node [fill=chromaygb3, rounded corners, fit=(restart7sc) (restart7ec)] { };
+        \coordinate (restart8sc) at ($(pic cs:restart8s) + (0.05, 0.15)$);
+        \coordinate (restart8ec) at ($(pic cs:restart8e) + (0, 0.00)$);
+        \node [fill=chromaygb3, rounded corners, fit=(restart8sc) (restart8ec)] { };
     \end{tikzpicture}
@@
 }
-\medskip
+\smallskip
     \nl $\FuncSty{search}$ (Domains $D$, Decisions $B$) $\rightarrow$ Bool
-    \KwSty{or} \KwSty{restart} \;
+    \tikzmark{restart8s}\KwSty{or} \KwSty{restart}\tikzmark{restart8e} \;
 \nl \Begin{
@@
             \nl \tikzmark{restart5s}\ElseIf{$S = \KwSty{restart}$}{
-                \nl Post nogoods $\{B \wedge (v \mapsto c) \rightarrow \bot : c \in C \}$\tikzmark{restart5r}\;
+                \nl Post nogoods $\{B \wedge (v \mapsto c) \Rightarrow \bot : c \in C \}$\tikzmark{restart5r}\;
                 \nl $\KwSty{return}~\KwSty{restart}$\tikzmark{restart5e} \;
@@
 }
-\medskip
+\smallskip
 \nl $\FuncSty{propagate}$ (Domains $D$) $\rightarrow$ Bool \;
@@
         \nl $v' \gets $ the single value in $D_v$ \;
-        \nl Trigger nogoods containing $(v \mapsto v')$ \;
+        \nl \tikzmark{restart7s}Trigger nogoods containing $(v \mapsto v')$\tikzmark{restart7e} \;
         \nl \ForEach{$D_w \in D - D_v$}{
@@
         permuting the graph at the top of search).
-    \item[Anti-Tailored] Where we select vertices from lowest degree to highest degree.
-    \item[Extra Tiebreaking] Where we select vertices from highest degree to lowest degree,
+    \item[Anti] Where we select vertices from lowest degree to highest degree.
+    \item[Tiebreaking] Where we select vertices from highest degree to lowest degree,
         tiebreaking on highest sum of neighbour degrees.
@@
         top of search).
-    \item[Tailor-Weighted Random] Where we shuffle the vertices before each for loop, but instead of
+    \item[Weighted] Where we shuffle the vertices before each for loop, but instead of
         picking a vertex uniformly at random, we use a weight of
@@
 
-?? Non-induced, compare to Glasgow, LAD, PathLAD, SND?, VF2
+?? Non-induced, compare to Glasgow, LAD, PathLAD, SND?, VF2, using tailored
 
-?? Induced, compare to Glasgow, LAD, PathLAD, VF2, VF3, McSplit.
+?? Possibly also do induced, compare to Glasgow, LAD, PathLAD, VF2, VF3, McSplit.
+
+\subsection{Value Ordering Heuristics}
+
+?? \cref{figure:value-ordering-heuristics}
+
+\begin{figure}
+    \centering
+    \includegraphics*{gen-graph-value-ordering-heuristics.pdf}
+
+    \caption{Cumulative number of instances solved over time, using the five different
+    value-ordering heuristics. The top three lines are very close together, with the weighted
+    heuristic being worse for low runtimes and slightly better towards the end.}
+    \label{figure:value-ordering-heuristics}
+\end{figure}
 
@@
 
-?? DDS counting bias both ways, LDS
+?? Results for DDS counting discrepancies both ways, maybe LDS too. Also look just at satisfiable
+instances.
 
@@
 
-?? Random, tailored, antitailored, random with restarts, tailored biased, tailored biased with
-restarts. Satisfiable and unsatisfiable split.
+\begin{figure}
+    \centering
+    \includegraphics*{gen-graph-restarts.pdf}
 
-?? Breakdown by family
+    \caption{Cumulative number of instances solved over time, with and without restarts.}
+    \label{figure:restarts}
+\end{figure}
 
-\subsection{Near the Phase Transition}
+\begin{figure}
+    \centering
+    \includegraphics*{gen-graph-scatter.pdf}
 
-?? Closer look at SAT instances near the phase transition
+    \caption{Instance by instance comparison of the basic algorithm with the tailored heuristic,
+    versus weighted with restarts.}
+    \label{figure:restarts}
+\end{figure}
+
+?? \cref{figure:restarts}
+
+?? breakdown by family
 

### 2017-09-25 paper/paper.tex
So, it has come to this

@@
 
-\definecolor{chromaygb1}{rgb}{1.0, 1.0, 0.878431}
-\definecolor{chromaygb1b}{rgb}{0.8, 0.972549, 0.72549}
-\definecolor{chromaygb2}{rgb}{0.560784,0.929412,0.564706}
-\definecolor{chromaygb3}{rgb}{0.407843,0.776471,0.560784}
-\definecolor{chromaygb4}{rgb}{0.227451,0.631373,0.552941}
-\definecolor{chromaygb5}{rgb}{0.301961,0.431373,0.670588}
-\definecolor{chromaygb6}{rgb}{0.317647,0.254902,0.588235}
-\definecolor{chromaygb7}{rgb}{0.294118,0.0,0.509804}
+\definecolor{magma8}{rgb}{0.996078, 0.760784, 0.529412}
 
@@
         \coordinate (restart1ec) at ($(pic cs:restart1e) + (0, 0.00)$);
-        \node [fill=chromaygb3, rounded corners, fit=(restart1sc) (restart1ec)] { };
+        \node [fill=magma8, rounded corners, fit=(restart1sc) (restart1ec)] { };
         \coordinate (restart2sc) at ($(pic cs:restart2s) + (0, 0.15)$);
         \coordinate (restart2ec) at ($(pic cs:restart2e) + (0, 0.00)$);
-        \node [fill=chromaygb3, rounded corners, fit=(restart2sc) (restart2ec)] { };
+        \node [fill=magma8, rounded corners, fit=(restart2sc) (restart2ec)] { };
         \coordinate (restart3sc) at ($(pic cs:restart3s) + (0, 0.15)$);
         \coordinate (restart3ec) at ($(pic cs:restart3e) + (0, 0.00)$);
-        \node [fill=chromaygb3, rounded corners, fit=(restart3sc) (restart3ec)] { };
+        \node [fill=magma8, rounded corners, fit=(restart3sc) (restart3ec)] { };
         \coordinate (restart4sc) at ($(pic cs:restart4s) + (0, 0.15)$);
         \coordinate (restart4ec) at ($(pic cs:restart4e) + (0, 0.00)$);
-        \node [fill=chromaygb3, rounded corners, fit=(restart4sc) (restart4ec)] { };
+        \node [fill=magma8, rounded corners, fit=(restart4sc) (restart4ec)] { };
         \coordinate (restart5sc) at ($(pic cs:restart5s) + (0, 0.15)$);
@@
         \coordinate (restart5rc) at ($(pic cs:restart5r) + (0, 0.00)$);
-        \node [fill=chromaygb3, rounded corners, fit=(restart5sc) (restart5ec) (restart5rc)] { };
+        \node [fill=magma8, rounded corners, fit=(restart5sc) (restart5ec) (restart5rc)] { };
         \coordinate (restart6sc) at ($(pic cs:restart6s) + (0, 0.15)$);
         \coordinate (restart6ec) at ($(pic cs:restart6e) + (0, 0.00)$);
-        \node [fill=chromaygb3, rounded corners, fit=(restart6sc) (restart6ec)] { };
+        \node [fill=magma8, rounded corners, fit=(restart6sc) (restart6ec)] { };
         \coordinate (restart7sc) at ($(pic cs:restart7s) + (0, 0.15)$);
         \coordinate (restart7ec) at ($(pic cs:restart7e) + (0, 0.00)$);
-        \node [fill=chromaygb3, rounded corners, fit=(restart7sc) (restart7ec)] { };
+        \node [fill=magma8, rounded corners, fit=(restart7sc) (restart7ec)] { };
         \coordinate (restart8sc) at ($(pic cs:restart8s) + (0.05, 0.15)$);
         \coordinate (restart8ec) at ($(pic cs:restart8e) + (0, 0.00)$);
-        \node [fill=chromaygb3, rounded corners, fit=(restart8sc) (restart8ec)] { };
+        \node [fill=magma8, rounded corners, fit=(restart8sc) (restart8ec)] { };
     \end{tikzpicture}
@@
 
-\begin{figure}
+\begin{figure}[p]
     \centering
@@
 
-\subsection{Discrepancies Don't Help}
-
-?? Results for DDS counting discrepancies both ways, maybe LDS too. Also look just at satisfiable
-instances.
-
-\subsection{Restarts Make it Better}
-
-\begin{figure}
+\begin{figure}[p]
     \centering
@@
     versus weighted with restarts.}
-    \label{figure:restarts}
+    \label{figure:scatter}
 \end{figure}
+
+\subsection{Discrepancies Don't Help}
+
+?? Results for DDS counting discrepancies both ways, maybe LDS too. Also look just at satisfiable
+instances.
+
+\subsection{Restarts Make it Better}
 

### 2017-09-25 paper/paper.tex
More plots

@@
 
-\begin{figure}[p]
+\begin{figure}[tb]
     \centering
@@
 \end{figure}
+
+\subsection{Discrepancies Don't Help}
+
+?? Results for DDS counting discrepancies both ways, maybe LDS too. Also look just at satisfiable
+instances.
+
+\subsection{Restarts Make it Better}
+
+?? \cref{figure:restarts}
+
+?? breakdown by family
 
@@
 
+    \bigskip
+    \centering
+    \includegraphics*{gen-graph-scatter-random.pdf}
+
     \caption{Instance by instance comparison of the basic algorithm with the tailored heuristic,
-    versus weighted with restarts.}
+    versus weighted with restarts. Below, comparing instead with simple random with restarts.}
     \label{figure:scatter}
 \end{figure}
-
-\subsection{Discrepancies Don't Help}
-
-?? Results for DDS counting discrepancies both ways, maybe LDS too. Also look just at satisfiable
-instances.
-
-\subsection{Restarts Make it Better}
-
-?? \cref{figure:restarts}
-
-?? breakdown by family
 

### 2017-09-26 paper/paper.tex
More thingies

@@
 
-\title{Value-Ordering, Discrepancies, and Restarts in Subgraph Solvers: \\ Some Half-Baked Mucking
+% cref style
+\crefname{algorithm}{Algorithm}{Algorithms}
+\Crefname{algorithm}{Algorithm}{Algorithms}
+\crefname{figure}{Figure}{Figures}
+\Crefname{figure}{Figure}{Figures}
+\crefname{table}{Table}{Tables}
+\Crefname{table}{Table}{Tables}
+\crefname{section}{Section}{Sections}
+\Crefname{section}{Section}{Sections}
+
+\title{Value Ordering, Discrepancies, and Restarts in Subgraph Solvers: \\ Some Half-Baked Mucking
 Around}
@@
 \begin{abstract}
-    Modern subgraph isomorphism solvers use reasonably good value-ordering heuristics to direct
-    search, but they are not perfect, particularly for early choices. We investigate two mechanisms
-    for offsetting this weakness: discrepancy searches, and restarts. Our results suggest that
-    discrepancy search is too expensive to be effective, but that restarts with a random
-    value-ordering heuristic beat tailored heuristics. We then design a slightly-random
-    value-ordering heuristic which, when combined with restarts, performs ?? times better on
-    satisfiable instances without substantially weakening performance on unsatisfiable instances.
+    At least one modern subgraph isomorphism solver use reasonably good value-ordering heuristics to
+    direct search, but these heuristics are not perfect, particularly for early choices. We
+    investigate two mechanisms for offsetting this weakness: discrepancy searches, and restarts. Our
+    results suggest that discrepancy search is too expensive to be effective, but that restarts with
+    a random value-ordering heuristic beats the best tailored heuristic. We also introduce a
+    slightly-random value-ordering heuristic with similar behaviour to the tailored heuristic.  When
+    combined with restarts, this new heuristic is even more effective on satisfiable instances, and
+    does not substantially affect performance on unsatisfiable instances.
 \end{abstract}
@@
 
-The current strongest subgraph isomorphism solver uses degree as a value-ordering heuristic. This is
-much better than not having a value-ordering heuristic, but it is far from perfect, particularly for
-early branching choices. There are two reasons we suspect we can do better. Firstly, . Secondly, even if degree
-were a perfect source of information, many graphs do not have a large degree distribution, and so we
-still need to determine \emph{which} vertex of highest degree would be best.
+The current single strongest subgraph isomorphism solver uses ``highest degree first'' as a
+value-ordering heuristic.  This is much better than not having a value-ordering heuristic at all,
+but it is far from perfect, particularly for early branching choices. There are two reasons to
+suspect we could do better.  Firstly, even if degree were a perfect source of information, many
+graphs do not have a large degree distribution, and so we would still need to determine \emph{which}
+vertex of highest degree would be best.  Secondly, experiments with a parallel search strategy which
+explicitly steals early branching choices often gives strongly super-linear speedups on satisfiable
+instances.
 
-We could try to introduce tiebreaking mechanisms. However, given the extreme cost of making an
-incorrect branching choice early on in search, we believe it is more fruitful to investigate
-alternatives to simple backtracking (depth-first) search.
+We could try to introduce tiebreaking mechanisms to further refine this heuristic. However, given
+the extreme cost of making an incorrect branching choice early on in search, we believe it is more
+fruitful to investigate alternatives to simple backtracking (depth-first) search. This paper
+evaluates two such approaches: discrepancy search, and restarts.
+
+?? Quick intro to discrepancy search. Quick intro to restarts, mentioning how it's important to
+change something when restarting.
+
+?? And concludes what?
 
@@
 and ended up with a simpler algorithm that performs as well in practice. In particular, large
-graphs, no ILF, and not doing supplemental graphs of length 3.
+graphs, no ILF, and not doing supplemental graphs of length 3. This is \cref{algorithm:sip},
+ignoring the shaded parts, which are introduced in the context of restarts.
+
+?? Briefly describe this, probably referring elsewhere as far as possible to save space.
 
 \newcommand{\siplabel}[1]{\label{line:sip:#1}}
-\begin{algorithm}[p]\DontPrintSemicolon\small
+\newcommand{\siplineref}[1]{line~\ref{line:sip:#1}}
+\begin{algorithm}[p]\DontPrintSemicolon\small\SetInd{0.1em}{1em}
     \begin{tikzpicture}[remember picture,overlay]
@@
     }
+    \nl \While{$\KwSty{true}$}{
         \nl \lIf{$\KwSty{not}~\FuncSty{propagate}(D)$}{$\KwSty{return}~\KwSty{false}$}
-    \nl \While{$\KwSty{true}$}{
         \nl \tikzmark{restart1s}$\KwSty{global}~(R, R') \gets (0, 666 \times \textnormal{the next Luby value})$\tikzmark{restart1e} \;
@@
         \nl \lElseIf{$S = \KwSty{false}$}{$\KwSty{return}~\KwSty{false}$}
-        \nl \tikzmark{restart2s}Trigger nogoods containing only one clause\tikzmark{restart2e} \;
+        \nl \tikzmark{restart2s}trigger nogoods containing only one clause\tikzmark{restart2e} \;
     }
 }
-\smallskip
+\medskip
     \nl $\FuncSty{search}$ (Domains $D$, Decisions $B$) $\rightarrow$ Bool
@@
     \nl \lIf{$D = \emptyset$}{$\KwSty{return}~\KwSty{true}$}
-    \nl $D_v \gets \textnormal{a domain from}~D~\textnormal{chosen by variable heuristic}$ \siplabel{pickd} \;
+    \nl $D_v \gets \textnormal{a domain from}~D~\textnormal{chosen by variable heuristic}$\siplabel{variableordering} \;
     \nl \tikzmark{restart3s}$C \gets \emptyset$\tikzmark{restart3e} \;
-    \nl \ForEach {$v' \in D_v~\textnormal{ordered by chosen value heuristic}$\siplabel{parforeach}}{
+    \nl \ForEach {$v' \in D_v~\textnormal{ordered by value heuristic}$\siplabel{valueordering}}{
         \nl \tikzmark{restart4s}$C \gets C \cup \{ v' \}$\tikzmark{restart4e} \;
@@
         \nl \If{\FuncSty{propagate}(D')}{
-            \nl $S \gets \FuncSty{search}(\{ E \in D' : \left|E\right| > 1 \},$ \\\hspace*{3em} $B
-            \wedge (v \mapsto v'))$ \;
+            \nl $S \gets \FuncSty{search}(\{ E \in D' : \left|E\right| > 1 \}, B \wedge (v \mapsto v'))$ \;
             \nl \lIf{$S = \KwSty{true}$}{$\KwSty{return}~\KwSty{true}$}
             \nl \tikzmark{restart5s}\ElseIf{$S = \KwSty{restart}$}{
-                \nl Post nogoods $\{B \wedge (v \mapsto c) \Rightarrow \bot : c \in C \}$\tikzmark{restart5r}\;
+                \nl post nogoods $\{B \wedge (v \mapsto c) \Rightarrow \bot : c \in C \}$\tikzmark{restart5r}\;
                 \nl $\KwSty{return}~\KwSty{restart}$\tikzmark{restart5e} \;
@@
 }
-\smallskip
+\medskip
 \nl $\FuncSty{propagate}$ (Domains $D$) $\rightarrow$ Bool \;
@@
         \nl $v' \gets $ the single value in $D_v$ \;
-        \nl \tikzmark{restart7s}Trigger nogoods containing $(v \mapsto v')$\tikzmark{restart7e} \;
+        \nl \tikzmark{restart7s}trigger nogoods with watch $(v \mapsto v')$\tikzmark{restart7e} \;
         \nl \ForEach{$D_w \in D - D_v$}{
@@
 }
-\caption{Subgraph isomorphism. Highlighted lines are ignored for the non-restarting algorithm.}
+\caption{Subgraph isomorphism. Shaded code is to be ignored for the non-restarting algorithm.}
+\label{algorithm:sip}
 \end{algorithm}
@@
 
-?? Variable ordering is smallest domain first with degree tiebreaking. This is not the focus of this
-paper.
+For the variable-ordering heuristic on \siplineref{variableordering} of \cref{algorithm:sip}, we use
+smallest domain first, tiebreaking on ?? degree. Variable ordering is not the focus of this
+paper---as far as possible, we will leave it intact, so as not to affect the performance of the
+algorithm on unsatisfiable instances.
 
-In our experiments, we compare five value-ordering heuristics.
+For value ordering, which occurs on \siplineref{valueordering} of \cref{algorithm:sip}, we will
+compare four strategies:
 
-\begin{description}
-    \item[Tailored] Where we select vertices from highest degree to lowest degree (which is done by
-        permuting the graph at the top of search).
-    \item[Anti] Where we select vertices from lowest degree to highest degree.
-    \item[Tiebreaking] Where we select vertices from highest degree to lowest degree,
-        tiebreaking on highest sum of neighbour degrees.
-    \item[Random] Where we select a branching vertex at random by (dynamically, that is, we shuffle
-        the vertices before entering the for loop, rather than by randomly permuting the graph once at the
-        top of search).
-    \item[Weighted] Where we shuffle the vertices before each for loop, but instead of
-        picking a vertex uniformly at random, we use a weight of
-        $2^{\left|\neighbourhood(v)\right|}$ when deciding whether to branch on vertex $v$
-        next.\footnote{Actually, we cheat a bit to avoid range issues with random number generation:
-        we give the vertex with highest degree in the entire graph weight $2^{50}$, and then give
-        lower degree vertices smaller powers, giving every vertex weight at least 1. This is a bit
-        of hack but getting it perfectly right is hard.} This scheme heavily prefer vertices of high
-        degree, but will give the same chance of being selected next to two vertices of equal
-        degree. This also introduces an element of randomness, which will be important for restarts.
-\end{description}
+\paragraph{Random} We select a vertex dynamically at random (that is, we shuffle the vertices before
+entering the $\KwSty{foreach}$ loop on \siplineref{valueordering} of \cref{algorithm:sip}). This is
+\emph{usually} effectively equivalent to picking vertices in the order in which they appear in the
+input files, although some instances from some benchmark suites are naturally in a special order
+(which may or may not be a good order for searching).
+
+\paragraph{Tailored} We select vertices from highest degree to lowest degree. This is done by
+permuting the target graph at the top of search. This heuristic is the current state of the art.
+
+\paragraph{Anti} We select vertices from lowest degree to highest degree. This heuristic is used as
+a sanity check: we expect tailored to beat random, and random to beat anti.
+
+\paragraph{Weighted} As with the random heuristic, we shuffle the vertices before we enter the
+$\KwSty{foreach}$ loop. Instead of picking a vertex uniformly at random, we use a weight of
+$2^{\left|\neighbourhood(v)\right|}$ when deciding whether to branch on vertex $v$ next.\footnote{We
+cheat slightly to avoid range issues with random number generation: we give the vertex with highest
+degree in the entire graph weight $2^{50}$, and then give lower degree vertices smaller powers,
+giving every vertex weight at least 1.  This is a bit of hack but programming is hard.} This
+heuristic heavily prefers vertices of high degree, similar to the tailored heuristic, but will give
+the same chance of being selected next to two vertices of equal degree. This heuristic also
+introduces an element of randomness, which will be important for restarts.
 
@@
 
-?? Not dealing with binary search. Make it binary. Other versions where we increase the discrepancy
-each time the degree changes, and where we count every ``against'' branch as one discrepancy.
+?? Talk about how we're not dealing with a binary search tree, so we could either make it binary, or
+count one discrepancy each time the degree changes, or count every ``against'' branch as being just
+one discrepancy regardless.
 
@@
 
-?? Aimed primarily at value-ordering, not value-ordering.
+?? We use the Luby scheme, because everyone else does. We count the number of backtracks (that is,
+when we reach the end of the main $\KwSty{foreach}$ loop) to decide when to restart. Following
+established wisdom, we multiply each item in the Luby sequence by a magic constant. Preliminary
+experiments demonstrated that this is useful, but we failed to determine a principled way of
+selecting the constant's value, so we fell back on divine revelation and set it to 666. As with
+everyone else, we are not entirely clear why we are doing this: it could be to allow the solver to
+spend more time deep in search, it could be to reduce the overheads from repeatedly visiting inner
+nodes, or it could be to keep the nogoods small.
 
-?? Value-ordering options
+?? We just use two watched literals. Could use increasing nogoods etc. We are not confident enough
+in our programming abilities to implement these correctly, and we are not maintaining arc
+consistency anyway. We also really do not want to propagate on negative decisions, because we do not
+want to iterate over every removed value.
 
-?? Propagation
-
-?? We use the Luby scheme, because everyone else does. We count the number of backtracks (that is,
-when we reach the end of the main for loop) to decide when to restart. Following established wisdom,
-we multiply each item in the sequence by a magic constant. Preliminary experiments demonstrated that
-this is useful, but we failed to determine a principled way of selecting the constant's value, so we
-fell back on divine revelation and set it to 666. As with everyone else, we are not entirely clear
-why we are doing this: it could be to allow the solver to spend more time deep in search, it could
-be to reduce the overheads from repeatedly visiting inner nodes, or it could be to keep the nogoods
-small.
-
-?? Increasing nogoods etc. We are not confident enough in our programming abilities to implement
-these correctly. Cannot propagate on negative decisions.
+?? Talk a bit about how it fits into \cref{algorithm:sip}.
 
@@
 
-\subsection{Instances}
+?? Dual Xeon E5-2697A v4 CPUs, 512GBytes RAM, running Ubuntu 17.04, implemented in C++ and compiled
+using GCC 6.3.0. Our code is heavily based upon ??.
 
-?? Other studies use a random selection of 200 of the mesh instances, partly because some earlier
-solvers find many of these instances extremely hard, and partly to avoid skewing a portfolio. We
-want larger numbers of satisfiable instances, so we include all pattern / target pairs.
+?? Deterministic RNG.
+
+?? Bitsets.
+
+?? The 5,725 instances.
+
+?? Other studies use a random selection of 200 of each of the mesh and images instances, partly
+because some earlier solvers find many of these instances extremely hard, and partly to avoid
+skewing a portfolio. We want larger numbers of satisfiable instances, so we include all pattern /
+target pairs. This gives a total of 14,621 instances, ?? of which we know are satisfiable, and ??
+unsatisfiable.
+
+?? Note that an awful lot of these instances are very easy. We are trying to tackle the remaining
+hard satisfiable instances.
 
@@
 
-?? Non-induced, compare to Glasgow, LAD, PathLAD, SND?, VF2, using tailored
+?? Establish a baseline. Compare to older Glasgow2, Glasgow3, LAD, PathLAD, SND if we can get it to
+run on the larger set of instances, VF2, maybe SAT just to have something to cite showing how bad
+SAT is at this. Just use tailored in this plot. Possibly just use the 5,725 instances here to avoid
+pain with lousy solvers.
 
@@
 
-?? \cref{figure:value-ordering-heuristics}
+In \cref{figure:value-ordering-heuristics} we show the cumulative number of instances solved over
+time using the four different value-ordering heuristics. As expected, the tailored heuristic is much
+better (?? give aggregate speedups for both all and just satisfiable instances) than random
+ordering, which is in turn ?? times better than the anti-heuristic.
+
+The performance of the weighted and tailored heuristics is very similar: the weighted heuristic has
+slightly worse performance below a 100s timeout, and effectively equal performance above it. This
+shows that we can introduce an element of randomness into the tailored value-ordering heuristic
+without adversely affecting its performance.
 
@@
 
-    \caption{Cumulative number of instances solved over time, using the five different
-    value-ordering heuristics. The top three lines are very close together, with the weighted
-    heuristic being worse for low runtimes and slightly better towards the end.}
+    \caption{Cumulative number of instances solved over time, using the four different
+    value-ordering heuristics.}
     \label{figure:value-ordering-heuristics}
@@
 
-?? Results for DDS counting discrepancies both ways, maybe LDS too. Also look just at satisfiable
-instances.
+?? Code up LDS and DDS and the like properly, and plot them being disappointing here.
 
@@
 
-?? \cref{figure:restarts}
-
-?? breakdown by family
+In \cref{figure:restarts} we show the effects of adding restarts to the algorithm. With restarts,
+the random variable-ordering heuristic comfortably beats the tailored strategy without restarts. In
+other words, although having a good variable-ordering heuristic is beneficial, introducing
+randomness into the search is better, if it is done alongside a mechanism to avoid heavy commitment
+to any particular random choice.  However, the weighted heuristic together with restarts is better
+still---that is, if we are introducing restarts, then it is better to add a small amount of
+randomness to a tailored deterministic heuristic than it is to throw away the heuristic altogether.
 
@@
     \caption{Instance by instance comparison of the basic algorithm with the tailored heuristic,
-    versus weighted with restarts. Below, comparing instead with simple random with restarts.}
+    versus weighted + restarts. Below, comparing instead with simple random + restarts.}
     \label{figure:scatter}
 \end{figure}
+
+Cumulative plots only show aggregate behaviour. What happens on an instance by instance basis?  In
+the top of \cref{figure:scatter} we scatter plot performance, showing satisfiable and unsatisfiable
+instances using dots and crosses respectively. On the $x$-axis we show the tailored heuristic,
+whilst on the $y$-axis we show weighted + restarts. Points on the outer axes represent instances
+which timed out with one heuristic but not the other.
+
+Nearly all of the unsatisfiable instances are on or slightly above the $x-y$ diagonal, showing that
+their performance is either unchanged or reduced slightly. On the other hand, there are ?? large
+numbers of satisfiable instances well below the diagonal line, indicating large speedups. Better
+yet, there are only ?? satisfiable instances that are more than a factor of ?? times worse.
+
+?? Figure out what's going on with the unsat instances not on the line. There are a few options.
+Could be SDF picking something else, could be the hacky all-diff missing something, could be how
+we're doing unit propagation, could be a horrible bug.
+
+For contrast, the bottom plot of \cref{figure:scatter} compares the tailored heuristic with random +
+restarts. The performance on unsatisfiable instances is the same. However, although better in
+aggregate, we see many more satisfiable instances performing substantially worse. ?? Quantify this a
+bit.
+
+?? do a more detailed breakdown by family
+
+?? Explain why this is so amazingly awesome and the most important idea ever to happen in the entire
+history of science.
 
@@
 
+?? This stuff is quite good.
+
+?? Bigger CP picture
+
 ?? Parallel

### 2017-09-26 paper/paper.tex
Zoom, LAD, etc

@@
 
-?? Establish a baseline. Compare to older Glasgow2, Glasgow3, LAD, PathLAD, SND if we can get it to
-run on the larger set of instances, VF2, maybe SAT just to have something to cite showing how bad
-SAT is at this. Just use tailored in this plot. Possibly just use the 5,725 instances here to avoid
-pain with lousy solvers.
+?? In \cref{figure:others} we show that our solver is the bestest.
+
+\begin{figure}[tb]
+    \centering
+    \includegraphics*{gen-graph-others.pdf}
+
+    \centering
+    \includegraphics*{gen-graph-others-zoom.pdf}
+
+    \caption{Cumulative number of instances solved over time, comparing our algorithm to other
+    solvers. Below, zoomed in a bit ?? make this pretty.}
+    \label{figure:others}
+\end{figure}
+
+?? Stick in Glasgow2, Glasgow3, VF2, SND if we can get it to run on the larger set of instances,
+maybe SAT just to have something to cite showing how bad SAT is at this.
 

### 2017-09-26 paper/paper.tex
Exciting

@@
 \newcommand{\nds}{\operatorname{S}}
+
+\newcommand{\citet}[1]{\citeauthor{#1} \shortcite{#1}}
+\newcommand{\citep}[1]{\cite{#1}}
 
@@
 \begin{abstract}
-    At least one modern subgraph isomorphism solver use reasonably good value-ordering heuristics to
-    direct search, but these heuristics are not perfect, particularly for early choices. We
-    investigate two mechanisms for offsetting this weakness: discrepancy searches, and restarts. Our
-    results suggest that discrepancy search is too expensive to be effective, but that restarts with
-    a random value-ordering heuristic beats the best tailored heuristic. We also introduce a
-    slightly-random value-ordering heuristic with similar behaviour to the tailored heuristic.  When
-    combined with restarts, this new heuristic is even more effective on satisfiable instances, and
-    does not substantially affect performance on unsatisfiable instances.
+    Modern subgraph isomorphism solvers use strong value-ordering heuristics to direct search, but
+    these heuristics are not perfect, particularly for early choices. Our results suggest that
+    discrepancy-based searches are too expensive to be effective at offsetting this weakness.
+    However, restarts with a random value-ordering heuristic beats conventional search with the best
+    tailored heuristic. This motivates the introduction of a slightly-random value-ordering
+    heuristic with similar behaviour to the tailored heuristic: when combined with restarts, this
+    new heuristic is even more effective on satisfiable instances, whilst not significantly
+    decreasing performance on unsatisfiable instances.
 \end{abstract}
@@
 The current single strongest subgraph isomorphism solver uses ``highest degree first'' as a
-value-ordering heuristic.  This is much better than not having a value-ordering heuristic at all,
-but it is far from perfect, particularly for early branching choices. There are two reasons to
-suspect we could do better.  Firstly, even if degree were a perfect source of information, many
-graphs do not have a large degree distribution, and so we would still need to determine \emph{which}
-vertex of highest degree would be best.  Secondly, experiments with a parallel search strategy which
-explicitly steals early branching choices often gives strongly super-linear speedups on satisfiable
-instances.
+value-ordering heuristic \citep{dblp:conf/cp/McCreeshP15,dblp:conf/ijcai/McCreeshPT16}.  This is
+much better than not having a value-ordering heuristic at all, but it is far from perfect,
+particularly for early branching choices. There are two reasons to suspect we could do better.
+Firstly, even if degree were a perfect source of information, many graphs do not have a large degree
+distribution, and so we would still need to determine \emph{which} vertex of highest degree would be
+best.  Secondly, experiments with a parallel search strategy which explicitly steals early branching
+choices often gives strongly super-linear speedups on satisfiable instances
+\citep{dblp:conf/cp/McCreeshP15}.
 
@@
 evaluates two such approaches: discrepancy search, and restarts.
-
-?? Quick intro to discrepancy search. Quick intro to restarts, mentioning how it's important to
-change something when restarting.
 
@@
 For the variable-ordering heuristic on \siplineref{variableordering} of \cref{algorithm:sip}, we use
-smallest domain first, tiebreaking on ?? degree. Variable ordering is not the focus of this
+smallest domain first, tiebreaking on highest degree. Variable ordering is not the focus of this
 paper---as far as possible, we will leave it intact, so as not to affect the performance of the
@@
 
-\paragraph{Weighted} As with the random heuristic, we shuffle the vertices before we enter the
+\paragraph{Biased} As with the random heuristic, we shuffle the vertices before we enter the
 $\KwSty{foreach}$ loop. Instead of picking a vertex uniformly at random, we use a weight of
-$2^{\left|\neighbourhood(v)\right|}$ when deciding whether to branch on vertex $v$ next.\footnote{We
-cheat slightly to avoid range issues with random number generation: we give the vertex with highest
-degree in the entire graph weight $2^{50}$, and then give lower degree vertices smaller powers,
-giving every vertex weight at least 1.  This is a bit of hack but programming is hard.} This
-heuristic heavily prefers vertices of high degree, similar to the tailored heuristic, but will give
-the same chance of being selected next to two vertices of equal degree. This heuristic also
-introduces an element of randomness, which will be important for restarts.
+$2^{\left|\neighbourhood(v)\right|}$ when deciding whether to branch on vertex $v$ next. (We cheat
+slightly to avoid range issues with random number generation: we give the vertex with highest degree
+in the entire graph weight $2^{50}$, and then give lower degree vertices smaller powers, giving
+every vertex weight at least 1.  This is a bit of hack but programming is hard.) This heuristic
+heavily prefers vertices of high degree, similar to the tailored heuristic, but will give the same
+chance of being selected next to two vertices of equal degree. This heuristic also introduces an
+element of randomness, and so can be used with restarts.
 
@@
 
+?? Quick intro to discrepancy search
+\citep{DBLP:conf/ijcai/HarveyG95,DBLP:conf/aaai/Korf96,DBLP:conf/ijcai/Walsh97,DBLP:conf/cpaior/KarouiHLN07,DBLP:journals/jea/ProsserU11}.
+Discrepancy search is about value ordering.
+
 ?? Talk about how we're not dealing with a binary search tree, so we could either make it binary, or
 count one discrepancy each time the degree changes, or count every ``against'' branch as being just
-one discrepancy regardless.
+one discrepancy regardless. ?? Has anyone done a proper job of this?
 
@@
 
-?? We use the Luby scheme, because everyone else does. We count the number of backtracks (that is,
-when we reach the end of the main $\KwSty{foreach}$ loop) to decide when to restart. Following
-established wisdom, we multiply each item in the Luby sequence by a magic constant. Preliminary
-experiments demonstrated that this is useful, but we failed to determine a principled way of
-selecting the constant's value, so we fell back on divine revelation and set it to 666. As with
-everyone else, we are not entirely clear why we are doing this: it could be to allow the solver to
-spend more time deep in search, it could be to reduce the overheads from repeatedly visiting inner
-nodes, or it could be to keep the nogoods small.
+?? Quick intro to restarts, mentioning how it's important to change something when restarting.
+Interestingly usually that thing is variable-ordering heuristics, not value-ordering
+\citep{DBLP:conf/ijcai/LecoutreSTV07,DBLP:journals/jsat/LecoutreSTV07,DBLP:conf/cp/GayHLS15,DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17}.
+?? Does anyone at all do this on value ordering explicitly? And if not, why not? There's
+\citet{DBLP:conf/flairs/RazgonOP07} at least. ?? There's a connection here to confidence-based work
+stealing \citep{DBLP:conf/cp/ChuSS09} too. But what? They talk about confidence in a heuristic, but
+not that different values could be ranked differently.
 
-?? We just use two watched literals. Could use increasing nogoods etc. We are not confident enough
+?? We use the Luby scheme \citep{DBLP:journals/ipl/LubySZ93}, because everyone else does. We count
+the number of backtracks (that is, when we reach the end of the main $\KwSty{foreach}$ loop) to
+decide when to restart. Following established wisdom, we multiply each item in the Luby sequence by
+a magic constant. Preliminary experiments demonstrated that this is useful, but we failed to
+determine a principled way of selecting the constant's value, so we fell back on divine revelation
+and set it to 666. As with everyone else, we are not entirely clear why we are doing this: it could
+be to allow the solver to spend more time deep in search, it could be to reduce the overheads from
+repeatedly visiting inner nodes, or it could be to keep the nogoods small.
+
+?? We just use two watched literals \citep{DBLP:conf/dac/MoskewiczMZZM01}. Could use increasing
+nogoods \citep{DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17} etc, but we are not confident enough
 in our programming abilities to implement these correctly, and we are not maintaining arc
@@
 ?? Dual Xeon E5-2697A v4 CPUs, 512GBytes RAM, running Ubuntu 17.04, implemented in C++ and compiled
-using GCC 6.3.0. Our code is heavily based upon ??.
+using GCC 6.3.0. Our code is heavily based upon \citet{dblp:conf/cp/McCreeshP15}.
 
@@
 
-?? The 5,725 instances.
+?? The 5,725 instances \citep{dblp:conf/lion/KotthoffMS16}.
 
@@
 
-The performance of the weighted and tailored heuristics is very similar: the weighted heuristic has
+The performance of the biased and tailored heuristics is very similar: the biased heuristic has
 slightly worse performance below a 100s timeout, and effectively equal performance above it. This
@@
 randomness into the search is better, if it is done alongside a mechanism to avoid heavy commitment
-to any particular random choice.  However, the weighted heuristic together with restarts is better
+to any particular random choice.  However, the biased heuristic together with restarts is better
 still---that is, if we are introducing restarts, then it is better to add a small amount of
@@
     \caption{Instance by instance comparison of the basic algorithm with the tailored heuristic,
-    versus weighted + restarts. Below, comparing instead with simple random + restarts.}
+    versus biased + restarts. Below, comparing instead with simple random + restarts.}
     \label{figure:scatter}
@@
 instances using dots and crosses respectively. On the $x$-axis we show the tailored heuristic,
-whilst on the $y$-axis we show weighted + restarts. Points on the outer axes represent instances
+whilst on the $y$-axis we show biased + restarts. Points on the outer axes represent instances
 which timed out with one heuristic but not the other.
@@
 
+?? maybe some engineering aspects: look at performance with restarts on tailored, and at restarts
+without nogood recording
+
 ?? Explain why this is so amazingly awesome and the most important idea ever to happen in the entire
@@
 
-?? Bigger CP picture
+?? We can keep good value-ordering heuristics, and get large improvement. Importantly, we can do
+this with almost no effect on unsatisfiable instances, and rare to make satisfiable much worse.
+
+?? Variable ordering
+
+?? Bigger CP picture.
 

### 2017-09-27 paper/paper.tex
More things

@@
 fruitful to investigate alternatives to simple backtracking (depth-first) search. This paper
-evaluates two such approaches: discrepancy search, and restarts.
+evaluates two such approaches: discrepancy search (which perform poorly), and restarts (which are
+extremely beneficial). To make restarts work, we introduce a new value-ordering heuristic, which is
+nearly the same as ``highest degree first'' except with small amounts of heavily biased randomness.
+The effects of the changes we propose are shown in the top plot of \cref{figure:scatter}: we can
+make a state-of-the-art algorithm's perform much better on a large number of satisfiable instances,
+without substantially degrading its performance on unsatisfiable or the remaining satisfiable
+instances.
 
@@
 \section{The Basic Algorithm}
+
+?? Notation: $\vertexset(G)$, $\neighbourhood_G(v)$, $\nds_G(v)$, $\preceq$, $\sim_G$, $G^{a,b}$, $\mapsto$.
 
@@
 count one discrepancy each time the degree changes, or count every ``against'' branch as being just
-one discrepancy regardless. ?? Has anyone done a proper job of this?
+one discrepancy regardless.
 
@@
 
-?? In \cref{figure:others} we show that our solver is the bestest.
+?? In \cref{figure:others} we show that our solver is the bestest. For each algorithm, the $y$ value
+gives the cumulative number of instances which (individually) can be solved in no more than $x$
+milliseconds.  The vertical distance between two lines therefore shows how many more instances can
+be solved by one solver than another, if every instance is run separately with the chosen $x$
+timeout; the horizontal distance shows how many times longer the per-instance timeout would need to
+be to allow the rightmost algorithm to succeed on $y$ out of the 14,621 instances (bearing in mind
+that the two sets of $y$ instances solved by the two algorithms could be completely different).
 
@@
 
-?? Stick in Glasgow2, Glasgow3, VF2, SND if we can get it to run on the larger set of instances,
+?? Add in Glasgow2, Glasgow3, SND if we can get it to run on the larger set of instances,
 maybe SAT just to have something to cite showing how bad SAT is at this.
@@
 In \cref{figure:restarts} we show the effects of adding restarts to the algorithm. With restarts,
-the random variable-ordering heuristic comfortably beats the tailored strategy without restarts. In
+the random value-ordering heuristic comfortably beats the tailored strategy without restarts. In
 other words, although having a good variable-ordering heuristic is beneficial, introducing

### 2017-09-27 paper/paper.tex
Fixing

@@
 The effects of the changes we propose are shown in the top plot of \cref{figure:scatter}: we can
-make a state-of-the-art algorithm's perform much better on a large number of satisfiable instances,
+make a state-of-the-art algorithm perform much better on a large number of satisfiable instances,
 without substantially degrading its performance on unsatisfiable or the remaining satisfiable
@@
 ?? Started with the Glasgow algorithm, removed parts which did not contribute hugely to speedups,
-and ended up with a simpler algorithm that performs as well in practice. In particular, large
-graphs, no ILF, and not doing supplemental graphs of length 3. This is \cref{algorithm:sip},
-ignoring the shaded parts, which are introduced in the context of restarts.
+and ended up with a simpler algorithm that performs as well in practice. In particular, because we
+will be working with reasonably large target graphs, no ILF, and not doing supplemental graphs of
+length 3. This is \cref{algorithm:sip}, ignoring the shaded parts, which are introduced in the
+context of restarts.
 
@@
 }
-\caption{Subgraph isomorphism. Shaded code is to be ignored for the non-restarting algorithm.}
+\caption{Non-induced subgraph isomorphism. Shaded code is added for the restarting variant.}
 \label{algorithm:sip}

### 2017-09-27 paper/paper.tex
Cite

@@
 
-?? Other studies use a random selection of 200 of each of the mesh and images instances, partly
-because some earlier solvers find many of these instances extremely hard, and partly to avoid
-skewing a portfolio. We want larger numbers of satisfiable instances, so we include all pattern /
-target pairs. This gives a total of 14,621 instances, ?? of which we know are satisfiable, and ??
-unsatisfiable.
+?? Other studies use a random selection of 200 of each of the mesh and images instances
+\citep{DBLP:journals/cviu/DamiandSHJS11}, partly because some earlier solvers find many of these
+instances extremely hard, and partly to avoid skewing a portfolio. We want larger numbers of
+satisfiable instances, so we include all pattern / target pairs. This gives a total of 14,621
+instances, ?? of which we know are satisfiable, and ??  unsatisfiable.
 

### 2017-09-30 paper/paper.tex
Tweaking

@@
 
+?? These results show our algorithm clearly beating PathLAD and VF2, except for very low choices of
+timeout. Our algorithm also beats Glasgow2 and Glasgow3, despite being slightly simpler---on an
+instance by instance basis, sometimes the additional inference performed by these algorithms pays
+off, but in aggregate, there is no difference for longer choices of timeout, and a clear gain for
+shorter choices.
+
 \begin{figure}[tb]
@@
 
-?? Add in Glasgow2, Glasgow3, SND if we can get it to run on the larger set of instances,
-maybe SAT just to have something to cite showing how bad SAT is at this.
+?? Add in SND if we can get it to run on the larger set of instances, maybe SAT just to have
+something to cite showing how bad SAT is at this.
 

### 2017-10-01 paper/paper.tex
Tweaks

@@
 \usepackage{tikz}
+
+\usepackage{hyperref}
 \usepackage{cleveref}
@@
 
+\begin{figure}[tb]
+    \centering
+    \includegraphics*{gen-graph-others.pdf}
+
+    \bigskip
+
+    \centering
+    \includegraphics*{gen-graph-others-zoom.pdf}
+
+    \caption{Cumulative number of instances solved over time, comparing our algorithm to other
+    solvers. Below, zoomed in a bit ?? make this pretty.}
+    \label{figure:others}
+\end{figure}
+
 ?? In \cref{figure:others} we show that our solver is the bestest. For each algorithm, the $y$ value
@@
 
-\begin{figure}[tb]
-    \centering
-    \includegraphics*{gen-graph-others.pdf}
-
-    \centering
-    \includegraphics*{gen-graph-others-zoom.pdf}
-
-    \caption{Cumulative number of instances solved over time, comparing our algorithm to other
-    solvers. Below, zoomed in a bit ?? make this pretty.}
-    \label{figure:others}
-\end{figure}
-
 ?? Add in SND if we can get it to run on the larger set of instances, maybe SAT just to have
@@
 
+\begin{figure}[tb]
+    \centering
+    \includegraphics*{gen-graph-value-ordering-heuristics.pdf}
+
+    \caption{The cumulative number of instances solved over time, using the four different
+    value-ordering heuristics.}
+    \label{figure:value-ordering-heuristics}
+\end{figure}
+
 In \cref{figure:value-ordering-heuristics} we show the cumulative number of instances solved over
@@
 
+\begin{figure}[tb]
+    \centering
+    \includegraphics*{gen-graph-scatter-heuristics.pdf}
+
+    \caption{An instance by instance comparison of the tailored and biased heuristics.}
+    \label{figure:scatter-heuristics}
+\end{figure}
+
 The performance of the biased and tailored heuristics is very similar: the biased heuristic has
@@
 shows that we can introduce an element of randomness into the tailored value-ordering heuristic
-without adversely affecting its performance.
-
-\begin{figure}[tb]
-    \centering
-    \includegraphics*{gen-graph-value-ordering-heuristics.pdf}
-
-    \caption{Cumulative number of instances solved over time, using the four different
-    value-ordering heuristics.}
-    \label{figure:value-ordering-heuristics}
-\end{figure}
+without adversely affecting its performance \emph{in aggregate}. \Cref{figure:scatter-heuristics}
+gives a more detailed look at these results. In this plot, satisfiable and unsatisfiable instances
+are shown using dots and crosses respectively. On the $x$-axis we measure the tailored heuristic,
+whilst on the $y$-axis we measure random + restarts. Points on the outer axes represent instances
+which timed out with one heuristic but not the other. This plot shows that on a case by case basis,
+randomness can make a large difference to individual satisfiable instances.
 
@@
 randomness into the search is better, if it is done alongside a mechanism to avoid heavy commitment
-to any particular random choice.  However, the biased heuristic together with restarts is better
-still---that is, if we are introducing restarts, then it is better to add a small amount of
-randomness to a tailored deterministic heuristic than it is to throw away the heuristic altogether.
+to any particular random choice. \Cref{figure:scatter-random} looks at this in more detail.  The
+performance on unsatisfiable instances is the same. However, although better in aggregate, we see
+many satisfiable instances performing substantially worse. ?? Quantify this a bit.
 
@@
     \centering
+    \includegraphics*{gen-graph-scatter-random.pdf}
+    \caption{Instance by instance comparison of the basic algorithm with the tailored heuristic,
+    versus random + restarts.}
+    \label{figure:scatter-random}
+\end{figure}
+
+\begin{figure}[p]
+    \centering
     \includegraphics*{gen-graph-scatter.pdf}
 
-    \bigskip
-    \centering
-    \includegraphics*{gen-graph-scatter-random.pdf}
-
     \caption{Instance by instance comparison of the basic algorithm with the tailored heuristic,
-    versus biased + restarts. Below, comparing instead with simple random + restarts.}
+    versus biased + restarts.}
     \label{figure:scatter}
@@
 
-Cumulative plots only show aggregate behaviour. What happens on an instance by instance basis?  In
-the top of \cref{figure:scatter} we scatter plot performance, showing satisfiable and unsatisfiable
-instances using dots and crosses respectively. On the $x$-axis we show the tailored heuristic,
-whilst on the $y$-axis we show biased + restarts. Points on the outer axes represent instances
-which timed out with one heuristic but not the other.
-
-Nearly all of the unsatisfiable instances are on or slightly above the $x-y$ diagonal, showing that
+\Cref{figure:restarts} shows that the biased heuristic together with restarts is better
+still---that is, if we are introducing restarts, then it is better to add a small amount of
+randomness to a tailored deterministic heuristic than it is to throw away the heuristic altogether.
+In the more detailed view in \cref{figure:scatter},
+nearly all of the unsatisfiable instances are on or slightly above the $x-y$ diagonal, showing that
 their performance is either unchanged or reduced slightly. On the other hand, there are ?? large
@@
 
+?? This has made up the consistency we lost from introducing randomness.
+
 ?? Figure out what's going on with the unsat instances not on the line. There are a few options.
@@
 we're doing unit propagation, could be a horrible bug.
-
-For contrast, the bottom plot of \cref{figure:scatter} compares the tailored heuristic with random +
-restarts. The performance on unsatisfiable instances is the same. However, although better in
-aggregate, we see many more satisfiable instances performing substantially worse. ?? Quantify this a
-bit.
 

### 2017-10-03 code/unit.cc
I hate std::map<>

@@ -8,8 +8,6 @@
 #include <numeric>
 #include <limits>
 #include <random>
-#include <set>
-#include <map>
 #include <list>
 
 #include <iostream>
@@ -79,13 +77,45 @@ namespace
         }
     };
 
+    // A nogood, aways of the form (list of assignments) -> false, where the
+    // last part is implicit. If there are at least two assignments, then the
+    // first two assignments are the watches (and the literals are permuted
+    // when the watches are updates).
     struct Nogood
     {
         std::vector<Assignment> literals;
     };
 
+    // nogoods stored here
     using Nogoods = std::list<Nogood>;
-    using Watches = std::map<Assignment, std::list<Nogoods::iterator> >;
+
+    // Two watched literals for our nogoods store.
+    struct Watches
+    {
+        // for each watched literal, we have a list of watched things, each of
+        // which is an iterator into the global watch list (so we can reorder
+        // the literal to keep the watches as the first two elements)
+        using WatchList = std::list<Nogoods::iterator>;
+
+        // two dimensional array, indexed by (target_size * p + t)
+        std::vector<WatchList> data;
+
+        unsigned pattern_size = 0, target_size = 0;
+
+        // not a ctor to avoid annoyingness with isolated vertices altering the
+        // pattern size
+        void initialise(unsigned p, unsigned t)
+        {
+            pattern_size = p;
+            target_size = t;
+            data.resize(p * t);
+        }
+
+        WatchList & operator[] (const Assignment & a)
+        {
+            return data[target_size * a.first + a.second];
+        }
+    };
 
     template <unsigned n_words_, int k_, int l_>
     struct SequentialSubgraphIsomorphism
@@ -147,6 +177,10 @@ namespace
 
             // determine ordering for target graph vertices
             std::iota(target_order.begin(), target_order.end(), 0);
+
+            // set up space for watches
+            if (params.restarts)
+                watches.initialise(pattern_size, target_size);
 
             if (params.tiebreaking)
                 tiebreaking_degree_sort(target, target_order, params.antiheuristic);
@@ -235,9 +269,9 @@ namespace
                 this_is_a_decision = false;
 
                 // propagate watches
-                auto watches_to_update = watches.find(current_assignment);
-                if (watches_to_update != watches.end()) {
-                    for (auto watch_to_update = watches_to_update->second.begin() ; watch_to_update != watches_to_update->second.end() ; ) {
+                if (params.restarts) {
[…]

### 2017-10-03 code/solve_subgraph_isomorphism.cc
Witchcraft

@@ -88,16 +88,17 @@ auto main(int argc, char * argv[]) -> int
     try {
         po::options_description display_options{ "Program options" };
         display_options.add_options()
-            ("help",                                  "Display help information")
-            ("timeout",            po::value<int>(),  "Abort after this many seconds")
+            ("help",                                         "Display help information")
+            ("timeout",            po::value<int>(),         "Abort after this many seconds")
             ("format",             po::value<std::string>(), "Specify the format of the input")
-            ("dds",                                   "Use depth-bounded discrepancy search")
-            ("restarts",                              "Use restarts (not with dds)")
-            ("shuffle",                               "Use shuffling (not with dds)")
-            ("biased-shuffle",                        "Use biased shuffling")
-            ("antiheuristic",                         "Use antiheuristic")
-            ("tiebreaking",                           "Use tiebreaking")
-            ("input-order",                           "Use input order")
+            ("dds",                                          "Use depth-bounded discrepancy search")
+            ("restarts",                                     "Use restarts (not with dds)")
+            ("shuffle",                                      "Use shuffling (not with dds)")
+            ("biased-shuffle",                               "Use biased shuffling")
+            ("antiheuristic",                                "Use antiheuristic")
+            ("tiebreaking",                                  "Use tiebreaking")
+            ("input-order",                                  "Use input order")
+            ("luby-multiplier",    po::value<unsigned>(),    "Specify a Luby multiplier")
             ;
 
         po::options_description all_options{ "All options" };
@@ -162,6 +163,9 @@ auto main(int argc, char * argv[]) -> int
         params.antiheuristic = options_vars.count("antiheuristic");
         params.tiebreaking = options_vars.count("tiebreaking");
         params.input_order = options_vars.count("input-order");
+
+        if (options_vars.count("luby-multiplier"))
+            params.luby_multiplier = options_vars["luby-multiplier"].as<unsigned>();
 
         /* Read in the graphs */
         auto graphs = std::make_pair(

### 2017-10-03 code/unit.cc

@@ -14,8 +14,6 @@
 
 namespace
 {
-    constexpr long long dodgy_magic_luby_multiplier = 666; // chosen by divine revelation
-
     enum class Search
     {
         Aborted,
@@ -789,7 +787,7 @@ namespace
                     std::list<long long> luby = {{ 1 }};
                     auto current_luby = luby.begin();
                     while (! done) {
-                        long long backtracks_until_restart = *current_luby * dodgy_magic_luby_multiplier;
+                        long long backtracks_until_restart = *current_luby * params.luby_multiplier;
                         if (std::next(current_luby) == luby.end()) {
                             luby.insert(luby.end(), luby.begin(), luby.end());
                             luby.push_back(*luby.rbegin() * 2);

### 2017-10-03 code/solve_subgraph_isomorphism.cc
Position shuffle

@@ -95,6 +95,7 @@ auto main(int argc, char * argv[]) -> int
             ("restarts",                                     "Use restarts (not with dds)")
             ("shuffle",                                      "Use shuffling (not with dds)")
             ("biased-shuffle",                               "Use biased shuffling")
+            ("position-shuffle",                             "Use position shuffling")
             ("antiheuristic",                                "Use antiheuristic")
             ("tiebreaking",                                  "Use tiebreaking")
             ("input-order",                                  "Use input order")
@@ -160,6 +161,7 @@ auto main(int argc, char * argv[]) -> int
         params.restarts = options_vars.count("restarts");
         params.shuffle = options_vars.count("shuffle");
         params.biased_shuffle = options_vars.count("biased-shuffle");
+        params.position_shuffle = options_vars.count("position-shuffle");
         params.antiheuristic = options_vars.count("antiheuristic");
         params.tiebreaking = options_vars.count("tiebreaking");
         params.input_order = options_vars.count("input-order");

### 2017-10-03 code/unit.cc

@@ -481,7 +481,7 @@ namespace
                     // go over the list until we've used up bias values totalling our
                     // random number
                     unsigned select_element = start;
-                    for ( ; select_element < branch_v.size() ; ++select_element) {
+                    for ( ; select_element < branch_v_end ; ++select_element) {
                         if (select_score <= target_vertex_biases[branch_v[select_element]])
                             break;
                         select_score -= target_vertex_biases[branch_v[select_element]];
@@ -489,6 +489,30 @@ namespace
 
                     // move to front, and update remaining_score
                     remaining_score -= target_vertex_biases[branch_v[select_element]];
+                    std::swap(branch_v[select_element], branch_v[start]);
+                }
+            }
+            else if (params.position_shuffle) {
+                // repeatedly pick a position-biased vertex, move it to the front of branch_v,
+                // and then only consider items further to the right in the next iteration.
+                for (unsigned start = 0 ; start < branch_v_end ; ++start) {
+                    // pick a random number between 0 and 1 inclusive
+                    std::uniform_real_distribution<double> dist(0, 1);
+                    double select_score = dist(global_rand);
+
+                    // this divides by two on each iteration, so we're twice as
+                    // likely to take the first vertex as the second and so on
+                    double select_if_score_ge = 1.0;
+
+                    // go over the list until we hit the score
+                    unsigned select_element = start;
+                    for ( ; select_element + 1 < branch_v_end ; ++select_element) {
+                        select_if_score_ge /= 2.0;
+                        if (select_score >= select_if_score_ge)
+                            break;
+                    }
+
+                    // move to front
                     std::swap(branch_v[select_element], branch_v[start]);
                 }
             }
@@ -832,7 +856,7 @@ namespace
                         }
                     }
                 }
-                else if (params.shuffle || params.biased_shuffle) {
+                else if (params.shuffle || params.biased_shuffle || params.position_shuffle) {
                     // still need to use the restarts variant
                     long long backtracks_until_restart = -1;
                     switch (restarting_search(assignments, domains, result.nodes, 0, backtracks_until_restart)) {

### 2017-10-04 code/solve_subgraph_isomorphism.cc
Nonogoods support

@@ -100,6 +100,7 @@ auto main(int argc, char * argv[]) -> int
             ("tiebreaking",                                  "Use tiebreaking")
             ("input-order",                                  "Use input order")
             ("luby-multiplier",    po::value<unsigned>(),    "Specify a Luby multiplier")
+            ("goods",                                        "No nogoods")
             ;
 
         po::options_description all_options{ "All options" };
@@ -165,6 +166,7 @@ auto main(int argc, char * argv[]) -> int
         params.antiheuristic = options_vars.count("antiheuristic");
         params.tiebreaking = options_vars.count("tiebreaking");
         params.input_order = options_vars.count("input-order");
+        params.goods = options_vars.count("goods");
 
         if (options_vars.count("luby-multiplier"))
             params.luby_multiplier = options_vars["luby-multiplier"].as<unsigned>();

### 2017-10-04 code/unit.cc

@@ -425,6 +425,9 @@ namespace
         auto post_nogood(
                 const Assignments & assignments)
         {
+            if (params.goods)
+                return;
+
             Nogood nogood;
 
             for (auto & a : assignments.values)

### 2017-10-04 paper/paper.tex
A bit cleaner
More things

@@
         \node [fill=magma8, rounded corners, fit=(restart8sc) (restart8ec)] { };
+        \coordinate (restart9sc) at ($(pic cs:restart9s) + (0, 0.15)$);
+        \coordinate (restart9ec) at ($(pic cs:restart9e) + (-0.15, 0.00)$);
+        \node [fill=magma8, rounded corners, fit=(restart9sc) (restart9ec)] { };
     \end{tikzpicture}
@@
         \nl $S \gets \FuncSty{search}(\{ E \in D : \left|E\right| > 1 \}, \top)$ \;
-        \nl \lIf{$S = \KwSty{true}$}{$\KwSty{return}~\KwSty{true}$}
-        \nl \lElseIf{$S = \KwSty{false}$}{$\KwSty{return}~\KwSty{false}$}
+        \nl \tikzmark{restart9s}\lIf{$S \ne \KwSty{restart}$}{\tikzmark{restart9e}$\KwSty{return}~S$}
         \nl \tikzmark{restart2s}trigger nogoods containing only one clause\tikzmark{restart2e} \;
@@
 }
-\medskip
+\bigskip
     \nl $\FuncSty{search}$ (Domains $D$, Decisions $B$) $\rightarrow$ Bool
@@
 }
-\medskip
+\bigskip
 \nl $\FuncSty{propagate}$ (Domains $D$) $\rightarrow$ Bool \;
@@
 For value ordering, which occurs on \siplineref{valueordering} of \cref{algorithm:sip}, we will
-compare four strategies:
+compare five strategies:
 
@@
 
-\paragraph{Tailored} We select vertices from highest degree to lowest degree. This is done by
+\paragraph{Degree} We select vertices from highest degree to lowest degree. This is done by
 permuting the target graph at the top of search. This heuristic is the current state of the art.
@@
 \paragraph{Anti} We select vertices from lowest degree to highest degree. This heuristic is used as
-a sanity check: we expect tailored to beat random, and random to beat anti.
+a sanity check: we expect degree to beat random, and random to beat anti.
 
-\paragraph{Biased} As with the random heuristic, we shuffle the vertices before we enter the
-$\KwSty{foreach}$ loop. Instead of picking a vertex uniformly at random, we use a weight of
+\paragraph{Position-biased} We permute the graph at the top of search, as in the degree heuristic.
+Then, as with the random heuristic, we shuffle the vertices before we enter the $\KwSty{foreach}$
+loop. Instead of picking a vertex uniformly at random, we select select the first (that is, highest
+degree) vertex with probability 0.5, the second vertex with probability 0.25, and the $n$th with
+probability $2^{-n}$.  This heuristic heavily prefers vertices of high degree, similar to the
+degree heuristic, but also introduces an element of randomness, and so can be used with restarts.
+
+\paragraph{Degree-biased} Refining position-biased, we shuffle the vertices, using a weight of
 $2^{\left|\neighbourhood(v)\right|}$ when deciding whether to branch on vertex $v$ next. (We cheat
@@
 every vertex weight at least 1.  This is a bit of hack but programming is hard.) This heuristic
-heavily prefers vertices of high degree, similar to the tailored heuristic, but will give the same
-chance of being selected next to two vertices of equal degree. This heuristic also introduces an
-element of randomness, and so can be used with restarts.
+continues to prefer vertices of high degree, but will give the same chance of being selected next to
+two vertices of equal degree.
 
@@
 In \cref{figure:value-ordering-heuristics} we show the cumulative number of instances solved over
-time using the four different value-ordering heuristics. As expected, the tailored heuristic is much
+time using the four different value-ordering heuristics. As expected, the degree heuristic is much
 better (?? give aggregate speedups for both all and just satisfiable instances) than random
@@
 
-    \caption{An instance by instance comparison of the tailored and biased heuristics.}
+    \caption{An instance by instance comparison of the degree and degree-biased heuristics.}
     \label{figure:scatter-heuristics}
@@
 
-The performance of the biased and tailored heuristics is very similar: the biased heuristic has
+?? Also talk about position-biased
+
+The performance of the degree-biased and degree heuristics is very similar: the degree-biased heuristic has
 slightly worse performance below a 100s timeout, and effectively equal performance above it. This
-shows that we can introduce an element of randomness into the tailored value-ordering heuristic
+shows that we can introduce an element of randomness into the degree value-ordering heuristic
 without adversely affecting its performance \emph{in aggregate}. \Cref{figure:scatter-heuristics}
 gives a more detailed look at these results. In this plot, satisfiable and unsatisfiable instances
-are shown using dots and crosses respectively. On the $x$-axis we measure the tailored heuristic,
+are shown using dots and crosses respectively. On the $x$-axis we measure the degree heuristic,
 whilst on the $y$-axis we measure random + restarts. Points on the outer axes represent instances
@@
 In \cref{figure:restarts} we show the effects of adding restarts to the algorithm. With restarts,
-the random value-ordering heuristic comfortably beats the tailored strategy without restarts. In
+the random value-ordering heuristic comfortably beats the degree strategy without restarts. In
 other words, although having a good variable-ordering heuristic is beneficial, introducing
@@
     \includegraphics*{gen-graph-scatter-random.pdf}
-    \caption{Instance by instance comparison of the basic algorithm with the tailored heuristic,
+    \caption{Instance by instance comparison of the basic algorithm with the degree heuristic,
     versus random + restarts.}
@@
 
-    \caption{Instance by instance comparison of the basic algorithm with the tailored heuristic,
-    versus biased + restarts.}
+    \caption{Instance by instance comparison of the basic algorithm with the degree heuristic,
+    versus degree-biased + restarts.}
     \label{figure:scatter}
@@
 
-\Cref{figure:restarts} shows that the biased heuristic together with restarts is better
+\Cref{figure:restarts} shows that either biased heuristic together with restarts is better
 still---that is, if we are introducing restarts, then it is better to add a small amount of
 randomness to a tailored deterministic heuristic than it is to throw away the heuristic altogether.
-In the more detailed view in \cref{figure:scatter},
-nearly all of the unsatisfiable instances are on or slightly above the $x-y$ diagonal, showing that
-their performance is either unchanged or reduced slightly. On the other hand, there are ?? large
-numbers of satisfiable instances well below the diagonal line, indicating large speedups. Better
-yet, there are only ?? satisfiable instances that are more than a factor of ?? times worse.
+The degree-biased heuristics is stronger than position-biased, showing that exploiting ties and
+relative sizes in the heuristic ordering is also worthwhile.
+
+In the more detailed view in \cref{figure:scatter}, comparing the basic algorithm with the degree
+heuristic to the degree-biased algorithm with restarts, nearly all of the unsatisfiable instances
+are on the $x-y$ diagonal, showing that their performance is unchanged. On the other hand, there are
+?? large numbers of satisfiable instances well below the diagonal line, indicating large speedups.
+Better yet, there are only ?? satisfiable instances that are more than a factor of ?? times worse.
 
@@
 
-?? maybe some engineering aspects: look at performance with restarts on tailored, and at restarts
+?? maybe some engineering aspects: look at performance with restarts on degree, and at restarts
 without nogood recording

### 2017-10-04 code/unit.cc
Track and store decisions explicitly

@@ -252,8 +252,6 @@ namespace
 
         auto propagate(Domains & new_domains, Assignments & assignments) -> bool
         {
-            bool this_is_a_decision = true;
-
             // whilst we've got a unit domain...
             for (typename Domains::iterator branch_domain = find_unit_domain(new_domains) ;
                     branch_domain != new_domains.end() ;
@@ -263,8 +261,7 @@ namespace
 
                 // ok, make the assignment
                 branch_domain->fixed = true;
-                assignments.values.push_back({ { current_assignment.first, current_assignment.second }, this_is_a_decision });
-                this_is_a_decision = false;
+                assignments.values.push_back({ { current_assignment.first, current_assignment.second }, false });
 
                 // propagate watches
                 if (params.restarts) {
@@ -402,6 +399,7 @@ namespace
                 remaining.unset(f_v);
 
                 auto assignments_size = assignments.values.size();
+                assignments.values.push_back({ { branch_domain->v, f_v }, true });
 
                 /* set up new domains */
                 Domains new_domains = prepare_domains(domains, branch_domain->v, f_v);
@@ -525,6 +523,9 @@ namespace
                 // modified in-place by appending, we can restore by shrinking
                 auto assignments_size = assignments.values.size();
 
+                // make the assignment
+                assignments.values.push_back({ { branch_domain->v, *f_v }, true });
+
                 // set up new domains
                 Domains new_domains = prepare_domains(domains, branch_domain->v, *f_v);
 
@@ -599,6 +600,8 @@ namespace
                         || (1 == discrepancies_allowed && 0 != count)
                         || (discrepancies_allowed > 1)) {
                     auto assignments_size = assignments.values.size();
+
+                    assignments.values.push_back({ { branch_domain->v, f_v }, true });
 
                     /* set up new domains */
                     Domains new_domains = prepare_domains(domains, branch_domain->v, f_v);

### 2017-10-04 code/unit.cc
Move around

@@ -804,48 +804,50 @@ namespace
 
             Assignments assignments;
             assignments.values.reserve(pattern_size);
-            if (propagate(domains, assignments)) {
-                if (params.dds) {
+            if (params.dds) {
+                if (propagate(domains, assignments)) {
                     for (unsigned discrepancies_allowed = 0 ; discrepancies_allowed <= pattern_size ; ++discrepancies_allowed)
                         if (dds_search(assignments, domains, result.nodes, 0, discrepancies_allowed) == Search::Satisfiable) {
                             save_result(assignments, result);
                             break;
                         }
                 }
-                else if (params.restarts) {
-                    bool done = false;
-                    std::list<long long> luby = {{ 1 }};
-                    auto current_luby = luby.begin();
-                    while (! done) {
-                        long long backtracks_until_restart = *current_luby * params.luby_multiplier;
-                        if (std::next(current_luby) == luby.end()) {
-                            luby.insert(luby.end(), luby.begin(), luby.end());
-                            luby.push_back(*luby.rbegin() * 2);
+            }
+            else if (params.restarts) {
+                bool done = false;
+                std::list<long long> luby = {{ 1 }};
+                auto current_luby = luby.begin();
+                while (! done) {
+                    long long backtracks_until_restart = *current_luby * params.luby_multiplier;
+                    if (std::next(current_luby) == luby.end()) {
+                        luby.insert(luby.end(), luby.begin(), luby.end());
+                        luby.push_back(*luby.rbegin() * 2);
+                    }
+                    ++current_luby;
+
+                    auto assignments_copy = assignments;
+
+                    // start watching new nogoods. we're not backjumping so this is a bit icky.
+                    for (auto & n : need_to_watch) {
+                        if (n->literals.empty()) {
+                            done = true;
+                            break;
                         }
-                        ++current_luby;
-
-                        auto assignments_copy = assignments;
-
-                        // start watching new nogoods. we're not backjumping so this is a bit icky.
-                        for (auto & n : need_to_watch) {
-                            if (n->literals.empty()) {
-                                done = true;
-                                break;
-                            }
-                            else if (1 == n->literals.size()) {
-                                for (auto & d : domains)
-                                    if (d.v == n->literals[0].first) {
-                                        d.values.unset(n->literals[0].second);
-                                        d.popcount = d.values.popcount();
-                                    }
-                            }
-                            else {
-                                watches[n->literals[0]].push_back(n);
-                                watches[n->literals[1]].push_back(n);
[…]

### 2017-10-04 code/solve_subgraph_isomorphism.cc
Be a bit noisier

@@ -210,7 +210,7 @@ auto main(int argc, char * argv[]) -> int
                 for (int j = 0 ; j < graphs.first.size() ; ++j) {
                     if (graphs.first.adjacent(i, j)) {
                         if (! graphs.second.adjacent(result.isomorphism.find(i)->second, result.isomorphism.find(j)->second)) {
-                            std::cerr << "Oops! not an isomorphism" << std::endl;
+                            std::cerr << "Oops! not an isomorphism: " << i << ", " << j << std::endl;
                             return EXIT_FAILURE;
                         }
                     }

### 2017-10-04 code/unit.cc
Put this in its proper place

@@ -825,8 +825,6 @@ namespace
                     }
                     ++current_luby;
 
-                    auto assignments_copy = assignments;
-
                     // start watching new nogoods. we're not backjumping so this is a bit icky.
                     for (auto & n : need_to_watch) {
                         if (n->literals.empty()) {
@@ -848,6 +846,8 @@ namespace
                     need_to_watch.clear();
 
                     if (propagate(domains, assignments)) {
+                        auto assignments_copy = assignments;
+
                         switch (restarting_search(assignments_copy, domains, result.nodes, 0, backtracks_until_restart)) {
                             case RestartingSearch::Satisfiable:
                                 save_result(assignments_copy, result);

### 2017-10-04 code/unit.cc
Uh, oops

@@ -863,6 +863,8 @@ namespace
                                 break;
                         }
                     }
+                    else
+                        done = true;
                 }
             }
             else if (params.shuffle || params.biased_shuffle || params.position_shuffle) {

### 2017-10-04 paper/paper.tex
Thingies

@@
 
-\definecolor{magma8}{rgb}{0.996078, 0.760784, 0.529412}
+\definecolor{inferno9}{rgb}{0.992157, 0.929412, 0.501961}
 
@@
         \coordinate (restart1ec) at ($(pic cs:restart1e) + (0, 0.00)$);
-        \node [fill=magma8, rounded corners, fit=(restart1sc) (restart1ec)] { };
+        \node [fill=inferno9, rounded corners, fit=(restart1sc) (restart1ec)] { };
         \coordinate (restart2sc) at ($(pic cs:restart2s) + (0, 0.15)$);
         \coordinate (restart2ec) at ($(pic cs:restart2e) + (0, 0.00)$);
-        \node [fill=magma8, rounded corners, fit=(restart2sc) (restart2ec)] { };
+        \node [fill=inferno9, rounded corners, fit=(restart2sc) (restart2ec)] { };
         \coordinate (restart3sc) at ($(pic cs:restart3s) + (0, 0.15)$);
         \coordinate (restart3ec) at ($(pic cs:restart3e) + (0, 0.00)$);
-        \node [fill=magma8, rounded corners, fit=(restart3sc) (restart3ec)] { };
+        \node [fill=inferno9, rounded corners, fit=(restart3sc) (restart3ec)] { };
         \coordinate (restart4sc) at ($(pic cs:restart4s) + (0, 0.15)$);
         \coordinate (restart4ec) at ($(pic cs:restart4e) + (0, 0.00)$);
-        \node [fill=magma8, rounded corners, fit=(restart4sc) (restart4ec)] { };
+        \node [fill=inferno9, rounded corners, fit=(restart4sc) (restart4ec)] { };
         \coordinate (restart5sc) at ($(pic cs:restart5s) + (0, 0.15)$);
@@
         \coordinate (restart5rc) at ($(pic cs:restart5r) + (0, 0.00)$);
-        \node [fill=magma8, rounded corners, fit=(restart5sc) (restart5ec) (restart5rc)] { };
+        \node [fill=inferno9, rounded corners, fit=(restart5sc) (restart5ec) (restart5rc)] { };
         \coordinate (restart6sc) at ($(pic cs:restart6s) + (0, 0.15)$);
         \coordinate (restart6ec) at ($(pic cs:restart6e) + (0, 0.00)$);
-        \node [fill=magma8, rounded corners, fit=(restart6sc) (restart6ec)] { };
+        \node [fill=inferno9, rounded corners, fit=(restart6sc) (restart6ec)] { };
         \coordinate (restart7sc) at ($(pic cs:restart7s) + (0, 0.15)$);
         \coordinate (restart7ec) at ($(pic cs:restart7e) + (0, 0.00)$);
-        \node [fill=magma8, rounded corners, fit=(restart7sc) (restart7ec)] { };
+        \node [fill=inferno9, rounded corners, fit=(restart7sc) (restart7ec)] { };
         \coordinate (restart8sc) at ($(pic cs:restart8s) + (0.05, 0.15)$);
         \coordinate (restart8ec) at ($(pic cs:restart8e) + (0, 0.00)$);
-        \node [fill=magma8, rounded corners, fit=(restart8sc) (restart8ec)] { };
+        \node [fill=inferno9, rounded corners, fit=(restart8sc) (restart8ec)] { };
         \coordinate (restart9sc) at ($(pic cs:restart9s) + (0, 0.15)$);
         \coordinate (restart9ec) at ($(pic cs:restart9e) + (-0.15, 0.00)$);
-        \node [fill=magma8, rounded corners, fit=(restart9sc) (restart9ec)] { };
+        \node [fill=inferno9, rounded corners, fit=(restart9sc) (restart9ec)] { };
     \end{tikzpicture}
@@
 
-?? Also talk about position-biased
-
-The performance of the degree-biased and degree heuristics is very similar: the degree-biased heuristic has
-slightly worse performance below a 100s timeout, and effectively equal performance above it. This
-shows that we can introduce an element of randomness into the degree value-ordering heuristic
-without adversely affecting its performance \emph{in aggregate}. \Cref{figure:scatter-heuristics}
-gives a more detailed look at these results. In this plot, satisfiable and unsatisfiable instances
-are shown using dots and crosses respectively. On the $x$-axis we measure the degree heuristic,
-whilst on the $y$-axis we measure random + restarts. Points on the outer axes represent instances
-which timed out with one heuristic but not the other. This plot shows that on a case by case basis,
-randomness can make a large difference to individual satisfiable instances.
+The performance of both the biased heuristics and the degree heuristic are very similar: the
+degree-biased heuristic has slightly worse performance below a 100s timeout, and effectively equal
+performance above it. This shows that we can introduce an element of randomness into the degree
+value-ordering heuristic without adversely affecting its performance \emph{in aggregate}.
+\Cref{figure:scatter-heuristics} gives a more detailed look at these results. In this plot,
+satisfiable and unsatisfiable instances are shown using dots and crosses respectively. On the
+$x$-axis we measure the degree heuristic, whilst on the $y$-axis we measure random + restarts.
+Points on the outer axes represent instances which timed out with one heuristic but not the other.
+This plot shows that on a case by case basis, randomness can make a large difference to individual
+satisfiable instances.
 
@@
 
-    \caption{Cumulative number of instances solved over time, with and without restarts.}
+    \caption{Cumulative number of instances solved over time, with (dotted lines) and without (solid
+    lines) restarts.}
     \label{figure:restarts}
@@
 randomness to a tailored deterministic heuristic than it is to throw away the heuristic altogether.
-The degree-biased heuristics is stronger than position-biased, showing that exploiting ties and
-relative sizes in the heuristic ordering is also worthwhile.
+The degree-biased heuristic is also stronger than position-biased, showing that exploiting ties and
+relative sizes in the heuristic ordering is also worthwhile. Interestingly, the gap between degree-
+and position-based is much larger when using restarts than it is when not using them.
 
@@
 ?? maybe some engineering aspects: look at performance with restarts on degree, and at restarts
-without nogood recording
+without nogood recording, and what happens without nogoods in \cref{figure:scatter-nogoods}.
+
+\begin{figure}[tb]
+    \centering
+    \includegraphics*{gen-graph-scatter-nogoods.pdf}
+
+    \caption{Is nogood recording important?}
+    \label{figure:scatter-nogoods}
+\end{figure}
+
+\begin{figure}[tb]
+    \centering
+    \includegraphics*{gen-graph-scatter-overheads.pdf}
+
+    \caption{Is nogood recording and restarting expensive?}
+    \label{figure:scatter-overhead}
+\end{figure}
 

### 2017-10-09 paper/paper.tex
hyperref is annoying me
Plottery

@@
 
-\usepackage{hyperref}
 \usepackage{cleveref}
@@
 
-\definecolor{inferno9}{rgb}{0.992157, 0.929412, 0.501961}
+\definecolor{inferno7}{rgb}{0.976471, 0.584314, 0.039216}
 
@@
         \coordinate (restart1ec) at ($(pic cs:restart1e) + (0, 0.00)$);
-        \node [fill=inferno9, rounded corners, fit=(restart1sc) (restart1ec)] { };
+        \node [fill=inferno7, rounded corners, fit=(restart1sc) (restart1ec)] { };
         \coordinate (restart2sc) at ($(pic cs:restart2s) + (0, 0.15)$);
         \coordinate (restart2ec) at ($(pic cs:restart2e) + (0, 0.00)$);
-        \node [fill=inferno9, rounded corners, fit=(restart2sc) (restart2ec)] { };
+        \node [fill=inferno7, rounded corners, fit=(restart2sc) (restart2ec)] { };
         \coordinate (restart3sc) at ($(pic cs:restart3s) + (0, 0.15)$);
         \coordinate (restart3ec) at ($(pic cs:restart3e) + (0, 0.00)$);
-        \node [fill=inferno9, rounded corners, fit=(restart3sc) (restart3ec)] { };
+        \node [fill=inferno7, rounded corners, fit=(restart3sc) (restart3ec)] { };
         \coordinate (restart4sc) at ($(pic cs:restart4s) + (0, 0.15)$);
         \coordinate (restart4ec) at ($(pic cs:restart4e) + (0, 0.00)$);
-        \node [fill=inferno9, rounded corners, fit=(restart4sc) (restart4ec)] { };
+        \node [fill=inferno7, rounded corners, fit=(restart4sc) (restart4ec)] { };
         \coordinate (restart5sc) at ($(pic cs:restart5s) + (0, 0.15)$);
@@
         \coordinate (restart5rc) at ($(pic cs:restart5r) + (0, 0.00)$);
-        \node [fill=inferno9, rounded corners, fit=(restart5sc) (restart5ec) (restart5rc)] { };
+        \node [fill=inferno7, rounded corners, fit=(restart5sc) (restart5ec) (restart5rc)] { };
         \coordinate (restart6sc) at ($(pic cs:restart6s) + (0, 0.15)$);
         \coordinate (restart6ec) at ($(pic cs:restart6e) + (0, 0.00)$);
-        \node [fill=inferno9, rounded corners, fit=(restart6sc) (restart6ec)] { };
+        \node [fill=inferno7, rounded corners, fit=(restart6sc) (restart6ec)] { };
         \coordinate (restart7sc) at ($(pic cs:restart7s) + (0, 0.15)$);
         \coordinate (restart7ec) at ($(pic cs:restart7e) + (0, 0.00)$);
-        \node [fill=inferno9, rounded corners, fit=(restart7sc) (restart7ec)] { };
+        \node [fill=inferno7, rounded corners, fit=(restart7sc) (restart7ec)] { };
         \coordinate (restart8sc) at ($(pic cs:restart8s) + (0.05, 0.15)$);
         \coordinate (restart8ec) at ($(pic cs:restart8e) + (0, 0.00)$);
-        \node [fill=inferno9, rounded corners, fit=(restart8sc) (restart8ec)] { };
+        \node [fill=inferno7, rounded corners, fit=(restart8sc) (restart8ec)] { };
         \coordinate (restart9sc) at ($(pic cs:restart9s) + (0, 0.15)$);
         \coordinate (restart9ec) at ($(pic cs:restart9e) + (-0.15, 0.00)$);
-        \node [fill=inferno9, rounded corners, fit=(restart9sc) (restart9ec)] { };
+        \node [fill=inferno7, rounded corners, fit=(restart9sc) (restart9ec)] { };
     \end{tikzpicture}
@@
 In the more detailed view in \cref{figure:scatter}, comparing the basic algorithm with the degree
-heuristic to the degree-biased algorithm with restarts, nearly all of the unsatisfiable instances
+heuristic to the degree-biased algorithm with restarts, all of the unsatisfiable instances
 are on the $x-y$ diagonal, showing that their performance is unchanged. On the other hand, there are
@@
 ?? This has made up the consistency we lost from introducing randomness.
-
-?? Figure out what's going on with the unsat instances not on the line. There are a few options.
-Could be SDF picking something else, could be the hacky all-diff missing something, could be how
-we're doing unit propagation, could be a horrible bug.
 
@@
     \label{figure:scatter-overhead}
+\end{figure}
+
+\begin{figure}[tb]
+    \centering
+    \includegraphics*{gen-graph-restart-policies.pdf}
+
+    \caption{Different restart policies}
+    \label{figure:restart-policies}
+\end{figure}
+
+\begin{figure}[tb]
+    \centering
+    \includegraphics*{gen-graph-scatter-geometric.pdf}
+
+    \caption{Geometric instead of Luby}
+    \label{figure:scatter-geometric}
 \end{figure}

### 2017-10-22 paper/paper.tex
Useful things

@@
 
-?? do a more detailed breakdown by family
+\begin{figure}[tb]
+    \centering
+    \includegraphics*{gen-graph-scatter-by-family.pdf}
 
-?? maybe some engineering aspects: look at performance with restarts on degree, and at restarts
-without nogood recording, and what happens without nogoods in \cref{figure:scatter-nogoods}.
+    \caption{Instance by instance comparison of the basic algorithm with the degree heuristic,
+    versus degree-biased + restarts, showing only satisfiable instances, on a family by family
+    basis.}
+    \label{figure:scatter-by-family}
+\end{figure}
 
-\begin{figure}[tb]
+?? do a more detailed breakdown by family, \cref{figure:scatter-by-family}
+
+?? maybe some engineering aspects: look at what happens without nogoods in
+\cref{figure:scatter-nogoods}, and at performance with restarts on degree in
+\cref{figure:scatter-overheads}.
+
+\begin{figure}[p]
     \centering
@@
 
-\begin{figure}[tb]
+\begin{figure}[p]
     \centering
@@
 
-    \caption{Is nogood recording and restarting expensive?}
-    \label{figure:scatter-overhead}
+    \caption{Is nogood recording expensive?}
+    \label{figure:scatter-overheads}
 \end{figure}
 
-\begin{figure}[tb]
+\begin{figure}[p]
     \centering
@@
 
-    \caption{Different restart policies}
+    \caption{Different restart policies.}
     \label{figure:restart-policies}
-\end{figure}
-
-\begin{figure}[tb]
-    \centering
-    \includegraphics*{gen-graph-scatter-geometric.pdf}
-
-    \caption{Geometric instead of Luby}
-    \label{figure:scatter-geometric}
 \end{figure}

### 2017-10-22 code/fixed_bit_set.cc
Refactorification

@@ -1,6 +1,6 @@
 /* vim: set sw=4 sts=4 et foldmethod=syntax : */
 
-#include "bit_graph.hh"
+#include "fixed_bit_set.hh"
 
 GraphTooBig::GraphTooBig() throw () = default;
 

### 2017-10-22 code/unit.cc

@@ -1,7 +1,7 @@
 /* vim: set sw=4 sts=4 et foldmethod=syntax : */
 
 #include "unit.hh"
-#include "bit_graph.hh"
+#include "fixed_bit_set.hh"
 #include "template_voodoo.hh"
 
 #include <algorithm>
@@ -137,8 +137,10 @@ namespace
         unsigned pattern_size, full_pattern_size, target_size;
 
         static constexpr int max_graphs = 5;
-        vector<FixedBitGraph<n_words_> > target_graphs;
-        vector<FixedBitGraph<n_words_> > pattern_graphs;
+
+        vector<uint8_t> pattern_adjacencies_bits;
+        vector<FixedBitSet<n_words_> > pattern_graph_rows;
+        vector<FixedBitSet<n_words_> > target_graph_rows;
 
         vector<int> pattern_permutation, target_permutation, isolated_vertices;
         vector<vector<int> > patterns_degrees, targets_degrees;
@@ -156,8 +158,6 @@ namespace
             pattern_size(pattern.size()),
             full_pattern_size(pattern.size()),
             target_size(target.size()),
-            target_graphs(max_graphs),
-            pattern_graphs(max_graphs),
             target_permutation(target.size()),
             patterns_degrees(max_graphs),
             targets_degrees(max_graphs)
@@ -172,11 +172,11 @@ namespace
                     pattern_permutation.push_back(v);
 
             // recode pattern to a bit graph
-            pattern_graphs.at(0).resize(pattern_size);
+            pattern_graph_rows.resize(pattern_size * max_graphs);
             for (unsigned i = 0 ; i < pattern_size ; ++i)
                 for (unsigned j = 0 ; j < pattern_size ; ++j)
                     if (pattern.adjacent(pattern_permutation.at(i), pattern_permutation.at(j)))
-                        pattern_graphs.at(0).add_edge(i, j);
+                        pattern_graph_rows[i * max_graphs + 0].set(j);
 
             // determine ordering for target graph vertices
             iota(target_permutation.begin(), target_permutation.end(), 0);
@@ -189,20 +189,20 @@ namespace
                 degree_sort(target, target_permutation, params.antiheuristic);
 
             // recode target to a bit graph
-            target_graphs.at(0).resize(target_size);
+            target_graph_rows.resize(target_size * max_graphs);
             for (unsigned i = 0 ; i < target_size ; ++i)
-                for (unsigned j = i ; j < target_size ; ++j)
+                for (unsigned j = 0 ; j < target_size ; ++j)
[…]

### 2017-10-22 code/unit.cc
Move around

@@ -597,20 +597,6 @@ namespace
 
         auto initialise_domains(Domains & domains) -> bool
         {
-            for (int g = 0 ; g < max_graphs ; ++g) {
-                patterns_degrees.at(g).resize(pattern_size);
-                targets_degrees.at(g).resize(target_size);
-            }
-
-            /* pattern and target degree sequences */
-            for (int g = 0 ; g < max_graphs ; ++g) {
-                for (unsigned i = 0 ; i < pattern_size ; ++i)
-                    patterns_degrees.at(g).at(i) = pattern_graph_rows[i * max_graphs + g].popcount();
-
-                for (unsigned i = 0 ; i < target_size ; ++i)
-                    targets_degrees.at(g).at(i) = target_graph_rows[i * max_graphs + g].popcount();
-            }
-
             /* pattern and target neighbourhood degree sequences */
             vector<vector<vector<int> > > patterns_ndss(max_graphs);
             vector<vector<vector<int> > > targets_ndss(max_graphs);
@@ -776,6 +762,21 @@ namespace
             build_supplemental_graphs(pattern_graph_rows, pattern_size);
             build_supplemental_graphs(target_graph_rows, target_size);
 
+            // pattern and target degrees, including supplemental graphs
+            for (int g = 0 ; g < max_graphs ; ++g) {
+                patterns_degrees.at(g).resize(pattern_size);
+                targets_degrees.at(g).resize(target_size);
+            }
+
+            for (int g = 0 ; g < max_graphs ; ++g) {
+                for (unsigned i = 0 ; i < pattern_size ; ++i)
+                    patterns_degrees.at(g).at(i) = pattern_graph_rows[i * max_graphs + g].popcount();
+
+                for (unsigned i = 0 ; i < target_size ; ++i)
+                    targets_degrees.at(g).at(i) = target_graph_rows[i * max_graphs + g].popcount();
+            }
+
+            // pattern adjacencies, compressed
             pattern_adjacencies_bits.resize(pattern_size * pattern_size);
             for (int g = 0 ; g < max_graphs ; ++g)
                 for (unsigned i = 0 ; i < pattern_size ; ++i)
@@ -783,15 +784,19 @@ namespace
                         if (pattern_graph_rows[i * max_graphs + g].test(j))
                             pattern_adjacencies_bits[i * pattern_size + j] |= (1u << g);
 
+            // domains
             Domains domains(pattern_size);
-
             if (! initialise_domains(domains))
                 return result;
 
[…]

### 2017-10-22 code/unit.cc
Make max_graphs not a constant

@@ -133,10 +133,9 @@ namespace
         using Domains = vector<Domain>;
 
         const Params & params;
+        const int max_graphs;
 
         unsigned pattern_size, full_pattern_size, target_size;
-
-        static constexpr int max_graphs = 5;
 
         vector<uint8_t> pattern_adjacencies_bits;
         vector<FixedBitSet<n_words_> > pattern_graph_rows;
@@ -155,6 +154,7 @@ namespace
 
         SIP(const Graph & target, const Graph & pattern, const Params & a) :
             params(a),
+            max_graphs(5 + (params.induced ? 1 : 0)),
             pattern_size(pattern.size()),
             full_pattern_size(pattern.size()),
             target_size(target.size()),
@@ -304,6 +304,11 @@ namespace
             return true;
         }
 
+        // The max_graphs_ template parameter is so that the for each graph
+        // pair loop gets unrolled, which makes an annoyingly large difference
+        // to performance. Note that for larger target graphs, half of the
+        // total runtime is spent in the middle bit of this function.
+        template <int max_graphs_>
         auto propagate_simple_constraints(Domains & new_domains, const Assignment & current_assignment) -> bool
         {
             // propagate for each remaining domain...
@@ -317,11 +322,11 @@ namespace
                 auto pattern_adjacency_bits = pattern_adjacencies_bits[pattern_size * current_assignment.first + d.v];
 
                 // for each graph pair...
-                for (int g = 0 ; g < max_graphs ; ++g) {
+                for (int g = 0 ; g < max_graphs_ ; ++g) {
                     // if we're adjacent...
                     if (pattern_adjacency_bits & (1u << g)) {
                         // ...then we can only be mapped to adjacent vertices
-                        d.values.intersect_with(target_graph_rows[current_assignment.second * max_graphs + g]);
+                        d.values.intersect_with(target_graph_rows[current_assignment.second * max_graphs_ + g]);
                     }
                 }
 
@@ -353,8 +358,20 @@ namespace
                         return false;
 
                 // propagate simple all different and adjacency
-                if (! propagate_simple_constraints(new_domains, current_assignment))
-                    return false;
+                switch (max_graphs) {
+                    case 5:
+                        if (! propagate_simple_constraints<5>(new_domains, current_assignment))
+                            return false;
+                        break;
+
+                    case 6:
+                        if (! propagate_simple_constraints<6>(new_domains, current_assignment))
+                            return false;
+                        break;
+
+                    default:
[…]

### 2017-10-23 code/unit.cc
Cleaner

@@ -307,8 +307,22 @@ namespace
         // The max_graphs_ template parameter is so that the for each graph
         // pair loop gets unrolled, which makes an annoyingly large difference
         // to performance. Note that for larger target graphs, half of the
-        // total runtime is spent in the middle bit of this function.
+        // total runtime is spent in this function.
         template <int max_graphs_>
+        auto propagate_adjacency_constraints(Domain & d, const Assignment & current_assignment) -> void
+        {
+            auto pattern_adjacency_bits = pattern_adjacencies_bits[pattern_size * current_assignment.first + d.v];
+
+            // for each graph pair...
+            for (int g = 0 ; g < max_graphs_ ; ++g) {
+                // if we're adjacent...
+                if (pattern_adjacency_bits & (1u << g)) {
+                    // ...then we can only be mapped to adjacent vertices
+                    d.values.intersect_with(target_graph_rows[current_assignment.second * max_graphs_ + g]);
+                }
+            }
+        }
+
         auto propagate_simple_constraints(Domains & new_domains, const Assignment & current_assignment) -> bool
         {
             // propagate for each remaining domain...
@@ -319,15 +333,13 @@ namespace
                 // all different
                 d.values.unset(current_assignment.second);
 
-                auto pattern_adjacency_bits = pattern_adjacencies_bits[pattern_size * current_assignment.first + d.v];
+                // adjacency
+                switch (max_graphs) {
+                    case 5: propagate_adjacency_constraints<5>(d, current_assignment); break;
+                    case 6: propagate_adjacency_constraints<6>(d, current_assignment); break;
 
-                // for each graph pair...
-                for (int g = 0 ; g < max_graphs_ ; ++g) {
-                    // if we're adjacent...
-                    if (pattern_adjacency_bits & (1u << g)) {
-                        // ...then we can only be mapped to adjacent vertices
-                        d.values.intersect_with(target_graph_rows[current_assignment.second * max_graphs_ + g]);
-                    }
+                    default:
+                        throw "you forgot to update the ugly max_graphs hack";
                 }
 
                 // we might have removed values
@@ -358,20 +370,8 @@ namespace
                         return false;
 
                 // propagate simple all different and adjacency
-                switch (max_graphs) {
-                    case 5:
-                        if (! propagate_simple_constraints<5>(new_domains, current_assignment))
-                            return false;
-                        break;
-
-                    case 6:
-                        if (! propagate_simple_constraints<6>(new_domains, current_assignment))
-                            return false;
[…]

### 2017-10-23 code/solve_subgraph_isomorphism.cc
Induced

@@ -107,6 +107,7 @@ auto main(int argc, char * argv[]) -> int
             ("help",                                         "Display help information")
             ("timeout",            po::value<int>(),         "Abort after this many seconds")
             ("format",             po::value<string>(),      "Specify the format of the input")
+            ("induced",                                      "Solve the induced version")
             ("restarts",                                     "Use restarts")
             ("shuffle",                                      "Use shuffling")
             ("biased-shuffle",                               "Use biased shuffling")
@@ -174,6 +175,7 @@ auto main(int argc, char * argv[]) -> int
         /* Figure out what our options should be. */
         Params params;
 
+        params.induced = options_vars.count("induced");
         params.restarts = options_vars.count("restarts");
         params.shuffle = options_vars.count("shuffle");
         params.biased_shuffle = options_vars.count("biased-shuffle");
@@ -224,8 +226,9 @@ auto main(int argc, char * argv[]) -> int
         if (! result.isomorphism.empty()) {
             for (int i = 0 ; i < graphs.first.size() ; ++i) {
                 for (int j = 0 ; j < graphs.first.size() ; ++j) {
-                    if (graphs.first.adjacent(i, j)) {
-                        if (! graphs.second.adjacent(result.isomorphism.find(i)->second, result.isomorphism.find(j)->second)) {
+                    if (params.induced || graphs.first.adjacent(i, j)) {
+                        if (graphs.first.adjacent(i, j) !=
+                                graphs.second.adjacent(result.isomorphism.find(i)->second, result.isomorphism.find(j)->second)) {
                             cerr << "Oops! not an isomorphism: " << i << ", " << j << endl;
                             return EXIT_FAILURE;
                         }

### 2017-10-23 code/unit.cc

@@ -162,9 +162,9 @@ namespace
             patterns_degrees(max_graphs),
             targets_degrees(max_graphs)
         {
-            // strip out isolated vertices in the pattern
+            // strip out isolated vertices in the pattern, and build pattern_permutation
             for (unsigned v = 0 ; v < full_pattern_size ; ++v)
-                if (0 == pattern.degree(v)) {
+                if ((! params.induced) && (0 == pattern.degree(v))) {
                     isolated_vertices.push_back(v);
                     --pattern_size;
                 }
@@ -242,6 +242,14 @@ namespace
                     }
                 }
             }
+        }
+
+        auto build_complement_graphs(vector<FixedBitSet<n_words_> > & graph_rows, unsigned size) -> void
+        {
+            for (unsigned v = 0 ; v < size ; ++v)
+                for (unsigned w = 0 ; w < size ; ++w)
+                    if (! graph_rows[v * max_graphs + 0].test(w))
+                        graph_rows[v * max_graphs + 5].set(w);
         }
 
         auto find_unit_domain(Domains & domains) -> typename Domains::iterator
@@ -778,6 +786,12 @@ namespace
 
             build_supplemental_graphs(pattern_graph_rows, pattern_size);
             build_supplemental_graphs(target_graph_rows, target_size);
+
+            // build complement graphs
+            if (params.induced) {
+                build_complement_graphs(pattern_graph_rows, pattern_size);
+                build_complement_graphs(target_graph_rows, target_size);
+            }
 
             // pattern and target degrees, including supplemental graphs
             for (int g = 0 ; g < max_graphs ; ++g) {

### 2017-11-03 code/kless.cc
No kless for now

@@ -1,863 +1,0 @@
-/* vim: set sw=4 sts=4 et foldmethod=syntax : */
-
-#include "sip.hh"
-#include "kless.hh"
-
-#include <algorithm>
-#include <chrono>
-#include <functional>
-#include <limits>
-#include <list>
-#include <map>
-#include <numeric>
-#include <random>
-#include <tuple>
-#include <utility>
-#include <vector>
-
-#include <iostream>
-
-#include <boost/dynamic_bitset.hpp>
-
-using std::find_if;
-using std::get;
-using std::greater;
-using std::list;
-using std::make_pair;
-using std::map;
-using std::mt19937;
-using std::numeric_limits;
-using std::pair;
-using std::to_string;
-using std::tuple;
-using std::uniform_int_distribution;
-using std::unique;
-using std::vector;
-
-using std::chrono::duration_cast;
-using std::chrono::milliseconds;
-using std::chrono::steady_clock;
-
-using std::cerr;
-using std::endl;
-
-namespace
-{
-    /// We'll use an array of unsigned long longs to represent our bits.
-    using BitWord = unsigned long long;
-
-    /// Number of bits per word.
-    static const constexpr int bits_per_word = sizeof(BitWord) * 8;
-
-    /**
-     * A bitset with a fixed maximum size. This only provides the operations
-     * we actually use in the bitset algorithms: it's more readable this way
-     * than doing all the bit voodoo inline.
-     *
-     * Indices start at 0.
-     */
-    template <unsigned words_>
-    class FixedBitSet
-    {
-        private:
-            using Bits = std::array<BitWord, words_>;
-
-            Bits _bits = {{ }};
-
-        public:
-            FixedBitSet() = default;
-
-            FixedBitSet(unsigned size)
-            {
-                assert(size <= words_ * bits_per_word);
-            }
-
-            FixedBitSet(const FixedBitSet &) = default;
-
-            FixedBitSet & operator= (const FixedBitSet &) = default;
-
-            /**
-             * Set a given bit 'on'.
-             */
-            auto set(int a) -> void
-            {
-                // The 1 does have to be of type BitWord. If we just specify a
-                // literal, it ends up being an int, and it isn't converted
-                // upwards until after the shift is done.
-                _bits[a / bits_per_word] |= (BitWord{ 1 } << (a % bits_per_word));
-            }
-
-            /**
-             * Set a given bit 'off'.
-             */
-            auto reset(int a) -> void
-            {
[…]

### 2017-11-04 code/sip.cc
Scienceing

@@ -9,6 +9,7 @@
 #include <functional>
 #include <limits>
 #include <list>
+#include <map>
 #include <numeric>
 #include <random>
 #include <utility>
@@ -20,12 +21,15 @@ using std::find_if;
 using std::greater;
 using std::list;
 using std::max;
+using std::map;
 using std::move;
 using std::mt19937;
 using std::next;
 using std::pair;
 using std::sort;
+using std::string;
 using std::swap;
+using std::to_string;
 using std::uniform_int_distribution;
 using std::uniform_real_distribution;
 using std::vector;
@@ -380,18 +384,17 @@ namespace
                 // propagate simple all different and adjacency
                 if (! propagate_simple_constraints(new_domains, current_assignment))
                     return false;
-            }
 
-            // all unit domains done, try all different (and don't bother fixedpointing)
-            if (! cheap_all_different(new_domains))
-                return false;
+                // propagate all different
+                if (! cheap_all_different(new_domains))
+                    return false;
+            }
 
             return true;
         }
 
         auto find_branch_domain(const Domains & domains) -> const Domain *
         {
-            uniform_real_distribution<double> dist(0, 1);
             const Domain * result = nullptr;
             for (auto & d : domains)
                 if (! d.fixed)
@@ -825,7 +828,7 @@ namespace
             assignments.values.reserve(pattern_size);
 
             // start search timer
-            result.times.push_back(duration_cast<milliseconds>(steady_clock::now() - params.start_time));
+            auto search_start_time = steady_clock::now();
 
             // do the appropriate search variant
             if (params.restarts) {
@@ -861,6 +864,7 @@ namespace
                                 if (d.v == n->literals[0].first) {
                                     d.values.unset(n->literals[0].second);
                                     d.popcount = d.values.popcount();
+                                    break;
                                 }
                         }
                         else {
@@ -869,6 +873,9 @@ namespace
                         }
                     }
                     need_to_watch.clear();
+
+                    if (done)
+                        break;
 
                     if (propagate(domains, assignments)) {
                         auto assignments_copy = assignments;
@@ -921,6 +928,22 @@ namespace
                     }
                 }
             }
+
+            result.extra_stats.emplace_back("search_time = " + to_string(
+                        duration_cast<milliseconds>(steady_clock::now() - search_start_time).count()));
[…]

### 2017-11-04 code/solve_subgraph_isomorphism.cc

@@ -205,23 +205,28 @@ auto main(int argc, char * argv[]) -> int
         /* Stop the clock. */
         auto overall_time = duration_cast<milliseconds>(steady_clock::now() - params.start_time);
 
-        /* Display the results. */
-        cout << boolalpha << ! result.isomorphism.empty() << " " << result.nodes;
-
+        cout << "status = ";
         if (aborted)
-            cout << " aborted";
+            cout << "aborted";
+        else if (! result.isomorphism.empty())
+            cout << "true";
+        else
+            cout << "false";
         cout << endl;
 
-        for (auto v : result.isomorphism)
-            cout << "(" << v.first << " -> " << v.second << ") ";
-        cout << endl;
+        cout << "nodes = " << result.nodes << endl;
 
-        cout << overall_time.count();
-        if (! result.times.empty()) {
-            for (auto t : result.times)
-                cout << " " << t.count();
+        if (! result.isomorphism.empty()) {
+            cout << "mapping = ";
+            for (auto v : result.isomorphism)
+                cout << "(" << v.first << " -> " << v.second << ") ";
+            cout << endl;
         }
-        cout << endl;
+
+        cout << "runtime = " << overall_time.count() << endl;
+
+        for (const auto & s : result.extra_stats)
+            cout << s << endl;
 
         if (! result.isomorphism.empty()) {
             for (int i = 0 ; i < graphs.first.size() ; ++i) {

### 2017-11-09 code/lad.cc
Randomy

@@ -8,16 +8,6 @@
 using std::ifstream;
 using std::string;
 
-GraphFileError::GraphFileError(const string & filename, const string & message) throw () :
-    _what("Error reading graph file '" + filename + "': " + message)
-{
-}
-
-auto GraphFileError::what() const throw () -> const char *
-{
-    return _what.c_str();
-}
-
 namespace
 {
     auto read_word(ifstream & infile) -> int

### 2017-11-09 code/solve_subgraph_isomorphism.cc

@@ -1,6 +1,7 @@
 /* vim: set sw=4 sts=4 et foldmethod=syntax : */
 
 #include "lad.hh"
+#include "dimacs.hh"
 #include "sip.hh"
 
 #include <boost/program_options.hpp>
@@ -113,7 +114,7 @@ auto main(int argc, char * argv[]) -> int
         display_options.add_options()
             ("help",                                         "Display help information")
             ("timeout",            po::value<int>(),         "Abort after this many seconds")
-            ("format",             po::value<string>(),      "Specify the format of the input")
+            ("dimacs",                                       "Read DIMACS format instead of LAD")
             ("induced",                                      "Solve the induced version")
             ("restarts",                                     "Use restarts")
             ("dds",                                          "Use dds")
@@ -213,8 +214,8 @@ auto main(int argc, char * argv[]) -> int
 
         /* Read in the graphs */
         auto graphs = make_pair(
-            read_lad(options_vars["pattern-file"].as<string>()),
-            read_lad(options_vars["target-file"].as<string>()));
+            (options_vars.count("dimacs") ? read_dimacs : read_lad)(options_vars["pattern-file"].as<string>()),
+            (options_vars.count("dimacs") ? read_dimacs : read_lad)(options_vars["target-file"].as<string>()));
 
         cout << "pattern_file = " << options_vars["pattern-file"].as<std::string>() << endl;
         cout << "target_file = " << options_vars["target-file"].as<std::string>() << endl;

### 2017-11-09 paper/paper.tex
Search times
Progress?

@@
 \begin{abstract}
-    Modern subgraph isomorphism solvers use strong value-ordering heuristics to direct search, but
-    these heuristics are not perfect, particularly for early choices. Our results suggest that
-    discrepancy-based searches are too expensive to be effective at offsetting this weakness.
-    However, restarts with a random value-ordering heuristic beats conventional search with the best
-    tailored heuristic. This motivates the introduction of a slightly-random value-ordering
-    heuristic with similar behaviour to the tailored heuristic: when combined with restarts, this
-    new heuristic is even more effective on satisfiable instances, whilst not significantly
-    decreasing performance on unsatisfiable instances.
+    Modern subgraph isomorphism solvers use value-ordering heuristics to direct search, but these
+    heuristics are not perfect, particularly for early choices. We investigate discrepancy searches
+    and restarts to offset this weakness. Our results suggest that discrepancy-based searches are
+    too expensive to be effective.  However, restarts with a random value-ordering heuristic beats
+    conventional search with the best tailored heuristic. This motivates the introduction of a
+    slightly-random value-ordering heuristic with similar behaviour to the tailored heuristic: when
+    combined with restarts, this new heuristic is even more effective on satisfiable instances,
+    whilst not significantly decreasing performance on unsatisfiable instances.
 \end{abstract}
@@
 nearly the same as ``highest degree first'' except with small amounts of heavily biased randomness.
-The effects of the changes we propose are shown in the top plot of \cref{figure:scatter}: we can
-make a state-of-the-art algorithm perform much better on a large number of satisfiable instances,
-without substantially degrading its performance on unsatisfiable or the remaining satisfiable
-instances.
-
-?? And concludes what?
+The effects of the changes we propose are shown in \cref{figure:scatter}: we can make a
+state-of-the-art algorithm perform much better on a large number of satisfiable instances, without
+substantially degrading its performance on unsatisfiable instances.
 
@@
         }
-    }
 \nl $(H,\,A,\,n) \gets (\emptyset,\,\emptyset,\,0)$ \;
@@
     \nl \lIf{$|A| = n$}{$(H,\,A,\,n) \gets (H \cup A,\,\emptyset,\,0)$\siplabel{hall}}
+}
     }
@@
 
-\subsection{Discrepancy Searches}
-
-?? Quick intro to discrepancy search
-\citep{DBLP:conf/ijcai/HarveyG95,DBLP:conf/aaai/Korf96,DBLP:conf/ijcai/Walsh97,DBLP:conf/cpaior/KarouiHLN07,DBLP:journals/jea/ProsserU11}.
-Discrepancy search is about value ordering.
-
-?? Talk about how we're not dealing with a binary search tree, so we could either make it binary, or
-count one discrepancy each time the degree changes, or count every ``against'' branch as being just
-one discrepancy regardless.
-
-\subsection{Restarts}
-
-?? Quick intro to restarts, mentioning how it's important to change something when restarting.
-Interestingly usually that thing is variable-ordering heuristics, not value-ordering
-\citep{DBLP:conf/ijcai/LecoutreSTV07,DBLP:journals/jsat/LecoutreSTV07,DBLP:conf/cp/GayHLS15,DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17}.
-?? Does anyone at all do this on value ordering explicitly? And if not, why not? There's
-\citet{DBLP:conf/flairs/RazgonOP07} at least. ?? There's a connection here to confidence-based work
-stealing \citep{DBLP:conf/cp/ChuSS09} too. But what? They talk about confidence in a heuristic, but
-not that different values could be ranked differently.
-
-?? We use the Luby scheme \citep{DBLP:journals/ipl/LubySZ93}, because everyone else does. We count
-the number of backtracks (that is, when we reach the end of the main $\KwSty{foreach}$ loop) to
-decide when to restart. Following established wisdom, we multiply each item in the Luby sequence by
-a magic constant. Preliminary experiments demonstrated that this is useful, but we failed to
-determine a principled way of selecting the constant's value, so we fell back on divine revelation
-and set it to 666. As with everyone else, we are not entirely clear why we are doing this: it could
-be to allow the solver to spend more time deep in search, it could be to reduce the overheads from
-repeatedly visiting inner nodes, or it could be to keep the nogoods small.
-
-?? We just use two watched literals \citep{DBLP:conf/dac/MoskewiczMZZM01}. Could use increasing
-nogoods \citep{DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17} etc, but we are not confident enough
-in our programming abilities to implement these correctly, and we are not maintaining arc
-consistency anyway. We also really do not want to propagate on negative decisions, because we do not
-want to iterate over every removed value.
-
-?? Talk a bit about how it fits into \cref{algorithm:sip}.
-
 \section{Empirical Evaluation}
@@
 satisfiable instances, so we include all pattern / target pairs. This gives a total of 14,621
-instances, ?? of which we know are satisfiable, and ??  unsatisfiable.
+instances, at least 2,110 of which are known to be satisfiable, and at least 12,323 are
+unsatisfiable.
 
@@
 
-In \cref{figure:value-ordering-heuristics} we show the cumulative number of instances solved over
-time using the four different value-ordering heuristics. As expected, the degree heuristic is much
-better (?? give aggregate speedups for both all and just satisfiable instances) than random
-ordering, which is in turn ?? times better than the anti-heuristic.
-
-\begin{figure}[tb]
+\begin{figure}[p]
     \centering
@@
 \end{figure}
+
+\begin{figure}[p]
+    \centering
+    \includegraphics*{gen-graph-scatter-dds.pdf}
+
+    \caption{The effects of using depth-bounded discrepancy search.}\label{figure:dds}
+\end{figure}
+
+In \cref{figure:value-ordering-heuristics} we show the cumulative number of instances solved over
+time using the four different value-ordering heuristics. As expected, the degree heuristic is much
+better (?? give aggregate speedups for both all and just satisfiable instances) than random
+ordering, which is in turn ?? times better than the anti-heuristic.
 
@@
 
-\subsection{Discrepancies Don't Help}
+\subsection{Discrepancy Searches}
 
-?? Code up LDS and DDS and the like properly, and plot them being disappointing here.
+?? Quick intro to discrepancy search
+\citep{DBLP:conf/ijcai/HarveyG95,DBLP:conf/aaai/Korf96,DBLP:conf/ijcai/Walsh97,DBLP:conf/cpaior/KarouiHLN07,DBLP:journals/jea/ProsserU11}.
+Discrepancy search is about value ordering.
+
+?? Talk about how we're not dealing with a binary search tree, so we could either make it binary, or
+count one discrepancy each time the degree changes, or count every ``against'' branch as being just
+one discrepancy regardless.
+
+?? \cref{figure:dds}. Sometimes extremely beneficial on satisfiable instances. However, on both
+unsatisfiable and satisfiable instances, the overheads are often large, and the algorithm is not
+better in aggregate (even when only considering satisfiable instances).
+
+\subsection{Restarts}
+
+?? Quick intro to restarts, mentioning how it's important to change something when restarting.
+Interestingly usually that thing is variable-ordering heuristics, not value-ordering
+\citep{DBLP:conf/ijcai/LecoutreSTV07,DBLP:journals/jsat/LecoutreSTV07,DBLP:conf/cp/GayHLS15,DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17}.
+?? Does anyone at all do this on value ordering explicitly? And if not, why not? There's
+\citet{DBLP:conf/flairs/RazgonOP07} at least. ?? There's a connection here to confidence-based work
+stealing \citep{DBLP:conf/cp/ChuSS09} too. But what? They talk about confidence in a heuristic, but
+not that different values could be ranked differently.
+
+?? We use the Luby scheme \citep{DBLP:journals/ipl/LubySZ93}, because everyone else does. We count
+the number of backtracks (that is, when we reach the end of the main $\KwSty{foreach}$ loop) to
+decide when to restart. Following established wisdom, we multiply each item in the Luby sequence by
+a magic constant. Preliminary experiments demonstrated that this is useful, but we failed to
+determine a principled way of selecting the constant's value, so we fell back on divine revelation
+and set it to 666. As with everyone else, we are not entirely clear why we are doing this: it could
+be to allow the solver to spend more time deep in search, it could be to reduce the overheads from
+repeatedly visiting inner nodes, or it could be to keep the nogoods small.
+
+?? We just use two watched literals \citep{DBLP:conf/dac/MoskewiczMZZM01}. Could use increasing
+nogoods \citep{DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17} etc, but we are not confident enough
+in our programming abilities to implement these correctly, and we are not maintaining arc
+consistency anyway. We also really do not want to propagate on negative decisions, because we do not
+want to iterate over every removed value.
+
+?? Talk a bit about how it fits into \cref{algorithm:sip}.
 
@@
     \caption{Cumulative number of instances solved over time, with (dotted lines) and without (solid
-    lines) restarts.}
+    lines) restarts, and with depth-bounded discrepancy search (dashed line).}
     \label{figure:restarts}
@@
 
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-scatter-nogoods.pdf}
-
-    \caption{Is nogood recording important?}
-    \label{figure:scatter-nogoods}
-\end{figure}
-
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-scatter-overheads.pdf}
-
-    \caption{Is nogood recording expensive?}
-    \label{figure:scatter-overheads}
-\end{figure}
-
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-restart-policies.pdf}
-
-    \caption{Different restart policies.}
-    \label{figure:restart-policies}
-\end{figure}
-
 ?? Explain why this is so amazingly awesome and the most important idea ever to happen in the entire
 history of science.
-
-\section{Other Subgraph Solvers}
-
-?? Also do this in McSplit and kdown?
 
@@
 
+\appendix
+\clearpage
+\section{Further Experiments}
+
+A bit of text here.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics*{gen-graph-scatter-nogoods.pdf}
+
+    \caption{Is nogood recording important?}
+    \label{figure:scatter-nogoods}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics*{gen-graph-scatter-overheads.pdf}
+
+    \caption{Is nogood recording expensive?}
+    \label{figure:scatter-overheads}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics*{gen-graph-restart-policies.pdf}
+
+    \caption{Different restart policies.}
+    \label{figure:restart-policies}
+\end{figure}
+
 \end{document}

### 2017-11-09 paper/paper.tex
All by family

@@
 nearly the same as ``highest degree first'' except with small amounts of heavily biased randomness.
-The effects of the changes we propose are shown in \cref{figure:scatter}: we can make a
+The effects of the changes we propose are shown in \cref{figure:scatter-by-family}: we can make a
 state-of-the-art algorithm perform much better on a large number of satisfiable instances, without
@@
     \centering
-    \includegraphics*{gen-graph-scatter.pdf}
+    \includegraphics*{gen-graph-scatter-by-family.pdf}
 
@@
     versus degree-biased + restarts.}
-    \label{figure:scatter}
+    \label{figure:scatter-by-family}
 \end{figure}
@@
 
-In the more detailed view in \cref{figure:scatter}, comparing the basic algorithm with the degree
-heuristic to the degree-biased algorithm with restarts, all of the unsatisfiable instances
-are on the $x-y$ diagonal, showing that their performance is unchanged. On the other hand, there are
-?? large numbers of satisfiable instances well below the diagonal line, indicating large speedups.
-Better yet, there are only ?? satisfiable instances that are more than a factor of ?? times worse.
+In the more detailed view in \cref{figure:scatter-by-family}, comparing the basic algorithm with the
+degree heuristic to the degree-biased algorithm with restarts, all of the unsatisfiable instances
+are very close to the $x-y$ diagonal, showing that their performance is nearly unchanged. On the
+other hand, there are ?? large numbers of satisfiable instances well below the diagonal line,
+indicating large speedups.  Better yet, there are only ?? satisfiable instances that are more than a
+factor of ?? times worse.
+
+?? By family
 
 ?? This has made up the consistency we lost from introducing randomness.
-
-\begin{figure}[tb]
-    \centering
-    \includegraphics*{gen-graph-scatter-by-family.pdf}
-
-    \caption{Instance by instance comparison of the basic algorithm with the degree heuristic,
-    versus degree-biased + restarts, showing only satisfiable instances, on a family by family
-    basis.}
-    \label{figure:scatter-by-family}
-\end{figure}
-
-?? do a more detailed breakdown by family, \cref{figure:scatter-by-family}
 

### 2017-11-09 paper/paper.tex
Bits

@@
 The effects of the changes we propose are shown in \cref{figure:scatter-by-family}: we can make a
-state-of-the-art algorithm perform much better on a large number of satisfiable instances, without
-substantially degrading its performance on unsatisfiable instances.
+state-of-the-art algorithm perform much better on a large number of satisfiable instances, whilst
+performing worse only rarely on satisfiable instances, and never on unsatisfiable instances.
+Further, these properties are \emph{not} shared by discrepancy searches (\cref{figure:scatter-dds})
+or by random value-ordering with restarts (\cref{figure:scatter-random}).
 
@@
 
-    \caption{The effects of using depth-bounded discrepancy search.}\label{figure:dds}
+    \caption{The effects of using depth-bounded discrepancy search.}\label{figure:scatter-dds}
 \end{figure}
@@
 
-?? \cref{figure:dds}. Sometimes extremely beneficial on satisfiable instances. However, on both
+?? \cref{figure:scatter-dds}. Sometimes extremely beneficial on satisfiable instances. However, on both
 unsatisfiable and satisfiable instances, the overheads are often large, and the algorithm is not
@@
     \caption{Instance by instance comparison of the basic algorithm with the degree heuristic,
-    versus random + restarts.}
+    versus random value-ordering with restarts.}
     \label{figure:scatter-random}
@@
     \caption{Instance by instance comparison of the basic algorithm with the degree heuristic,
-    versus degree-biased + restarts.}
+    versus degree-biased value ordering with restarts.}
     \label{figure:scatter-by-family}

### 2017-11-09 paper/paper.tex
Tweaks

@@
 
-The performance of both the biased heuristics and the degree heuristic are very similar: the
+The performance of both of the biased heuristics and the degree heuristic are very similar: the
 degree-biased heuristic has slightly worse performance below a 100s timeout, and effectively equal
@@
 \Cref{figure:scatter-heuristics} gives a more detailed look at these results. In this plot,
-satisfiable and unsatisfiable instances are shown using dots and crosses respectively. On the
-$x$-axis we measure the degree heuristic, whilst on the $y$-axis we measure random + restarts.
-Points on the outer axes represent instances which timed out with one heuristic but not the other.
-This plot shows that on a case by case basis, randomness can make a large difference to individual
-satisfiable instances.
+satisfiable and unsatisfiable instances are shown using coloured shapes and black dots respectively.
+On the $x$-axis we measure the degree heuristic, whilst on the $y$-axis we measure the degree-biased
+heuristic. Points on the outer axes represent instances which timed out with one heuristic but not
+the other.  This plot shows that despite the aggregate performance being very similar, on an case by
+case basis, the two heuristics can make a large difference to the performance of satisfiable
+instances.
 
@@
 
-?? maybe some engineering aspects: look at what happens without nogoods in
-\cref{figure:scatter-nogoods}, and at performance with restarts on degree in
-\cref{figure:scatter-overheads}.
-
 ?? Explain why this is so amazingly awesome and the most important idea ever to happen in the entire
@@
 
-A bit of text here.
+?? engineering aspects: look at what happens without nogoods in \cref{figure:scatter-nogoods}, and
+at performance with restarts on degree in \cref{figure:scatter-overheads}.
 
@@
 
+?? effects of restart policies, \cref{figure:restart-policies}.
+
 \begin{figure}[h]

### 2017-12-07 code/sequential.cc
Remove existing nogoods that are subsumed by a new one

@@ -87,6 +87,11 @@ namespace
         }
     };
 
+    struct Nogood;
+
+    // nogoods stored here
+    using Nogoods = list<Nogood>;
+
     // A nogood, aways of the form (list of assignments) -> false, where the
     // last part is implicit. If there are at least two assignments, then the
     // first two assignments are the watches (and the literals are permuted
@@ -94,10 +99,34 @@ namespace
     struct Nogood
     {
         vector<Assignment> literals;
-    };
 
-    // nogoods stored here
-    using Nogoods = list<Nogood>;
+        struct ClauseMembership
+        {
+            list<Nogoods::iterator> & lst;
+            list<Nogoods::iterator>::iterator it;
+        };
+
+        // Keep track of lists and list nodes in literal_clause_membership that
+        // point to this Nogood, so that we can efficiently delete the list nodes
+        // if this clause is erased.
+        vector<ClauseMembership> clause_memberships;
+
+        // used in a quick, incomplete subsumption test
+        unsigned long long signature = 0;
+
+        void add_literal(Assignment literal) {
+            literals.emplace_back(literal);
+            signature |= 1ull << ((11400714819323198549ull * literal.first) >> 58);
+            signature |= 1ull << ((11400714819323198549ull * literal.second) >> 58);
+        }
+        void add_clause_membership(list<Nogoods::iterator> & lst, list<Nogoods::iterator>::iterator it) {
+            clause_memberships.push_back({lst, it});
+        }
+        void remove_from_clause_membership_lists() {
+            for (auto & cm : clause_memberships)
+                cm.lst.erase(cm.it);
+        }
+    };
 
     // Two watched literals for our nogoods store.
     struct Watches
@@ -160,6 +189,11 @@ namespace
         Watches watches;
         list<typename Nogoods::iterator> need_to_watch;
 
+        // For each literal lit, keep a list of nogoods in which lit appears.
+        // The `Watches` struct is used here as it does what we need; these
+        // aren't really watch lists.
+        Watches literal_clause_membership;
+
         vector<unsigned long long> target_vertex_biases;
 
         mt19937 global_rand;
@@ -194,9 +228,11 @@ namespace
             // determine ordering for target graph vertices
             iota(target_permutation.begin(), target_permutation.end(), 0);
 
-            // set up space for watches
-            if (params.restarts && ! params.goods)
+            // set up space for watches and clauses membership lists
+            if (params.restarts && ! params.goods) {
[…]

### 2017-12-07 paper/paper.tex
More
Tweak for no reason

@@
 
-\title{Value Ordering, Discrepancies, and Restarts in Subgraph Solvers: \\ Some Half-Baked Mucking
-Around}
+\title{Value Ordering, Discrepancies, and Restarts for Subgraph Isomorphism}
 \author{Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences
@@
 \begin{abstract}
-    Modern subgraph isomorphism solvers use value-ordering heuristics to direct search, but these
-    heuristics are not perfect, particularly for early choices. We investigate discrepancy searches
-    and restarts to offset this weakness. Our results suggest that discrepancy-based searches are
-    too expensive to be effective.  However, restarts with a random value-ordering heuristic beats
-    conventional search with the best tailored heuristic. This motivates the introduction of a
-    slightly-random value-ordering heuristic with similar behaviour to the tailored heuristic: when
-    combined with restarts, this new heuristic is even more effective on satisfiable instances,
-    whilst not significantly decreasing performance on unsatisfiable instances.
+    Modern subgraph isomorphism solvers use value-ordering heuristics to direct a constraint
+    programming style search. These heuristics are not perfect, particularly for early branching
+    choices. We show that discrepancy-based searches are too expensive to be effective in this
+    setting.  However, using restarts with a random value-ordering heuristic beats conventional
+    search with the best tailored heuristic.  This motivates the introduction of a slightly-random
+    value-ordering heuristic with similar behaviour to the tailored heuristic: when combined with
+    restarts, this new search strategy is even more effective on satisfiable instances, whilst not
+    significantly decreasing performance on unsatisfiable instances.
 \end{abstract}
@@
 much better than not having a value-ordering heuristic at all, but it is far from perfect,
-particularly for early branching choices. There are two reasons to suspect we could do better.
-Firstly, even if degree were a perfect source of information, many graphs do not have a large degree
-distribution, and so we would still need to determine \emph{which} vertex of highest degree would be
-best.  Secondly, experiments with a parallel search strategy which explicitly steals early branching
-choices often gives strongly super-linear speedups on satisfiable instances
+particularly for early branching choices. There are two reasons to suspect a better option may
+possible.  Firstly, even if degree were a perfect source of information, many graphs do not have a
+large degree distribution, and so we would still need to determine \emph{which} vertex of highest
+degree would be best.  Secondly, experiments with a parallel search strategy which explicitly steals
+early branching choices often gives strongly super-linear speedups on satisfiable instances
 \citep{dblp:conf/cp/McCreeshP15}.
@@
 satisfiable instances, so we include all pattern / target pairs. This gives a total of 14,621
-instances, at least 2,110 of which are known to be satisfiable, and at least 12,323 are
+instances, at least 2,126 of which are known to be satisfiable, and at least 12,323 are
 unsatisfiable.
@@
 
+\begin{figure}[tb]
+    \centering
+    \includegraphics*{gen-graph-scatter-parallel.pdf}
+
+    \caption{Parallel}
+    \label{figure:scatter-parallel}
+\end{figure}
+
+\begin{figure}[tb]
+    \centering
+    \includegraphics*{gen-graph-final.pdf}
+
+    \caption{Final results}
+    \label{figure:final}
+\end{figure}
+
 ?? Parallel

### 2017-12-07 code/sequential.cc
Softmax?

@@ -6,6 +6,7 @@
 
 #include <algorithm>
 #include <array>
+#include <cmath>
 #include <functional>
 #include <limits>
 #include <list>
@@ -17,6 +18,7 @@
 #include <utility>
 
 using std::array;
+using std::exp;
 using std::iota;
 using std::fill;
 using std::find_if;
@@ -607,6 +609,35 @@ namespace
                     swap(branch_v[select_element], branch_v[start]);
                 }
             }
+            else if (params.softmax_shuffle) {
+                // sum up the bias scores of every branch vertex
+                double total = 0.0;
+                for (unsigned v = 0 ; v < branch_v_end ; ++v)
+                    total += exp(double(targets_degrees[0][v]));
+
+                // repeatedly pick a softmax-biased vertex, move it to the front of branch_v,
+                // and then only consider items further to the right in the next iteration.
+                for (unsigned start = 0 ; start < branch_v_end ; ++start) {
+                    // pick a random number between 0 and 1 inclusive
+                    uniform_real_distribution<double> dist(0, 1);
+                    double select_score = dist(global_rand);
+
+                    // this decreases on each iteration by the sum of the
+                    // scores seen so far
+                    double select_if_score_ge = 1.0;
+
+                    // go over the list until we hit the score
+                    unsigned select_element = start;
+                    for ( ; select_element + 1 < branch_v_end ; ++select_element) {
+                        select_if_score_ge -= exp(double(targets_degrees[0][select_element]));
+                        if (select_score >= select_if_score_ge)
+                            break;
+                    }
+
+                    // move to front
+                    swap(branch_v[select_element], branch_v[start]);
+                }
+            }
             else if (params.position_shuffle) {
                 // repeatedly pick a position-biased vertex, move it to the front of branch_v,
                 // and then only consider items further to the right in the next iteration.
@@ -981,7 +1012,7 @@ namespace
 
                 result.extra_stats.emplace_back("restarts = " + to_string(number_of_restarts));
             }
-            else if (params.shuffle || params.biased_shuffle || params.position_shuffle) {
+            else if (params.shuffle || params.biased_shuffle || params.position_shuffle || params.softmax_shuffle) {
                 if (propagate(domains, assignments)) {
                     // still need to use the restarts variant
                     long long backtracks_until_restart = -1;

### 2017-12-07 code/solve_subgraph_isomorphism.cc

@@ -131,6 +131,7 @@ auto main(int argc, char * argv[]) -> int
             ("dds",                                          "Use dds")
             ("shuffle",                                      "Use shuffling")
             ("biased-shuffle",                               "Use biased shuffling")
+            ("softmax-shuffle",                              "Use softmax shuffling")
             ("position-shuffle",                             "Use position shuffling")
             ("antiheuristic",                                "Use antiheuristic")
             ("input-order",                                  "Use input order")
@@ -198,6 +199,7 @@ auto main(int argc, char * argv[]) -> int
         if (options_vars["algorithm"].as<string>() == "customisable-parallel") {
             if (! options_vars.count("restarts")
                     || ! options_vars.count("biased-shuffle")
+                    || options_vars.count("softmax-shuffle")
                     || ! options_vars.count("input-order")
                     || options_vars.count("dds")
                     || options_vars.count("shuffle")
@@ -206,7 +208,7 @@ auto main(int argc, char * argv[]) -> int
                     || options_vars.count("antiheuristic")) {
                 cerr << "Parallel algorithm currently requires --restarts --input-order" << endl;
                 cerr << "  --biased-shuffle, and cannot use any of --dds --shuffle " << endl;
-                cerr << "  --goods --position-shuffle --antiheuristic" << endl;
+                cerr << "  --softmax-shuffle --goods --position-shuffle --antiheuristic" << endl;
                 return EXIT_FAILURE;
             }
         }
@@ -221,6 +223,7 @@ auto main(int argc, char * argv[]) -> int
             params.dds = options_vars.count("dds");
             params.shuffle = options_vars.count("shuffle");
             params.biased_shuffle = options_vars.count("biased-shuffle");
+            params.softmax_shuffle = options_vars.count("softmax-shuffle");
             params.position_shuffle = options_vars.count("position-shuffle");
             params.antiheuristic = options_vars.count("antiheuristic");
             params.input_order = options_vars.count("input-order");

### 2017-12-07 code/sequential.cc
Revert "I hate James"
This reverts commit a166b808c99c5bb55fa5ec2c1033f8cfe431c499.

Revert "Some of James's hackery"

This reverts commit 0ad975897a79a5f549a496740f6e6f99ea597e34.

@@ -11,7 +11,6 @@
 #include <limits>
 #include <list>
 #include <map>
-#include <memory>
 #include <numeric>
 #include <random>
 #include <tuple>
@@ -147,13 +146,9 @@ namespace
 
         unsigned pattern_size, full_pattern_size, target_size;
 
-        std::vector<Domains> domains_by_depth;
-
         vector<uint8_t> pattern_adjacencies_bits;
         vector<FixedBitSet<n_words_> > pattern_graph_rows;
         vector<FixedBitSet<n_words_> > target_graph_rows;
-
-        vector<vector<FixedBitSet<n_words_> > > adjacency_masks;
 
         vector<int> pattern_permutation, target_permutation, isolated_vertices;
         vector<vector<int> > patterns_degrees, targets_degrees;
@@ -172,7 +167,6 @@ namespace
             pattern_size(pattern.size()),
             full_pattern_size(pattern.size()),
             target_size(target.size()),
-            domains_by_depth(pattern.size() + 2), // TODO: this is probably bigger than required?
             target_permutation(target.size()),
             patterns_degrees(max_graphs),
             targets_degrees(max_graphs)
@@ -259,24 +253,6 @@ namespace
             }
         }
 
-        auto build_adjacency_masks() {
-            for (unsigned i = 0 ; i < (2u << max_graphs) ; ++i) {
-                adjacency_masks.push_back({});
-                auto & am = adjacency_masks.back();
-                for (unsigned j = 0 ; j < target_size ; j++) {
-                    am.push_back({});
-                    am.back().set_all();
-                    for (int g = 0 ; g < max_graphs ; ++g) {
-                        // if we're adjacent...
-                        if (i & (1u << g)) {
-                            // ...then we can only be mapped to adjacent vertices
-                            am.back().intersect_with(target_graph_rows[j * max_graphs + g]);
-                        }
-                    }
-                }
-            }
-        }
-
         auto build_complement_graphs(vector<FixedBitSet<n_words_> > & graph_rows, unsigned size) -> void
         {
             for (unsigned v = 0 ; v < size ; ++v)
@@ -345,10 +321,27 @@ namespace
             return true;
         }
 
+        // The max_graphs_ template parameter is so that the for each graph
+        // pair loop gets unrolled, which makes an annoyingly large difference
+        // to performance. Note that for larger target graphs, half of the
+        // total runtime is spent in this function.
+        template <int max_graphs_>
[…]

### 2017-12-08 code/sequential.cc
A divisive commit

@@ -623,10 +623,13 @@ namespace
                     // go over the list until we hit the score
                     unsigned select_element = start;
                     for ( ; select_element + 1 < branch_v_end ; ++select_element) {
-                        select_if_score_ge -= exp(double(targets_degrees[0][select_element]));
+                        select_if_score_ge -= exp(double(targets_degrees[0][select_element])) / total;
                         if (select_score >= select_if_score_ge)
                             break;
                     }
+
+                    // total is now lower
+                    total -= exp(double(targets_degrees[0][select_element]));
 
                     // move to front
                     swap(branch_v[select_element], branch_v[start]);

### 2017-12-08 code/sequential.cc
A bit more flexibility

@@ -17,7 +17,6 @@
 #include <utility>
 
 using std::array;
-using std::exp;
 using std::iota;
 using std::fill;
 using std::find_if;
@@ -30,6 +29,7 @@ using std::move;
 using std::mt19937;
 using std::next;
 using std::pair;
+using std::pow;
 using std::sort;
 using std::string;
 using std::swap;
@@ -607,7 +607,7 @@ namespace
                 // sum up the bias scores of every branch vertex
                 double total = 0.0;
                 for (unsigned v = 0 ; v < branch_v_end ; ++v)
-                    total += exp(double(targets_degrees[0][v]));
+                    total += pow(params.softmax_base, double(targets_degrees[0][v]));
 
                 // repeatedly pick a softmax-biased vertex, move it to the front of branch_v,
                 // and then only consider items further to the right in the next iteration.
@@ -623,7 +623,7 @@ namespace
                     // go over the list until we hit the score
                     unsigned select_element = start;
                     for ( ; select_element + 1 < branch_v_end ; ++select_element) {
-                        select_if_score_ge -= exp(double(targets_degrees[0][select_element])) / total;
+                        select_if_score_ge -= pow(params.softmax_base, double(targets_degrees[0][select_element])) / total;
                         if (select_score >= select_if_score_ge)
                             break;
                     }

### 2017-12-08 code/solve_subgraph_isomorphism.cc

@@ -132,6 +132,7 @@ auto main(int argc, char * argv[]) -> int
             ("shuffle",                                      "Use shuffling")
             ("biased-shuffle",                               "Use biased shuffling")
             ("softmax-shuffle",                              "Use softmax shuffling")
+            ("softmax-base",         po::value<double>(),    "Specify the base for softmax shuffling")
             ("position-shuffle",                             "Use position shuffling")
             ("antiheuristic",                                "Use antiheuristic")
             ("input-order",                                  "Use input order")
@@ -229,6 +230,8 @@ auto main(int argc, char * argv[]) -> int
             params.input_order = options_vars.count("input-order");
             params.goods = options_vars.count("goods");
 
+            if (options_vars.count("softmax-base"))
+                params.softmax_base = options_vars["softmax-base"].as<double>();
             if (options_vars.count("luby-multiplier"))
                 params.luby_multiplier = options_vars["luby-multiplier"].as<unsigned>();
             if (options_vars.count("geometric-multiplier"))

### 2017-12-13 mcsplit-code/mcsp.c
One watched literal

@@ -252,6 +252,30 @@ struct Nogood
 
 using Nogoods = list<Nogood>;
 
+// One watched literal for our nogoods store.
+struct Watches
+{
+    // for each watched literal, we have a list of watched things, each of
+    // which is an iterator into the global watch list (so we can reorder
+    // the literal to keep the watch as the first element)
+    using WatchList = list<Nogoods::iterator>;
+
+    unsigned target_size_plus_1;
+
+    // two dimensional array, indexed by ((target_size+1) * p + t + 1)
+    vector<WatchList> data;
+
+    // not a ctor to avoid annoyingness with isolated vertices altering the
+    // pattern size
+    Watches(unsigned p, unsigned t)
+        : target_size_plus_1(t + 1), data(p * (t + 1)) {}
+
+    WatchList & operator[] (const Assignment & a)
+    {
+        return data[target_size_plus_1 * a.v + a.w + 1];
+    }
+};
+
 class MCS
 {
     const Graph & g0;
@@ -260,6 +284,8 @@ class MCS
     vector<int> right;
     vector<Assignment> incumbent;
     Nogoods nogoods;
+    Watches watches;
+    list<typename Nogoods::iterator> need_to_watch;
 
     vector<int> vtx_current_assignment;
 
@@ -467,18 +493,49 @@ class MCS
 
 //        std::cout << "literals size " << nogood.literals.size() << std::endl;
         nogoods.emplace_back(std::move(nogood));
+        need_to_watch.emplace_back(prev(nogoods.end()));
     }
 
     auto current_contains_nogood(
-            vector<int> & vtx_current_assignment,
-            Nogood & n) -> bool
+            Assignment most_recent_assignment,
+            vector<int> & vtx_current_assignment) -> bool
     {
-//        std::cout << n.literals.size() << std::endl;
-        for (const Assignment a : n.literals)
-            if (vtx_current_assignment[a.v] != a.w)
-                return false;
+        auto & watches_to_update = watches[most_recent_assignment];
+        for (auto watch_to_update = watches_to_update.begin() ; watch_to_update != watches_to_update.end() ; ) {
+            Nogood & nogood = **watch_to_update;
 
-        return true;
+            // can we find something else to watch?
+            bool success = false;
+            for (auto new_literal = next(nogood.literals.begin(), 1) ; new_literal != nogood.literals.end() ; ++new_literal) {
+                if (vtx_current_assignment[new_literal->v] != new_literal->w) {
[…]

### 2017-12-13 mcsplit-code/mcsp.c
Erase subsumed nogoods

@@ -245,12 +245,42 @@ bool check_sol(const Graph & g0, const Graph & g1 , const vector<Assignment> & s
     return true;
 }
 
+struct Nogood;
+
+using Nogoods = list<Nogood>;
+
 struct Nogood
 {
     vector<Assignment> literals;
-};
 
-using Nogoods = list<Nogood>;
+    struct ClauseMembership
+    {
+        list<Nogoods::iterator> & lst;
+        list<Nogoods::iterator>::iterator it;
+    };
+
+    // Keep track of lists and list nodes in literal_clause_membership that
+    // point to this Nogood, so that we can efficiently delete the list nodes
+    // if this clause is erased.
+    vector<ClauseMembership> clause_memberships;
+
+    // a hash of the nogood, used in a quick, incomplete subsumption test
+    unsigned long long signature = 0;
+
+    void add_literal(Assignment literal) {
+        literals.emplace_back(literal);
+        // Set a couple of bits in the signature.  I'm sure this could be done better!
+        signature |= 1ull << ((11400714819323198549ull * (literal.v + literal.w + 1)) >> 58);
+        signature |= 1ull << ((11400714819323198549ull * (literal.v * (literal.w + 2))) >> 58);
+    }
+    void add_clause_membership(list<Nogoods::iterator> & lst, list<Nogoods::iterator>::iterator it) {
+        clause_memberships.push_back({lst, it});
+    }
+    void remove_from_clause_membership_lists() {
+        for (auto & cm : clause_memberships)
+            cm.lst.erase(cm.it);
+    }
+};
 
 // One watched literal for our nogoods store.
 struct Watches
@@ -286,6 +316,7 @@ class MCS
     Nogoods nogoods;
     Watches watches;
     list<typename Nogoods::iterator> need_to_watch;
+    Watches literal_clause_membership;
 
     vector<int> vtx_current_assignment;
 
@@ -483,17 +514,63 @@ class MCS
         if (!arguments.quiet) cout << "Incumbent size: " << incumbent.size() << endl;
     }
 
+//    auto nogood_has_already_been_posted(Nogood & n) -> bool {
+//        // `literal_with_lowest_tally` will be an iterator to the literal in `n` that appears in
+//        // fewest members of `nogoods`
+//        auto literal_with_lowest_tally = std::min_element(
+//                n.literals.begin(),
+//                n.literals.end(),
+//                [&](const Assignment & a, const Assignment & b) {
+//                    return literal_clause_membership[a].size() < literal_clause_membership[b].size();
[…]

### 2017-12-13 mcsplit-code/mcsp.c
Linear-time subsumption test

@@ -320,6 +320,9 @@ class MCS
 
     vector<int> vtx_current_assignment;
 
+    // used in subsumption test; assigned once for speed
+    vector<int> assignments_in_nogood;
+
     void show(const vector<VarAssignment>& current, const vector<Bidomain> &domains)
     {
         cout << "Nodes: " << nodes << std::endl;
@@ -742,19 +745,10 @@ class MCS
             return false;
 
         // Return false if some literal in `a` does not appear in `b`.
-        //
-        // The reverse iteration seems to give a slight speed-up compared to
-        // forward iteration, due to literals that appear late in a clause appearing
-        // in few other clauses.  This is based on intuition and a tiny-scale
-        // experiment only; perhaps forward iteration would in fact be better.
-        //
-        // This quadratic-complexity algorithm could be replaced with a linear
-        // one.
-        for (auto it = a.literals.rbegin() ; it != a.literals.rend() ; ++it)
-            if (std::find(b.literals.begin(), b.literals.end(), *it) == b.literals.end())
-                return false;
-
-        return true;
+        unsigned count = 0;
+        for (auto lit : b.literals)
+            count += (assignments_in_nogood[lit.v] == lit.w);
+        return count == a.literals.size();
     }
 
     auto remove_subsumed_clauses(Nogoods::iterator n_it) -> void {
@@ -771,6 +765,10 @@ class MCS
         // `lst` is the list of clauses containing the
         // literal `*literal_with_lowest_tally`
         auto & lst = literal_clause_membership[*literal_with_lowest_tally];
+
+        std::fill(assignments_in_nogood.begin(), assignments_in_nogood.end(), -2);
+        for (auto lit : n.literals)
+            assignments_in_nogood[lit.v] = lit.w;
 
         // erase all nogoods in `lst` that are subsumed by `n`
         for (auto nogood_it_it = lst.begin() ; nogood_it_it != lst.end() ; ) {
@@ -894,7 +892,8 @@ public:
     MCS(Graph & g0, Graph & g1)
         : g0(g0), g1(g1), watches(g0.n, g1.n),
           literal_clause_membership(g0.n, g1.n),
-          vtx_current_assignment(g0.n)
+          vtx_current_assignment(g0.n),
+          assignments_in_nogood(g0.n)
     { }
 
     vector<Assignment> run() {

### 2017-12-13 mcsplit-code/mcsp.c
Biased shuffle instead of position shuffle

@@ -1,5 +1,5 @@
 #define RESTARTS 1
-#define LUBY_MULTIPLIER 500
+#define LUBY_MULTIPLIER 500 
 
 #include "graph.h"
 
@@ -47,7 +47,7 @@ static struct argp_option options[] = {
     {"verbose", 'v', 0, 0, "Verbose output"},
     {"dimacs", 'd', 0, 0, "Read DIMACS format"},
     {"lad", 'l', 0, 0, "Read LAD format"},
-    {"position-shuffle", 'p', 0, 0, "Position-shuffle value ordering heuristic"},
+    {"biased-shuffle", 'b', 0, 0, "Position-shuffle value ordering heuristic"},
     {"connected", 'c', 0, 0, "Solve max common CONNECTED subgraph problem"},
     {"directed", 'i', 0, 0, "Use directed graphs"},
     {"labelled", 'a', 0, 0, "Use edge and vertex labels"},
@@ -61,7 +61,7 @@ struct Arguments {
     bool verbose;
     bool dimacs;
     bool lad;
-    bool position_shuffle;
+    bool biased_shuffle;
     bool connected;
     bool directed;
     bool edge_labelled;
@@ -89,8 +89,8 @@ static error_t parse_opt (int key, char *arg, struct argp_state *state) {
                 fail("The -d and -l options cannot be used together.\n");
             arguments.lad = true;
             break;
-        case 'p':
-            arguments.position_shuffle = true;
+        case 'b':
+            arguments.biased_shuffle = true;
             break;
         case 'q':
             arguments.quiet = true;
@@ -245,6 +245,18 @@ bool check_sol(const Graph & g0, const Graph & g1 , const vector<Assignment> & s
     return true;
 }
 
+vector<int> calculate_degrees(const Graph & g) {
+    vector<int> degree(g.n, 0);
+    for (int v=0; v<g.n; v++) {
+        for (int w=0; w<g.n; w++) {
+            unsigned int mask = 0xFFFFu;
+            if (g.adjmat[v][w] & mask) degree[v]++;
+            if (g.adjmat[v][w] & ~mask) degree[v]++;  // inward edge, in directed case
+        }
+    }
+    return degree;
+}
+
 struct Nogood;
 
 using Nogoods = list<Nogood>;
@@ -322,6 +334,8 @@ class MCS
 
     // used in subsumption test; assigned once for speed
     vector<int> assignments_in_nogood;
+
[…]

### 2017-12-13 mcsplit-code/mcsp.c
Deleted some old commented-out code

@@ -523,21 +523,6 @@ class MCS
             remaining_score -= target_vertex_biases[vec[select_element]];
             std::swap(vec[select_element], vec[start]);
         }
-//        for (unsigned start=0; start<vec.size()-1; ++start) {
-//            std::uniform_real_distribution<double> dist(0, 1);
-//            double select_score = dist(global_rand);
-//
-//            double select_if_score_ge = 1.0;
-//
-//            unsigned select_element = start;
-//            for ( ; select_element + 1 < vec.size(); ++select_element) {
-//                select_if_score_ge /= 2.0;
-//                if (select_score >= select_if_score_ge)
-//                    break;
-//            }
-//            
-//            std::swap(vec[select_element], vec[start]);
-//        }
     }
 
     enum class Search

### 2017-12-13 mcsplit-code/mcsp.c
Command-line switch for restarts

@@ -1,4 +1,3 @@
-#define RESTARTS 1
 #define LUBY_MULTIPLIER 500 
 
 #include "graph.h"
@@ -47,6 +46,7 @@ static struct argp_option options[] = {
     {"verbose", 'v', 0, 0, "Verbose output"},
     {"dimacs", 'd', 0, 0, "Read DIMACS format"},
     {"lad", 'l', 0, 0, "Read LAD format"},
+    {"restarts", 'r', 0, 0, "Use restarts and nogood learning"},
     {"biased-shuffle", 'b', 0, 0, "Position-shuffle value ordering heuristic"},
     {"connected", 'c', 0, 0, "Solve max common CONNECTED subgraph problem"},
     {"directed", 'i', 0, 0, "Use directed graphs"},
@@ -61,6 +61,7 @@ struct Arguments {
     bool verbose;
     bool dimacs;
     bool lad;
+    bool restarts;
     bool biased_shuffle;
     bool connected;
     bool directed;
@@ -88,6 +89,9 @@ static error_t parse_opt (int key, char *arg, struct argp_state *state) {
             if (arguments.dimacs)
                 fail("The -d and -l options cannot be used together.\n");
             arguments.lad = true;
+            break;
+        case 'r':
+            arguments.restarts = true;
             break;
         case 'b':
             arguments.biased_shuffle = true;
@@ -969,7 +973,7 @@ public:
 
         VarAssignments current;
 
-        if (RESTARTS) {
+        if (arguments.restarts) {
             run_with_restarts(current, domains);
         } else {
             search(current, domains);

### 2017-12-13 mcsplit-code/mcsp.cc
Make filenames more C++-like

@@ -1,6 +1,6 @@
 #define LUBY_MULTIPLIER 500 
 
-#include "graph.h"
+#include "graph.hh"
 
 #include <algorithm>
 #include <atomic>

### 2017-12-13 mcsplit-code/graph.cc
Moved algorithm into separate file, and tweaked code

@@ -1,5 +1,6 @@
 #include "graph.hh"
 
+#include <limits.h>
 #include <stdio.h>
 #include <stdlib.h>
 

### 2017-12-13 mcsplit-code/mcsp.cc

@@ -1,39 +1,24 @@
-#define LUBY_MULTIPLIER 500 
-
 #include "graph.hh"
+#include "solve_mcs.hh"
 
-#include <algorithm>
 #include <atomic>
 #include <chrono>
 #include <condition_variable>
 #include <iostream>
-#include <list>
 #include <mutex>
-#include <numeric>
-#include <random>
-#include <set>
-#include <string>
 #include <thread>
-#include <utility>
 #include <vector>
 
 #include <argp.h>
-#include <limits.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
 
 using std::cout;
 using std::endl;
-using std::list;
 using std::vector;
 
 static void fail(std::string msg) {
     std::cerr << msg << std::endl;
     exit(1);
 }
-
-enum Heuristic { min_max, min_product };
 
 /*******************************************************************************
                              Command-line arguments
@@ -75,8 +60,6 @@ struct Arguments {
 };
 
 static Arguments arguments;
-
-static std::atomic<bool> abort_due_to_timeout;
 
 static error_t parse_opt (int key, char *arg, struct argp_state *state) {
     switch (key) {
@@ -129,9 +112,9 @@ static error_t parse_opt (int key, char *arg, struct argp_state *state) {
         case ARGP_KEY_ARG:
             if (arguments.arg_num == 0) {
                 if (std::string(arg) == "min_max")
-                    arguments.heuristic = min_max;
+                    arguments.heuristic = Heuristic::min_max;
                 else if (std::string(arg) == "min_product")
-                    arguments.heuristic = min_product;
+                    arguments.heuristic = Heuristic::min_product;
                 else
                     fail("Unknown heuristic (try min_max or min_product)");
             } else if (arguments.arg_num == 1) {
@@ -161,72 +144,7 @@ static struct argp argp = { options, parse_opt, args_doc, doc };
 unsigned long long nodes{ 0 };
 
 /*******************************************************************************
-                                 MCS functions
 *******************************************************************************/
-
-std::mt19937 global_rand;
-
-struct Assignment {
-    int v;
-    int w;
-
-    auto operator==(const Assignment & other) const -> bool
-    {
-        return v==other.v && w==other.w;
-    }
-};
-
-struct VarAssignment {
-    Assignment assignment;
-    bool is_decision;
-};
-
-class VarAssignments
-{
[…]

### 2017-12-13 mcsplit-code/solve_mcs.cc

@@ -1,0 +1,814 @@
+#define LUBY_MULTIPLIER 500 
+
+#include "graph.hh"
+#include "solve_mcs.hh"
+
+#include <algorithm>
+#include <iostream>
+#include <list>
+#include <numeric>
+#include <random>
+#include <set>
+#include <vector>
+
+using std::cout;
+using std::endl;
+using std::list;
+using std::vector;
+
+std::atomic<bool> abort_due_to_timeout;
+
+namespace {
+    std::mt19937 global_rand;
+
+    struct VarAssignment
+    {
+        Assignment assignment;
+        bool is_decision;
+    };
+
+    class VarAssignments
+    {
+        vector<VarAssignment> var_assignments;
+        unsigned num_vtx_assignments = 0;
+    public:
+        auto push(VarAssignment a) -> void
+        {
+            var_assignments.push_back(a);
+            if (a.assignment.w != -1)
+                ++num_vtx_assignments;
+        }
+
+        auto pop() -> void
+        {
+            if (var_assignments.back().assignment.w != -1)
+                --num_vtx_assignments;
+            var_assignments.pop_back();
+        }
+
+        auto get_var_assignments() const -> const vector<VarAssignment> &
+        {
+            return var_assignments;
+        }
+
+        auto get_num_vtx_assignments() -> unsigned
+        {
+            return num_vtx_assignments;
+        }
+
+        auto clear() -> void
+        {
+            var_assignments.clear();
+        }
+    };
+
+    struct Bidomain
+    {
+        int l,        r;        // start indices of left and right sets
+        int left_len, right_len;
+        bool is_adjacent;
+    };
+
+    auto calculate_degrees(const Graph & g) -> vector<int>
+    {
+        vector<int> degree(g.n, 0);
+        for (int v=0; v<g.n; v++) {
+            for (int w=0; w<g.n; w++) {
+                unsigned int mask = 0xFFFFu;
+                if (g.adjmat[v][w] & mask) degree[v]++;
+                if (g.adjmat[v][w] & ~mask) degree[v]++;  // inward edge, in directed case
+            }
+        }
+        return degree;
+    }
+
+    struct Nogood;
+
+    using Nogoods = list<Nogood>;
+
+    struct Nogood
+    {
+        vector<Assignment> literals;
+
+        struct ClauseMembership
+        {
+            list<Nogoods::iterator> & lst;
+            list<Nogoods::iterator>::iterator it;
+        };
+
[…]

### 2017-12-13 paper/paper.tex
More things

@@
 
-\definecolor{inferno7}{rgb}{0.976471, 0.584314, 0.039216}
+\definecolor{highlight}{rgb}{0.819608, 0.733333, 0.34902}
 
@@
     Modern subgraph isomorphism solvers use value-ordering heuristics to direct a constraint
-    programming style search. These heuristics are not perfect, particularly for early branching
-    choices. We show that discrepancy-based searches are too expensive to be effective in this
-    setting.  However, using restarts with a random value-ordering heuristic beats conventional
-    search with the best tailored heuristic.  This motivates the introduction of a slightly-random
-    value-ordering heuristic with similar behaviour to the tailored heuristic: when combined with
-    restarts, this new search strategy is even more effective on satisfiable instances, whilst not
-    significantly decreasing performance on unsatisfiable instances.
+    programming style search. These heuristics are far from perfect, and experiments show that using
+    a random value-ordering heuristic with restarts beats conventional search with the best known
+    tailored heuristic. This motivates the introduction of a slightly-random value-ordering
+    heuristic with similar behaviour to the tailored heuristic: when combined with restarts, this
+    new search strategy is over a hundred times more effective on satisfiable instances, and does
+    not significantly decrease performance on unsatisfiable instances. In contrast,
+    discrepancy-based searches are too expensive to be effective in this setting.
 \end{abstract}
@@
 
-?? Subgraph isomorphism is useful.
-
-The current single strongest subgraph isomorphism solver uses ``highest degree first'' as a
-value-ordering heuristic \citep{dblp:conf/cp/McCreeshP15,dblp:conf/ijcai/McCreeshPT16}.  This is
+The subgraph isomorphism problem is to find a copy of a small ``pattern'' graph inside a larger
+``target'' graph.  The current single strongest subgraph isomorphism solver uses ``highest degree
+first'' as a value-ordering heuristic to direct a constraint programming style search
+\citep{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:conf/ijcai/McCreeshPT16}.  This is
 much better than not having a value-ordering heuristic at all, but it is far from perfect,
 particularly for early branching choices. There are two reasons to suspect a better option may
-possible.  Firstly, even if degree were a perfect source of information, many graphs do not have a
-large degree distribution, and so we would still need to determine \emph{which} vertex of highest
-degree would be best.  Secondly, experiments with a parallel search strategy which explicitly steals
-early branching choices often gives strongly super-linear speedups on satisfiable instances
-\citep{dblp:conf/cp/McCreeshP15}.
+possible.  Firstly, even if degree were a perfect source of information, many target graphs do not
+have a large degree distribution, and so we would still need to determine \emph{which} vertex of
+highest degree would be best.  Secondly, experiments with a parallel search strategy which
+explicitly steals early branching choices often gives strongly super-linear speedups on satisfiable
+instances \citep{DBLP:conf/cp/McCreeshP15}.
 
@@
 fruitful to investigate alternatives to simple backtracking (depth-first) search. This paper
-evaluates two such approaches: discrepancy search (which perform poorly), and restarts (which are
-extremely beneficial). To make restarts work, we introduce a new value-ordering heuristic, which is
-nearly the same as ``highest degree first'' except with small amounts of heavily biased randomness.
-The effects of the changes we propose are shown in \cref{figure:scatter-by-family}: we can make a
-state-of-the-art algorithm perform much better on a large number of satisfiable instances, whilst
-performing worse only rarely on satisfiable instances, and never on unsatisfiable instances.
-Further, these properties are \emph{not} shared by discrepancy searches (\cref{figure:scatter-dds})
-or by random value-ordering with restarts (\cref{figure:scatter-random}).
+evaluates two such approaches: discrepancy search (which performs poorly), and restarts (which are
+extremely beneficial). To explore different parts of the search space when restarting, we introduce
+a new value-ordering heuristic, which adds a small amount of randomness to the ``highest degree
+first'' heuristic.  The effects of the changes we propose are shown in
+\cref{figure:scatter-by-family}: we can make a state-of-the-art algorithm perform much better on a
+large number of satisfiable instances, whilst performing worse only rarely on satisfiable instances,
+and never on unsatisfiable instances.  Further, these properties are \emph{not} shared by
+discrepancy searches (\cref{figure:scatter-dds}) or by random value-ordering with restarts
+(\cref{figure:scatter-random}).
 
 \section{The Basic Algorithm}
+
+?? Non-induced!
 
@@
         \coordinate (restart1ec) at ($(pic cs:restart1e) + (0, 0.00)$);
-        \node [fill=inferno7, rounded corners, fit=(restart1sc) (restart1ec)] { };
+        \node [fill=highlight, rounded corners, fit=(restart1sc) (restart1ec)] { };
         \coordinate (restart2sc) at ($(pic cs:restart2s) + (0, 0.15)$);
         \coordinate (restart2ec) at ($(pic cs:restart2e) + (0, 0.00)$);
-        \node [fill=inferno7, rounded corners, fit=(restart2sc) (restart2ec)] { };
+        \node [fill=highlight, rounded corners, fit=(restart2sc) (restart2ec)] { };
         \coordinate (restart3sc) at ($(pic cs:restart3s) + (0, 0.15)$);
         \coordinate (restart3ec) at ($(pic cs:restart3e) + (0, 0.00)$);
-        \node [fill=inferno7, rounded corners, fit=(restart3sc) (restart3ec)] { };
+        \node [fill=highlight, rounded corners, fit=(restart3sc) (restart3ec)] { };
         \coordinate (restart4sc) at ($(pic cs:restart4s) + (0, 0.15)$);
         \coordinate (restart4ec) at ($(pic cs:restart4e) + (0, 0.00)$);
-        \node [fill=inferno7, rounded corners, fit=(restart4sc) (restart4ec)] { };
+        \node [fill=highlight, rounded corners, fit=(restart4sc) (restart4ec)] { };
         \coordinate (restart5sc) at ($(pic cs:restart5s) + (0, 0.15)$);
@@
         \coordinate (restart5rc) at ($(pic cs:restart5r) + (0, 0.00)$);
-        \node [fill=inferno7, rounded corners, fit=(restart5sc) (restart5ec) (restart5rc)] { };
+        \node [fill=highlight, rounded corners, fit=(restart5sc) (restart5ec) (restart5rc)] { };
         \coordinate (restart6sc) at ($(pic cs:restart6s) + (0, 0.15)$);
         \coordinate (restart6ec) at ($(pic cs:restart6e) + (0, 0.00)$);
-        \node [fill=inferno7, rounded corners, fit=(restart6sc) (restart6ec)] { };
+        \node [fill=highlight, rounded corners, fit=(restart6sc) (restart6ec)] { };
         \coordinate (restart7sc) at ($(pic cs:restart7s) + (0, 0.15)$);
         \coordinate (restart7ec) at ($(pic cs:restart7e) + (0, 0.00)$);
-        \node [fill=inferno7, rounded corners, fit=(restart7sc) (restart7ec)] { };
+        \node [fill=highlight, rounded corners, fit=(restart7sc) (restart7ec)] { };
         \coordinate (restart8sc) at ($(pic cs:restart8s) + (0.05, 0.15)$);
         \coordinate (restart8ec) at ($(pic cs:restart8e) + (0, 0.00)$);
-        \node [fill=inferno7, rounded corners, fit=(restart8sc) (restart8ec)] { };
+        \node [fill=highlight, rounded corners, fit=(restart8sc) (restart8ec)] { };
         \coordinate (restart9sc) at ($(pic cs:restart9s) + (0, 0.15)$);
         \coordinate (restart9ec) at ($(pic cs:restart9e) + (-0.15, 0.00)$);
-        \node [fill=inferno7, rounded corners, fit=(restart9sc) (restart9ec)] { };
+        \node [fill=highlight, rounded corners, fit=(restart9sc) (restart9ec)] { };
     \end{tikzpicture}
@@
         \nl \lIf{$\KwSty{not}~\FuncSty{propagate}(D)$}{$\KwSty{return}~\KwSty{false}$}
-        \nl \tikzmark{restart1s}$\KwSty{global}~(R, R') \gets (0, 666 \times \textnormal{the next Luby value})$\tikzmark{restart1e} \;
+        \nl \tikzmark{restart1s}$\KwSty{global}~(R, R') \gets (0, \textnormal{the next restart sequence value})$\tikzmark{restart1e} \;
         \nl $S \gets \FuncSty{search}(\{ E \in D : \left|E\right| > 1 \}, \top)$ \;
@@
 paper---as far as possible, we will leave it intact, so as not to affect the performance of the
-algorithm on unsatisfiable instances.
+algorithm on unsatisfiable instances.  For value ordering, which occurs on
+\siplineref{valueordering} of \cref{algorithm:sip}, we will compare four strategies:
 
-For value ordering, which occurs on \siplineref{valueordering} of \cref{algorithm:sip}, we will
-compare five strategies:
+\paragraph{Random} We select vertices uniformly at random.
 
-\paragraph{Random} We select a vertex dynamically at random (that is, we shuffle the vertices before
-entering the $\KwSty{foreach}$ loop on \siplineref{valueordering} of \cref{algorithm:sip}). This is
-\emph{usually} effectively equivalent to picking vertices in the order in which they appear in the
-input files, although some instances from some benchmark suites are naturally in a special order
-(which may or may not be a good order for searching).
-
-\paragraph{Degree} We select vertices from highest degree to lowest degree. This is done by
-permuting the target graph at the top of search. This heuristic is the current state of the art.
+\paragraph{Degree} We select vertices from highest degree to lowest degree. This heuristic is the
+current state of the art.
 
@@
 
-\paragraph{Position-biased} We permute the graph at the top of search, as in the degree heuristic.
-Then, as with the random heuristic, we shuffle the vertices before we enter the $\KwSty{foreach}$
-loop. Instead of picking a vertex uniformly at random, we select select the first (that is, highest
-degree) vertex with probability 0.5, the second vertex with probability 0.25, and the $n$th with
-probability $2^{-n}$.  This heuristic heavily prefers vertices of high degree, similar to the
-degree heuristic, but also introduces an element of randomness, and so can be used with restarts.
+\paragraph{Biased} We branch by selecting a vertex $v'$ from the chosen domain $D_v$ with
+probability \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \text{.} \] This heuristic
+continues to prefer vertices of high degree, but will give the same chance of being selected next to
+two vertices of equal degree.  It also introduces an element of randomness, which is needed to make
+restarts have an effect.
 
-\paragraph{Degree-biased} Refining position-biased, we shuffle the vertices, using a weight of
-$2^{\left|\neighbourhood(v)\right|}$ when deciding whether to branch on vertex $v$ next. (We cheat
-slightly to avoid range issues with random number generation: we give the vertex with highest degree
-in the entire graph weight $2^{50}$, and then give lower degree vertices smaller powers, giving
-every vertex weight at least 1.  This is a bit of hack but programming is hard.) This heuristic
-continues to prefer vertices of high degree, but will give the same chance of being selected next to
-two vertices of equal degree.
+The biased heuristic resembles the well-known \emph{softmax} weighting scheme applied to the degree
+heuristic, although usually \emph{softmax} uses a base of $e$ rather than $2$.  Experiments showed
+that in this context, the choice of base is not important to the search space size; however, a base
+of $2$ avoids the need to use floating point arithmetic, which was a substantial overhead.
 
@@
 
-?? Dual Xeon E5-2697A v4 CPUs, 512GBytes RAM, running Ubuntu 17.04, implemented in C++ and compiled
-using GCC 6.3.0. Our code is heavily based upon \citet{dblp:conf/cp/McCreeshP15}.
+Our experiments are performed on systems with dual Intel Xeon E5-2697A v4 CPUs and 512GBytes RAM,
+running Ubuntu 17.04. We implemented\footnote{Institutional DOI URL removed for anonymous review}
+\cref{algorithm:sip} in C++ using the bit-parallel implementation of
+\citet{DBLP:conf/aaai/HoffmannMR17} as a starting point, and compiled it using GCC 6.3.0. We use a
+deterministic pseudo-random number generator for reproducibility.
 
-?? Deterministic RNG.
-
-?? Bitsets.
-
-?? The 5,725 instances \citep{dblp:conf/lion/KotthoffMS16}.
-
-?? Other studies use a random selection of 200 of each of the mesh and images instances
-\citep{DBLP:journals/cviu/DamiandSHJS11}, partly because some earlier solvers find many of these
-instances extremely hard, and partly to avoid skewing a portfolio. We want larger numbers of
-satisfiable instances, so we include all pattern / target pairs. This gives a total of 14,621
-instances, at least 2,126 of which are known to be satisfiable, and at least 12,323 are
-unsatisfiable.
-
-?? Note that an awful lot of these instances are very easy. We are trying to tackle the remaining
-hard satisfiable instances.
+We use the dataset introduced by \citet{DBLP:conf/lion/KotthoffMS16} for evaluation. This dataset
+brings together a range of randomly-generated and application instance families from earlier papers.
+Other studies use a random selection of 200 of each of the instances from the ``meshes'' and
+``images'' families \citep{DBLP:journals/cviu/DamiandSHJS11} because some earlier solvers find many
+of these instances extremely hard. We would like to have a larger number of satisfiable instances in
+our test set, and so we include all pattern / target pairs. This gives a total of 14,621 instances
+(rather than the original 5,725). At least ?? of these instances are known to be satisfiable, and at
+least ?? are unsatisfiable.
 
@@
 
-    \caption{Cumulative number of instances solved over time, comparing our algorithm to other
-    solvers. Below, zoomed in a bit ?? make this pretty.}
+    \caption{Above, the cumulative number of instances solved over time, comparing our algorithm
+    (both in its basic form, and with the improvements introduced in the remainder of the paper) to
+    other solvers. Below, the same, looking only at satisfiable instances.}
     \label{figure:others}
@@
 
-?? In \cref{figure:others} we show that our solver is the bestest. For each algorithm, the $y$ value
+We begin by verifying that our implementation of the basic \cref{algorithm:sip} is competitive. In
+\cref{figure:others} we show the cumulative number of instances solved over time, using our
+implementation (both with the degree heuristic, and with the modifications described in the
+remainder of this paper), the Glasgow2 and Glasgow3 algorithms from which \cref{algorithm:sip} is derived
+\citep{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}, the PathLAD variation of the LAD
+algorithm \citep{DBLP:journals/ai/Solnon10,DBLP:conf/lion/KotthoffMS16}, and VF2
+\citep{DBLP:journals/pami/CordellaFSV04}. For each algorithm, the $y$ value
 gives the cumulative number of instances which (individually) can be solved in no more than $x$
@@
 
-?? These results show our algorithm clearly beating PathLAD and VF2, except for very low choices of
-timeout. Our algorithm also beats Glasgow2 and Glasgow3, despite being slightly simpler---on an
-instance by instance basis, sometimes the additional inference performed by these algorithms pays
-off, but in aggregate, there is no difference for longer choices of timeout, and a clear gain for
-shorter choices.
+As expected, the performance of \cref{algorithm:sip} with the degree heuristic is very close to that
+of Glasgow2, although we perform better at lower runtimes because we use simpler preprocessing at
+the top of search. Our implementation also clearly beats PathLAD and VF2, except for very low
+choices of timeout.
 
-?? Add in SND if we can get it to run on the larger set of instances, maybe SAT just to have
-something to cite showing how bad SAT is at this.
-
-?? Possibly also do induced, compare to Glasgow, LAD, PathLAD, VF2, VF3, McSplit.
-
-\subsection{Value Ordering Heuristics}
-
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-value-ordering-heuristics.pdf}
-
-    \caption{The cumulative number of instances solved over time, using the four different
-    value-ordering heuristics.}
-    \label{figure:value-ordering-heuristics}
-\end{figure}
-
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-scatter-heuristics.pdf}
-
-    \caption{An instance by instance comparison of the degree and degree-biased heuristics.}
-    \label{figure:scatter-heuristics}
-\end{figure}
-
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-scatter-dds.pdf}
-
-    \caption{The effects of using depth-bounded discrepancy search.}\label{figure:scatter-dds}
-\end{figure}
-
-In \cref{figure:value-ordering-heuristics} we show the cumulative number of instances solved over
-time using the four different value-ordering heuristics. As expected, the degree heuristic is much
-better (?? give aggregate speedups for both all and just satisfiable instances) than random
-ordering, which is in turn ?? times better than the anti-heuristic.
-
-The performance of both of the biased heuristics and the degree heuristic are very similar: the
-degree-biased heuristic has slightly worse performance below a 100s timeout, and effectively equal
-performance above it. This shows that we can introduce an element of randomness into the degree
-value-ordering heuristic without adversely affecting its performance \emph{in aggregate}.
-\Cref{figure:scatter-heuristics} gives a more detailed look at these results. In this plot,
-satisfiable and unsatisfiable instances are shown using coloured shapes and black dots respectively.
-On the $x$-axis we measure the degree heuristic, whilst on the $y$-axis we measure the degree-biased
-heuristic. Points on the outer axes represent instances which timed out with one heuristic but not
-the other.  This plot shows that despite the aggregate performance being very similar, on an case by
-case basis, the two heuristics can make a large difference to the performance of satisfiable
-instances.
+The dataset includes many instances which are extremely easy for a good solver, and so it can be
+hard to see the differences between the stronger solvers at higher runtimes. This paper focusses
+upon improving the performance on the remaining hard satisfiable instances, and so in the second
+plot in \cref{figure:others} (and in subsequent cumulative plots) we show only satisfiable
+instances, and use a reduced range on both axes.
 
@@
 \citep{DBLP:conf/ijcai/HarveyG95,DBLP:conf/aaai/Korf96,DBLP:conf/ijcai/Walsh97,DBLP:conf/cpaior/KarouiHLN07,DBLP:journals/jea/ProsserU11}.
-Discrepancy search is about value ordering.
+Discrepancy search is about value ordering, and it mostly looks at optimisation problems and at
+problems where an optimality proof is not needed.
 
@@
 better in aggregate (even when only considering satisfiable instances).
+
+\subsection{Value Ordering Heuristics}
+
+\begin{figure}[p]
+    \centering
+    \includegraphics*{gen-graph-scatter-dds.pdf}
+
+    \caption{Comparing depth-bounded discrepancy search to backtracking search, both with the degree
+    heuristic.}\label{figure:scatter-dds}
+\end{figure}
+
+\begin{figure}[p]
+    \centering
+    \includegraphics*{gen-graph-value-ordering-heuristics.pdf}
+
+    \caption{The cumulative number of satisfiable instances solved over time, using four
+    different value-ordering heuristics. The degree and biased lines are close together.}
+    \label{figure:value-ordering-heuristics}
+\end{figure}
+
+\begin{figure}[p]
+    \centering
+    \includegraphics*{gen-graph-scatter-heuristics.pdf}
+
+    \caption{An instance by instance comparison of the degree and biased heuristics on all
+    instances. Points on the outer axes are timeouts, and point style shows instance family.}
+    \label{figure:scatter-heuristics}
+\end{figure}
+
+In \cref{figure:value-ordering-heuristics} we show the cumulative number of instances solved over
+time using the four different value-ordering heuristics. As expected, the degree heuristic is much
+better (?? give aggregate speedups for both all and just satisfiable instances) than random
+ordering, which is in turn ?? times better than the anti-heuristic.
+
+The performance of the two degree heuristics are very similar: the
+biased heuristic has slightly worse performance below a 100s timeout, and effectively equal
+performance above it. This shows that we can introduce an element of randomness into the degree
+value-ordering heuristic without adversely affecting its performance \emph{in aggregate}.
+\Cref{figure:scatter-heuristics} gives a more detailed look at these results. In this plot,
+satisfiable and unsatisfiable instances are shown using coloured shapes and black dots respectively.
+On the $x$-axis we measure the degree heuristic, whilst on the $y$-axis we measure the biased
+heuristic. Points on the outer axes represent instances which timed out with one heuristic but not
+the other.  This plot shows that despite the aggregate performance being very similar, on an case by
+case basis, the two heuristics can make a large difference to the performance of satisfiable
+instances.
 
@@
 
-    \caption{Cumulative number of instances solved over time, with (dotted lines) and without (solid
-    lines) restarts, and with depth-bounded discrepancy search (dashed line).}
+    \caption{The cumulative number of satisfiable instances solved over time, with and without
+    restarts.}
     \label{figure:restarts}
@@
     \includegraphics*{gen-graph-scatter-random.pdf}
-    \caption{Instance by instance comparison of the basic algorithm with the degree heuristic,
-    versus random value-ordering with restarts.}
+    \caption{Comparing basic backtracking with the degree heuristic, versus the Random heuristic
+    with restarts.}
     \label{figure:scatter-random}
@@
 
-    \caption{Instance by instance comparison of the basic algorithm with the degree heuristic,
-    versus degree-biased value ordering with restarts.}
+    \caption{Comparing basic backtracking with the degree heuristic, versus the biased
+    heuristic with restarts.}
     \label{figure:scatter-by-family}
@@
 
+?? Justify aggregate speedups on sat instances
+
+\subsection{Further Experiments}
+
+\begin{itemize}
+    \item The choice of Luby multiplier is not hugely important, although 100 is too small and 1,000
+        is too big.
+    \item The base of the exponential does not have to be $e$, and it may be easier on implementers
+        to use 2. Using a polynomial weighting is bad, and using a base of $|D_v|$ is bad.
+    \item Using geometric restarts instead of Luby does not give much of a benefit. A constant
+        restart policy is also fine (and terminates on unsatisfiable instances because we count
+        backtracks).
+    \item Using rank with restarts is better than random with restarts, but not as good as using
+        biased with restarts.
+    \item Nogood recording is important.
+    \item Nogood recording is cheap.
+    \item Discarding subsumed nogoods may be done cheaply, if memory space is a concern.
+\end{itemize}
+
 \section{Conclusion}
@@
 
+?? No theoretical justification for biased. Also tried polynomial weightings, exponential using
+$\left|D_v\right|$ as the base. Using $e$ (or for that matter, $2$ or $\pi$) as a base seems to be
+``just the right amount'' of randomness.
+
 ?? Variable ordering
@@
 
+?? Parallel future work
+
 \begin{figure}[tb]
     \centering
-    \includegraphics*{gen-graph-scatter-parallel.pdf}
-
-    \caption{Parallel}
-    \label{figure:scatter-parallel}
+    \includegraphics*{gen-graph-as.pdf}
+    \caption{?? Aggregate speedups on satisfiable instances}
 \end{figure}
@@
     \centering
-    \includegraphics*{gen-graph-final.pdf}
+    \includegraphics*{gen-graph-scatter-degree.pdf}
 
-    \caption{Final results}
-    \label{figure:final}
+    \caption{When is the degree heuristic good?}
+    \label{figure:scatter-degree}
 \end{figure}
-
-?? Parallel
 
@@
 
-\appendix
-\clearpage
-\section{Further Experiments}
-
-?? engineering aspects: look at what happens without nogoods in \cref{figure:scatter-nogoods}, and
-at performance with restarts on degree in \cref{figure:scatter-overheads}.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics*{gen-graph-scatter-nogoods.pdf}
-
-    \caption{Is nogood recording important?}
-    \label{figure:scatter-nogoods}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics*{gen-graph-scatter-overheads.pdf}
-
-    \caption{Is nogood recording expensive?}
-    \label{figure:scatter-overheads}
-\end{figure}
-
-?? effects of restart policies, \cref{figure:restart-policies}.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics*{gen-graph-restart-policies.pdf}
-
-    \caption{Different restart policies.}
-    \label{figure:restart-policies}
-\end{figure}
-
 \end{document}

### 2017-12-13 paper/asify.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+
+join -1 2 -2 2 <(grep 'i$' $1 | sort -k2 ) <(grep 'i$' $2 | sort -k2 ) | cut -d' ' -f1,2,4 | sort -n -k3,3

### 2017-12-13 paper/paper.tex
Edity

@@
 The performance of the two degree heuristics are very similar: the
-biased heuristic has slightly worse performance below a 100s timeout, and effectively equal
+biased heuristic has slightly worse performance below a 400s timeout, and effectively equal
 performance above it. This shows that we can introduce an element of randomness into the degree
@@
 randomness to a tailored deterministic heuristic than it is to throw away the heuristic altogether.
-The degree-biased heuristic is also stronger than position-biased, showing that exploiting ties and
-relative sizes in the heuristic ordering is also worthwhile. Interestingly, the gap between degree-
-and position-based is much larger when using restarts than it is when not using them.
 
@@
         is too big.
-    \item The base of the exponential does not have to be $e$, and it may be easier on implementers
-        to use 2. Using a polynomial weighting is bad, and using a base of $|D_v|$ is bad.
-    \item Using geometric restarts instead of Luby does not give much of a benefit. A constant
+    \item The base of the exponential can be $2$ or $e$ or $3$ or $1.5$ and it makes little
+        difference. Using a polynomial weighting is bad, and using a base of $|D_v|$ is bad.
+    \item Using geometric restarts instead of Luby does not give as much of a benefit. A constant
         restart policy is also fine (and terminates on unsatisfiable instances because we count

### 2017-12-13 mcsplit-code/solve_mcs.cc
Don't try to remove clauses subsumed by the empty clause

@@ -556,6 +556,7 @@ namespace {
             return count == a.literals.size();
         }
 
+        // precondition: the nogood that n_it points to is not empty
         auto remove_subsumed_clauses(Nogoods::iterator n_it) -> void {
             auto & n = *n_it;
             // `literal_with_lowest_tally` will be an iterator to the literal in `n` that appears in
@@ -601,11 +602,10 @@ namespace {
                 ++number_of_restarts;
 
                 for (auto & n : need_to_watch) {
-                    remove_subsumed_clauses(n);
-
                     if (n->literals.empty()) {
                         return;
                     } else {
+                        remove_subsumed_clauses(n);
                         watches[n->literals[0]].push_back(n);
                     }
                 }

### 2017-12-14 paper/paper.tex
I have faith in James!

@@
 
-\title{Value Ordering, Discrepancies, and Restarts for Subgraph Isomorphism}
+\title{Value Ordering, Discrepancies, and Restarts for Subgraph Algorithms}
 \author{Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences
@@
     Modern subgraph isomorphism solvers use value-ordering heuristics to direct a constraint
-    programming style search. These heuristics are far from perfect, and experiments show that using
-    a random value-ordering heuristic with restarts beats conventional search with the best known
-    tailored heuristic. This motivates the introduction of a slightly-random value-ordering
-    heuristic with similar behaviour to the tailored heuristic: when combined with restarts, this
-    new search strategy is over a hundred times more effective on satisfiable instances, and does
-    not significantly decrease performance on unsatisfiable instances. In contrast,
-    discrepancy-based searches are too expensive to be effective in this setting.
+    programming style search. Although these heuristics are far from perfect (particularly early on
+    in search), discrepancy-based searches are too expensive to be effective in this setting.
+    However, we show that a random value-ordering heuristic with restarts beats conventional search
+    with the best-known tailored heuristic. This motivates the introduction of a slightly-random
+    value-ordering heuristic with similar behaviour to the tailored heuristic: when combined with
+    restarts, this new search strategy is over ?? a hundred times more effective on satisfiable
+    instances, and does not significantly decrease performance on unsatisfiable instances. Finally,
+    we show that this strategy also improves two maximum common induced subgraph algorithms.
 \end{abstract}
@@
 discrepancy searches (\cref{figure:scatter-dds}) or by random value-ordering with restarts
-(\cref{figure:scatter-random}).
+(\cref{figure:scatter-random}). Finally, we show that this strategy is also effective in an
+optimisation setting, producing clear benefits in two maximum common induced subgraph algorithms.
 
@@
 }
-\bigskip
+\medskip
     \nl $\FuncSty{search}$ (Domains $D$, Decisions $B$) $\rightarrow$ Bool
@@
 }
-\bigskip
+\medskip
 \nl $\FuncSty{propagate}$ (Domains $D$) $\rightarrow$ Bool \;
@@
 
+\section{Maximum Common Subgraph Algorithms}
+
+\begin{figure}[p]
+    \centering
+    \includegraphics*{gen-graph-scatter-mcsplit.pdf}
+    \caption{McSplit speedups}
+\end{figure}
+
+\begin{figure}[p]
+    \centering
+    \includegraphics*{gen-graph-scatter-kdown.pdf}
+    \caption{k${\downarrow}$ speedups}
+\end{figure}
+
+\begin{figure}[p]
+    \centering
+    \includegraphics*{gen-graph-mcs.pdf}
+    \caption{k${\downarrow}$ and McSplit cumulative curves.}
+\end{figure}
+
 \section{Conclusion}
@@
 ?? Parallel future work
+
+\section{Extra Plots not for the Final Paper}
 

### 2017-12-14 paper/paper.tex
Definite faith

@@
 
-\title{Value Ordering, Discrepancies, and Restarts for Subgraph Algorithms}
-\author{Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences
-    Research Council [grant number EP/026842/1]}\\ University of Glasgow, Glasgow, Scotland \\
+\title{Value Ordering, Discrepancies, and Restarts for Subgraph Algorithms\thanks{This work was
+supported by the Engineering and Physical Sciences Research Council [grant numbers EP/026842/1 and EP/M508056/1]}}
+\author{Ciaran McCreesh \and Patrick Prosser \and James Trimble\\ University of Glasgow, Glasgow, Scotland \\
     ciaran.mccreesh@glasgow.ac.uk}

### 2017-12-14 paper/paper.tex
More thingies

@@
 first'' as a value-ordering heuristic to direct a constraint programming style search
-\citep{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:conf/ijcai/McCreeshPT16}.  This is
-much better than not having a value-ordering heuristic at all, but it is far from perfect,
-particularly for early branching choices. There are two reasons to suspect a better option may
-possible.  Firstly, even if degree were a perfect source of information, many target graphs do not
-have a large degree distribution, and so we would still need to determine \emph{which} vertex of
-highest degree would be best.  Secondly, experiments with a parallel search strategy which
-explicitly steals early branching choices often gives strongly super-linear speedups on satisfiable
-instances \citep{DBLP:conf/cp/McCreeshP15}.
+\citep{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:conf/ijcai/McCreeshPT16}. This
+heuristic is far from perfect. Furthermore, even if degree were a perfect source of information,
+many target graphs do not have a large degree distribution, so we would still need to determine
+\emph{which} vertex of highest degree would be best.
 
@@
 
+?? Two recent maximum common induced subgraph algorithms: k${\downarrow}$
+\citep{DBLP:conf/aaai/HoffmannMR17}, and McSplit \citep{DBLP:conf/ijcai/McCreeshPT17}. The former
+attempts to solve the problem by first trying to solve the induced subgraph isomorphism problem, and
+then if that fails, retries allowing a single unmatched vertex (and thus using weaker invariants),
+and so on. Due to its similarity to \cref{algorithm:sip}, we can introduce the same bias and restart
+strategy.
+
+The McSplit algorithm, meanwhile, uses a constraint programming style search, but with special
+propagators and backtrackable data structures that exploit special properties of the problem. The
+unconventional domain store used by McSplit precludes the use of arbitrary unit propagation, and so
+when introducing restarts, we cannot propagate using nogoods. Instead, we can only detect when we
+are inside an already-visited branch. We must therefore use the one watched literal scheme instead.
+
+?? Benchmark results on the same families of instances are shown in
+\cref{figure:mcsplit,figure:kdown,figure:mcs}. Clearly beneficial. Note that on McSplit we're doing
+extra work and this shows, but it's still worthwhile.
+
 \begin{figure}[p]
@@
     \includegraphics*{gen-graph-scatter-mcsplit.pdf}
-    \caption{McSplit speedups}
+    \caption{McSplit speedups ?? partial data}\label{figure:mcsplit}
 \end{figure}
@@
     \includegraphics*{gen-graph-scatter-kdown.pdf}
-    \caption{k${\downarrow}$ speedups}
+    \caption{k${\downarrow}$ speedups ?? missing data}\label{figure:kdown}
 \end{figure}
@@
     \includegraphics*{gen-graph-mcs.pdf}
-    \caption{k${\downarrow}$ and McSplit cumulative curves.}
+    \caption{k${\downarrow}$ and McSplit cumulative curves ?? partial and missing
+    data}\label{figure:mcs}
 \end{figure}
@@
 
+\clearpage
+\bibliographystyle{named}
+\bibliography{dblp}
+
+\clearpage
+\appendix
+
 \section{Extra Plots not for the Final Paper}
 
-\begin{figure}[tb]
+\begin{figure}[h]
     \centering
@@
 
-\begin{figure}[tb]
+\begin{figure}[h]
+    \centering
+    \includegraphics*{gen-graph-as-mcs.pdf}
+    \caption{?? Aggregate speedups on MCS instances}
+\end{figure}
+
+\begin{figure}[h]
     \centering
@@
 
-\bibliographystyle{named}
-\bibliography{dblp}
-
 \end{document}

### 2017-12-14 mcsplit-code/solve_mcs.cc
Random insertion point for "do not use this vertex" choice

@@ -483,8 +483,15 @@ namespace {
             if (params.biased_shuffle)
                 biased_shuffle(possible_values);
 
-            if (bound != incumbent.size() + 1 || bd.left_len > bd.right_len)
-                possible_values.push_back(-1);
+            if (bound != incumbent.size() + 1 || bd.left_len > bd.right_len) {
+                if (params.biased_shuffle) {
+                    std::uniform_int_distribution<unsigned> dist(0, possible_values.size());
+                    unsigned insertion_point = dist(global_rand);
+                    possible_values.insert(next(possible_values.begin(), insertion_point), -1);
+                } else {
+                    possible_values.push_back(-1);
+                }
+            }
 
             remove_vtx_from_left_domain(domains[bd_idx], v);
             bd.right_len--;
@@ -502,9 +509,11 @@ namespace {
                     search_result = restarting_search(current, new_domains, backtracks_until_restart);
                 } else {
                     bd.right_len++;
+                    auto copied_domains = domains;
                     if (bd.left_len == 0)
-                        remove_bidomain(domains, bd_idx);
-                    search_result = restarting_search(current, domains, backtracks_until_restart);
+                        remove_bidomain(copied_domains, bd_idx);
+                    search_result = restarting_search(current, copied_domains, backtracks_until_restart);
+                    bd.right_len--;
                 }
                 current.pop();
                 switch (search_result)

### 2017-12-14 mcsplit-code/solve_mcs.cc
Immediately on posting a nogood, start watching it and remove subsumed nogoods

@@ -159,7 +159,6 @@ namespace {
         vector<Assignment> incumbent;
         Nogoods nogoods;
         Watches watches;
-        list<typename Nogoods::iterator> need_to_watch;
         Watches literal_clause_membership;
 
         vector<int> vtx_current_assignment;
@@ -345,7 +344,7 @@ namespace {
             if (!params.quiet) cout << "Incumbent size: " << incumbent.size() << endl;
         }
 
-        auto post_nogood(const VarAssignments & current) -> void
+        auto post_nogood(const VarAssignments & current) -> bool
         {
             nogoods.push_back({});
             auto & nogood = nogoods.back();
@@ -354,15 +353,21 @@ namespace {
                 if (a.is_decision)
                     nogood.literals.emplace_back(a.assignment);
 
-            auto nogood_it = prev(nogoods.end());
+            if (nogood.literals.empty())
+                return false;
 
-            need_to_watch.emplace_back(nogood_it);
+            auto nogood_it = prev(nogoods.end());
 
             for (auto & a : nogood.literals) {
                 auto & lcm = literal_clause_membership[a];
                 lcm.emplace_back(nogood_it);
                 nogood.add_clause_membership(lcm, prev(lcm.end()));
             }
+
+            remove_subsumed_clauses(nogood_it);
+            watches[nogood.literals[0]].push_back(nogood_it);
+
+            return true;
         }
 
         auto erase_nogood(Nogoods::iterator nogood_it) -> void
@@ -523,7 +528,6 @@ namespace {
                         if (u == w)
                             break;
                         current.push({{v, u}, true});
-    //                        std::cout << "sz " << current.get_var_assignments().size() << endl;
                         post_nogood(current);
                         current.pop();
                     }
@@ -536,8 +540,8 @@ namespace {
             }
 
             if (backtracks_until_restart > 0 && 0 == --backtracks_until_restart) {
-    //            std::cout << "sz " << current.get_var_assignments().size() << endl;
-                post_nogood(current);
+                if (! post_nogood(current))
+                    return Search::Done;
                 return Search::Restart;
             } else {
                 return Search::Done;
@@ -609,16 +613,6 @@ namespace {
                 ++current_luby;
 
                 ++number_of_restarts;
-
-                for (auto & n : need_to_watch) {
-                    if (n->literals.empty()) {
-                        return;
-                    } else {
-                        remove_subsumed_clauses(n);
-                        watches[n->literals[0]].push_back(n);
-                    }
[…]

### 2017-12-14 mcsplit-code/solve_mcs.cc
Check for unsatisfied nogoods, even if last decision was forced

@@ -465,11 +465,11 @@ namespace {
                 update_incumbent(incumbent, current);
 
             unsigned int bound = current.get_num_vtx_assignments() + calc_bound(domains);
+
             if (bound <= incumbent.size())
                 return Search::Done;
 
-            if (current.get_var_assignments().size() != 0 &&
-                    current.get_var_assignments().back().is_decision &&
+            if (!current.get_var_assignments().empty() &&
                     contains_a_nogood(current, domains))
                 return Search::Done;
 

### 2017-12-14 mcsplit-code/solve_mcs.cc
Decrease backtracks_until_restart in more cases

@@ -424,6 +424,9 @@ namespace {
         auto contains_a_nogood(VarAssignments & current,
                 vector<Bidomain> & domains) -> bool
         {
+            if (current.get_var_assignments().empty())
+                return false;
+
             std::fill(vtx_current_assignment.begin(), vtx_current_assignment.end(), -1);
 
             for (auto & bd : domains) {
@@ -466,76 +469,73 @@ namespace {
 
             unsigned int bound = current.get_num_vtx_assignments() + calc_bound(domains);
 
-            if (bound <= incumbent.size())
-                return Search::Done;
+            int bd_idx;
 
-            if (!current.get_var_assignments().empty() &&
-                    contains_a_nogood(current, domains))
-                return Search::Done;
+            if (bound > incumbent.size() &&
+                    !contains_a_nogood(current, domains) &&
+                    -1 != (bd_idx = select_bidomain(domains, current.get_num_vtx_assignments())))
+            {
+                Bidomain &bd = domains[bd_idx];
 
-            int bd_idx = select_bidomain(domains, current.get_num_vtx_assignments());
-            if (bd_idx == -1)   // In the MCCS case, there may be nothing we can branch on
-                return Search::Done;
-            Bidomain &bd = domains[bd_idx];
+                int v = find_min_value(left, bd.l, bd.left_len);
 
-            int v = find_min_value(left, bd.l, bd.left_len);
+                // Try assigning v to each vertex w in the colour class beginning at bd.r, in turn
+                std::vector<int> possible_values(right.begin() + bd.r,  // the vertices in the colour class beginning at bd.r
+                        right.begin() + bd.r + bd.right_len);
+                std::sort(possible_values.begin(), possible_values.end());
 
-            // Try assigning v to each vertex w in the colour class beginning at bd.r, in turn
-            std::vector<int> possible_values(right.begin() + bd.r,  // the vertices in the colour class beginning at bd.r
-                    right.begin() + bd.r + bd.right_len);
-            std::sort(possible_values.begin(), possible_values.end());
+                if (params.biased_shuffle)
+                    biased_shuffle(possible_values);
 
-            if (params.biased_shuffle)
-                biased_shuffle(possible_values);
-
-            if (bound != incumbent.size() + 1 || bd.left_len > bd.right_len) {
-                if (params.biased_shuffle) {
-                    std::uniform_int_distribution<unsigned> dist(0, possible_values.size());
-                    unsigned insertion_point = dist(global_rand);
[…]

### 2017-12-14 paper/paper.tex
More writing

@@
 \newcommand{\citep}[1]{\cite{#1}}
+
+\newcommand{\siplabel}[1]{\label{line:sip:#1}}
+\newcommand{\siplineref}[1]{line~\ref{line:sip:#1}}
+\newcommand{\siplinerangeref}[2]{lines~\ref{line:sip:#1} to~\ref{line:sip:#2}}
 
@@
 \citep{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:conf/ijcai/McCreeshPT16}. This
-heuristic is far from perfect. Furthermore, even if degree were a perfect source of information,
+heuristic is far from perfect---and even if degree were a perfect source of information,
 many target graphs do not have a large degree distribution, so we would still need to determine
@@
 
-\section{The Basic Algorithm}
+\section{Definitions and Algorithms}
 
-?? Non-induced!
+Let $G$ be a graph. We denote its set of vertices by $\vertexset(G)$, and write $v \sim_G w$ to mean
+that vertices $v$ and $w$ are adjacent. The \emph{neighbourhood} of a vertex, $\neighbourhood_G(v)$,
+is the set of vertices adjacent to $v$, and the \emph{degree} of $v$, $\deg_G(v)$, is the
+cardinality of its neighbourhood; when the graph is clear from the context, we omit the $G$
+subscript. The \emph{neighbourhood degree sequence} of $v$, $\nds_G(v)$, is the sequence of degrees
+of the neighbours of $v$ in descending order, and we denote by $\preceq$ the inclusion relation of
+\citet{DBLP:journals/constraints/ZampelliDS10}. As per \citet{DBLP:conf/aaai/HoffmannMR17},
+$G^{n,\ell}$ denotes the graph with vertex set $\vertexset(G)$, and with edges between vertices $v$
+and $v$ if there are at least $n$ simple paths of length exactly $\ell$ between vertices $v$ and $w$
+in $G$.
 
-?? Notation: $\vertexset(G)$, $\neighbourhood_G(v)$, $\nds_G(v)$, $\preceq$, $\sim_G$, $G^{a,b}$, $\mapsto$.
+The non-induced subgraph isomorphism problem is to find an injective mapping from the vertices of a
+pattern graph $\mathcal{P}$ to a target graph $\mathcal{T}$, such that adjacent vertices in
+$\mathcal{P}$ are mapped to adjacent vertices in $\mathcal{T}$. We write $v \mapsto v'$ to mean that
+pattern vertex $v$ is mapped to target $v'$ under such a mapping.
 
-?? Started with the Glasgow algorithm, removed parts which did not contribute hugely to speedups,
-and ended up with a simpler algorithm that performs as well in practice. In particular, because we
-will be working with reasonably large target graphs, no ILF, and not doing supplemental graphs of
-length 3. This is \cref{algorithm:sip}, ignoring the shaded parts, which are introduced in the
-context of restarts.
+\subsection{A Subgraph Isomorphism Algorithm}
 
-?? Briefly describe this, probably referring elsewhere as far as possible to save space.
+We now briefly outline the non-shaded parts of \cref{algorithm:sip}, which is how we will solve the
+subgraph isomorphism problem. This algorithm is very closely based upon the $k{\downarrow}$
+algorithm of \citet{DBLP:conf/aaai/HoffmannMR17} with $k = 0$, and we refer the reader to that paper
+for full technical details; that algorithm, in turn, is a simplification of the Glasgow2 algorithm
+\citep{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}.
 
-\newcommand{\siplabel}[1]{\label{line:sip:#1}}
-\newcommand{\siplineref}[1]{line~\ref{line:sip:#1}}
+\Cref{algorithm:sip} works by recursively building up a valid mapping of pattern vertices to target
+vertices. For each pattern vertex $v$, we have a domain $D_v$ containing a set of feasible target
+vertices. Each domain is initialised at the top of search (\siplinerangeref{initstart}{initend}),
+using the invariants described by \citet{DBLP:conf/aaai/HoffmannMR17} to restrict its initial
+values.
+
+The $\FuncSty{search}$ procedure recursively selects an unassigned domain $D$
+(\siplineref{variableordering}), and tries assigning
+it each of its remaining value in turn (\siplineref{valueordering}). The effects of this assignment
+are propagated (\siplineref{dopropagate}). If an
+inconsistency is detected, we continue and try another value; otherwise, we recurse
+(\siplineref{recurse}) until a solution is found. Finally, if we run out of values, we backtrack
+(\siplineref{backtrack}).
+
+Our $\FuncSty{propagate}$ routine is similarly close to that of \citet{DBLP:conf/aaai/HoffmannMR17},
+differing only in that we carry out the filtering for injectivity
+(\siplinerangeref{alldiffstart}{alldiffend}) after every assignment rather than once at the end of
+the unit propagation loop. (We make this change because the all-different propagator used is not
+idempotent, and we wish to avoid an odd interaction when we introduce nogood filtering.)
+
 \begin{algorithm}[p]\DontPrintSemicolon\small\SetInd{0.1em}{1em}
@@
         (\mathcal{P}^{4,2},\ \mathcal{T}^{4,2}) \big]$ \;
-    \nl \ForEach{$v \in \vertexset(\mathcal{P})$}{
+    \nl \ForEach{\siplabel{initstart}$v \in \vertexset(\mathcal{P})$}{
         \nl $D_v \gets \vertexset(\mathcal{T})$\;
@@
             \nl $D_v \gets \{ w \in D_v : $ \siplabel{degreefiltering} \\\hspace*{4em}$v \sim_P v \Rightarrow w \sim_T w~\wedge$
-            \\\hspace*{4em}$\nds_P(v) \preceq \nds_T(w)\}$ \;
+            \\\hspace*{4em}$\nds_P(v) \preceq \nds_T(w)\}$\siplabel{initend} \;
         }
@@
         \nl $D'_v \gets \{ v' \} $ \;
-        \nl \If{\FuncSty{propagate}(D')}{
-            \nl $S \gets \FuncSty{search}(\{ E \in D' : \left|E\right| > 1 \}, B \wedge (v \mapsto v'))$ \;
+        \nl \If{\FuncSty{propagate}(D')\siplabel{dopropagate}}{
+            \nl $S \gets \FuncSty{search}(\{ E \in D'\,{:}\,\left|E\right|{>}\,1 \}, B \wedge (v\,{\mapsto}\,v'))$\siplabel{recurse} \;
             \nl \lIf{$S = \KwSty{true}$}{$\KwSty{return}~\KwSty{true}$}
@@
     \nl \tikzmark{restart6s}\lIf{$(R \gets R + 1) = R'$}{$\KwSty{return}~\KwSty{restart}\tikzmark{restart6e}$}
-    \nl $\KwSty{return}~\KwSty{false}$
+    \nl $\KwSty{return}~\KwSty{false}$\siplabel{backtrack}
 }
@@
         }
-\nl $(H,\,A,\,n) \gets (\emptyset,\,\emptyset,\,0)$ \;
+\nl $(H,\,A,\,n) \gets (\emptyset,\,\emptyset,\,0)$\siplabel{alldiffstart} \;
 \nl \ForEach{$D_v \in D$ \textnormal{from smallest cardinality upwards\siplabel{eachdomain}}}{
@@
     \nl \lIf{$D_v = \emptyset~\vee~|A| < n$}{$\KwSty{return}~\KwSty{false}$\siplabel{failhall}}
-    \nl \lIf{$|A| = n$}{$(H,\,A,\,n) \gets (H \cup A,\,\emptyset,\,0)$\siplabel{hall}}
+    \nl \lIf{$|A| = n$}{$(H,\,A,\,n) \gets (H \cup A,\,\emptyset,\,0)$\siplabel{hall}}\siplabel{alldiffend}
 }
@@
 }
-\caption{Non-induced subgraph isomorphism. Shaded code is added for the restarting variant.}
+\caption{Non-induced subgraph isomorphism. The shaded code is added for the restarting variant.}
 \label{algorithm:sip}

### 2017-12-15 paper/paper.tex
More things

@@
     Modern subgraph isomorphism solvers use value-ordering heuristics to direct a constraint
-    programming style search. Although these heuristics are far from perfect (particularly early on
-    in search), discrepancy-based searches are too expensive to be effective in this setting.
-    However, we show that a random value-ordering heuristic with restarts beats conventional search
-    with the best-known tailored heuristic. This motivates the introduction of a slightly-random
-    value-ordering heuristic with similar behaviour to the tailored heuristic: when combined with
-    restarts, this new search strategy is over ?? a hundred times more effective on satisfiable
-    instances, and does not significantly decrease performance on unsatisfiable instances. Finally,
-    we show that this strategy also improves two maximum common induced subgraph algorithms.
+    programming style search. Although these heuristics are far from perfect, discrepancy-based
+    searches are too expensive to be effective in this setting.  However, we show that a random
+    value-ordering heuristic with restarts beats conventional search with the best-known tailored
+    heuristic. This motivates the introduction of a slightly-random value-ordering heuristic with
+    similar behaviour to the tailored heuristic: when combined with restarts, this new search
+    strategy is over ?? times more effective on satisfiable instances, and does not
+    significantly decrease performance on unsatisfiable instances. Finally, we show that this
+    strategy also improves two maximum common induced subgraph algorithms.
 \end{abstract}
@@
 \citep{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:conf/ijcai/McCreeshPT16}. This
-heuristic is far from perfect---and even if degree were a perfect source of information,
-many target graphs do not have a large degree distribution, so we would still need to determine
-\emph{which} vertex of highest degree would be best.
-
-We could try to introduce tiebreaking mechanisms to further refine this heuristic. However, given
+heuristic is much better than branching randomly, but is still far from perfect.  We could try to
+introduce tiebreaking mechanisms and similar refinements to strengthen the heuristic. However, given
 the extreme cost of making an incorrect branching choice early on in search, we believe it is more
@@
     }
-    \nl \While{$\KwSty{true}$}{
+    \nl \While{$\KwSty{true}$\siplabel{restartsloop}}{
         \nl \lIf{$\KwSty{not}~\FuncSty{propagate}(D)$}{$\KwSty{return}~\KwSty{false}$}
-        \nl \tikzmark{restart1s}$\KwSty{global}~(R, R') \gets (0, \textnormal{the next restart sequence value})$\tikzmark{restart1e} \;
+        \nl \tikzmark{restart1s}$\KwSty{global}~(R, R') \gets (0, \textnormal{the next Luby sequence value})$\tikzmark{restart1e} \;
         \nl $S \gets \FuncSty{search}(\{ E \in D : \left|E\right| > 1 \}, \top)$ \;
@@
     }
-    \nl \tikzmark{restart6s}\lIf{$(R \gets R + 1) = R'$}{$\KwSty{return}~\KwSty{restart}\tikzmark{restart6e}$}
+    \nl \tikzmark{restart6s}\lIf{$(R \gets R + 1) = R'$}{$\KwSty{return}~\KwSty{restart}\tikzmark{restart6e}$\siplabel{countrestarts}}
     \nl $\KwSty{return}~\KwSty{false}$\siplabel{backtrack}
@@
         \nl $v' \gets $ the single value in $D_v$ \;
-        \nl \tikzmark{restart7s}trigger nogoods with watch $(v \mapsto v')$\tikzmark{restart7e} \;
+        \nl \tikzmark{restart7s}trigger nogoods with watch $(v \mapsto v')$\tikzmark{restart7e}\siplabel{2wl} \;
         \nl \ForEach{$D_w \in D - D_v$}{
@@
 
-\subsection{Ordering Heuristics}
-
 For the variable-ordering heuristic on \siplineref{variableordering} of \cref{algorithm:sip}, we use
@@
 algorithm on unsatisfiable instances.  For value ordering, which occurs on
-\siplineref{valueordering} of \cref{algorithm:sip}, we will compare four strategies:
+\siplineref{valueordering} of \cref{algorithm:sip}, we will initially select vertices from highest
+degree to lowest degree; we return to value ordering in the second half of this paper.
 
-\paragraph{Random} We select vertices uniformly at random.
-
-\paragraph{Degree} We select vertices from highest degree to lowest degree. This heuristic is the
-current state of the art.
-
-\paragraph{Anti} We select vertices from lowest degree to highest degree. This heuristic is used as
-a sanity check: we expect degree to beat random, and random to beat anti.
-
-\paragraph{Biased} We branch by selecting a vertex $v'$ from the chosen domain $D_v$ with
-probability \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \text{.} \] This heuristic
-continues to prefer vertices of high degree, but will give the same chance of being selected next to
-two vertices of equal degree.  It also introduces an element of randomness, which is needed to make
-restarts have an effect.
-
-The biased heuristic resembles the well-known \emph{softmax} weighting scheme applied to the degree
-heuristic, although usually \emph{softmax} uses a base of $e$ rather than $2$.  Experiments showed
-that in this context, the choice of base is not important to the search space size; however, a base
-of $2$ avoids the need to use floating point arithmetic, which was a substantial overhead.
+?? note that variable ordering affects unsat, whereas value ordering affects only sat.
 
@@
 Other studies use a random selection of 200 of each of the instances from the ``meshes'' and
-``images'' families \citep{DBLP:journals/cviu/DamiandSHJS11} because some earlier solvers find many
-of these instances extremely hard. We would like to have a larger number of satisfiable instances in
-our test set, and so we include all pattern / target pairs. This gives a total of 14,621 instances
-(rather than the original 5,725). At least ?? of these instances are known to be satisfiable, and at
-least ?? are unsatisfiable.
+``images'' families because some earlier solvers find many of these instances extremely hard. We
+would like to have a larger number of satisfiable instances in our test set, and so we include all
+pattern / target pairs. This gives a total of 14,621 instances (rather than the original 5,725). At
+least ?? of these instances are known to be satisfiable, and at least ?? are unsatisfiable.
 
-\subsection{Our Solver is Good}
+\subsection{Baseline Performance}
 
@@
 
-?? Quick intro to discrepancy search
-\citep{DBLP:conf/ijcai/HarveyG95,DBLP:conf/aaai/Korf96,DBLP:conf/ijcai/Walsh97,DBLP:conf/cpaior/KarouiHLN07,DBLP:journals/jea/ProsserU11}.
-Discrepancy search is about value ordering, and it mostly looks at optimisation problems and at
-problems where an optimality proof is not needed.
+A \emph{discrepancy} is where search goes against a value-ordering heuristic.  Discrepancy searches
+\citep{DBLP:conf/ijcai/HarveyG95,DBLP:conf/aaai/Korf96,DBLP:conf/ijcai/Walsh97} are alternatives to
+backtracking search that initially search disallowing all discrepancies, and then retry search
+allowing an increasing number of discrepancies at each iteration until either a solution is found or
+unsatisfiability is proven. These schemes assume that value-ordering heuristics are usually
+reliable, and that most solutions can be found with only a small number of discrepancies; in such
+cases, the heavy commitment to early branching choices made by conventional backtracking search can
+be extremely costly.
 
-?? Talk about how we're not dealing with a binary search tree, so we could either make it binary, or
-count one discrepancy each time the degree changes, or count every ``against'' branch as being just
-one discrepancy regardless.
+\Cref{figure:scatter-dds} shows the effects of adding \citet{DBLP:conf/ijcai/Walsh97}'s
+depth-bounded discrepancy search (DDS) to \cref{algorithm:sip}. Each point represents the solving
+time for one instance---to avoid noise for easier instances, we measure only time spent during
+search, and exclude time spent in preprocessing and initialisation.  Points below the $x-y$ diagonal
+are speedups, whilst points on the top and right axes represent instances which timed out after one
+thousand seconds with one algorithm, but not the other. For satisfiable instances, the different
+point styles show the different families, whilst all unsatisfiable instances are shown as dark dots.
 
-?? \cref{figure:scatter-dds}. Sometimes extremely beneficial on satisfiable instances. However, on both
-unsatisfiable and satisfiable instances, the overheads are often large, and the algorithm is not
-better in aggregate (even when only considering satisfiable instances).
+The points well below the diagonal line and along the right-hand axis show that DDS can sometimes be
+extremely beneficial on satisfiable instances.  However, on both unsatisfiable and most satisfiable
+instances, the overheads can be extremely large, and DDS is much worse in aggregate (even when only
+considering satisfiable instances).
 
-\subsection{Value Ordering Heuristics}
+These large overheads are to be expected: discrepancy searches are aimed primarily at getting better
+feasible solutions in optimisation problems which are too large for a proof of optimality to be a
+realistic prospect, and they are not well-suited for unsatisfiable decision problems. However, the
+extremely large gains on some satisfiable instances confirm our suspicions that we should look at
+value-ordering heuristics in more detail.
+
+\subsection{Value-Ordering Heuristics}
 
@@
 
-In \cref{figure:value-ordering-heuristics} we show the cumulative number of instances solved over
-time using the four different value-ordering heuristics. As expected, the degree heuristic is much
-better (?? give aggregate speedups for both all and just satisfiable instances) than random
-ordering, which is in turn ?? times better than the anti-heuristic.
+In \cref{figure:value-ordering-heuristics} we show the cumulative number of satisfiable instances
+solved over time using four different value-ordering heuristics:
 
-The performance of the two degree heuristics are very similar: the
+\paragraph{Degree} We select vertices from highest degree to lowest degree. This heuristic is the
+current state of the art.
+
+\paragraph{Random} We select vertices uniformly at random.
+
+\paragraph{Anti} We select vertices from lowest degree to highest degree. This heuristic is used as
+a sanity check: as expected, degree beats random, and random beats anti.
+
+\paragraph{Biased} We branch by selecting a vertex $v'$ from the chosen domain $D_v$ with
+probability \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \text{.} \] This heuristic
+continues to prefer vertices of high degree, but will give the same chance of being selected next to
+two vertices of equal degree.  It also introduces an element of randomness, which is needed to make
+restarts have an effect.
+
+The biased heuristic resembles the well-known \emph{softmax} weighting scheme applied to the degree
+heuristic, although usually \emph{softmax} uses a base of $e$ rather than $2$.  Experiments showed
+that in this context, the choice of base is not important to the search space size; however, a base
+of $2$ avoids the need to use floating point arithmetic, which was a substantial overhead.
+
+The performance of the degree and biased heuristics are very similar: the
 biased heuristic has slightly worse performance below a 400s timeout, and effectively equal
@@
 value-ordering heuristic without adversely affecting its performance \emph{in aggregate}.
-\Cref{figure:scatter-heuristics} gives a more detailed look at these results. In this plot,
-satisfiable and unsatisfiable instances are shown using coloured shapes and black dots respectively.
-On the $x$-axis we measure the degree heuristic, whilst on the $y$-axis we measure the biased
-heuristic. Points on the outer axes represent instances which timed out with one heuristic but not
-the other.  This plot shows that despite the aggregate performance being very similar, on an case by
-case basis, the two heuristics can make a large difference to the performance of satisfiable
-instances.
+\Cref{figure:scatter-heuristics} gives a more detailed look at these results. This scatter plot
+shows that despite the aggregate performance being very similar, on an case by case basis, the two
+heuristics can make a large difference to the performance of satisfiable instances.
 
-\subsection{Restarts}
+\subsection{Adding Restarts to the Algorithm}
 
-?? Quick intro to restarts, mentioning how it's important to change something when restarting.
-Interestingly usually that thing is variable-ordering heuristics, not value-ordering
-\citep{DBLP:conf/ijcai/LecoutreSTV07,DBLP:journals/jsat/LecoutreSTV07,DBLP:conf/cp/GayHLS15,DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17}.
-?? Does anyone at all do this on value ordering explicitly? And if not, why not? There's
-\citet{DBLP:conf/flairs/RazgonOP07} at least. ?? There's a connection here to confidence-based work
-stealing \citep{DBLP:conf/cp/ChuSS09} too. But what? They talk about confidence in a heuristic, but
-not that different values could be ranked differently.
+Another alternative to plain backtracking search is provided by \emph{restarts}. The general idea is
+to perform a certain amount of search, and then if no solution has been found (and unsatisfiability
+has not been proven), to abandon search and restart from the beginning. Obviously, such an approach
+can only be beneficial if something changes after restarting---in a constraint programming setting,
+this is usually the variable-ordering heuristic
+\citep{DBLP:journals/jsat/LecoutreSTV07,DBLP:conf/cp/GayHLS15,DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17}.
+In this paper, we instead rely upon randomness in the \emph{value}-ordering heuristic, and keep the
+variable ordering fixed.\footnote{It may be possible to further improve \cref{algorithm:sip} by also
+introducing randomness or some form of learning into its variable-ordering heuristic. However,
+introducing a second change would considerably complicate the empirical analysis. We therefore leave
+this change as future work.} Using restarts on value-ordering heuristics appears to be uncommon,
+although \citet{DBLP:conf/flairs/RazgonOP07} look at learning value-ordering heuristics from
+restarts, and \citet{DBLP:conf/cp/ChuSS09} use a similar scheme in the context of parallel search.
 
-?? We use the Luby scheme \citep{DBLP:journals/ipl/LubySZ93}, because everyone else does. We count
-the number of backtracks (that is, when we reach the end of the main $\KwSty{foreach}$ loop) to
-decide when to restart. Following established wisdom, we multiply each item in the Luby sequence by
-a magic constant. Preliminary experiments demonstrated that this is useful, but we failed to
-determine a principled way of selecting the constant's value, so we fell back on divine revelation
-and set it to 666. As with everyone else, we are not entirely clear why we are doing this: it could
-be to allow the solver to spend more time deep in search, it could be to reduce the overheads from
-repeatedly visiting inner nodes, or it could be to keep the nogoods small.
+The shaded code shows how to introduce restarts to \cref{algorithm:sip}. The recursive
+$\FuncSty{search}$ procedure is modified to return either true (if the instance is satisfiable),
+false (unsatisfiable), or a third value if some threshold was reached without the solution being
+known.  The main function then calls $\FuncSty{search}$ in a loop (\siplineref{restartsloop}), until
+a solution is found.
 
-?? We just use two watched literals \citep{DBLP:conf/dac/MoskewiczMZZM01}. Could use increasing
-nogoods \citep{DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17} etc, but we are not confident enough
-in our programming abilities to implement these correctly, and we are not maintaining arc
-consistency anyway. We also really do not want to propagate on negative decisions, because we do not
-want to iterate over every removed value.
+We use the Luby scheme \citep{DBLP:journals/ipl/LubySZ93} to determine when to restart. We count the
+number of backtracks (that is, when we reach the end of the main $\KwSty{foreach}$ loop on
+\siplineref{countrestarts}) to decide when to restart, rather than counting recursive calls.
+Following established wisdom, we multiply each item in the Luby sequence by a magic constant.
+Preliminary experiments demonstrated that a constant 100 is slightly too small and 1,000 is slightly
+too large, but lacking a more principled way of selecting the constant's value, we fell back on
+divine revelation and use a multiplier of 666. (We also tried a geometric scheme. The results were
+less favourable---it appears that value-ordering heuristics benefit from a very aggressive restart
+strategy.)
 
-?? Talk a bit about how it fits into \cref{algorithm:sip}.
+The remainder of the changes are to introduce nogoods. To avoid exploring portions of the search
+space that we have already visitied, every time we restart, we add new constraints to the problem
+which eliminate already-explored subtrees. We generate simple decision nogoods, as follows. The $B$
+argument to $\FuncSty{search}$ tracks the branching choices made so far, in the form of a CNF clause
+with each literal being a guessed $(v \mapsto v')$ assignment. Upon backtracking due to a decision to
+restart, we post a nogood of the form $(v \mapsto v') \wedge (w \mapsto w') \wedge (x \mapsto x')
+\Rightarrow \bot$ for the current branch, and for every branch to its left in the search tree.
 
-\subsection{Restarts Make it Better}
+Finally, we use the two watched literals technique \citep{DBLP:conf/dac/MoskewiczMZZM01} to
+propagate stored nogoods (\siplineref{2wl}). This has two benefits: the propagation complexity does
+not particularly depend upon the number of stored nogoods, and it does not require any work upon
+backtracking.  Other more sophisticated nogood generation and propagation schemes exist
+\citep{DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17}, but it is not clear whether these will be
+beneficial in a setting where we are not maintaining arc consistency.
+
+\subsection{The Effects of Restarts}
 
 In \cref{figure:restarts} we show the effects of adding restarts to the algorithm. With restarts,
-the random value-ordering heuristic comfortably beats the degree strategy without restarts. In
-other words, although having a good variable-ordering heuristic is beneficial, introducing
-randomness into the search is better, if it is done alongside a mechanism to avoid heavy commitment
-to any particular random choice. \Cref{figure:scatter-random} looks at this in more detail.  The
-performance on unsatisfiable instances is the same. However, although better in aggregate, we see
-many satisfiable instances performing substantially worse. ?? Quantify this a bit.
+the random value-ordering heuristic comfortably beats the degree strategy without restarts. In other
+words, although having a good variable-ordering heuristic is beneficial, introducing randomness into
+the search is better, if it is done alongside a mechanism to avoid heavy commitment to any
+particular random choice. \Cref{figure:scatter-random} looks at this in more detail. The performance
+on unsatisfiable instances, drawn as dark dots, remains unchanged as expected (which shows that the
+overheads of nogoods and restarting are negligible). However, although better in aggregate on
+satisfiable instances, on a case by case basis many instances perform substantially worse with
+randomisation and restarts. ?? Quantify this a bit
 
@@
 
-\subsection{Further Experiments}
-
-\begin{itemize}
-    \item The choice of Luby multiplier is not hugely important, although 100 is too small and 1,000
-        is too big.
-    \item The base of the exponential can be $2$ or $e$ or $3$ or $1.5$ and it makes little
-        difference. Using a polynomial weighting is bad, and using a base of $|D_v|$ is bad.
-    \item Using geometric restarts instead of Luby does not give as much of a benefit. A constant
-        restart policy is also fine (and terminates on unsatisfiable instances because we count
-        backtracks).
-    \item Using rank with restarts is better than random with restarts, but not as good as using
-        biased with restarts.
-    \item Nogood recording is important.
-    \item Nogood recording is cheap.
-    \item Discarding subsumed nogoods may be done cheaply, if memory space is a concern.
-\end{itemize}
-
 \section{Maximum Common Subgraph Algorithms}
@@
     \includegraphics*{gen-graph-scatter-mcsplit.pdf}
-    \caption{McSplit speedups ?? partial data}\label{figure:mcsplit}
+    \caption{McSplit speedups ?? partial data, can cut for space and just use
+    cumulatives if necessary}\label{figure:mcsplit}
 \end{figure}
@@
     \includegraphics*{gen-graph-scatter-kdown.pdf}
-    \caption{k${\downarrow}$ speedups ?? missing data}\label{figure:kdown}
+    \caption{k${\downarrow}$ speedups ?? missing data, can cut for space and just use
+    cumulatives if necessary}\label{figure:kdown}
 \end{figure}
@@
 
-\clearpage
 \bibliographystyle{named}

### 2017-12-15 mcsplit-code/solve_mcs.cc
Revert to always choosing "do not use this vertex" last

@@ -487,15 +487,8 @@ namespace {
                 if (params.biased_shuffle)
                     biased_shuffle(possible_values);
 
-                if (bound != incumbent.size() + 1 || bd.left_len > bd.right_len) {
-                    if (params.biased_shuffle) {
-                        std::uniform_int_distribution<unsigned> dist(0, possible_values.size());
-                        unsigned insertion_point = dist(global_rand);
-                        possible_values.insert(next(possible_values.begin(), insertion_point), -1);
-                    } else {
-                        possible_values.push_back(-1);
-                    }
-                }
+                if (bound != incumbent.size() + 1 || bd.left_len > bd.right_len)
+                    possible_values.push_back(-1);
 
                 remove_vtx_from_left_domain(domains[bd_idx], v);
                 bd.right_len--;
@@ -513,11 +506,9 @@ namespace {
                         search_result = restarting_search(current, new_domains, backtracks_until_restart);
                     } else {
                         bd.right_len++;
-                        auto copied_domains = domains;
                         if (bd.left_len == 0)
-                            remove_bidomain(copied_domains, bd_idx);
-                        search_result = restarting_search(current, copied_domains, backtracks_until_restart);
-                        bd.right_len--;
+                            remove_bidomain(domains, bd_idx);
+                        search_result = restarting_search(current, domains, backtracks_until_restart);
                     }
                     current.pop();
                     switch (search_result)

### 2017-12-15 mcsplit-code/solve_mcs.cc
Revert to linear scan to check if literal is in current assignment

@@ -60,6 +60,12 @@ namespace {
         {
             var_assignments.clear();
         }
+
+        bool contains(const Assignment & assignment) const
+        {
+            return var_assignments.end() != find_if(var_assignments.begin(), var_assignments.end(),
+                    [&] (const VarAssignment & a) { return a.assignment == assignment; });
+        }
     };
 
     struct Bidomain
@@ -160,8 +166,6 @@ namespace {
         Nogoods nogoods;
         Watches watches;
         Watches literal_clause_membership;
-
-        vector<int> vtx_current_assignment;
 
         // used in subsumption test; assigned once to avoid a vector
         // allocation on each function call
@@ -382,11 +386,12 @@ namespace {
 
         auto can_find_another_watch(Nogood & nogood,
                 list<Nogoods::iterator> & watches_to_update,
-                list<Nogoods::iterator>::iterator watch_to_update) -> bool
+                list<Nogoods::iterator>::iterator watch_to_update,
+                VarAssignments & current) -> bool
         {
             // can we find something else to watch?
             for (auto new_literal = next(nogood.literals.begin(), 1) ; new_literal != nogood.literals.end() ; ++new_literal) {
-                if (vtx_current_assignment[new_literal->v] != new_literal->w) {
+                if (!current.contains(*new_literal)) {
                     // move the new watch to be the first item in the nogood
                     std::swap(nogood.literals[0], *new_literal);
 
@@ -404,7 +409,7 @@ namespace {
 
         auto current_contains_nogood(
                 Assignment most_recent_assignment,
-                vector<int> & vtx_current_assignment) -> bool
+                VarAssignments & current) -> bool
         {
             auto & watches_to_update = watches[most_recent_assignment];
             for (auto watch_to_update = watches_to_update.begin() ; watch_to_update != watches_to_update.end() ; ) {
@@ -412,7 +417,7 @@ namespace {
 
                 auto next_watch_to_update = next(watch_to_update);
 
-                if (!can_find_another_watch(nogood, watches_to_update, watch_to_update))
+                if (!can_find_another_watch(nogood, watches_to_update, watch_to_update, current))
                     return true;
                 
                 watch_to_update = next_watch_to_update;
@@ -427,21 +432,9 @@ namespace {
             if (current.get_var_assignments().empty())
                 return false;
 
-            std::fill(vtx_current_assignment.begin(), vtx_current_assignment.end(), -1);
-
[…]

### 2017-12-15 mcsplit-code/solve_mcs.cc
Partition using std::partition

@@ -236,14 +236,11 @@ namespace {
         // Returns length of left half of array
         auto partition(vector<int>& all_vv, int start, int len, const vector<unsigned int> & adjrow) -> int
         {
-            int i=0;
-            for (int j=0; j<len; j++) {
-                if (adjrow[all_vv[start+j]]) {
-                    std::swap(all_vv[start+i], all_vv[start+j]);
-                    i++;
-                }
-            }
-            return i;
+            auto it = std::partition(
+                    all_vv.begin() + start,
+                    all_vv.begin() + start + len,
+                    [&](const int elem){ return 0 != adjrow[elem]; });
+            return it - (all_vv.begin() + start);
         }
 
         // multiway is for directed and/or labelled graphs

### 2017-12-15 mcsplit-code/solve_mcs.cc
Use signatures for quick subsumption check

@@ -352,7 +352,7 @@ namespace {
 
             for (auto & a : current.get_var_assignments())
                 if (a.is_decision)
-                    nogood.literals.emplace_back(a.assignment);
+                    nogood.add_literal(a.assignment);
 
             if (nogood.literals.empty())
                 return false;

### 2017-12-15 mcsplit-code/solve_mcs.cc
Minor tidying

@@ -520,9 +520,7 @@ namespace {
                 }
             }
 
-            if (backtracks_until_restart > 0 && 0 == --backtracks_until_restart) {
-                if (! post_nogood(current))
-                    return Search::Done;
+            if (backtracks_until_restart > 0 && 0 == --backtracks_until_restart && post_nogood(current)) {
                 return Search::Restart;
             } else {
                 return Search::Done;

### 2017-12-15 mcsplit-code/solve_mcs.cc
Decrement backtracks_until_restart less frequently.
Decrement backtracks_until_restart only when there was more than one
option.  This more-or-less reverses commit
a7ee253e0c0299968437d04a3e88a2c1ecc83e78.

@@ -461,6 +461,11 @@ namespace {
 
             int bd_idx;
 
+            // Is more than one value choice available at this search node?
+            // This will be set to true if more than one possible value is found to exist
+            // later in the function.
+            bool is_decision = false;
+
             if (bound > incumbent.size() &&
                     !contains_a_nogood(current, domains) &&
                     -1 != (bd_idx = select_bidomain(domains, current.get_num_vtx_assignments())))
@@ -480,11 +485,14 @@ namespace {
                 if (bound != incumbent.size() + 1 || bd.left_len > bd.right_len)
                     possible_values.push_back(-1);
 
+                if (possible_values.size() > 1)
+                    is_decision = true;
+
                 remove_vtx_from_left_domain(domains[bd_idx], v);
                 bd.right_len--;
 
                 for (int w : possible_values) {
-                    current.push({{v, w}, possible_values.size() > 1});
+                    current.push({{v, w}, is_decision});
                     Search search_result;
                     if (w != -1) {
                         // swap w to the end of its colour class
@@ -520,7 +528,7 @@ namespace {
                 }
             }
 
-            if (backtracks_until_restart > 0 && 0 == --backtracks_until_restart && post_nogood(current)) {
+            if (is_decision && backtracks_until_restart > 0 && 0 == --backtracks_until_restart && post_nogood(current)) {
                 return Search::Restart;
             } else {
                 return Search::Done;

### 2017-12-15 mcsplit-code/mcsp.cc
Re-added McSplit-down

@@ -36,6 +36,7 @@ static struct argp_option options[] = {
     {"connected", 'c', 0, 0, "Solve max common CONNECTED subgraph problem"},
     {"directed", 'i', 0, 0, "Use directed graphs"},
     {"labelled", 'a', 0, 0, "Use edge and vertex labels"},
+    {"mcsplit-down", 'm', 0, 0, "Use the McSplit-down algorithm rather than branch and bound"},
     {"vertex-labelled-only", 'x', 0, 0, "Use vertex labels, but not edge labels"},
     {"timeout", 't', "timeout", 0, "Specify a timeout (seconds)"},
     { 0 }
@@ -52,6 +53,7 @@ struct Arguments {
     bool directed;
     bool edge_labelled;
     bool vertex_labelled;
+    bool mcsplit_down;
     Heuristic heuristic;
     char *filename1;
     char *filename2;
@@ -105,6 +107,9 @@ static error_t parse_opt (int key, char *arg, struct argp_state *state) {
             if (arguments.edge_labelled)
                 fail("The -a and -x options can't be used together.");
             arguments.vertex_labelled = true;
+            break;
+        case 'm':
+            arguments.mcsplit_down = true;
             break;
         case 't':
             arguments.timeout = std::stoi(arg);
@@ -212,6 +217,7 @@ int main(int argc, char** argv) {
         arguments.directed,
         arguments.edge_labelled,
         arguments.vertex_labelled,
+        arguments.mcsplit_down,
         arguments.heuristic
     };
 

### 2017-12-15 mcsplit-code/solve_mcs.cc

@@ -152,6 +152,12 @@ namespace {
         {
             return data[target_size_plus_1 * a.v + a.w + 1];
         }
+
+        auto clear_all() -> void
+        {
+            for (auto & watch_list : data)
+                watch_list.clear();
+        }
     };
 
     class MCS
@@ -172,6 +178,13 @@ namespace {
         vector<int> assignments_in_nogood;
 
         vector<unsigned long long> target_vertex_biases;
+
+        auto clear_nogoods() -> void
+        {
+            nogoods.clear();
+            watches.clear_all();
+            literal_clause_membership.clear_all();
+        }
 
         auto show(const vector<VarAssignment>& current, const vector<Bidomain> &domains) -> void
         {
@@ -336,7 +349,7 @@ namespace {
             }
         }
 
-        auto update_incumbent(vector<Assignment> & incumbent, VarAssignments current) -> void
+        auto update_incumbent(VarAssignments current) -> void
         {
             incumbent.clear();
             for (auto a : current.get_var_assignments())
@@ -438,13 +451,15 @@ namespace {
         {
             Aborted,
             Done,
-            Restart
+            Restart,
+            ReachedGoal
         };
 
         auto restarting_search(
                 VarAssignments & current,
                 vector<Bidomain> & domains,
-                long long & backtracks_until_restart) -> Search
+                long long & backtracks_until_restart,
+                unsigned int matching_size_goal) -> Search
         {
             if (abort_due_to_timeout)
                 return Search::Aborted;
@@ -454,8 +469,11 @@ namespace {
             if (params.verbose)
                 show(current.get_var_assignments(), domains);
 
-            if (current.get_num_vtx_assignments() > incumbent.size())
-                update_incumbent(incumbent, current);
+            if (current.get_num_vtx_assignments() > incumbent.size()) {
+                update_incumbent(current);
+                if (incumbent.size() == matching_size_goal)
+                    return Search::ReachedGoal;
+            }
 
             unsigned int bound = current.get_num_vtx_assignments() + calc_bound(domains);
 
@@ -467,6 +485,7 @@ namespace {
             bool is_decision = false;
 
             if (bound > incumbent.size() &&
+                    (!params.mcsplit_down || bound >= matching_size_goal) &&
                     !contains_a_nogood(current, domains) &&
                     -1 != (bd_idx = select_bidomain(domains, current.get_num_vtx_assignments())))
             {
@@ -482,7 +501,10 @@ namespace {
[…]

### 2017-12-15 mcsplit-code/mcsp.cc
Show time when new incumbent is found

@@ -218,7 +218,8 @@ int main(int argc, char** argv) {
         arguments.edge_labelled,
         arguments.vertex_labelled,
         arguments.mcsplit_down,
-        arguments.heuristic
+        arguments.heuristic,
+        start
     };
 
     auto result = solve_mcs(g0, g1, params);

### 2018-01-07 paper/paper.tex
Betterer

@@
 and so on. Due to its similarity to \cref{algorithm:sip}, we can introduce the same bias and restart
-strategy.
+strategy, requiring only a small amount of jiggery-pokery to handle wildcards in nogoods.
 
@@
     \includegraphics*{gen-graph-scatter-mcsplit.pdf}
-    \caption{McSplit speedups ?? partial data, can cut for space and just use
-    cumulatives if necessary}\label{figure:mcsplit}
+    \caption{McSplit speedups}\label{figure:mcsplit}
 \end{figure}
@@
     \includegraphics*{gen-graph-scatter-kdown.pdf}
-    \caption{k${\downarrow}$ speedups ?? missing data, can cut for space and just use
-    cumulatives if necessary}\label{figure:kdown}
+    \caption{k${\downarrow}$ speedups}\label{figure:kdown}
 \end{figure}
@@
     \includegraphics*{gen-graph-mcs.pdf}
-    \caption{k${\downarrow}$ and McSplit cumulative curves ?? partial and missing
-    data}\label{figure:mcs}
+    \caption{k${\downarrow}$ and McSplit cumulative curves}\label{figure:mcs}
 \end{figure}

### 2018-01-12 paper/paper.tex
Bettererer

@@
 \newcommand{\citep}[1]{\cite{#1}}
+\newcommand{\citepos}[1]{\citeauthor{#1}'s \shortcite{#1}}
 
@@
 \begin{abstract}
-    Modern subgraph isomorphism solvers use value-ordering heuristics to direct a constraint
-    programming style search. Although these heuristics are far from perfect, discrepancy-based
-    searches are too expensive to be effective in this setting.  However, we show that a random
-    value-ordering heuristic with restarts beats conventional search with the best-known tailored
-    heuristic. This motivates the introduction of a slightly-random value-ordering heuristic with
-    similar behaviour to the tailored heuristic: when combined with restarts, this new search
-    strategy is over ?? times more effective on satisfiable instances, and does not
-    significantly decrease performance on unsatisfiable instances. Finally, we show that this
-    strategy also improves two maximum common induced subgraph algorithms.
+    Modern subgraph isomorphism solvers use degree-based heuristics to direct a constraint
+    programming style search. We show that discrepancy-based searches are too expensive to be
+    effective in offsetting mistakes made by value-ordering heuristic in this setting.  We then
+    demonstrate that a random heuristic with restarts beats conventional search with the best-known
+    tailored heuristic. This motivates the introduction of a slightly-random heuristic with similar
+    behaviour to the tailored heuristic: when combined with restarts, this new search strategy is
+    over a hundred times more effective on satisfiable instances, and does not affect performance on
+    unsatisfiable instances. Finally, we show that the same strategy also improves two maximum
+    common induced subgraph algorithms.
 \end{abstract}
@@
 a new value-ordering heuristic, which adds a small amount of randomness to the ``highest degree
-first'' heuristic.  The effects of the changes we propose are shown in
+first'' heuristic.  The effects of these changes are shown in
 \cref{figure:scatter-by-family}: we can make a state-of-the-art algorithm perform much better on a
@@
 (\cref{figure:scatter-random}). Finally, we show that this strategy is also effective in an
-optimisation setting, producing clear benefits in two maximum common induced subgraph algorithms.
+optimisation setting, producing benefits in two maximum common induced subgraph algorithms.
 
@@
 Let $G$ be a graph. We denote its set of vertices by $\vertexset(G)$, and write $v \sim_G w$ to mean
-that vertices $v$ and $w$ are adjacent. The \emph{neighbourhood} of a vertex, $\neighbourhood_G(v)$,
-is the set of vertices adjacent to $v$, and the \emph{degree} of $v$, $\deg_G(v)$, is the
-cardinality of its neighbourhood; when the graph is clear from the context, we omit the $G$
-subscript. The \emph{neighbourhood degree sequence} of $v$, $\nds_G(v)$, is the sequence of degrees
-of the neighbours of $v$ in descending order, and we denote by $\preceq$ the inclusion relation of
-\citet{DBLP:journals/constraints/ZampelliDS10}. As per \citet{DBLP:conf/aaai/HoffmannMR17},
-$G^{n,\ell}$ denotes the graph with vertex set $\vertexset(G)$, and with edges between vertices $v$
-and $v$ if there are at least $n$ simple paths of length exactly $\ell$ between vertices $v$ and $w$
-in $G$.
+that vertices $v$ and $w$ are adjacent; we allow a vertex to be adjacent to itself. The
+\emph{neighbourhood} of a vertex, $\neighbourhood_G(v)$, is the set of vertices adjacent to $v$, and
+the \emph{degree} of $v$, $\deg_G(v)$, is the cardinality of its neighbourhood; when the graph is
+clear from the context, we omit the $G$ subscript. The \emph{neighbourhood degree sequence} of $v$,
+$\nds_G(v)$, is the sequence of degrees of the neighbours of $v$ in descending order, and we denote
+by $\preceq$ the inclusion relation of \citet{DBLP:journals/constraints/ZampelliDS10}. As per
+\citet{DBLP:conf/aaai/HoffmannMR17}, $G^{n,\ell}$ denotes the graph with vertex set $\vertexset(G)$,
+and with edges between vertices $v$ and $v$ if there are at least $n$ simple paths of length exactly
+$\ell$ between vertices $v$ and $w$ in $G$.
 
@@
 \siplineref{valueordering} of \cref{algorithm:sip}, we will initially select vertices from highest
-degree to lowest degree; we return to value ordering in the second half of this paper.
-
-?? note that variable ordering affects unsat, whereas value ordering affects only sat.
+degree to lowest degree; we return to value ordering in the second half of this paper. Note that the
+variable-ordering heuristic affects both unsatisfiable and satisfiable instances, whilst the
+value-ordering heuristic is relevant only for satisfiable instances.
 
@@
 pattern / target pairs. This gives a total of 14,621 instances (rather than the original 5,725). At
-least ?? of these instances are known to be satisfiable, and at least ?? are unsatisfiable.
+least \input{gen-n-sat.tex}\unskip\ of these instances are known to be satisfiable, and at least
+\input{gen-n-unsat.tex}\unskip\ are unsatisfiable.
 
@@
 algorithm \citep{DBLP:journals/ai/Solnon10,DBLP:conf/lion/KotthoffMS16}, and VF2
-\citep{DBLP:journals/pami/CordellaFSV04}. For each algorithm, the $y$ value
+\citep{DBLP:journals/pami/CordellaFSV04}, in each case using the original implementation provided
+by the algorithm's authors. For each algorithm, the $y$ value
 gives the cumulative number of instances which (individually) can be solved in no more than $x$
@@
 
-\Cref{figure:scatter-dds} shows the effects of adding \citet{DBLP:conf/ijcai/Walsh97}'s
+\Cref{figure:scatter-dds} shows the effects of adding \citepos{DBLP:conf/ijcai/Walsh97}
 depth-bounded discrepancy search (DDS) to \cref{algorithm:sip}. Each point represents the solving
@@
 
-\Cref{figure:restarts} shows that either biased heuristic together with restarts is better
-still---that is, if we are introducing restarts, then it is better to add a small amount of
-randomness to a tailored deterministic heuristic than it is to throw away the heuristic altogether.
+\Cref{figure:restarts} shows that the biased heuristic together with restarts is better still---that
+is, if we are introducing restarts, then it is better to add a small amount of randomness to a
+tailored deterministic heuristic than it is to throw away the heuristic altogether.
 
@@
 ?? Benchmark results on the same families of instances are shown in
-\cref{figure:mcsplit,figure:kdown,figure:mcs}. Clearly beneficial. Note that on McSplit we're doing
+\cref{figure:kdown,figure:mcsplit,figure:mcs}. Clearly beneficial. Note that on McSplit we're doing
 extra work and this shows, but it's still worthwhile.
+
+\begin{figure}[p]
+    \centering
+    \includegraphics*{gen-graph-scatter-kdown.pdf}
+    \caption{k${\downarrow}$ speedups}\label{figure:kdown}
+\end{figure}
 
@@
     \caption{McSplit speedups}\label{figure:mcsplit}
-\end{figure}
-
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-scatter-kdown.pdf}
-    \caption{k${\downarrow}$ speedups}\label{figure:kdown}
 \end{figure}

### 2018-01-12 code/parallel.cc
Updateify

@@ -36,6 +36,7 @@ using std::move;
 using std::mt19937;
 using std::mutex;
 using std::next;
+using std::numeric_limits;
 using std::pair;
 using std::sort;
 using std::string;
@@ -164,8 +165,7 @@ namespace
 
         vector<int> pattern_permutation, target_permutation, isolated_vertices;
         vector<vector<int> > patterns_degrees, targets_degrees;
-
-        vector<unsigned long long> target_vertex_biases;
+        int largest_target_degree;
 
         mutex new_nogoods_mutex;
         Nogoods nogoods;
@@ -185,7 +185,8 @@ namespace
             target_size(target.size()),
             target_permutation(target.size()),
             patterns_degrees(max_graphs),
-            targets_degrees(max_graphs)
+            targets_degrees(max_graphs),
+            largest_target_degree(0)
         {
             // strip out isolated vertices in the pattern, and build pattern_permutation
             for (unsigned v = 0 ; v < full_pattern_size ; ++v)
@@ -212,19 +213,6 @@ namespace
                 for (unsigned j = 0 ; j < target_size ; ++j)
                     if (target.adjacent(target_permutation.at(i), target_permutation.at(j)))
                         target_graph_rows[i * max_graphs + 0].set(j);
-
-            // build up vertex selection biases
-            unsigned max_degree = 0;
-            for (unsigned j = 0 ; j < target_size ; ++j)
-                max_degree = max(max_degree, target_graph_rows[j * max_graphs + 0].popcount());
-
-            for (unsigned j = 0 ; j < target_size ; ++j) {
-                unsigned degree = target_graph_rows[j * max_graphs + 0].popcount();
-                if (max_degree - degree >= 50)
-                    target_vertex_biases.push_back(1);
-                else
-                    target_vertex_biases.push_back(1ull << (50 - (max_degree - degree)));
-            }
         }
 
         auto build_supplemental_graphs(vector<FixedBitSet<n_words_> > & graph_rows, unsigned size) -> void
@@ -488,30 +476,43 @@ namespace
                 branch_v[branch_v_end++] = f_v;
             }
 
-            // sum up the bias scores of every branch vertex
-            unsigned long long remaining_score = 0;
-            for (unsigned v = 0 ; v < branch_v_end ; ++v)
-                remaining_score += target_vertex_biases[branch_v[v]];
+            if (params.shuffle) {
+                shuffle(branch_v.begin(), branch_v.begin() + branch_v_end, my_thread_data.rand);
+            }
[…]

### 2018-01-12 code/sequential.cc

@@ -26,8 +26,8 @@ using std::max;
 using std::map;
 using std::move;
 using std::mt19937;
-using std::numeric_limits;
 using std::next;
+using std::numeric_limits;
 using std::pair;
 using std::sort;
 using std::string;
@@ -157,8 +157,6 @@ namespace
         Watches watches;
         list<typename Nogoods::iterator> need_to_watch;
 
-        vector<unsigned long long> target_vertex_biases;
-
         mt19937 global_rand;
 
         SIP(const Graph & target, const Graph & pattern, const Params & a) :
@@ -204,21 +202,6 @@ namespace
                 for (unsigned j = 0 ; j < target_size ; ++j)
                     if (target.adjacent(target_permutation.at(i), target_permutation.at(j)))
                         target_graph_rows[i * max_graphs + 0].set(j);
-
-            // build up vertex selection biases
-            if (params.biased_shuffle) {
-                unsigned max_degree = 0;
-                for (unsigned j = 0 ; j < target_size ; ++j)
-                    max_degree = max(max_degree, target_graph_rows[j * max_graphs + 0].popcount());
-
-                for (unsigned j = 0 ; j < target_size ; ++j) {
-                    unsigned degree = target_graph_rows[j * max_graphs + 0].popcount();
-                    if (max_degree - degree >= 50)
-                        target_vertex_biases.push_back(1);
-                    else
-                        target_vertex_biases.push_back(1ull << (50 - (max_degree - degree)));
-                }
-            }
         }
 
         auto build_supplemental_graphs(vector<FixedBitSet<n_words_> > & graph_rows, unsigned size) -> void
@@ -576,33 +559,6 @@ namespace
 
             if (params.shuffle) {
                 shuffle(branch_v.begin(), branch_v.begin() + branch_v_end, global_rand);
-            }
-            else if (params.biased_shuffle) {
-                // sum up the bias scores of every branch vertex
-                unsigned long long remaining_score = 0;
-                for (unsigned v = 0 ; v < branch_v_end ; ++v)
-                    remaining_score += target_vertex_biases[branch_v[v]];
-
-                // now repeatedly pick a biased-random vertex, move it to the front of branch_v,
-                // and then only consider items further to the right in the next iteration.
-                for (unsigned start = 0 ; start < branch_v_end ; ++start) {
-                    // pick a random number between 0 and remaining_score inclusive
-                    uniform_int_distribution<unsigned long long> dist(1, remaining_score);
[…]

### 2018-01-12 code/solve_subgraph_isomorphism.cc

@@ -131,7 +131,6 @@ auto main(int argc, char * argv[]) -> int
             ("restarts",                                     "Use restarts")
             ("dds",                                          "Use dds")
             ("shuffle",                                      "Use shuffling")
-            ("biased-shuffle",                               "Use biased shuffling")
             ("softmax-shuffle",                              "Use softmax shuffling")
             ("antiheuristic",                                "Use antiheuristic")
             ("input-order",                                  "Use input order")
@@ -198,7 +197,6 @@ auto main(int argc, char * argv[]) -> int
         // Some sanity checking
         if (options_vars["algorithm"].as<string>() == "customisable-parallel") {
             if (! options_vars.count("restarts")
-                    || ! options_vars.count("biased-shuffle")
                     || options_vars.count("softmax-shuffle")
                     || ! options_vars.count("input-order")
                     || options_vars.count("dds")
@@ -206,7 +204,7 @@ auto main(int argc, char * argv[]) -> int
                     || options_vars.count("goods")
                     || options_vars.count("antiheuristic")) {
                 cerr << "Parallel algorithm currently requires --restarts --input-order" << endl;
-                cerr << "  --biased-shuffle, and cannot use any of --dds --shuffle " << endl;
+                cerr << "  --softmax-shuffle, and cannot use any of --dds --shuffle " << endl;
                 cerr << "  --softmax-shuffle --goods --antiheuristic" << endl;
                 return EXIT_FAILURE;
             }
@@ -221,7 +219,6 @@ auto main(int argc, char * argv[]) -> int
             params.restarts = options_vars.count("restarts");
             params.dds = options_vars.count("dds");
             params.shuffle = options_vars.count("shuffle");
-            params.biased_shuffle = options_vars.count("biased-shuffle");
             params.softmax_shuffle = options_vars.count("softmax-shuffle");
             params.antiheuristic = options_vars.count("antiheuristic");
             params.input_order = options_vars.count("input-order");
@@ -233,6 +230,11 @@ auto main(int argc, char * argv[]) -> int
                 params.geometric_multiplier = options_vars["geometric-multiplier"].as<double>();
             if (options_vars.count("geometric-start"))
                 params.geometric_start = options_vars["geometric-start"].as<unsigned>();
+        }
+        else if (options_vars["algorithm"].as<std::string>() == "parallel") {
+            params.restarts = true;
+            params.softmax_shuffle = true;
+            params.input_order = true;
         }
[…]

### 2018-01-12 code/solve_subgraph_isomorphism.cc
Tweaks

@@ -205,7 +205,7 @@ auto main(int argc, char * argv[]) -> int
                     || options_vars.count("antiheuristic")) {
                 cerr << "Parallel algorithm currently requires --restarts --input-order" << endl;
                 cerr << "  --softmax-shuffle, and cannot use any of --dds --shuffle " << endl;
-                cerr << "  --softmax-shuffle --goods --antiheuristic" << endl;
+                cerr << "  --goods --antiheuristic" << endl;
                 return EXIT_FAILURE;
             }
         }

### 2018-01-12 code/solve_subgraph_isomorphism.cc
That

@@ -197,7 +197,7 @@ auto main(int argc, char * argv[]) -> int
         // Some sanity checking
         if (options_vars["algorithm"].as<string>() == "customisable-parallel") {
             if (! options_vars.count("restarts")
-                    || options_vars.count("softmax-shuffle")
+                    || ! options_vars.count("softmax-shuffle")
                     || ! options_vars.count("input-order")
                     || options_vars.count("dds")
                     || options_vars.count("shuffle")

### 2018-01-16 long-talk/talk.tex
Long talk, for StA

% vim: set spell spelllang=en tw=100 et sw=4 sts=4 foldmethod=marker foldmarker={{{,}}} :

\documentclass{beamer}

\usepackage{tikz}
\usepackage{xcolor}
\usepackage{complexity}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{amsmath}                   % \operatorname
\usepackage{amsfonts}                  % \mathcal
\usepackage{amssymb}                   % \nexists
\usepackage{gnuplot-lua-tikz}          % graphs
\usepackage[vlined]{algorithm2e} % algorithms
\usepackage{centernot}
\usepackage{graphicx}

\usetikzlibrary{shapes, arrows, shadows, calc, positioning, fit}
\usetikzlibrary{decorations.pathreplacing, decorations.pathmorphing, shapes.misc}
\usetikzlibrary{tikzmark}

\definecolor{uofguniversityblue}{rgb}{0, 0.219608, 0.396078}

\definecolor{uofgheather}{rgb}{0.356863, 0.32549, 0.490196}
\definecolor{uofgaquamarine}{rgb}{0.603922, 0.72549, 0.678431}
\definecolor{uofgslate}{rgb}{0.309804, 0.34902, 0.380392}
\definecolor{uofgrose}{rgb}{0.823529, 0.470588, 0.709804}
\definecolor{uofgmocha}{rgb}{0.709804, 0.564706, 0.47451}
\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
\definecolor{uofgforest}{rgb}{0, 0.2, 0.129412}
\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
\definecolor{uofgturquoise}{rgb}{0, 0.709804, 0.819608}
\definecolor{uofgsunshine}{rgb}{1.0, 0.862745, 0.211765}
\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}
\definecolor{uofgrust}{rgb}{0.603922, 0.227451, 0.023529}
\definecolor{uofgburgundy}{rgb}{0.490196, 0.133333, 0.223529}
\definecolor{uofgpillarbox}{rgb}{0.701961, 0.047059, 0}
\definecolor{uofglavendar}{rgb}{0.356863, 0.301961, 0.580392}

\tikzset{vertex/.style={draw, circle, inner sep=0pt, minimum size=0.5cm, font=\small\bfseries}}
\tikzset{notvertex/.style={vertex, color=white, text=black}}
\tikzset{plainvertex/.style={vertex}}
\tikzset{vertexc1/.style={vertex, fill=uofgburgundy, text=white}}
\tikzset{vertexc2/.style={vertex, fill=uofgsandstone, text=white}}
\tikzset{vertexc3/.style={vertex, fill=uofgforest, text=white}}
\tikzset{vertexc4/.style={vertex, fill=uofgheather, text=white}}
\tikzset{edge/.style={color=black!50!white}}
\tikzset{bedge/.style={ultra thick}}

% {{{ theme things
\useoutertheme[footline=authortitle]{miniframes}
\useinnertheme{rectangles}

\setbeamerfont{block title}{size={}}
\setbeamerfont{title}{size=\large,series=\bfseries}
\setbeamerfont{section title}{size=\large,series=\mdseries}
\setbeamerfont{author}{size=\normalsize,series=\mdseries}
\setbeamercolor*{structure}{fg=uofguniversityblue}
\setbeamercolor*{palette primary}{use=structure,fg=black,bg=white}
\setbeamercolor*{palette secondary}{use=structure,fg=white,bg=uofgcobalt}
\setbeamercolor*{palette tertiary}{use=structure,fg=white,bg=uofguniversityblue}
\setbeamercolor*{palette quaternary}{fg=white,bg=black}

\setbeamercolor*{titlelike}{parent=palette primary}

\beamertemplatenavigationsymbolsempty

\setbeamertemplate{title page}
{
    \begin{tikzpicture}[remember picture, overlay]
        \node at (current page.north west) {
            \begin{tikzpicture}[remember picture, overlay]
                \fill [fill=uofguniversityblue, anchor=north west] (0, 0) rectangle (\paperwidth, -2.6cm);
            \end{tikzpicture}
        };

        \node (logo) [anchor=north east, shift={(-0.6cm,-0.6cm)}] at (current page.north east) {
            \includegraphics[keepaspectratio=true,scale=0.7]{UoG_keyline.pdf}
        };

        \node [anchor=west, xshift=0.2cm] at (current page.west |- logo.west) {
            \begin{minipage}{0.65\paperwidth}\raggedright
                {\usebeamerfont{title}\usebeamercolor[white]{}\inserttitle}\\[0.1cm]
                {\usebeamerfont{author}\usebeamercolor[white]{}\insertauthor}
            \end{minipage}
        };
    \end{tikzpicture}
}

\setbeamertemplate{section page}
{
    \begin{centering}
        \begin{beamercolorbox}[sep=12pt,center]{part title}
            \usebeamerfont{section title}\insertsection\par
        \end{beamercolorbox}
    \end{centering}
}

\newcommand{\frameofframes}{/}
\newcommand{\setframeofframes}[1]{\renewcommand{\frameofframes}{#1}}

\makeatletter
\setbeamertemplate{footline}
{%
    \begin{beamercolorbox}[colsep=1.5pt]{upper separation line foot}
    \end{beamercolorbox}
    \begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,%
        leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \leavevmode{\usebeamerfont{author in head/foot}\insertshortauthor}%
        \hfill%
        {\usebeamerfont{institute in head/foot}\usebeamercolor[fg]{institute in head/foot}\insertshortinstitute}%
    \end{beamercolorbox}%
    \begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,%
        leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        {\usebeamerfont{title in head/foot}\insertshorttitle}%
        \hfill%
        {\usebeamerfont{frame number}\usebeamercolor[fg]{frame number}\insertframenumber~\frameofframes~\inserttotalframenumber}
    \end{beamercolorbox}%
    \begin{beamercolorbox}[colsep=1.5pt]{lower separation line foot}
    \end{beamercolorbox}
}

\makeatletter
\newenvironment{nearlyplainframe}[2][]{
    \def\beamer@entrycode{\vspace*{-\headheight}\vspace*{3pt}}
    \setbeamertemplate{headline}
    {%
        \begin{beamercolorbox}[colsep=1.5pt]{upper separation line head}
        \end{beamercolorbox}
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[colsep=1.5pt]{lower separation line head}
        \end{beamercolorbox}
        \vspace*{\headheight}
    }

    \setbeamertemplate{footline}
    {%
        \begin{beamercolorbox}[colsep=1.5pt]{upper separation line foot}
        \end{beamercolorbox}
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[colsep=1.5pt]{lower separation line foot}
        \end{beamercolorbox}
    }

    \begin{frame}[#1]{#2}
    }{
    \end{frame}
}
\makeatother

% }}}

\title[Value Ordering, Discrepancies, and Restarts for Subgraph Algorithms]{Value Ordering,
Discrepancies, and Restarts for Subgraph Algorithms}
\author[Ciaran McCreesh, Patrick Prosser and James Trimble]{\textbf{Ciaran McCreesh}, Patrick
Prosser and \\ James Trimble}

\begin{document}

{
    \usebackgroundtemplate{
        \tikz[overlay, remember picture]
        \node[at=(current page.south), anchor=south, inner sep=0pt]{\includegraphics[keepaspectratio=true, width=\paperwidth]{background.jpg}};
    }
    \begin{frame}[plain,noframenumbering]
        \titlepage
    \end{frame}
}

\section{Subgraph Isomorphism}

\begin{frame}{Non-Induced Subgraph Isomorphism}

    \begin{center}\begin{tikzpicture}
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {1};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {2};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {3};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {4};

        \node <2> [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {1};
        \node <2> [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {2};
        \node <2> [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {3};
        \node <2> [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {4};

        \draw <1> [thick, color=uofgsandstone!50] (Na) -- (Nb);
        \draw <1> [thick, color=uofgsandstone!50] (Nb) -- (Nc);
        \draw <1> [thick, color=uofgsandstone!50] (Nc) -- (Nd);
        \draw <1> [thick, color=uofgsandstone!50] (Nb) -- (Nd);

        \draw <2> [ultra thick] (Na) -- (Nb);
        \draw <2> [ultra thick] (Nb) -- (Nc);
        \draw <2> [ultra thick] (Nc) -- (Nd);
        \draw <2> [ultra thick] (Nb) -- (Nd);

        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {1};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {2};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {3};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {4};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {5};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {6};

        \node <2> [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {1};
        \node <2> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {2};
        \node <2> [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {3};
        \node <2> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {4};
        \node <2> [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {5};
        \node <2> [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {6};

        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N2);
        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N3);
        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N4);
        \draw <1> [thick, color=uofgsandstone!50] (N2) -- (N4);
        \draw <1> [thick, color=uofgsandstone!50] (N3) -- (N5);
        \draw <1> [thick, color=uofgsandstone!50] (N3) -- (N6);
        \draw <1> [thick, color=uofgsandstone!50] (N4) -- (N6);
        \draw <1> [thick, color=uofgsandstone!50] (N5) -- (N6);
        \draw <1> [thick, color=uofgsandstone!50] (N2) to [in=45, out=315] (N6);
        \draw <1> [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);

        \draw <2> [thick, color=uofgsandstone!50] (N1) -- (N2);
        \draw <2> [ultra thick] (N1) -- (N3);
        \draw <2> [thick, color=uofgsandstone!50] (N1) -- (N4);
        \draw <2> [thick, color=uofgsandstone!50] (N2) -- (N4);
        \draw <2> [ultra thick] (N3) -- (N5);
        \draw <2> [ultra thick] (N3) -- (N6);
        \draw <2> [thick, color=uofgsandstone!50] (N4) -- (N6);
        \draw <2> [ultra thick] (N5) -- (N6);
        \draw <2> [thick, color=uofgsandstone!50] (N2) to [in=45, out=315] (N6);
        \draw <2> [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
    \end{tikzpicture}\end{center}

\end{frame}

\begin{frame}{Benchmark Instances}

    \begin{itemize}
        \item Other papers: a subset of 5,725 instances from Christine Solnon's selection:
            \begin{itemize}
                \item Randomly generated with different models.
                \item Real-world graphs.
                \item Computer vision problems.
                \item Biochemistry problems.
                \item Phase transition instances.
            \end{itemize}
        \item Here: all 14,621 instances:
            \begin{itemize}
                \item $\ge \input{../paper/gen-n-sat.tex}$ satisfiable.
                \item $\ge \input{../paper/gen-n-unsat.tex}$ unsatisfiable.
            \end{itemize}
        \item A lot of them are very easy for good algorithms.
    \end{itemize}

\end{frame}

\begin{frame}{The Algorithm}

    \begin{itemize}
        \item A variable for every pattern vertex.
        \item Domains range over target vertices.
        \item Clever filtering at the top of search using neighbourhood degree sequences and paths.
        \item Bit-parallel forward checking search:
            \begin{itemize}
                \item Adjacency and pre-computed path count constraints.
                \item Cheap hacky all-different propagator.
                \item Smallest domain first, tie-breaking on highest degree.
            \end{itemize}
    \end{itemize}

\end{frame}

\begin{frame}{Is It Any Good?}

    \includegraphics<1>{gen-graph-others.pdf}%
    \includegraphics<2>{gen-graph-others-zoom.pdf}

\end{frame}

\begin{frame}{Value-Ordering Heuristics}
    \begin{itemize}
        \item Highest static degree first.
        \item But degree spread is usually quite low\ldots
    \end{itemize}
\end{frame}

\begin{frame}{Sanity Check}

    \includegraphics<1>{gen-graph-value-ordering-heuristics.pdf}

\end{frame}

\begin{frame}{Discrepancy Search}

    \includegraphics<1>{gen-graph-scatter-dds.pdf}

\end{frame}

\begin{frame}{Restarts and Nogoods}

    \begin{itemize}
        \item Restarts to affect \emph{value}-ordering heuristics.
        \item Two watched literals to avoid repeating ourselves.
        \item Luby with a magic constant multiplier.
            \begin{itemize}
                \item 666, chosen by divine revelation.
            \end{itemize}
    \end{itemize}

\end{frame}

\begin{frame}{Random Value-Ordering with Restarts}

    \includegraphics<1>{gen-graph-random-restarts.pdf}%
    \includegraphics<2>{gen-graph-scatter-random.pdf}

\end{frame}

\begin{frame}{Biased Value-Ordering}

    \only<1>{
    \begin{itemize}
        \item Select a vertex $v'$ from the chosen domain $D_v$ with
            probability \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \text{.} \]
        \item Looks a lot like $\mathit{softmax}$, which uses base $e$.
    \end{itemize}
    }%
    \includegraphics<2>{gen-graph-value-ordering-heuristics-biased.pdf}

\end{frame}

\begin{frame}{Biased Value-Ordering with Restarts}

    \includegraphics<1>{gen-graph-restarts.pdf}%
    \includegraphics<2>{gen-graph-scatter-final.pdf}

\end{frame}

\section{Maximum Common Subgraph}

\begin{frame}{Maximum Common Subgraph Algorithms}

    \includegraphics<1>{gen-graph-mcs.pdf}%
    \includegraphics<2>{gen-graph-scatter-kdown.pdf}%
    \includegraphics<3>{gen-graph-scatter-mcsplit.pdf}%

\end{frame}

\section{Parallel}

\begin{frame}{Parallel}

    \includegraphics<1>{gen-graph-parallel.pdf}%
    \includegraphics<2>{gen-graph-as-parallel.pdf}%
    \includegraphics<3>{gen-graph-scatter-parallel.pdf}%

\end{frame}

\begin{frame}{Properties of This Sort of Parallel Search}
    \begin{itemize}
        \item Reproducible (assuming each thread has its own deterministic random number generator)
        \item Can perform worse than sequential on satisfiable instances:
            \begin{itemize}
                \item Nogoods can knock out values from domains, leading to a different random
                    search order, even if the first thread uses the same seed.
                \item Challenge for you: design a sane and fast biased random selection algorithm
                    that is immune to this, and that is statistically the same sequentially. It
                    feels like this should be possible\ldots
                \item Also, our all different propagator isn't monotonic or idempotent.
            \end{itemize}

        \item Similarly, adding extra threads can make it worse.
    \end{itemize}
\end{frame}

\begin{frame}[plain,noframenumbering]
    \begin{center}
        \vspace*{4em}
        \url{http://www.dcs.gla.ac.uk/~ciaran} \\
        \vspace{1em}
        \href{mailto:ciaran.mccreesh@glasgow.ac.uk}{\nolinkurl{ciaran.mccreesh@glasgow.ac.uk}} \\
    \end{center}

    \begin{tikzpicture}[remember picture, overlay]
        \node at (current page.north west) {
            \begin{tikzpicture}[remember picture, overlay]
                \fill [fill=uofguniversityblue, anchor=north west] (0, 0) rectangle (\paperwidth, -2.6cm);
            \end{tikzpicture}
        };

        \node (logo) [anchor=north east, shift={(-0.6cm,-0.6cm)}] at (current page.north east) {
            \includegraphics[keepaspectratio=true,scale=0.7]{UoG_keyline.pdf}
        };
    \end{tikzpicture}
\end{frame}

\end{document}

### 2018-01-18 long-talk/talk.tex
Just a little bit more
Another plot

@@
 
-    \includegraphics<1>{gen-graph-value-ordering-heuristics.pdf}
+    \includegraphics<1>{gen-graph-value-ordering-heuristics.pdf}%
+    \includegraphics<2>{gen-graph-value-ordering-heuristics-unsat.pdf}%
 
@@
     }%
-    \includegraphics<2>{gen-graph-value-ordering-heuristics-biased.pdf}
+    \includegraphics<2>{gen-graph-value-ordering-heuristics-biased.pdf}%
+    \includegraphics<3>{gen-graph-scatter-biased.pdf}
 

### 2018-01-18 paper/paper.tex
Things
Patrick's changes

@@
     programming style search. We show that discrepancy-based searches are too expensive to be
-    effective in offsetting mistakes made by value-ordering heuristic in this setting.  We then
+    effective in offsetting mistakes made by value-ordering heuristics in this situation.  We then
     demonstrate that a random heuristic with restarts beats conventional search with the best-known
@@
 \citet{DBLP:conf/aaai/HoffmannMR17}, $G^{n,\ell}$ denotes the graph with vertex set $\vertexset(G)$,
-and with edges between vertices $v$ and $v$ if there are at least $n$ simple paths of length exactly
+and with edges between vertices $v$ and $w$ if there are at least $n$ simple paths of length exactly
 $\ell$ between vertices $v$ and $w$ in $G$.
@@
 (\siplineref{variableordering}), and tries assigning
-it each of its remaining value in turn (\siplineref{valueordering}). The effects of this assignment
+it each of its remaining values in turn (\siplineref{valueordering}). The effects of this assignment
 are propagated (\siplineref{dopropagate}). If an
@@
 \paragraph{Degree} We select vertices from highest degree to lowest degree. This heuristic is the
-current state of the art.
+current state of the art \citep{DBLP:conf/ijcai/McCreeshPT16}.
 
@@
 \Cref{figure:scatter-heuristics} gives a more detailed look at these results. This scatter plot
-shows that despite the aggregate performance being very similar, on an case by case basis, the two
+shows that despite the aggregate performance being very similar, on a case by case basis, the two
 heuristics can make a large difference to the performance of satisfiable instances.
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics*{gen-graph-scatter-constant.pdf}
+
+    \caption{Constant restarts?}
+    \label{figure:scatter-constant}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics*{gen-graph-parallel.pdf}
+
+    \caption{Parallel}
+    \label{figure:parallel}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics*{gen-graph-scatter-parallel.pdf}
+
+    \caption{Parallel Luby}
+    \label{figure:parallel-luby}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics*{gen-graph-scatter-parallel-constant.pdf}
+
+    \caption{Parallel Constant}
+    \label{figure:parallel-constant}
+\end{figure}
+
 \end{document}

### 2018-01-23 parameter-tuning/scenario/wrapper.py
Initial version of parameter tuning using SMAC

@@ -1,0 +1,33 @@
+#!/usr/bin/python
+
+import sys, os, time, re
+from subprocess import Popen, PIPE
+
+# algo, instance, instance information, cutoff time, cutoff length, seed,
+# param...
+
+# Read in first 5 arguments.
+instance = sys.argv[1]
+ignored_1 = sys.argv[2]
+cutoff = int(float(sys.argv[3]) + 1)
+seed = int(sys.argv[4])
+ignored_2 = int(sys.argv[5])
+
+params = sys.argv[6:]
+configMap = dict((name, value) for name, value in zip(params[::2], params[1::2]))
+
+filename1, filename2 = instance.strip().split("@")
+cmd = "../../code/solve_subgraph_isomorphism --luby-multiplier {} --timeout {} --restarts --softmax-shuffle --input-order customisable-sequential ../{} ../{}".format(
+        configMap['-luby-multiplier'], cutoff, filename1, filename2)
+
+# Execute the call and track its runtime.
+start_time = time.time()
+res = os.system(cmd)
+runtime = time.time() - start_time
+
+status = "SUCCESS"
+if res != 0:
+    status = "CRASHED"
+
+# status, runtime, runlength, quality, seed, additional stuff
+print("Result of algorithm run: %s, %f, 0, 0, %i" %(status, runtime, seed))

### 2018-01-23 parameter-tuning/scenario/wrapper.py
Get correct runtime

@@ -21,13 +21,23 @@ cmd = "../../code/solve_subgraph_isomorphism --luby-multiplier {} --timeout {} -
         configMap['-luby-multiplier'], cutoff, filename1, filename2)
 
 # Execute the call and track its runtime.
-start_time = time.time()
-res = os.system(cmd)
-runtime = time.time() - start_time
+#start_time = time.time()
+#res = os.system(cmd)
+#runtime = time.time() - start_time
+
+io = Popen(cmd.split(" "), stdout=PIPE, stderr=PIPE)
+(stdout_, stderr_) = io.communicate()
 
 status = "SUCCESS"
-if res != 0:
+if io.returncode != 0:
     status = "CRASHED"
+elif re.search('status = aborted', stdout_):
+    status = "TIMEOUT"
+
+lines = stdout_.split("\n")
+for line in lines:
+    if re.search('runtime =', line):
+        runtime = float(line.split()[2]) / 1000
 
 # status, runtime, runlength, quality, seed, additional stuff
 print("Result of algorithm run: %s, %f, 0, 0, %i" %(status, runtime, seed))

### 2018-01-23 paper/paper.tex
Little bits

@@
 \begin{abstract}
-    Modern subgraph isomorphism solvers use degree-based heuristics to direct a constraint
-    programming style search. We show that discrepancy-based searches are too expensive to be
-    effective in offsetting mistakes made by value-ordering heuristics in this situation.  We then
-    demonstrate that a random heuristic with restarts beats conventional search with the best-known
-    tailored heuristic. This motivates the introduction of a slightly-random heuristic with similar
-    behaviour to the tailored heuristic: when combined with restarts, this new search strategy is
-    over a hundred times more effective on satisfiable instances, and does not affect performance on
-    unsatisfiable instances. Finally, we show that the same strategy also improves two maximum
-    common induced subgraph algorithms.
+    Modern subgraph isomorphism solvers use degree-based value-ordering heuristics to direct
+    backtracking search. We show that discrepancy-based searches are too expensive to be effective
+    in offsetting heuristic mistakes in this situation.  We then demonstrate that a random heuristic
+    with restarts beats conventional search with the best-known tailored heuristic. This motivates
+    the introduction of a slightly-random heuristic with similar behaviour to the tailored
+    heuristic: when combined with restarts, this new search strategy is over a hundred times more
+    effective on satisfiable instances, and does not affect performance on unsatisfiable instances.
+    Finally, we show that this strategy also improves two maximum common induced subgraph
+    algorithms.
 \end{abstract}
@@
 
+\subsection{Value-Ordering Heuristics}
+
+\Cref{figure:value-ordering-heuristics} shows the cumulative number of satisfiable instances
+solved over time using different value-ordering heuristics:
+
+\paragraph{Degree} We select vertices from highest degree to lowest degree
+\citep{DBLP:conf/ijcai/McCreeshPT16}.
+
+\paragraph{Random} We select vertices uniformly at random.
+
+\paragraph{Anti} We select vertices from lowest degree to highest degree. This heuristic is used as
+a sanity check: as expected, degree beats random, and random beats anti.
+
+\paragraph{Biased} We branch by selecting a vertex $v'$ from the chosen domain $D_v$ with
+probability \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \text{.} \] This heuristic
+continues to prefer vertices of high degree, but will give the same chance of being selected next to
+two vertices of equal degree.  It also introduces an element of randomness, which is needed to make
+restarts have an effect.
+
+The biased heuristic resembles the well-known \emph{softmax} weighting scheme applied to the degree
+heuristic, although usually \emph{softmax} uses a base of $e$ rather than $2$.  Experiments showed
+that in this context, the choice of base is not important to the search space size; however, a base
+of $2$ avoids the need to use floating point arithmetic, which was a substantial overhead.
+
+The performance of the degree and biased heuristics are very similar: the
+biased heuristic has slightly worse performance below a 400s timeout, and effectively equal
+performance above it. This shows that we can introduce an element of randomness into the degree
+value-ordering heuristic without adversely affecting its performance \emph{in aggregate}.
+\Cref{figure:scatter-heuristics} gives a more detailed look at these results. This scatter plot
+shows that despite the aggregate performance being very similar, on a case by case basis, the two
+heuristics can make a large difference to the performance of satisfiable instances.
+
+\begin{figure}[p]
+    \centering
+    \includegraphics*{gen-graph-scatter-dds.pdf}
+
+    \caption{Comparing depth-bounded discrepancy search to backtracking search, both with the degree
+    heuristic.}\label{figure:scatter-dds}
+\end{figure}
+
+\begin{figure}[p]
+    \centering
+    \includegraphics*{gen-graph-value-ordering-heuristics.pdf}
+
+    \caption{The cumulative number of satisfiable instances solved over time, using four
+    different value-ordering heuristics. The degree and biased lines are close together.}
+    \label{figure:value-ordering-heuristics}
+\end{figure}
+
+\begin{figure}[p]
+    \centering
+    \includegraphics*{gen-graph-scatter-heuristics.pdf}
+
+    \caption{An instance by instance comparison of the degree and biased heuristics on all
+    instances. Points on the outer axes are timeouts, and point style shows instance family.}
+    \label{figure:scatter-heuristics}
+\end{figure}
+
 \subsection{Discrepancy Searches}
@@
 realistic prospect, and they are not well-suited for unsatisfiable decision problems. However, the
-extremely large gains on some satisfiable instances confirm our suspicions that we should look at
-value-ordering heuristics in more detail.
-
-\subsection{Value-Ordering Heuristics}
-
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-scatter-dds.pdf}
-
-    \caption{Comparing depth-bounded discrepancy search to backtracking search, both with the degree
-    heuristic.}\label{figure:scatter-dds}
-\end{figure}
-
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-value-ordering-heuristics.pdf}
-
-    \caption{The cumulative number of satisfiable instances solved over time, using four
-    different value-ordering heuristics. The degree and biased lines are close together.}
-    \label{figure:value-ordering-heuristics}
-\end{figure}
-
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-scatter-heuristics.pdf}
-
-    \caption{An instance by instance comparison of the degree and biased heuristics on all
-    instances. Points on the outer axes are timeouts, and point style shows instance family.}
-    \label{figure:scatter-heuristics}
-\end{figure}
-
-In \cref{figure:value-ordering-heuristics} we show the cumulative number of satisfiable instances
-solved over time using four different value-ordering heuristics:
-
-\paragraph{Degree} We select vertices from highest degree to lowest degree. This heuristic is the
-current state of the art \citep{DBLP:conf/ijcai/McCreeshPT16}.
-
-\paragraph{Random} We select vertices uniformly at random.
-
-\paragraph{Anti} We select vertices from lowest degree to highest degree. This heuristic is used as
-a sanity check: as expected, degree beats random, and random beats anti.
-
-\paragraph{Biased} We branch by selecting a vertex $v'$ from the chosen domain $D_v$ with
-probability \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \text{.} \] This heuristic
-continues to prefer vertices of high degree, but will give the same chance of being selected next to
-two vertices of equal degree.  It also introduces an element of randomness, which is needed to make
-restarts have an effect.
-
-The biased heuristic resembles the well-known \emph{softmax} weighting scheme applied to the degree
-heuristic, although usually \emph{softmax} uses a base of $e$ rather than $2$.  Experiments showed
-that in this context, the choice of base is not important to the search space size; however, a base
-of $2$ avoids the need to use floating point arithmetic, which was a substantial overhead.
-
-The performance of the degree and biased heuristics are very similar: the
-biased heuristic has slightly worse performance below a 400s timeout, and effectively equal
-performance above it. This shows that we can introduce an element of randomness into the degree
-value-ordering heuristic without adversely affecting its performance \emph{in aggregate}.
-\Cref{figure:scatter-heuristics} gives a more detailed look at these results. This scatter plot
-shows that despite the aggregate performance being very similar, on a case by case basis, the two
-heuristics can make a large difference to the performance of satisfiable instances.
+extremely large gains on some satisfiable instances confirm our suspicions that we should find an
+alternative to heavy-commitment backtracking search.
 
@@
 indicating large speedups.  Better yet, there are only ?? satisfiable instances that are more than a
-factor of ?? times worse.
-
-?? By family
-
-?? This has made up the consistency we lost from introducing randomness.
-
-?? Explain why this is so amazingly awesome and the most important idea ever to happen in the entire
-history of science.
-
-?? Justify aggregate speedups on sat instances
+factor of ten times worse. In other words, as well as improving performance, we have made up most of
+the consistency we lost by introducing randomness.
 
@@
 
-?? Two recent maximum common induced subgraph algorithms: k${\downarrow}$
-\citep{DBLP:conf/aaai/HoffmannMR17}, and McSplit \citep{DBLP:conf/ijcai/McCreeshPT17}. The former
-attempts to solve the problem by first trying to solve the induced subgraph isomorphism problem, and
-then if that fails, retries allowing a single unmatched vertex (and thus using weaker invariants),
-and so on. Due to its similarity to \cref{algorithm:sip}, we can introduce the same bias and restart
-strategy, requiring only a small amount of jiggery-pokery to handle wildcards in nogoods.
-
-The McSplit algorithm, meanwhile, uses a constraint programming style search, but with special
+Two recent maximum common induced subgraph algorithms also make use of backtracking search with
+degree as a value-ordering heuristic. The k${\downarrow}$ algorithm
+\citep{DBLP:conf/aaai/HoffmannMR17} attempts to solve the problem by first trying to solve the
+induced subgraph isomorphism problem, and then if that fails, retries allowing a single unmatched
+vertex (and thus using weaker invariants), and so on. Due to its similarity to \cref{algorithm:sip},
+we can introduce the same bias and restart strategy, requiring only a small amount of jiggery-pokery
+to handle wildcards in nogoods.  Meanwhile, the McSplit algorithm
+\citep{DBLP:conf/ijcai/McCreeshPT17} uses a constraint programming style search, but with special
 propagators and backtrackable data structures that exploit special properties of the problem. The
@@
 
-\clearpage
-\appendix
-
-\section{Extra Plots not for the Final Paper}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics*{gen-graph-as.pdf}
-    \caption{?? Aggregate speedups on satisfiable instances}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics*{gen-graph-as-mcs.pdf}
-    \caption{?? Aggregate speedups on MCS instances}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics*{gen-graph-scatter-degree.pdf}
-
-    \caption{When is the degree heuristic good?}
-    \label{figure:scatter-degree}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics*{gen-graph-scatter-constant.pdf}
-
-    \caption{Constant restarts?}
-    \label{figure:scatter-constant}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics*{gen-graph-parallel.pdf}
-
-    \caption{Parallel}
-    \label{figure:parallel}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics*{gen-graph-scatter-parallel.pdf}
-
-    \caption{Parallel Luby}
-    \label{figure:parallel-luby}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics*{gen-graph-scatter-parallel-constant.pdf}
-
-    \caption{Parallel Constant}
-    \label{figure:parallel-constant}
-\end{figure}
-
 \end{document}

### 2018-01-29 paper/paper.tex
Are we nearly there yet?

@@
 \documentclass{article}
-\usepackage{ijcai17}
+\pdfpagewidth=8.5in
+\pdfpageheight=11in
+\usepackage{ijcai18}
 \usepackage{times}
+\usepackage{xcolor}
+\usepackage{soul}
+\usepackage[utf8]{inputenc}
+\usepackage[small]{caption}
 \usepackage{microtype}
@@
 
-\title{Value Ordering, Discrepancies, and Restarts for Subgraph Algorithms\thanks{This work was
-supported by the Engineering and Physical Sciences Research Council [grant numbers EP/026842/1 and EP/M508056/1]}}
-\author{Ciaran McCreesh \and Patrick Prosser \and James Trimble\\ University of Glasgow, Glasgow, Scotland \\
-    ciaran.mccreesh@glasgow.ac.uk}
+\title{Value Ordering, Discrepancies, and Restarts for Subgraph Algorithms\thanks{This footnote
+leaves room for a piece of text saying how grateful we are to our funding agency. This includes a
+magic grant number, so it takes up three precious lines of space.}}
+\author{Anonymous \and Anonymous \and Anonymous \\ Somewhere \\ Some Email Address}
 
@@
 \emph{neighbourhood} of a vertex, $\neighbourhood_G(v)$, is the set of vertices adjacent to $v$, and
-the \emph{degree} of $v$, $\deg_G(v)$, is the cardinality of its neighbourhood; when the graph is
-clear from the context, we omit the $G$ subscript. The \emph{neighbourhood degree sequence} of $v$,
-$\nds_G(v)$, is the sequence of degrees of the neighbours of $v$ in descending order, and we denote
-by $\preceq$ the inclusion relation of \citet{DBLP:journals/constraints/ZampelliDS10}. As per
-\citet{DBLP:conf/aaai/HoffmannMR17}, $G^{n,\ell}$ denotes the graph with vertex set $\vertexset(G)$,
-and with edges between vertices $v$ and $w$ if there are at least $n$ simple paths of length exactly
-$\ell$ between vertices $v$ and $w$ in $G$.
+the \emph{degree} of $v$, $\deg_G(v)$, is the cardinality of its neighbourhood. The
+\emph{neighbourhood degree sequence} of $v$, $\nds_G(v)$, is the sequence of degrees of the
+neighbours of $v$ in descending order, and we denote by $\preceq$ the inclusion relation of
+\citet{DBLP:journals/constraints/ZampelliDS10}. When the graph is clear from the context, we omit
+the $G$ subscripts. As per \citet{DBLP:conf/aaai/HoffmannMR17}, $G^{n,\ell}$ denotes the graph with
+vertex set $\vertexset(G)$, and with edges between vertices $v$ and $w$ if there are at least $n$
+simple paths of length exactly $\ell$ between vertices $v$ and $w$ in $G$.
 
@@
     \centering
-    \includegraphics*{gen-graph-scatter-dds.pdf}
-
-    \caption{Comparing depth-bounded discrepancy search to backtracking search, both with the degree
-    heuristic.}\label{figure:scatter-dds}
-\end{figure}
-
-\begin{figure}[p]
-    \centering
     \includegraphics*{gen-graph-value-ordering-heuristics.pdf}
@@
     \label{figure:scatter-heuristics}
+\end{figure}
+
+\begin{figure}[p]
+    \centering
+    \includegraphics*{gen-graph-scatter-dds.pdf}
+
+    \caption{Comparing depth-bounded discrepancy search to backtracking search, both with the degree
+    heuristic.}\label{figure:scatter-dds}
 \end{figure}
@@
 introducing randomness or some form of learning into its variable-ordering heuristic. However,
-introducing a second change would considerably complicate the empirical analysis. We therefore leave
-this change as future work.} Using restarts on value-ordering heuristics appears to be uncommon,
-although \citet{DBLP:conf/flairs/RazgonOP07} look at learning value-ordering heuristics from
-restarts, and \citet{DBLP:conf/cp/ChuSS09} use a similar scheme in the context of parallel search.
+simultaneously introducing a second change would considerably complicate the empirical analysis.}
+Using restarts on value-ordering heuristics appears to be uncommon, although
+\citet{DBLP:conf/flairs/RazgonOP07} look at learning value-ordering heuristics from restarts, and
+\citet{DBLP:conf/cp/ChuSS09} use a similar scheme in the context of parallel search.
 
@@
 \siplineref{countrestarts}) to decide when to restart, rather than counting recursive calls.
-Following established wisdom, we multiply each item in the Luby sequence by a magic constant.
-Preliminary experiments demonstrated that a constant 100 is slightly too small and 1,000 is slightly
-too large, but lacking a more principled way of selecting the constant's value, we fell back on
-divine revelation and use a multiplier of 666. (We also tried a geometric scheme. The results were
-less favourable---it appears that value-ordering heuristics benefit from a very aggressive restart
-strategy.)
+Following established wisdom, we multiply each item in the Luby sequence by a magic constant---we
+used SMAC \citep{DBLP:conf/lion/HutterHL11} to select a value of 660. (We also tried a geometric
+restart scheme. The results were less favourable---it appears that value-ordering heuristics benefit
+from a very aggressive restart strategy.)
 
@@
 restart, we post a nogood of the form $(v \mapsto v') \wedge (w \mapsto w') \wedge (x \mapsto x')
-\Rightarrow \bot$ for the current branch, and for every branch to its left in the search tree.
+\Rightarrow \bot$ for every branch to the left of the current (incomplete) branch in the search tree.
 
@@
 satisfiable instances, on a case by case basis many instances perform substantially worse with
-randomisation and restarts. ?? Quantify this a bit
+randomisation and restarts.
 
@@
 
-\Cref{figure:restarts} shows that the biased heuristic together with restarts is better still---that
-is, if we are introducing restarts, then it is better to add a small amount of randomness to a
-tailored deterministic heuristic than it is to throw away the heuristic altogether.
+\Cref{figure:restarts} also shows that the biased heuristic together with restarts is better
+still---that is, if we are introducing restarts, then it is better to add a small amount of
+randomness to a tailored deterministic heuristic than it is to throw away the heuristic altogether.
+Indeed, the original algorithm can solve 1,983 satisfiable instances by 993.0 seconds, whilst the
+biased and random restarting algorithms require only 9.2 seconds and 33.8 seconds respectively to
+solve the same number.
 
@@
 are very close to the $x-y$ diagonal, showing that their performance is nearly unchanged. On the
-other hand, there are ?? large numbers of satisfiable instances well below the diagonal line,
-indicating large speedups.  Better yet, there are only ?? satisfiable instances that are more than a
-factor of ten times worse. In other words, as well as improving performance, we have made up most of
-the consistency we lost by introducing randomness.
+other hand, there are large numbers of satisfiable instances well below the diagonal line,
+indicating large speedups.  Better yet, there are only a handful of satisfiable instances that are
+more than a factor of ten times worse. In other words, as well as improving performance, we have
+made up most of the consistency we lost by introducing randomness.
 
@@
 we can introduce the same bias and restart strategy, requiring only a small amount of jiggery-pokery
-to handle wildcards in nogoods.  Meanwhile, the McSplit algorithm
+to handle wildcards in nogoods.  Meanwhile, the McSplit${\downarrow}$ algorithm
 \citep{DBLP:conf/ijcai/McCreeshPT17} uses a constraint programming style search, but with special
 propagators and backtrackable data structures that exploit special properties of the problem. The
-unconventional domain store used by McSplit precludes the use of arbitrary unit propagation, and so
+unconventional domain store used by McSplit${\downarrow}$ precludes the use of arbitrary unit propagation, and so
 when introducing restarts, we cannot propagate using nogoods.  Instead, we can only detect when we
@@
 
-?? Benchmark results on the same families of instances are shown in
-\cref{figure:kdown,figure:mcsplit,figure:mcs}. Clearly beneficial. Note that on McSplit we're doing
-extra work and this shows, but it's still worthwhile.
+Performance results from these two modified algorithms, using the same families of instances as in
+the previous section, are shown in \cref{figure:kdown,figure:mcsplit,figure:mcs}. Although we have
+moved from a decision problem to an optimisation problem, the same changes remain clearly
+beneficial. For the $k{\downarrow}$ algorithm, the change has a minimal effect on many instances
+(typically, where the $k = 0$ subproblem is unsatisfiable and hard, and the $k = 1$ subproblem is
+satisfiable and easy), but gives large benefits on many more instances than it penalises: it is over
+an order of magnitude better on nearly a hundred instances, whilst being an order of magnitude worse
+on only five.
+
+With McSplit${\downarrow}$, the inability to use two watched literals means that in many cases we
+introduce a small slowdown. However, the overall pattern is the same: when introducing restarts and
+a biased value ordering heuristic, it is much more common to see a large speedup than it is a large
+slowdown.
+
+\begin{figure}[p]
+    \centering
+    \includegraphics*{gen-graph-mcs.pdf}
+    \caption{The cumulative number of maximum common induced subgraph instances solved by
+    k${\downarrow}$ and McSplit${\downarrow}$ over time, with and without restarts.}\label{figure:mcs}
+\end{figure}
 
@@
     \includegraphics*{gen-graph-scatter-kdown.pdf}
-    \caption{k${\downarrow}$ speedups}\label{figure:kdown}
+    \caption{Comparing k${\downarrow}$ with and without restarts on an instance by instance basis.}\label{figure:kdown}
 \end{figure}
@@
     \includegraphics*{gen-graph-scatter-mcsplit.pdf}
-    \caption{McSplit speedups}\label{figure:mcsplit}
+    \caption{Comparing McSplit${\downarrow}$ with and without restarts on an instance by instance
+    basis.}\label{figure:mcsplit}
 \end{figure}
 
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-mcs.pdf}
-    \caption{k${\downarrow}$ and McSplit cumulative curves}\label{figure:mcs}
-\end{figure}
+\section{Conclusion and Future Work}
 
-\section{Conclusion}
+We have shown that it is both possible and beneficial to introduce a small amount of randomness into
+backtracking search algorithms for hard subgraph problems, without compromising existing search
+order heuristics, and without paying a penalty for unsatisfiable instances.
 
-?? This stuff is quite good.
+The success of this approach raises a number of interesting questions and avenues for possible
+future research. For example, will this same technique work for other problems where we have a good
+value-ordering heuristic? Are there further benefits to be had from also modifying variable-ordering
+heuristics? Are there theoretical rather than just empirical reasons for using the
+\emph{softmax}-like bias scheme? And how can such a search process be parallelised?
 
-?? We can keep good value-ordering heuristics, and get large improvement. Importantly, we can do
-this with almost no effect on unsatisfiable instances, and rare to make satisfiable much worse.
+\section*{Acknowledgments}
 
-?? No theoretical justification for biased. Also tried polynomial weightings, exponential using
-$\left|D_v\right|$ as the base. Using $e$ (or for that matter, $2$ or $\pi$) as a base seems to be
-``just the right amount'' of randomness.
+We would like to thank some anonymous people for their comments. These people have names that take
+up approximately this amount of space, which is helpful for us getting the page layout right when
+anonymising the paper for review.
 
-?? Variable ordering
-
-?? Bigger CP picture.
-
-?? Parallel future work
+% Ruth Hoffmann, Christophe Lecoutre, Christine Solnon, Craig Reilly.
 

### 2018-01-29 paper/paper.tex
660

@@
 an order of magnitude better on nearly a hundred instances, whilst being an order of magnitude worse
-on only five.
+on only seven.
 

### 2018-01-31 paper/paper.tex
James's comments

@@
 introduce tiebreaking mechanisms and similar refinements to strengthen the heuristic. However, given
-the extreme cost of making an incorrect branching choice early on in search, we believe it is more
+the extreme cost of making an incorrect branching choice early in search, we believe it is more
 fruitful to investigate alternatives to simple backtracking (depth-first) search. This paper
@@
 
-Our $\FuncSty{propagate}$ routine is similarly close to that of \citet{DBLP:conf/aaai/HoffmannMR17},
+Our $\FuncSty{propagate}$ routine is also close to that of \citet{DBLP:conf/aaai/HoffmannMR17},
 differing only in that we carry out the filtering for injectivity
@@
 be to allow the rightmost algorithm to succeed on $y$ out of the 14,621 instances (bearing in mind
-that the two sets of $y$ instances solved by the two algorithms could be completely different).
+that the two sets of $y$ instances could be different).
 
@@
 Following established wisdom, we multiply each item in the Luby sequence by a magic constant---we
-used SMAC \citep{DBLP:conf/lion/HutterHL11} to select a value of 660. (We also tried a geometric
+used SMAC \citep{DBLP:conf/lion/HutterHL11} to select the value 660. (We also tried a geometric
 restart scheme. The results were less favourable---it appears that value-ordering heuristics benefit
@@
 
-Two recent maximum common induced subgraph algorithms also make use of backtracking search with
-degree as a value-ordering heuristic. The k${\downarrow}$ algorithm
+Having looked at subgraph isomorphism in detail, we now briefly discuss the more general maximum
+common induced subgraph problem.  Two recent algorithms for this problem also make use of
+backtracking search with degree as a value-ordering heuristic. The k${\downarrow}$ algorithm
 \citep{DBLP:conf/aaai/HoffmannMR17} attempts to solve the problem by first trying to solve the
@@
 propagators and backtrackable data structures that exploit special properties of the problem. The
-unconventional domain store used by McSplit${\downarrow}$ precludes the use of arbitrary unit propagation, and so
-when introducing restarts, we cannot propagate using nogoods.  Instead, we can only detect when we
-are inside an already-visited branch.  We must therefore use the one watched literal scheme instead.
+unconventional domain store used by McSplit${\downarrow}$ precludes the use of arbitrary unit
+propagation, and so when introducing restarts, we cannot propagate using nogoods.  Instead, we can
+only detect when we are inside an already-visited branch.  We must therefore use the one watched
+literal scheme instead, and we also introduce a basic subsumption scheme to prune redundant clauses.
 
@@
 
-We would like to thank some anonymous people for their comments. These people have names that take
-up approximately this amount of space, which is helpful for us getting the page layout right when
-anonymising the paper for review.
+We would like to thank some anonymous people for their comments.
 

### 2018-01-31 paper/paper.tex
Some of Patrick's changes

@@
 smallest domain first, tiebreaking on highest degree. Variable ordering is not the focus of this
-paper---as far as possible, we will leave it intact, so as not to affect the performance of the
-algorithm on unsatisfiable instances.  For value ordering, which occurs on
-\siplineref{valueordering} of \cref{algorithm:sip}, we will initially select vertices from highest
-degree to lowest degree; we return to value ordering in the second half of this paper. Note that the
-variable-ordering heuristic affects both unsatisfiable and satisfiable instances, whilst the
-value-ordering heuristic is relevant only for satisfiable instances.
+paper---as far as possible, we use this same dynamic ordering throughout the paper, so as not to
+affect the performance of the algorithm on unsatisfiable instances.  For value ordering, which
+occurs on \siplineref{valueordering} of \cref{algorithm:sip}, we will initially select vertices from
+highest degree to lowest degree; we return to value ordering in the second half of this paper. Note
+that the variable-ordering heuristic affects both unsatisfiable and satisfiable instances, whilst
+the value-ordering heuristic is relevant only for satisfiable instances.
 
@@
 \citep{DBLP:journals/jsat/LecoutreSTV07,DBLP:conf/cp/GayHLS15,DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17}.
-In this paper, we instead rely upon randomness in the \emph{value}-ordering heuristic, and keep the
-variable ordering fixed.\footnote{It may be possible to further improve \cref{algorithm:sip} by also
-introducing randomness or some form of learning into its variable-ordering heuristic. However,
-simultaneously introducing a second change would considerably complicate the empirical analysis.}
-Using restarts on value-ordering heuristics appears to be uncommon, although
-\citet{DBLP:conf/flairs/RazgonOP07} look at learning value-ordering heuristics from restarts, and
-\citet{DBLP:conf/cp/ChuSS09} use a similar scheme in the context of parallel search.
+In this paper, we instead rely upon randomness in the \emph{value}-ordering heuristic, and continue
+to use smallest domain first with static tiebreaking for variable-ordering.\footnote{It may be
+possible to further improve \cref{algorithm:sip} by also introducing randomness or some form of
+learning into its variable-ordering heuristic. However, simultaneously introducing a second change
+would considerably complicate the empirical analysis.} Using restarts on value-ordering heuristics
+appears to be uncommon, although \citet{DBLP:conf/flairs/RazgonOP07} look at learning value-ordering
+heuristics from restarts, and \citet{DBLP:conf/cp/ChuSS09} use a similar scheme in the context of
+parallel search.
 
@@
 used SMAC \citep{DBLP:conf/lion/HutterHL11} to select the value 660. (We also tried a geometric
-restart scheme. The results were less favourable---it appears that value-ordering heuristics benefit
-from a very aggressive restart strategy.)
+restart scheme, which gave less favourable results.)
 
@@
 the random value-ordering heuristic comfortably beats the degree strategy without restarts. In other
-words, although having a good variable-ordering heuristic is beneficial, introducing randomness into
+words, although having a good value-ordering heuristic is beneficial, introducing randomness into
 the search is better, if it is done alongside a mechanism to avoid heavy commitment to any
@@
     \includegraphics*{gen-graph-scatter-random.pdf}
-    \caption{Comparing basic backtracking with the degree heuristic, versus the Random heuristic
+    \caption{Comparing basic backtracking with the degree heuristic, versus the random heuristic
     with restarts.}
@@
 vertex (and thus using weaker invariants), and so on. Due to its similarity to \cref{algorithm:sip},
-we can introduce the same bias and restart strategy, requiring only a small amount of jiggery-pokery
-to handle wildcards in nogoods.  Meanwhile, the McSplit${\downarrow}$ algorithm
+we can introduce the same bias and restart strategy, requiring only small changes
+to handle wildcards in nogoods.
+
+Meanwhile, the McSplit${\downarrow}$ algorithm
 \citep{DBLP:conf/ijcai/McCreeshPT17} uses a constraint programming style search, but with special

### 2018-04-16 paper/paper.tex
IJCAI doesn't love us. ESA?

@@
 
-\documentclass{article}
-\pdfpagewidth=8.5in
-\pdfpageheight=11in
-\usepackage{ijcai18}
-\usepackage{times}
-\usepackage{xcolor}
-\usepackage{soul}
-\usepackage[utf8]{inputenc}
-\usepackage[small]{caption}
+\documentclass[a4paper,UKenglish]{lipics-v2018}
+
 \usepackage{microtype}
+\usepackage{complexity}
+\usepackage{subcaption}
 \usepackage{amsmath}
@@
 
-\usepackage{cleveref}
+% \usepackage{showframe}
 
-% \usepackage{showframe}
+\bibliographystyle{plainurl}
 
@@
 
-\newcommand{\citet}[1]{\citeauthor{#1} \shortcite{#1}}
-\newcommand{\citep}[1]{\cite{#1}}
-\newcommand{\citepos}[1]{\citeauthor{#1}'s \shortcite{#1}}
-
 \newcommand{\siplabel}[1]{\label{line:sip:#1}}
@@
 \newcommand{\siplinerangeref}[2]{lines~\ref{line:sip:#1} to~\ref{line:sip:#2}}
+
+\newcommand{\figureref}[1]{Figure~\ref{#1}}
+\newcommand{\Figureref}[1]{Figure~\ref{#1}}
+\newcommand{\threefigureref}[3]{Figures~\ref{#1}, \ref{#2} and \ref{#3}}
+\newcommand{\algorithmref}[1]{Algorithm~\ref{#1}}
+\newcommand{\Algorithmref}[1]{Algorithm~\ref{#1}}
 
@@
 
-% cref style
-\crefname{algorithm}{Algorithm}{Algorithms}
-\Crefname{algorithm}{Algorithm}{Algorithms}
-\crefname{figure}{Figure}{Figures}
-\Crefname{figure}{Figure}{Figures}
-\crefname{table}{Table}{Tables}
-\Crefname{table}{Table}{Tables}
-\crefname{section}{Section}{Sections}
-\Crefname{section}{Section}{Sections}
+\title{Value Ordering, Discrepancies, and Restarts for Subgraph Algorithms}
+\author{Ciaran McCreesh}{University of Glasgow}{ciaran.mccreesh@glasgow.ac.uk}{}{}
+\author{Patrick Prosser}{University of Glasgow}{patrick.prosser@glasgow.ac.uk}{}{}
+\author{James Trimble}{University of Glasgow}{j.trimble.1@research.gla.ac.uk}{}{}
 
-\title{Value Ordering, Discrepancies, and Restarts for Subgraph Algorithms\thanks{This footnote
-leaves room for a piece of text saying how grateful we are to our funding agency. This includes a
-magic grant number, so it takes up three precious lines of space.}}
-\author{Anonymous \and Anonymous \and Anonymous \\ Somewhere \\ Some Email Address}
+\authorrunning{C. McCreesh, P. Prosser and J. Trimble}
+
+\Copyright{Ciaran McCreesh, Patrick Prosser and James Trimble}
+
+\subjclass{
+\ccsdesc[500]{Mathematics of computing~Graph algorithms},
+\ccsdesc[500]{Theory of computation~Discrete optimization},
+\ccsdesc[500]{Theory of computation~Backtracking}}
+
+\keywords{Subgraph isomorphism, combinatorial search, rich search}
+
+\funding{This work was supported by the Engineering and Physical Sciences Research Council [grant
+numbers EP/026842/1 and EP/M508056/1]}
+
+\acknowledgements{The authors would like to thank Ruth Hoffmann, Christophe Lecoutre, Christine
+Solnon and Craig Reilly for their comments.}
+
+%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\EventEditors{John Q. Open and Joan R. Access}
+\EventNoEds{2}
+\EventLongTitle{26th Annual European Symposium on Algorithms (ESA 2018)}
+\EventShortTitle{ESA 2018}
+\EventAcronym{ESA}
+\EventYear{2018}
+\EventDate{August 20--22, 2018}
+\EventLocation{Helsinki, Finland}
+\EventLogo{}
+\SeriesVolume{42}
+\ArticleNo{23}
+%\nolinenumbers %uncomment to disable line numbering
+\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 first'' as a value-ordering heuristic to direct a constraint programming style search
-\citep{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:conf/ijcai/McCreeshPT16}. This
+\cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:conf/ijcai/McCreeshPT16}. This
 heuristic is much better than branching randomly, but is still far from perfect.  We could try to
@@
 first'' heuristic.  The effects of these changes are shown in
-\cref{figure:scatter-by-family}: we can make a state-of-the-art algorithm perform much better on a
+\figureref{figure:scatter-by-family}: we can make a state-of-the-art algorithm perform much better on a
 large number of satisfiable instances, whilst performing worse only rarely on satisfiable instances,
 and never on unsatisfiable instances.  Further, these properties are \emph{not} shared by
-discrepancy searches (\cref{figure:scatter-dds}) or by random value-ordering with restarts
-(\cref{figure:scatter-random}). Finally, we show that this strategy is also effective in an
+discrepancy searches (\figureref{figure:scatter-dds}) or by random value-ordering with restarts
+(\figureref{figure:scatter-random}). Finally, we show that this strategy is also effective in an
 optimisation setting, producing benefits in two maximum common induced subgraph algorithms.
@@
 neighbours of $v$ in descending order, and we denote by $\preceq$ the inclusion relation of
-\citet{DBLP:journals/constraints/ZampelliDS10}. When the graph is clear from the context, we omit
-the $G$ subscripts. As per \citet{DBLP:conf/aaai/HoffmannMR17}, $G^{n,\ell}$ denotes the graph with
+Zampelli et al.\ \cite{DBLP:journals/constraints/ZampelliDS10}. When the graph is clear from the context, we omit
+the $G$ subscripts. As per Hoffmann et al.\ \cite{DBLP:conf/aaai/HoffmannMR17}, $G^{n,\ell}$ denotes the graph with
 vertex set $\vertexset(G)$, and with edges between vertices $v$ and $w$ if there are at least $n$
@@
 
-We now briefly outline the non-shaded parts of \cref{algorithm:sip}, which is how we will solve the
+We now briefly outline the non-shaded parts of \algorithmref{algorithm:sip}, which is how we will solve the
 subgraph isomorphism problem. This algorithm is very closely based upon the $k{\downarrow}$
-algorithm of \citet{DBLP:conf/aaai/HoffmannMR17} with $k = 0$, and we refer the reader to that paper
+algorithm of Hoffmann et al.\ \cite{DBLP:conf/aaai/HoffmannMR17} with $k = 0$, and we refer the reader to that paper
 for full technical details; that algorithm, in turn, is a simplification of the Glasgow2 algorithm
-\citep{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}.
+\cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}.
 
-\Cref{algorithm:sip} works by recursively building up a valid mapping of pattern vertices to target
+\Algorithmref{algorithm:sip} works by recursively building up a valid mapping of pattern vertices to target
 vertices. For each pattern vertex $v$, we have a domain $D_v$ containing a set of feasible target
 vertices. Each domain is initialised at the top of search (\siplinerangeref{initstart}{initend}),
-using the invariants described by \citet{DBLP:conf/aaai/HoffmannMR17} to restrict its initial
+using the invariants described by Hoffmann et al.\ \cite{DBLP:conf/aaai/HoffmannMR17} to restrict its initial
 values.
@@
 
-Our $\FuncSty{propagate}$ routine is also close to that of \citet{DBLP:conf/aaai/HoffmannMR17},
+Our $\FuncSty{propagate}$ routine is also close to that of Hoffmann et al.\ \cite{DBLP:conf/aaai/HoffmannMR17},
 differing only in that we carry out the filtering for injectivity
@@
         (\mathcal{P}^{1,2}, \mathcal{T}^{1,2}),
-        (\mathcal{P}^{2,2},\ \mathcal{T}^{2,2}), $\\\hspace*{6em}$ (\mathcal{P}^{3,2},\ \mathcal{T}^{3,2}),
+        (\mathcal{P}^{2,2},\ \mathcal{T}^{2,2}), (\mathcal{P}^{3,2},\ \mathcal{T}^{3,2}),
         (\mathcal{P}^{4,2},\ \mathcal{T}^{4,2}) \big]$ \;
@@
         \nl $D_v \gets \vertexset(\mathcal{T})$\;
-        \nl \ForEach{$(P,\,T) \in L$}{
-            \nl $D_v \gets \{ w \in D_v : $ \siplabel{degreefiltering} \\\hspace*{4em}$v \sim_P v \Rightarrow w \sim_T w~\wedge$
-            \\\hspace*{4em}$\nds_P(v) \preceq \nds_T(w)\}$\siplabel{initend} \;
+        \nl \lForEach{$(P,\,T) \in L$}{$D_v \gets \{ w \in D_v : $ \siplabel{degreefiltering} $v \sim_P v \Rightarrow w \sim_T w~\wedge$
+            $\nds_P(v) \preceq \nds_T(w)\}$\siplabel{initend}
         }
@@
             \nl $D_w \gets D_w - v'$ \siplabel{removev} \;
-            \nl \ForEach{$(P,\,T) \in L$}{
-                \nl \lIf{$v \sim_P w$}{$D_w \gets D_w \cap \neighbourhood_T(v')$
-                }
-            }
+            \nl \lForEach{$(P,\,T) \in L~\KwSty{where}~v \sim_P w$}{$D_w \gets D_w \cap \neighbourhood_T(v')$}
             \nl \lIf{$D_w = \emptyset$}{$\KwSty{return}~\KwSty{false}$}
@@
 \nl \ForEach{$D_v \in D$ \textnormal{from smallest cardinality upwards\siplabel{eachdomain}}}{
-    \nl $D_v \gets D_v \setminus H$ \siplabel{elimhall} \;
-    \nl $(A,\,n) \gets (A \cup D_v,\,n + 1)$ \siplabel{acc} \;
+    \nl $(D_v, A,\,n) \gets (D_v \gets D_v \setminus H\siplabel{elimhall}, A \cup D_v,\,n + 1)$ \siplabel{acc} \;
     \nl \lIf{$D_v = \emptyset~\vee~|A| < n$}{$\KwSty{return}~\KwSty{false}$\siplabel{failhall}}
@@
 
-For the variable-ordering heuristic on \siplineref{variableordering} of \cref{algorithm:sip}, we use
+For the variable-ordering heuristic on \siplineref{variableordering} of \algorithmref{algorithm:sip}, we use
 smallest domain first, tiebreaking on highest degree. Variable ordering is not the focus of this
@@
 affect the performance of the algorithm on unsatisfiable instances.  For value ordering, which
-occurs on \siplineref{valueordering} of \cref{algorithm:sip}, we will initially select vertices from
+occurs on \siplineref{valueordering} of \algorithmref{algorithm:sip}, we will initially select vertices from
 highest degree to lowest degree; we return to value ordering in the second half of this paper. Note
@@
 Our experiments are performed on systems with dual Intel Xeon E5-2697A v4 CPUs and 512GBytes RAM,
-running Ubuntu 17.04. We implemented\footnote{Institutional DOI URL removed for anonymous review}
-\cref{algorithm:sip} in C++ using the bit-parallel implementation of
-\citet{DBLP:conf/aaai/HoffmannMR17} as a starting point, and compiled it using GCC 6.3.0. We use a
+running Ubuntu 17.04. We
+implemented\footnote{\url{https://github.com/ciaranm/ijcai2018-restarting-sip-paper}, DOI to be issued pre-publication}
+\algorithmref{algorithm:sip} in C++ using the bit-parallel implementation of
+Hoffmann et al.\ \cite{DBLP:conf/aaai/HoffmannMR17} as a starting point, and compiled it using GCC 6.3.0. We use a
 deterministic pseudo-random number generator for reproducibility.
 
-We use the dataset introduced by \citet{DBLP:conf/lion/KotthoffMS16} for evaluation. This dataset
+We use the dataset introduced by Kotthoff et al.\ \cite{DBLP:conf/lion/KotthoffMS16} for evaluation. This dataset
 brings together a range of randomly-generated and application instance families from earlier papers.
@@
 
-We begin by verifying that our implementation of the basic \cref{algorithm:sip} is competitive. In
-\cref{figure:others} we show the cumulative number of instances solved over time, using our
+We begin by verifying that our implementation of the basic \algorithmref{algorithm:sip} is competitive. In
+\figureref{figure:others} we show the cumulative number of instances solved over time, using our
 implementation (both with the degree heuristic, and with the modifications described in the
-remainder of this paper), the Glasgow2 and Glasgow3 algorithms from which \cref{algorithm:sip} is derived
-\citep{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}, the PathLAD variation of the LAD
-algorithm \citep{DBLP:journals/ai/Solnon10,DBLP:conf/lion/KotthoffMS16}, and VF2
-\citep{DBLP:journals/pami/CordellaFSV04}, in each case using the original implementation provided
+remainder of this paper), the Glasgow2 and Glasgow3 algorithms from which \algorithmref{algorithm:sip} is derived
+\cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}, the PathLAD variation of the LAD
+algorithm \cite{DBLP:journals/ai/Solnon10,DBLP:conf/lion/KotthoffMS16}, and VF2
+\cite{DBLP:journals/pami/CordellaFSV04}, in each case using the original implementation provided
 by the algorithm's authors. For each algorithm, the $y$ value
@@
 
-As expected, the performance of \cref{algorithm:sip} with the degree heuristic is very close to that
+As expected, the performance of \algorithmref{algorithm:sip} with the degree heuristic is very close to that
 of Glasgow2, although we perform better at lower runtimes because we use simpler preprocessing at
@@
 upon improving the performance on the remaining hard satisfiable instances, and so in the second
-plot in \cref{figure:others} (and in subsequent cumulative plots) we show only satisfiable
+plot in \figureref{figure:others} (and in subsequent cumulative plots) we show only satisfiable
 instances, and use a reduced range on both axes.
@@
 
-\Cref{figure:value-ordering-heuristics} shows the cumulative number of satisfiable instances
+\Figureref{figure:value-ordering-heuristics} shows the cumulative number of satisfiable instances
 solved over time using different value-ordering heuristics:
 
-\paragraph{Degree} We select vertices from highest degree to lowest degree
-\citep{DBLP:conf/ijcai/McCreeshPT16}.
+\begin{description}
+\item[Degree] We select vertices from highest degree to lowest degree
+\cite{DBLP:conf/ijcai/McCreeshPT16}.
 
-\paragraph{Random} We select vertices uniformly at random.
+\item[Random] We select vertices uniformly at random.
 
-\paragraph{Anti} We select vertices from lowest degree to highest degree. This heuristic is used as
+\item[Anti] We select vertices from lowest degree to highest degree. This heuristic is used as
 a sanity check: as expected, degree beats random, and random beats anti.
 
-\paragraph{Biased} We branch by selecting a vertex $v'$ from the chosen domain $D_v$ with
-probability \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \text{.} \] This heuristic
-continues to prefer vertices of high degree, but will give the same chance of being selected next to
-two vertices of equal degree.  It also introduces an element of randomness, which is needed to make
-restarts have an effect.
+\item[Biased] We branch by selecting a vertex $v'$ from the chosen domain $D_v$ with
+probability \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \text{.} \]\end{description}
 
-The biased heuristic resembles the well-known \emph{softmax} weighting scheme applied to the degree
-heuristic, although usually \emph{softmax} uses a base of $e$ rather than $2$.  Experiments showed
-that in this context, the choice of base is not important to the search space size; however, a base
-of $2$ avoids the need to use floating point arithmetic, which was a substantial overhead.
+The biased heuristic continues to prefer vertices of high degree, but will give the same chance of
+being selected next to two vertices of equal degree.  It also introduces an element of randomness,
+which is needed to make restarts have an effect.  We note that it resembles the well-known
+\emph{softmax} weighting scheme applied to the degree heuristic, although usually \emph{softmax}
+uses a base of $e$ rather than $2$.  Experiments showed that in this context, the choice of base is
+not important to the search space size; however, a base of $2$ avoids the need to use floating point
+arithmetic, which was a substantial overhead.
 
@@
 value-ordering heuristic without adversely affecting its performance \emph{in aggregate}.
-\Cref{figure:scatter-heuristics} gives a more detailed look at these results. This scatter plot
+\Figureref{figure:scatter-heuristics} gives a more detailed look at these results. This scatter plot
 shows that despite the aggregate performance being very similar, on a case by case basis, the two
@@
 A \emph{discrepancy} is where search goes against a value-ordering heuristic.  Discrepancy searches
-\citep{DBLP:conf/ijcai/HarveyG95,DBLP:conf/aaai/Korf96,DBLP:conf/ijcai/Walsh97} are alternatives to
+\cite{DBLP:conf/ijcai/HarveyG95,DBLP:conf/aaai/Korf96,DBLP:conf/ijcai/Walsh97} are alternatives to
 backtracking search that initially search disallowing all discrepancies, and then retry search
@@
 
-\Cref{figure:scatter-dds} shows the effects of adding \citepos{DBLP:conf/ijcai/Walsh97}
-depth-bounded discrepancy search (DDS) to \cref{algorithm:sip}. Each point represents the solving
+\Figureref{figure:scatter-dds} shows the effects of adding Walsh's {DBLP:conf/ijcai/Walsh97}
+depth-bounded discrepancy search (DDS) to \algorithmref{algorithm:sip}. Each point represents the solving
 time for one instance---to avoid noise for easier instances, we measure only time spent during
@@
 this is usually the variable-ordering heuristic
-\citep{DBLP:journals/jsat/LecoutreSTV07,DBLP:conf/cp/GayHLS15,DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17}.
+\cite{DBLP:journals/jsat/LecoutreSTV07,DBLP:conf/cp/GayHLS15,DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17}.
 In this paper, we instead rely upon randomness in the \emph{value}-ordering heuristic, and continue
 to use smallest domain first with static tiebreaking for variable-ordering.\footnote{It may be
-possible to further improve \cref{algorithm:sip} by also introducing randomness or some form of
+possible to further improve \algorithmref{algorithm:sip} by also introducing randomness or some form of
 learning into its variable-ordering heuristic. However, simultaneously introducing a second change
 would considerably complicate the empirical analysis.} Using restarts on value-ordering heuristics
-appears to be uncommon, although \citet{DBLP:conf/flairs/RazgonOP07} look at learning value-ordering
-heuristics from restarts, and \citet{DBLP:conf/cp/ChuSS09} use a similar scheme in the context of
+appears to be uncommon, although Razgon et al.\ \cite{DBLP:conf/flairs/RazgonOP07} look at learning value-ordering
+heuristics from restarts, and Chu et al.\ \cite{DBLP:conf/cp/ChuSS09} use a similar scheme in the context of
 parallel search.
 
-The shaded code shows how to introduce restarts to \cref{algorithm:sip}. The recursive
+The shaded code shows how to introduce restarts to \algorithmref{algorithm:sip}. The recursive
 $\FuncSty{search}$ procedure is modified to return either true (if the instance is satisfiable),
@@
 
-We use the Luby scheme \citep{DBLP:journals/ipl/LubySZ93} to determine when to restart. We count the
+We use the Luby scheme \cite{DBLP:journals/ipl/LubySZ93} to determine when to restart. We count the
 number of backtracks (that is, when we reach the end of the main $\KwSty{foreach}$ loop on
@@
 Following established wisdom, we multiply each item in the Luby sequence by a magic constant---we
-used SMAC \citep{DBLP:conf/lion/HutterHL11} to select the value 660. (We also tried a geometric
+used SMAC \cite{DBLP:conf/lion/HutterHL11} to select the value 660. (We also tried a geometric
 restart scheme, which gave less favourable results.)
@@
 
-Finally, we use the two watched literals technique \citep{DBLP:conf/dac/MoskewiczMZZM01} to
+Finally, we use the two watched literals technique \cite{DBLP:conf/dac/MoskewiczMZZM01} to
 propagate stored nogoods (\siplineref{2wl}). This has two benefits: the propagation complexity does
@@
 backtracking.  Other more sophisticated nogood generation and propagation schemes exist
-\citep{DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17}, but it is not clear whether these will be
+\cite{DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17}, but it is not clear whether these will be
 beneficial in a setting where we are not maintaining arc consistency.
@@
 
-In \cref{figure:restarts} we show the effects of adding restarts to the algorithm. With restarts,
+In \figureref{figure:restarts} we show the effects of adding restarts to the algorithm. With restarts,
 the random value-ordering heuristic comfortably beats the degree strategy without restarts. In other
@@
 the search is better, if it is done alongside a mechanism to avoid heavy commitment to any
-particular random choice. \Cref{figure:scatter-random} looks at this in more detail. The performance
+particular random choice. \Figureref{figure:scatter-random} looks at this in more detail. The performance
 on unsatisfiable instances, drawn as dark dots, remains unchanged as expected (which shows that the
@@
 
-\Cref{figure:restarts} also shows that the biased heuristic together with restarts is better
-still---that is, if we are introducing restarts, then it is better to add a small amount of
-randomness to a tailored deterministic heuristic than it is to throw away the heuristic altogether.
-Indeed, the original algorithm can solve 1,983 satisfiable instances by 993.0 seconds, whilst the
-biased and random restarting algorithms require only 9.2 seconds and 33.8 seconds respectively to
-solve the same number.
-
-In the more detailed view in \cref{figure:scatter-by-family}, comparing the basic algorithm with the
-degree heuristic to the degree-biased algorithm with restarts, all of the unsatisfiable instances
-are very close to the $x-y$ diagonal, showing that their performance is nearly unchanged. On the
-other hand, there are large numbers of satisfiable instances well below the diagonal line,
-indicating large speedups.  Better yet, there are only a handful of satisfiable instances that are
-more than a factor of ten times worse. In other words, as well as improving performance, we have
-made up most of the consistency we lost by introducing randomness.
-
-\section{Maximum Common Subgraph Algorithms}
-
-Having looked at subgraph isomorphism in detail, we now briefly discuss the more general maximum
-common induced subgraph problem.  Two recent algorithms for this problem also make use of
-backtracking search with degree as a value-ordering heuristic. The k${\downarrow}$ algorithm
-\citep{DBLP:conf/aaai/HoffmannMR17} attempts to solve the problem by first trying to solve the
-induced subgraph isomorphism problem, and then if that fails, retries allowing a single unmatched
-vertex (and thus using weaker invariants), and so on. Due to its similarity to \cref{algorithm:sip},
-we can introduce the same bias and restart strategy, requiring only small changes
-to handle wildcards in nogoods.
-
-Meanwhile, the McSplit${\downarrow}$ algorithm
-\citep{DBLP:conf/ijcai/McCreeshPT17} uses a constraint programming style search, but with special
-propagators and backtrackable data structures that exploit special properties of the problem. The
-unconventional domain store used by McSplit${\downarrow}$ precludes the use of arbitrary unit
-propagation, and so when introducing restarts, we cannot propagate using nogoods.  Instead, we can
-only detect when we are inside an already-visited branch.  We must therefore use the one watched
-literal scheme instead, and we also introduce a basic subsumption scheme to prune redundant clauses.
-
-Performance results from these two modified algorithms, using the same families of instances as in
-the previous section, are shown in \cref{figure:kdown,figure:mcsplit,figure:mcs}. Although we have
-moved from a decision problem to an optimisation problem, the same changes remain clearly
-beneficial. For the $k{\downarrow}$ algorithm, the change has a minimal effect on many instances
-(typically, where the $k = 0$ subproblem is unsatisfiable and hard, and the $k = 1$ subproblem is
-satisfiable and easy), but gives large benefits on many more instances than it penalises: it is over
-an order of magnitude better on nearly a hundred instances, whilst being an order of magnitude worse
-on only seven.
-
-With McSplit${\downarrow}$, the inability to use two watched literals means that in many cases we
-introduce a small slowdown. However, the overall pattern is the same: when introducing restarts and
-a biased value ordering heuristic, it is much more common to see a large speedup than it is a large
-slowdown.
-
 \begin{figure}[p]
@@
 
+\Figureref{figure:restarts} also shows that the biased heuristic together with restarts is better
+still---that is, if we are introducing restarts, then it is better to add a small amount of
+randomness to a tailored deterministic heuristic than it is to throw away the heuristic altogether.
+Indeed, the original algorithm can solve 1,983 satisfiable instances by 993.0 seconds, whilst the
+biased and random restarting algorithms require only 9.2 seconds and 33.8 seconds respectively to
+solve the same number.
+
+In the more detailed view in \figureref{figure:scatter-by-family}, comparing the basic algorithm with the
+degree heuristic to the degree-biased algorithm with restarts, all of the unsatisfiable instances
+are very close to the $x-y$ diagonal, showing that their performance is nearly unchanged. On the
+other hand, there are large numbers of satisfiable instances well below the diagonal line,
+indicating large speedups.  Better yet, there are only a handful of satisfiable instances that are
+more than a factor of ten times worse. In other words, as well as improving performance, we have
+made up most of the consistency we lost by introducing randomness.
+
+\section{Maximum Common Subgraph Algorithms}
+
+Having looked at subgraph isomorphism in detail, we now briefly discuss the more general maximum
+common induced subgraph problem.  Two recent algorithms for this problem also make use of
+backtracking search with degree as a value-ordering heuristic. The k${\downarrow}$ algorithm
+\cite{DBLP:conf/aaai/HoffmannMR17} attempts to solve the problem by first trying to solve the
+induced subgraph isomorphism problem, and then if that fails, retries allowing a single unmatched
+vertex (and thus using weaker invariants), and so on. Due to its similarity to \algorithmref{algorithm:sip},
+we can introduce the same bias and restart strategy, requiring only small changes
+to handle wildcards in nogoods.
+
+Meanwhile, the McSplit${\downarrow}$ algorithm
+\cite{DBLP:conf/ijcai/McCreeshPT17} uses a constraint programming style search, but with special
+propagators and backtrackable data structures that exploit special properties of the problem. The
+unconventional domain store used by McSplit${\downarrow}$ precludes the use of arbitrary unit
+propagation, and so when introducing restarts, we cannot propagate using nogoods.  Instead, we can
+only detect when we are inside an already-visited branch.  We must therefore use the one watched
+literal scheme instead, and we also introduce a basic subsumption scheme to prune redundant clauses.
+
+Performance results from these two modified algorithms, using the same families of instances as in
+the previous section, are shown in \threefigureref{figure:kdown}{figure:mcsplit}{figure:mcs}. Although we have
+moved from a decision problem to an optimisation problem, the same changes remain clearly
+beneficial. For the $k{\downarrow}$ algorithm, the change has a minimal effect on many instances
+(typically, where the $k = 0$ subproblem is unsatisfiable and hard, and the $k = 1$ subproblem is
+satisfiable and easy), but gives large benefits on many more instances than it penalises: it is over
+an order of magnitude better on nearly a hundred instances, whilst being an order of magnitude worse
+on only seven.
+
+With McSplit${\downarrow}$, the inability to use two watched literals means that in many cases we
+introduce a small slowdown. However, the overall pattern is the same: when introducing restarts and
+a biased value ordering heuristic, it is much more common to see a large speedup than it is a large
+slowdown.
+
 \section{Conclusion and Future Work}
@@
 
-\section*{Acknowledgments}
-
-We would like to thank some anonymous people for their comments.
-
-% Ruth Hoffmann, Christophe Lecoutre, Christine Solnon, Craig Reilly.
-
-\bibliographystyle{named}
 \bibliography{dblp}

### 2018-04-16 paper/paper.tex
More abstract

@@
     Modern subgraph isomorphism solvers use degree-based value-ordering heuristics to direct
-    backtracking search. We show that discrepancy-based searches are too expensive to be effective
-    in offsetting heuristic mistakes in this situation.  We then demonstrate that a random heuristic
-    with restarts beats conventional search with the best-known tailored heuristic. This motivates
-    the introduction of a slightly-random heuristic with similar behaviour to the tailored
-    heuristic: when combined with restarts, this new search strategy is over a hundred times more
-    effective on satisfiable instances, and does not affect performance on unsatisfiable instances.
-    Finally, we show that this strategy also improves two maximum common induced subgraph
-    algorithms.
+    backtracking search. This kind of search makes a heavy commitment to the first branching choice,
+    which is often incorrect. We evaluate discrepancy-based searches as an alternative, but find
+    them too expensive to be effective in this situation, even on satisfiable instances. We then
+    demonstrate that a random value-ordering heuristic with restarts beats conventional search with
+    the best-known tailored heuristic. This motivates the introduction of a slightly-random
+    heuristic with similar behaviour to the tailored heuristic: when combined with restarts and
+    nogood recording, this new search strategy is over a hundred times more effective on satisfiable
+    instances, and does not affect performance on unsatisfiable instances.  Finally, we show that
+    this strategy also improves two maximum common induced subgraph algorithms.
 \end{abstract}

### 2018-04-16 paper/paper.tex
More plots

@@
 
-\Figureref{figure:scatter-dds} shows the effects of adding Walsh's {DBLP:conf/ijcai/Walsh97}
+\Figureref{figure:scatter-dds} shows the effects of adding Walsh's \cite{DBLP:conf/ijcai/Walsh97}
 depth-bounded discrepancy search (DDS) to \algorithmref{algorithm:sip}. Each point represents the solving
@@
 
+\begin{figure}[tb]
+    \centering
+    \includegraphics*{gen-graph-restarts.pdf}
+
+    \caption{The cumulative number of satisfiable instances solved over time, with and without
+    restarts and nogood recording.}
+    \label{figure:restarts}
+\end{figure}
+
+\begin{figure}[p]
+    \centering
+    \includegraphics*{gen-graph-scatter-random-goods.pdf}
+    \caption{Comparing basic backtracking with the degree heuristic, versus the random heuristic
+    with restarts, but without nogood recording.}
+    \label{figure:scatter-random-goods}
+\end{figure}
+
+\begin{figure}[p]
+    \centering
+    \includegraphics*{gen-graph-scatter-random.pdf}
+    \caption{Comparing basic backtracking with the degree heuristic, versus the random heuristic
+    with restarts and nogood recording.}
+    \label{figure:scatter-random}
+\end{figure}
+
+\begin{figure}[p]
+    \centering
+    \includegraphics*{gen-graph-scatter-by-family.pdf}
+
+    \caption{Comparing basic backtracking with the degree heuristic, versus the biased
+    heuristic with restarts and nogood recording.}
+    \label{figure:scatter-by-family}
+\end{figure}
+
 In \figureref{figure:restarts} we show the effects of adding restarts to the algorithm. With restarts,
@@
     \centering
-    \includegraphics*{gen-graph-restarts.pdf}
-
-    \caption{The cumulative number of satisfiable instances solved over time, with and without
-    restarts.}
-    \label{figure:restarts}
-\end{figure}
-
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-scatter-random.pdf}
-    \caption{Comparing basic backtracking with the degree heuristic, versus the random heuristic
-    with restarts.}
-    \label{figure:scatter-random}
-\end{figure}
-
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-scatter-by-family.pdf}
-
-    \caption{Comparing basic backtracking with the degree heuristic, versus the biased
-    heuristic with restarts.}
-    \label{figure:scatter-by-family}
-\end{figure}
-
-\begin{figure}[p]
-    \centering
     \includegraphics*{gen-graph-mcs.pdf}
     \caption{The cumulative number of maximum common induced subgraph instances solved by
-    k${\downarrow}$ and McSplit${\downarrow}$ over time, with and without restarts.}\label{figure:mcs}
+    k${\downarrow}$ and McSplit${\downarrow}$ over time, with and without restarts and nogood
+    recording.}\label{figure:mcs}
 \end{figure}
@@
     \includegraphics*{gen-graph-scatter-kdown.pdf}
-    \caption{Comparing k${\downarrow}$ with and without restarts on an instance by instance basis.}\label{figure:kdown}
+    \caption{Comparing k${\downarrow}$ with and without restarts and nogood recording on an instance
+    by instance basis.}\label{figure:kdown}
 \end{figure}
@@
     \includegraphics*{gen-graph-scatter-mcsplit.pdf}
-    \caption{Comparing McSplit${\downarrow}$ with and without restarts on an instance by instance
-    basis.}\label{figure:mcsplit}
+    \caption{Comparing McSplit${\downarrow}$ with and without restarts and nogood recording on an
+    instance by instance basis.}\label{figure:mcsplit}
 \end{figure}

### 2018-04-17 paper/paper.tex
A bit better

@@
 \newcommand{\Figureref}[1]{Figure~\ref{#1}}
-\newcommand{\threefigureref}[3]{Figures~\ref{#1}, \ref{#2} and \ref{#3}}
+\newcommand{\twofigureref}[2]{Figures~\ref{#1} and~\ref{#2}}
 \newcommand{\algorithmref}[1]{Algorithm~\ref{#1}}
@@
 \EventLongTitle{26th Annual European Symposium on Algorithms (ESA 2018)}
-\EventShortTitle{ESA 2018}
+\EventShortTitle{Preprint}
 \EventAcronym{ESA}
@@
 vertices. Each domain is initialised at the top of search (\siplinerangeref{initstart}{initend}),
-using the invariants described by Hoffmann et al.\ \cite{DBLP:conf/aaai/HoffmannMR17} to restrict its initial
-values.
+using the invariants described by Zampelli et al.\ \cite{DBLP:journals/constraints/ZampelliDS10} and
+Hoffmann et al.\ \cite{DBLP:conf/aaai/HoffmannMR17} to restrict its initial values.
 
@@
         \coordinate (restart6ec) at ($(pic cs:restart6e) + (0, 0.00)$);
-        \node [fill=highlight, rounded corners, fit=(restart6sc) (restart6ec)] { };
+        \coordinate (restart6rc) at ($(pic cs:restart6r) + (0.8, 0.00)$);
+        \node [fill=highlight, rounded corners, fit=(restart6sc) (restart6ec) (restart6rc)] { };
         \coordinate (restart7sc) at ($(pic cs:restart7s) + (0, 0.15)$);
@@
             \nl \tikzmark{restart5s}\ElseIf{$S = \KwSty{restart}$}{
-                \nl post nogoods $\{B \wedge (v \mapsto c) \Rightarrow \bot : c \in C \}$\tikzmark{restart5r}\;
+                \nl \lForEach{$c \in C$}{$\textnormal{post nogood}~B \wedge (v \mapsto c) \Rightarrow \bot$\tikzmark{restart5r}}
                 \nl $\KwSty{return}~\KwSty{restart}$\tikzmark{restart5e} \;
@@
     }
-    \nl \tikzmark{restart6s}\lIf{$(R \gets R + 1) = R'$}{$\KwSty{return}~\KwSty{restart}\tikzmark{restart6e}$\siplabel{countrestarts}}
+    \nl \tikzmark{restart6s}\If{$(R \gets R + 1) = R'$\tikzmark{restart6r}}{
+        \nl post nogood $B$ \;
+        \nl $\KwSty{return}~\KwSty{restart}\tikzmark{restart6e}$\siplabel{countrestarts}}
     \nl $\KwSty{return}~\KwSty{false}$\siplabel{backtrack}
@@
 \nl \Begin{
-    \nl \While{$D_v \gets$~\textnormal{a unit domain from}~$D$}{
-        \nl $v' \gets $ the single value in $D_v$ \;
+    \nl \While{$(D_v, v') \gets ($\textnormal{a unit domain from}~$D, \textnormal{the single value
+    in that domain})$}{
         \nl \tikzmark{restart7s}trigger nogoods with watch $(v \mapsto v')$\tikzmark{restart7e}\siplabel{2wl} \;
@@
 }
-\caption{Non-induced subgraph isomorphism. The shaded code is added for the restarting variant.}
+\caption{Subgraph isomorphism. The shaded code is added for restarting.}
 \label{algorithm:sip}
@@
 We use the dataset introduced by Kotthoff et al.\ \cite{DBLP:conf/lion/KotthoffMS16} for evaluation. This dataset
-brings together a range of randomly-generated and application instance families from earlier papers.
+brings together a range of randomly-generated and application instance families from earlier papers:
+
+\begin{description}
+    \item[BVG(r), M4D(r), and Rand] are randomly generated graphs using different models (bounded
+        degree, regular mesh, and uniform), where each pattern is a permuted random connected subgraph
+        of the target (and so each instance is satisfiable) \cite{DBLP:journals/pami/CordellaFSV04}.
+        These benchmark instances are widely used, but have unusual properties and so broad
+        conclusions should not be drawn based solely upon behaviour on these instances
+        \cite{DBLP:journals/jair/McCreeshPST18}.
+    \item[SF] contains randomly generated scale-free graphs using a similar method
+        \cite{DBLP:journals/constraints/ZampelliDS10}.
+   \item[LV] consists of various kinds of graph gathered by Larrosa and Valiente
+       \cite{DBLP:journals/mscs/LarrosaV02} from the Stanford Graph Database. We include both the
+        50 small graphs, and the 50 larger graphs.
+    \item[Phase] contains hard crafted instances that lie near the satisfiable / unsatisfiable phase
+        transition \cite{DBLP:journals/jair/McCreeshPST18}.
+    \item[PR] contains graphs generated from segmented images, corresponding to a computer vision
+        problem \cite{DBLP:journals/pr/SolnonDHJ15}.
+    \item[Images and Meshes] contain graphs representing 2D segmented images and 3D object models
+        respectively, again representing a computer vision problem
+        \cite{DBLP:journals/cviu/DamiandSHJS11}.
+\end{description}
+
 Other studies use a random selection of 200 of each of the instances from the ``meshes'' and
@@
 would considerably complicate the empirical analysis.} Using restarts on value-ordering heuristics
-appears to be uncommon, although Razgon et al.\ \cite{DBLP:conf/flairs/RazgonOP07} look at learning value-ordering
-heuristics from restarts, and Chu et al.\ \cite{DBLP:conf/cp/ChuSS09} use a similar scheme in the context of
-parallel search.
+appears to be uncommon---although Razgon et al.\ \cite{DBLP:conf/flairs/RazgonOP07} look at learning
+value-ordering heuristics from restarts, Chu et al.\ \cite{DBLP:conf/cp/ChuSS09} use a similar
+scheme in the context of parallel search, and an early approach by Gomes et
+al.\ \cite{DBLP:conf/aaai/GomesSK98} does so in an optimisation context.
 
@@
 
-\subsection{The Effects of Restarts}
+\subsection{The Effects of Restarts and Nogood Recording}
 
-\begin{figure}[tb]
+\begin{figure}[p]
     \centering
@@
 \begin{figure}[p]
-    \centering
     \includegraphics*{gen-graph-scatter-random-goods.pdf}
-    \caption{Comparing basic backtracking with the degree heuristic, versus the random heuristic
-    with restarts, but without nogood recording.}
-    \label{figure:scatter-random-goods}
-\end{figure}
-
-\begin{figure}[p]
-    \centering
+    \hfill
     \includegraphics*{gen-graph-scatter-random.pdf}
     \caption{Comparing basic backtracking with the degree heuristic, versus the random heuristic
-    with restarts and nogood recording.}
+    with restarts. The plot on the right uses nogood recording, whilst the plot on the left does not.}
     \label{figure:scatter-random}
@@
 
-In \figureref{figure:restarts} we show the effects of adding restarts to the algorithm. With restarts,
-the random value-ordering heuristic comfortably beats the degree strategy without restarts. In other
+In \figureref{figure:restarts} we show the effects of adding restarts and nogood recording to the
+algorithm. With restarts and nogood recording,
+the random value-ordering heuristic comfortably beats the degree strategy with simple backtracking
+search. In other
 words, although having a good value-ordering heuristic is beneficial, introducing randomness into
 the search is better, if it is done alongside a mechanism to avoid heavy commitment to any
-particular random choice. \Figureref{figure:scatter-random} looks at this in more detail. The performance
-on unsatisfiable instances, drawn as dark dots, remains unchanged as expected (which shows that the
-overheads of nogoods and restarting are negligible). However, although better in aggregate on
-satisfiable instances, on a case by case basis many instances perform substantially worse with
-randomisation and restarts.
-
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-mcs.pdf}
-    \caption{The cumulative number of maximum common induced subgraph instances solved by
-    k${\downarrow}$ and McSplit${\downarrow}$ over time, with and without restarts and nogood
-    recording.}\label{figure:mcs}
-\end{figure}
-
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-scatter-kdown.pdf}
-    \caption{Comparing k${\downarrow}$ with and without restarts and nogood recording on an instance
-    by instance basis.}\label{figure:kdown}
-\end{figure}
-
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-scatter-mcsplit.pdf}
-    \caption{Comparing McSplit${\downarrow}$ with and without restarts and nogood recording on an
-    instance by instance basis.}\label{figure:mcsplit}
-\end{figure}
+particular random choice. \Figureref{figure:scatter-random} looks at
+this in more detail. On the left-hand plot, without nogood recording, unsatisfiable instances
+experience a slowdown of more than an order of magnitude, whilst on the right-hand plot the
+performance remains unchanged (which shows that the overheads of nogoods and restarting are
+negligible). We therefore use nogood recording in all subsequent experiments with restarts. However,
+although better in aggregate on satisfiable instances, on a case by case basis many instances
+perform substantially worse with randomisation (either with or without restarts).
 
@@
 \section{Maximum Common Subgraph Algorithms}
+
+\begin{figure}[tb]
+    \centering
+    \includegraphics*{gen-graph-mcs.pdf}
+    \caption{The cumulative number of maximum common induced subgraph instances solved by
+    k${\downarrow}$ and McSplit${\downarrow}$ over time, with and without restarts and nogood
+    recording.}\label{figure:mcs}
+\end{figure}
+
+\begin{figure}[tb]
+    \includegraphics*{gen-graph-scatter-kdown.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-mcsplit.pdf}
+    \caption{On the left, comparing k${\downarrow}$ with and without restarts and nogood recording on an instance
+    by instance basis. On the right, the same, for McSplit${\downarrow}$.}\label{figure:mcsplitandkdown}
+\end{figure}
 
@@
 Performance results from these two modified algorithms, using the same families of instances as in
-the previous section, are shown in \threefigureref{figure:kdown}{figure:mcsplit}{figure:mcs}. Although we have
+the previous section, are shown in \twofigureref{figure:mcs}{figure:mcsplitandkdown}. Although we have
 moved from a decision problem to an optimisation problem, the same changes remain clearly

### 2018-04-20 paper/paper.tex
Tweaks

@@
 
-The subgraph isomorphism problem is to find a copy of a small ``pattern'' graph inside a larger
-``target'' graph.  The current single strongest subgraph isomorphism solver uses ``highest degree
-first'' as a value-ordering heuristic to direct a constraint programming style search
+The subgraph isomorphism problem is to decide whether a copy of a small ``pattern'' graph inside a
+larger ``target'' graph. Although \NP-complete, by combining design techniques from artificial
+intelligence with careful algorithm engineering, modern subgraph isomorphism solvers can often
+produce exact solutions quickly even on graphs with thousands of vertices. The current single
+strongest subgraph isomorphism solver uses ``highest degree first'' as a value-ordering heuristic to
+direct a constraint programming style search
 \cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:conf/ijcai/McCreeshPT16}. This
@@
 fruitful to investigate alternatives to simple backtracking (depth-first) search. This paper
-evaluates two such approaches: discrepancy search (which performs poorly), and restarts (which are
-extremely beneficial). To explore different parts of the search space when restarting, we introduce
-a new value-ordering heuristic, which adds a small amount of randomness to the ``highest degree
-first'' heuristic.  The effects of these changes are shown in
-\figureref{figure:scatter-by-family}: we can make a state-of-the-art algorithm perform much better on a
-large number of satisfiable instances, whilst performing worse only rarely on satisfiable instances,
-and never on unsatisfiable instances.  Further, these properties are \emph{not} shared by
-discrepancy searches (\figureref{figure:scatter-dds}) or by random value-ordering with restarts
-(\figureref{figure:scatter-random}). Finally, we show that this strategy is also effective in an
-optimisation setting, producing benefits in two maximum common induced subgraph algorithms.
+evaluates two such approaches: discrepancy search (which performs poorly), and a new form of search
+(which is extremely beneficial) based around restarts and nogood recording. To explore different
+parts of the search space when restarting, we introduce a novel value-ordering heuristic, which adds
+a small amount of randomness to the ``highest degree first'' heuristic.  The effects of these
+changes are shown in \figureref{figure:scatter-by-family}: we can make a state-of-the-art algorithm
+perform much better on a large number of satisfiable instances, whilst performing worse only rarely
+on satisfiable instances, and never on unsatisfiable instances.  Further, these properties are
+\emph{not} shared by discrepancy searches (\figureref{figure:scatter-dds}) or by random
+value-ordering with restarts (\figureref{figure:scatter-random}). Finally, we show that this
+strategy is also effective in an optimisation setting, producing benefits in two maximum common
+induced subgraph algorithms.
 
@@
 
-For the variable-ordering heuristic on \siplineref{variableordering} of \algorithmref{algorithm:sip}, we use
-smallest domain first, tiebreaking on highest degree. Variable ordering is not the focus of this
-paper---as far as possible, we use this same dynamic ordering throughout the paper, so as not to
-affect the performance of the algorithm on unsatisfiable instances.  For value ordering, which
-occurs on \siplineref{valueordering} of \algorithmref{algorithm:sip}, we will initially select vertices from
-highest degree to lowest degree; we return to value ordering in the second half of this paper. Note
-that the variable-ordering heuristic affects both unsatisfiable and satisfiable instances, whilst
-the value-ordering heuristic is relevant only for satisfiable instances.
+For the variable-ordering heuristic on \siplineref{variableordering} of
+\algorithmref{algorithm:sip}, we use smallest domain first, tiebreaking on highest degree
+\cite{DBLP:journals/ai/HaralickE80,DBLP:journals/jair/McCreeshPST18}. Variable ordering is not the
+focus of this paper---as far as possible, we use this same dynamic ordering throughout the paper, so
+as not to affect the performance of the algorithm on unsatisfiable instances.  For value ordering,
+which occurs on \siplineref{valueordering} of \algorithmref{algorithm:sip}, we will initially select
+vertices from highest degree to lowest degree; we return to value ordering in the second half of
+this paper. Note that the variable-ordering heuristic affects both unsatisfiable and satisfiable
+instances, whilst the value-ordering heuristic is relevant only for satisfiable instances.
 
@@
 running Ubuntu 17.04. We
-implemented\footnote{\url{https://github.com/ciaranm/ijcai2018-restarting-sip-paper}, DOI to be issued pre-publication}
+implemented\footnote{\url{https://github.com/ciaranm/restarting-sip-paper}, DOI to be issued pre-publication}
 \algorithmref{algorithm:sip} in C++ using the bit-parallel implementation of
@@
 be solved by one solver than another, if every instance is run separately with the chosen $x$
-timeout; the horizontal distance shows how many times longer the per-instance timeout would need to
+timeout. The horizontal distance shows how many times longer the per-instance timeout would need to
 be to allow the rightmost algorithm to succeed on $y$ out of the 14,621 instances (bearing in mind
@@
     \caption{The cumulative number of satisfiable instances solved over time, using four
-    different value-ordering heuristics. The degree and biased lines are close together.}
+    different value-ordering heuristics.}
     \label{figure:value-ordering-heuristics}
@@
 
-\subsection{Adding Restarts to the Algorithm}
+\subsection{Adding Restarts and Nogood Recording to the Algorithm}
 
@@
 Following established wisdom, we multiply each item in the Luby sequence by a magic constant---we
-used SMAC \cite{DBLP:conf/lion/HutterHL11} to select the value 660. (We also tried a geometric
-restart scheme, which gave less favourable results.)
+used the SMAC automatic parameter tuner \cite{DBLP:conf/lion/HutterHL11} to select the value 660.
+(We also tried a geometric restart scheme, which gave less favourable results.)
 
@@
 space that we have already visitied, every time we restart, we add new constraints to the problem
-which eliminate already-explored subtrees. We generate simple decision nogoods, as follows. The $B$
-argument to $\FuncSty{search}$ tracks the branching choices made so far, in the form of a CNF clause
-with each literal being a guessed $(v \mapsto v')$ assignment. Upon backtracking due to a decision to
-restart, we post a nogood of the form $(v \mapsto v') \wedge (w \mapsto w') \wedge (x \mapsto x')
-\Rightarrow \bot$ for every branch to the left of the current (incomplete) branch in the search tree.
+which eliminate already-explored subtrees---such a constraint is called a nogood. We generate simple
+decision nogoods, as follows. The $B$ argument to $\FuncSty{search}$ tracks the branching choices
+made so far, in the form of a CNF clause with each literal being a guessed $(v \mapsto v')$
+assignment. Upon backtracking due to a decision to restart, we post a nogood of the form $(v \mapsto
+v') \wedge (w \mapsto w') \wedge (x \mapsto x') \Rightarrow \bot$ for every branch to the left of
+the current (incomplete) branch in the search tree, and when we first make a decision to restart
+before backtracking, we post a similar nogood eliminating the entire subtree explored.
 
@@
 introduce a small slowdown. However, the overall pattern is the same: when introducing restarts and
-a biased value ordering heuristic, it is much more common to see a large speedup than it is a large
+a biased value ordering heuristic, it is much more common to see a large speedup than a large
 slowdown.
@@
 backtracking search algorithms for hard subgraph problems, without compromising existing search
-order heuristics, and without paying a penalty for unsatisfiable instances.
+order heuristics, and without paying a penalty for unsatisfiable instances. In principle it is
+well-known that introducing some form of randomness \cite{DBLP:conf/aaai/GomesSK98} or reduced
+commitment \cite{DBLP:conf/ijcai/HarveyG95} into backtracking search can be beneficial, but until
+now it has not been clear how to do so efficiently in practice, particularly when we already have
+strong search ordering heuristics and when are expecting to encounter large numbers of unsatisfiable
+instances. Our results show that the key to success is a combination of aggressive restarts, nogood
+recording, and biasing an existing value-ordering heuristic with just a small amount of randomness.
 
-The success of this approach raises a number of interesting questions and avenues for possible
+The effectiveness of this approach raises a number of interesting questions and avenues for possible
 future research. For example, will this same technique work for other problems where we have a good

### 2018-07-25 paper/paper.tex
AAAAAAAAAAI

@@
 
-\documentclass[a4paper,UKenglish]{lipics-v2018}
+\def\year{2019}\relax
+%File: formatting-instruction.tex
+\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
+\usepackage{aaai18}  %Required
+\usepackage{times}  %Required
+\usepackage{helvet}  %Required
+\usepackage{courier}  %Required
+\usepackage{url}  %Required
+\usepackage{graphicx}  %Required
+\frenchspacing  %Required
+\setlength{\pdfpagewidth}{8.5in}  %Required
+\setlength{\pdfpageheight}{11in}  %Required
+%PDF Info Is Required:
+  \pdfinfo{
+/Title (Value Ordering, Discrepancies, and Restarts for Subgraph Algorithms)
+/Author (Anonymous and Anonymous and Anonymous)}
+\setcounter{secnumdepth}{0}
 
-\usepackage{microtype}
 \usepackage{complexity}
-\usepackage{subcaption}
 \usepackage{amsmath}
@@
 \usepackage{tikz}
+\usepackage{cleveref}
 
-% \usepackage{showframe}
-
-\bibliographystyle{plainurl}
+\usepackage{showframe}
 
@@
 \newcommand{\vertexset}{\operatorname{V}}
-\newcommand{\nds}{\operatorname{S}}
 
@@
 
-\newcommand{\figureref}[1]{Figure~\ref{#1}}
-\newcommand{\Figureref}[1]{Figure~\ref{#1}}
-\newcommand{\twofigureref}[2]{Figures~\ref{#1} and~\ref{#2}}
-\newcommand{\algorithmref}[1]{Algorithm~\ref{#1}}
-\newcommand{\Algorithmref}[1]{Algorithm~\ref{#1}}
+\newcommand{\citet}[1]{\citeauthor{#1} \shortcite{#1}}
 
@@
 \title{Value Ordering, Discrepancies, and Restarts for Subgraph Algorithms}
-\author{Ciaran McCreesh}{University of Glasgow}{ciaran.mccreesh@glasgow.ac.uk}{}{}
-\author{Patrick Prosser}{University of Glasgow}{patrick.prosser@glasgow.ac.uk}{}{}
-\author{James Trimble}{University of Glasgow}{j.trimble.1@research.gla.ac.uk}{}{}
-
-\authorrunning{C. McCreesh, P. Prosser and J. Trimble}
-
-\Copyright{Ciaran McCreesh, Patrick Prosser and James Trimble}
-
-\subjclass{
-\ccsdesc[500]{Mathematics of computing~Graph algorithms},
-\ccsdesc[500]{Theory of computation~Discrete optimization},
-\ccsdesc[500]{Theory of computation~Backtracking}}
-
-\keywords{Subgraph isomorphism, combinatorial search, rich search}
-
-\funding{This work was supported by the Engineering and Physical Sciences Research Council [grant
-numbers EP/026842/1 and EP/M508056/1]}
-
-\acknowledgements{The authors would like to thank Ruth Hoffmann, Christophe Lecoutre, Christine
-Solnon and Craig Reilly for their comments.}
-
-%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\EventEditors{John Q. Open and Joan R. Access}
-\EventNoEds{2}
-\EventLongTitle{26th Annual European Symposium on Algorithms (ESA 2018)}
-\EventShortTitle{Preprint}
-\EventAcronym{ESA}
-\EventYear{2018}
-\EventDate{August 20--22, 2018}
-\EventLocation{Helsinki, Finland}
-\EventLogo{}
-\SeriesVolume{42}
-\ArticleNo{23}
-%\nolinenumbers %uncomment to disable line numbering
-\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\author{Anonymous \and Anonymous \and Anonymous \\ email address \\ address}
 
@@
 a small amount of randomness to the ``highest degree first'' heuristic.  The effects of these
-changes are shown in \figureref{figure:scatter-by-family}: we can make a state-of-the-art algorithm
+changes are shown in the final plot of \cref{figure:scatter-random}: we can make a state-of-the-art algorithm
 perform much better on a large number of satisfiable instances, whilst performing worse only rarely
 on satisfiable instances, and never on unsatisfiable instances.  Further, these properties are
-\emph{not} shared by discrepancy searches (\figureref{figure:scatter-dds}) or by random
-value-ordering with restarts (\figureref{figure:scatter-random}). Finally, we show that this
-strategy is also effective in an optimisation setting, producing benefits in two maximum common
-induced subgraph algorithms.
+\emph{not} shared by discrepancy searches (\cref{figure:scatter-dds}) or by random
+value-ordering with restarts (the first two plots of \cref{figure:scatter-random}). Finally, we
+show that this strategy is also effective in an optimisation setting, producing benefits in two
+maximum common induced subgraph algorithms.
 
@@
 \emph{neighbourhood} of a vertex, $\neighbourhood_G(v)$, is the set of vertices adjacent to $v$, and
-the \emph{degree} of $v$, $\deg_G(v)$, is the cardinality of its neighbourhood. The
-\emph{neighbourhood degree sequence} of $v$, $\nds_G(v)$, is the sequence of degrees of the
-neighbours of $v$ in descending order, and we denote by $\preceq$ the inclusion relation of
-Zampelli et al.\ \cite{DBLP:journals/constraints/ZampelliDS10}. When the graph is clear from the context, we omit
-the $G$ subscripts. As per Hoffmann et al.\ \cite{DBLP:conf/aaai/HoffmannMR17}, $G^{n,\ell}$ denotes the graph with
+the \emph{degree} of $v$, $\deg_G(v)$, is the cardinality of its neighbourhood. When the graph is clear from the context, we omit
+the $G$ subscripts. As per \citet{DBLP:conf/aaai/HoffmannMR17}, $G^{n,\ell}$ denotes the graph with
 vertex set $\vertexset(G)$, and with edges between vertices $v$ and $w$ if there are at least $n$
-simple paths of length exactly $\ell$ between vertices $v$ and $w$ in $G$.
+simple paths of length exactly $\ell$ between vertices $v$ and $w$ in $G$. Finally, we define a
+function $\FuncSty{compatible}(v, w)$ to mean that mapping pattern vertex $v$ to target vertex $w$
+is feasible based upon both loop constraints and the neighbourhood degree sequence relation of
+\citet{DBLP:journals/constraints/ZampelliDS10}.
 
@@
 
-We now briefly outline the non-shaded parts of \algorithmref{algorithm:sip}, which is how we will solve the
+We now briefly outline the non-shaded parts of \cref{algorithm:sip}, which is how we will solve the
 subgraph isomorphism problem. This algorithm is very closely based upon the $k{\downarrow}$
-algorithm of Hoffmann et al.\ \cite{DBLP:conf/aaai/HoffmannMR17} with $k = 0$, and we refer the reader to that paper
+algorithm of \citet{DBLP:conf/aaai/HoffmannMR17} with $k = 0$, and we refer the reader to that paper
 for full technical details; that algorithm, in turn, is a simplification of the Glasgow2 algorithm
@@
 
-\Algorithmref{algorithm:sip} works by recursively building up a valid mapping of pattern vertices to target
+\Cref{algorithm:sip} works by recursively building up a valid mapping of pattern vertices to target
 vertices. For each pattern vertex $v$, we have a domain $D_v$ containing a set of feasible target
 vertices. Each domain is initialised at the top of search (\siplinerangeref{initstart}{initend}),
-using the invariants described by Zampelli et al.\ \cite{DBLP:journals/constraints/ZampelliDS10} and
-Hoffmann et al.\ \cite{DBLP:conf/aaai/HoffmannMR17} to restrict its initial values.
+using the invariants described by \citet{DBLP:journals/constraints/ZampelliDS10} and
+\citet{DBLP:conf/aaai/HoffmannMR17} to restrict its initial values.
 
@@
 
-Our $\FuncSty{propagate}$ routine is also close to that of Hoffmann et al.\ \cite{DBLP:conf/aaai/HoffmannMR17},
+Our $\FuncSty{propagate}$ routine is also close to that of \citet{DBLP:conf/aaai/HoffmannMR17},
 differing only in that we carry out the filtering for injectivity
@@
         (\mathcal{P}^{1,2}, \mathcal{T}^{1,2}),
-        (\mathcal{P}^{2,2},\ \mathcal{T}^{2,2}), (\mathcal{P}^{3,2},\ \mathcal{T}^{3,2}),
+        (\mathcal{P}^{2,2},\ \mathcal{T}^{2,2}),$ \\
+        $\hspace*{3em}(\mathcal{P}^{3,2},\ \mathcal{T}^{3,2}),
         (\mathcal{P}^{4,2},\ \mathcal{T}^{4,2}) \big]$ \;
@@
         \nl $D_v \gets \vertexset(\mathcal{T})$\;
-        \nl \lForEach{$(P,\,T) \in L$}{$D_v \gets \{ w \in D_v : $ \siplabel{degreefiltering} $v \sim_P v \Rightarrow w \sim_T w~\wedge$
-            $\nds_P(v) \preceq \nds_T(w)\}$\siplabel{initend}
+        \nl \ForEach{$(P,\,T) \in L$}{\nl$D_v \gets \{ w \in D_v : \FuncSty{compatible}(v, w)\}$\siplabel{initend}
         }
@@
         \nl \If{\FuncSty{propagate}(D')\siplabel{dopropagate}}{
-            \nl $S \gets \FuncSty{search}(\{ E \in D'\,{:}\,\left|E\right|{>}\,1 \}, B \wedge (v\,{\mapsto}\,v'))$\siplabel{recurse} \;
+            \nl $S \gets \FuncSty{search}(\{ E \in D'\,{:}\,\left|E\right|{>}\,1 \}, B \wedge (v\,{\mapsto}\,v'))$\siplabel{recurse}\;
             \nl \lIf{$S = \KwSty{true}$}{$\KwSty{return}~\KwSty{true}$}
@@
             \nl $D_w \gets D_w - v'$ \siplabel{removev} \;
-            \nl \lForEach{$(P,\,T) \in L~\KwSty{where}~v \sim_P w$}{$D_w \gets D_w \cap \neighbourhood_T(v')$}
+            \nl \ForEach{$(P,\,T) \in L~\KwSty{where}~v \sim_P w$}{\nl$D_w \gets D_w \cap \neighbourhood_T(v')$}
             \nl \lIf{$D_w = \emptyset$}{$\KwSty{return}~\KwSty{false}$}
@@
 \nl $(H,\,A,\,n) \gets (\emptyset,\,\emptyset,\,0)$\siplabel{alldiffstart} \;
-\nl \ForEach{$D_v \in D$ \textnormal{from smallest cardinality upwards\siplabel{eachdomain}}}{
+\nl \ForEach{$D_v \in D$ \textnormal{in ascending cardinality order\siplabel{eachdomain}}}{
     \nl $(D_v, A,\,n) \gets (D_v \gets D_v \setminus H\siplabel{elimhall}, A \cup D_v,\,n + 1)$ \siplabel{acc} \;
@@
 For the variable-ordering heuristic on \siplineref{variableordering} of
-\algorithmref{algorithm:sip}, we use smallest domain first, tiebreaking on highest degree
+\cref{algorithm:sip}, we use smallest domain first, tiebreaking on highest degree
 \cite{DBLP:journals/ai/HaralickE80,DBLP:journals/jair/McCreeshPST18}. Variable ordering is not the
@@
 as not to affect the performance of the algorithm on unsatisfiable instances.  For value ordering,
-which occurs on \siplineref{valueordering} of \algorithmref{algorithm:sip}, we will initially select
+which occurs on \siplineref{valueordering} of \cref{algorithm:sip}, we will initially select
 vertices from highest degree to lowest degree; we return to value ordering in the second half of
@@
 running Ubuntu 17.04. We
-implemented\footnote{\url{https://github.com/ciaranm/restarting-sip-paper}, DOI to be issued pre-publication}
-\algorithmref{algorithm:sip} in C++ using the bit-parallel implementation of
-Hoffmann et al.\ \cite{DBLP:conf/aaai/HoffmannMR17} as a starting point, and compiled it using GCC 6.3.0. We use a
+implemented\footnote{DOI removed for anonymous review}
+\cref{algorithm:sip} in C++ using the bit-parallel implementation of
+\citet{DBLP:conf/aaai/HoffmannMR17} as a starting point, and compiled it using GCC 6.3.0. We use a
 deterministic pseudo-random number generator for reproducibility.
 
-We use the dataset introduced by Kotthoff et al.\ \cite{DBLP:conf/lion/KotthoffMS16} for evaluation. This dataset
+We use the dataset introduced by \citet{DBLP:conf/lion/KotthoffMS16} for evaluation. This dataset
 brings together a range of randomly-generated and application instance families from earlier papers:
@@
 
-We begin by verifying that our implementation of the basic \algorithmref{algorithm:sip} is competitive. In
-\figureref{figure:others} we show the cumulative number of instances solved over time, using our
+We begin by verifying that our implementation of the basic \cref{algorithm:sip} is competitive. In
+\cref{figure:others} we show the cumulative number of instances solved over time, using our
 implementation (both with the degree heuristic, and with the modifications described in the
-remainder of this paper), the Glasgow2 and Glasgow3 algorithms from which \algorithmref{algorithm:sip} is derived
+remainder of this paper), the Glasgow2 and Glasgow3 algorithms from which \cref{algorithm:sip} is derived
 \cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}, the PathLAD variation of the LAD
@@
 
-As expected, the performance of \algorithmref{algorithm:sip} with the degree heuristic is very close to that
+As expected, the performance of \cref{algorithm:sip} with the degree heuristic is very close to that
 of Glasgow2, although we perform better at lower runtimes because we use simpler preprocessing at
@@
 upon improving the performance on the remaining hard satisfiable instances, and so in the second
-plot in \figureref{figure:others} (and in subsequent cumulative plots) we show only satisfiable
+plot in \cref{figure:others} (and in subsequent cumulative plots) we show only satisfiable
 instances, and use a reduced range on both axes.
@@
 
-\Figureref{figure:value-ordering-heuristics} shows the cumulative number of satisfiable instances
+\Cref{figure:value-ordering-heuristics} shows the cumulative number of satisfiable instances
 solved over time using different value-ordering heuristics:
@@
 value-ordering heuristic without adversely affecting its performance \emph{in aggregate}.
-\Figureref{figure:scatter-heuristics} gives a more detailed look at these results. This scatter plot
+\Cref{figure:scatter-heuristics} gives a more detailed look at these results. This scatter plot
 shows that despite the aggregate performance being very similar, on a case by case basis, the two
@@
 
-\Figureref{figure:scatter-dds} shows the effects of adding Walsh's \cite{DBLP:conf/ijcai/Walsh97}
-depth-bounded discrepancy search (DDS) to \algorithmref{algorithm:sip}. Each point represents the solving
+\Cref{figure:scatter-dds} shows the effects of adding Walsh's \cite{DBLP:conf/ijcai/Walsh97}
+depth-bounded discrepancy search (DDS) to \cref{algorithm:sip}. Each point represents the solving
 time for one instance---to avoid noise for easier instances, we measure only time spent during
@@
 
-\subsection{Adding Restarts and Nogood Recording to the Algorithm}
+\subsection{Adding Restarts and Nogood Recording}
 
@@
 to use smallest domain first with static tiebreaking for variable-ordering.\footnote{It may be
-possible to further improve \algorithmref{algorithm:sip} by also introducing randomness or some form of
+possible to further improve \cref{algorithm:sip} by also introducing randomness or some form of
 learning into its variable-ordering heuristic. However, simultaneously introducing a second change
 would considerably complicate the empirical analysis.} Using restarts on value-ordering heuristics
-appears to be uncommon---although Razgon et al.\ \cite{DBLP:conf/flairs/RazgonOP07} look at learning
-value-ordering heuristics from restarts, Chu et al.\ \cite{DBLP:conf/cp/ChuSS09} use a similar
-scheme in the context of parallel search, and an early approach by Gomes et
-al.\ \cite{DBLP:conf/aaai/GomesSK98} does so in an optimisation context.
+appears to be uncommon---although \citet{DBLP:conf/flairs/RazgonOP07} look at learning
+value-ordering heuristics from restarts, \citet{DBLP:conf/cp/ChuSS09} use a similar
+scheme in the context of parallel search, and an early approach by \citet{DBLP:conf/aaai/GomesSK98} does so in an optimisation context.
 
-The shaded code shows how to introduce restarts to \algorithmref{algorithm:sip}. The recursive
+The shaded code shows how to introduce restarts to \cref{algorithm:sip}. The recursive
 $\FuncSty{search}$ procedure is modified to return either true (if the instance is satisfiable),
@@
 
-\begin{figure}[p]
+\begin{figure*}[tb]
+    \includegraphics*{gen-graph-scatter-random-goods.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-random.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-by-family.pdf}
+    \caption{Comparing basic backtracking with the degree heuristic, versus the random and biased
+    heuristics with restarts. The first plot does not use nogood recording, whilst the second and
+    third do.}
+    \label{figure:scatter-random}
+\end{figure*}
+
+\begin{figure}[tb]
     \centering
@@
 
-\begin{figure}[p]
-    \includegraphics*{gen-graph-scatter-random-goods.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-random.pdf}
-    \caption{Comparing basic backtracking with the degree heuristic, versus the random heuristic
-    with restarts. The plot on the right uses nogood recording, whilst the plot on the left does not.}
-    \label{figure:scatter-random}
-\end{figure}
-
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-scatter-by-family.pdf}
-
-    \caption{Comparing basic backtracking with the degree heuristic, versus the biased
-    heuristic with restarts and nogood recording.}
-    \label{figure:scatter-by-family}
-\end{figure}
-
-In \figureref{figure:restarts} we show the effects of adding restarts and nogood recording to the
+In \cref{figure:restarts} we show the effects of adding restarts and nogood recording to the
 algorithm. With restarts and nogood recording,
@@
 the search is better, if it is done alongside a mechanism to avoid heavy commitment to any
-particular random choice. \Figureref{figure:scatter-random} looks at
-this in more detail. On the left-hand plot, without nogood recording, unsatisfiable instances
-experience a slowdown of more than an order of magnitude, whilst on the right-hand plot the
+particular random choice. \Cref{figure:scatter-random} looks at
+this in more detail. On the first plot, without nogood recording, unsatisfiable instances
+experience a slowdown of more than an order of magnitude, whilst on the second plot the
 performance remains unchanged (which shows that the overheads of nogoods and restarting are
@@
 
-\Figureref{figure:restarts} also shows that the biased heuristic together with restarts is better
+\Cref{figure:restarts} also shows that the biased heuristic together with restarts is better
 still---that is, if we are introducing restarts, then it is better to add a small amount of
@@
 
-In the more detailed view in \figureref{figure:scatter-by-family}, comparing the basic algorithm with the
-degree heuristic to the degree-biased algorithm with restarts, all of the unsatisfiable instances
-are very close to the $x-y$ diagonal, showing that their performance is nearly unchanged. On the
-other hand, there are large numbers of satisfiable instances well below the diagonal line,
-indicating large speedups.  Better yet, there are only a handful of satisfiable instances that are
-more than a factor of ten times worse. In other words, as well as improving performance, we have
-made up most of the consistency we lost by introducing randomness.
+In the more detailed view in the final plot of \cref{figure:scatter-random}, comparing the
+basic algorithm with the degree heuristic to the degree-biased algorithm with restarts, all of the
+unsatisfiable instances are very close to the $x-y$ diagonal, showing that their performance is
+nearly unchanged. On the other hand, there are large numbers of satisfiable instances well below the
+diagonal line, indicating large speedups.  Better yet, there are only a handful of satisfiable
+instances that are more than a factor of ten times worse. In other words, as well as improving
+performance, we have made up most of the consistency we lost by introducing randomness.
 
 \section{Maximum Common Subgraph Algorithms}
-
-\begin{figure}[tb]
-    \centering
-    \includegraphics*{gen-graph-mcs.pdf}
-    \caption{The cumulative number of maximum common induced subgraph instances solved by
-    k${\downarrow}$ and McSplit${\downarrow}$ over time, with and without restarts and nogood
-    recording.}\label{figure:mcs}
-\end{figure}
-
-\begin{figure}[tb]
-    \includegraphics*{gen-graph-scatter-kdown.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-mcsplit.pdf}
-    \caption{On the left, comparing k${\downarrow}$ with and without restarts and nogood recording on an instance
-    by instance basis. On the right, the same, for McSplit${\downarrow}$.}\label{figure:mcsplitandkdown}
-\end{figure}
 
@@
 induced subgraph isomorphism problem, and then if that fails, retries allowing a single unmatched
-vertex (and thus using weaker invariants), and so on. Due to its similarity to \algorithmref{algorithm:sip},
+vertex (and thus using weaker invariants), and so on. Due to its similarity to \cref{algorithm:sip},
 we can introduce the same bias and restart strategy, requiring only small changes
@@
 Performance results from these two modified algorithms, using the same families of instances as in
-the previous section, are shown in \twofigureref{figure:mcs}{figure:mcsplitandkdown}. Although we have
+the previous section, are shown in \cref{figure:mcs,figure:kdown,figure:mcsplit}. Although we have
 moved from a decision problem to an optimisation problem, the same changes remain clearly
@@
 on only seven.
+
+\begin{figure}[p]
+    \centering
+    \includegraphics*{gen-graph-mcs.pdf}
+    \caption{The cumulative number of maximum common induced subgraph instances solved by
+    k${\downarrow}$ and McSplit${\downarrow}$ over time, with and without restarts and nogood
+    recording.}\label{figure:mcs}
+\end{figure}
+
+\begin{figure}[p]
+    \centering
+    \includegraphics*{gen-graph-scatter-kdown.pdf}
+    \caption{Comparing k${\downarrow}$ with and without restarts and nogood recording on an instance
+    by instance basis.}\label{figure:kdown}
+\end{figure}
+
+\begin{figure}[p]
+    \centering
+    \includegraphics*{gen-graph-scatter-mcsplit.pdf}
+    \caption{Comparing McSplit${\downarrow}$ with and without restarts and nogood recording on an instance
+    by instance basis.}\label{figure:mcsplit}
+\end{figure}
 
@@
 
+\section*{Acknowledgements}
+
+% The authors would like to thank Ruth Hoffmann, Christophe Lecoutre, Christine Solnon and Craig
+% Reilly for their comments.  This work was supported by the Engineering and Physical Sciences
+% Research Council [grant numbers EP/026842/1 and EP/M508056/1].
+
+The authors would like to thank Anonymous, Anonymous, Anonymous and Anonymous for their comments.
+This work was supported by an anonymous funding agency.
+
+\bibliographystyle{aaai}
 \bibliography{dblp}

### 2018-07-25 paper/paper.tex
Betterer

@@
 
-\usepackage{showframe}
+% \usepackage{showframe}
 
@@
 
+% cref style
+\crefname{figure}{Figure}{Figures}
+\Crefname{figure}{Figure}{Figures}
+\crefname{section}{Section}{Sections}
+\Crefname{section}{Section}{Sections}
+\crefname{proposition}{Proposition}{Propositions}
+\Crefname{proposition}{Proposition}{Propositions}
+\crefname{corollary}{Corollary}{Corollaries}
+\Crefname{corollary}{Corollary}{Corollaries}
+
+\crefname{algocf}{Algorithm}{Algorithms}
+\Crefname{algocf}{Algorithm}{Algorithms}
+
 \usetikzlibrary{decorations, decorations.pathreplacing, decorations.pathmorphing,
@@
 
-\definecolor{highlight}{rgb}{0.819608, 0.733333, 0.34902}
+\definecolor{highlight}{rgb}{0.529412, 0.74902, 0.466667}
 
@@
 
-The subgraph isomorphism problem is to decide whether a copy of a small ``pattern'' graph inside a
-larger ``target'' graph. Although \NP-complete, by combining design techniques from artificial
-intelligence with careful algorithm engineering, modern subgraph isomorphism solvers can often
-produce exact solutions quickly even on graphs with thousands of vertices. The current single
+The subgraph isomorphism problem is to decide whether a copy of a small ``pattern'' graph occurs
+inside a larger ``target'' graph. Although \NP-complete, by combining design techniques from
+artificial intelligence with careful algorithm engineering, modern subgraph isomorphism solvers can
+often produce exact solutions quickly even on graphs with thousands of vertices. The current single
 strongest subgraph isomorphism solver uses ``highest degree first'' as a value-ordering heuristic to
@@
 (which is extremely beneficial) based around restarts and nogood recording. To explore different
-parts of the search space when restarting, we introduce a novel value-ordering heuristic, which adds
-a small amount of randomness to the ``highest degree first'' heuristic.  The effects of these
-changes are shown in the final plot of \cref{figure:scatter-random}: we can make a state-of-the-art algorithm
-perform much better on a large number of satisfiable instances, whilst performing worse only rarely
-on satisfiable instances, and never on unsatisfiable instances.  Further, these properties are
-\emph{not} shared by discrepancy searches (\cref{figure:scatter-dds}) or by random
-value-ordering with restarts (the first two plots of \cref{figure:scatter-random}). Finally, we
-show that this strategy is also effective in an optimisation setting, producing benefits in two
-maximum common induced subgraph algorithms.
+parts of the search space when restarting, we introduce a novel way of adapting a value-ordering
+heuristic to add a very small amount of randomness. The effects of these changes are shown in the
+final plot of \cref{figure:scatter-random}: we can make a state-of-the-art algorithm perform much
+better on a large number of satisfiable instances, whilst performing worse only rarely on
+satisfiable instances, and never on unsatisfiable instances.  Further, these properties are
+\emph{not} shared by discrepancy searches (\cref{figure:scatter-dds}) or by random value-ordering
+with restarts (the first two plots of \cref{figure:scatter-random}). Finally, we show that this
+strategy is also effective in an optimisation setting, producing benefits in two maximum common
+induced subgraph algorithms.
 
@@
 vertex set $\vertexset(G)$, and with edges between vertices $v$ and $w$ if there are at least $n$
-simple paths of length exactly $\ell$ between vertices $v$ and $w$ in $G$. Finally, we define a
-function $\FuncSty{compatible}(v, w)$ to mean that mapping pattern vertex $v$ to target vertex $w$
-is feasible based upon both loop constraints and the neighbourhood degree sequence relation of
-\citet{DBLP:journals/constraints/ZampelliDS10}.
+simple paths of length $\ell$ between vertices $v$ and $w$ in $G$.
 
@@
 pattern graph $\mathcal{P}$ to a target graph $\mathcal{T}$, such that adjacent vertices in
-$\mathcal{P}$ are mapped to adjacent vertices in $\mathcal{T}$. We write $v \mapsto v'$ to mean that
-pattern vertex $v$ is mapped to target $v'$ under such a mapping.
+$\mathcal{P}$ are mapped to adjacent vertices in $\mathcal{T}$ (including that vertices with loops
+in $\mathcal{P}$ may only be mapped to vertices with loops in $\mathcal{T}$). We write $v \mapsto
+v'$ to mean that pattern vertex $v$ is mapped to target $v'$ under such a mapping. Finally, we
+define a function $\FuncSty{compatible}(v, w)$ to mean that mapping $v \mapsto w$ is feasible based
+upon both the loop rule, and the neighbourhood degree sequence relation of
+\citet{DBLP:journals/constraints/ZampelliDS10}.
 
@@
 \begin{description}
-    \item[BVG(r), M4D(r), and Rand] are randomly generated graphs using different models (bounded
+    \item[BVG(r), M4D(r), and Rand] are families of randomly generated graphs using different models (bounded
         degree, regular mesh, and uniform), where each pattern is a permuted random connected subgraph
@@
 \item[Biased] We branch by selecting a vertex $v'$ from the chosen domain $D_v$ with
-probability \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \text{.} \]\end{description}
-
-The biased heuristic continues to prefer vertices of high degree, but will give the same chance of
+probability \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \text{.} \] This is a new
+heuristic, which continues to prefer vertices of high degree, but will give the same chance of
 being selected next to two vertices of equal degree.  It also introduces an element of randomness,
-which is needed to make restarts have an effect.  We note that it resembles the well-known
-\emph{softmax} weighting scheme applied to the degree heuristic, although usually \emph{softmax}
-uses a base of $e$ rather than $2$.  Experiments showed that in this context, the choice of base is
-not important to the search space size; however, a base of $2$ avoids the need to use floating point
-arithmetic, which was a substantial overhead.
+which is needed to make restarts have an effect. \end{description}
 
@@
 value-ordering heuristic? Are there further benefits to be had from also modifying variable-ordering
-heuristics? Are there theoretical rather than just empirical reasons for using the
-\emph{softmax}-like bias scheme? And how can such a search process be parallelised?
+heuristics? Are there theoretical rather than just empirical reasons for using an exponential bias
+scheme? And how can such a search process be parallelised?
 

### 2018-07-26 paper/paper.tex
A bit of editoring

@@
             \nl \tikzmark{restart5s}\ElseIf{$S = \KwSty{restart}$}{
-                \nl \lForEach{$c \in C$}{$\textnormal{post nogood}~B \wedge (v \mapsto c) \Rightarrow \bot$\tikzmark{restart5r}}
+                \nl \ForEach{$c \in C$}{\nl$\textnormal{post nogood}~B \wedge (v \mapsto c) \Rightarrow \bot$\tikzmark{restart5r}}
                 \nl $\KwSty{return}~\KwSty{restart}$\tikzmark{restart5e} \;
@@
 \nl \Begin{
-    \nl \While{$(D_v, v') \gets ($\textnormal{a unit domain from}~$D, \textnormal{the single value
-    in that domain})$}{
+    \nl \While{$D_v \gets $\textnormal{a unit domain from}~$D$}{
+        \nl $v' \gets \textnormal{the single value in}~D_v$ \;
         \nl \tikzmark{restart7s}trigger nogoods with watch $(v \mapsto v')$\tikzmark{restart7e}\siplabel{2wl} \;
@@
 
+\noindent
 Other studies use a random selection of 200 of each of the instances from the ``meshes'' and
@@
 being selected next to two vertices of equal degree.  It also introduces an element of randomness,
-which is needed to make restarts have an effect. \end{description}
+which will be needed to make restarts have an effect. \end{description}
 
+\noindent
 The performance of the degree and biased heuristics are very similar: the
@@
 value-ordering heuristic without adversely affecting its performance \emph{in aggregate}.
-\Cref{figure:scatter-heuristics} gives a more detailed look at these results. This scatter plot
-shows that despite the aggregate performance being very similar, on a case by case basis, the two
+\Cref{figure:scatter-heuristics} gives a detailed look at these results. This scatter plot
+shows that despite the aggregate performance being similar, on a case by case basis, the two
 heuristics can make a large difference to the performance of satisfiable instances.
@@
 
+\begin{figure}[tb]
+    \centering
+    \includegraphics*{gen-graph-restarts.pdf}
+
+    \caption{The cumulative number of satisfiable instances solved over time, with and without
+    restarts and nogood recording.}
+    \label{figure:restarts}
+\end{figure}
+
 \begin{figure*}[tb]
@@
 \end{figure*}
-
-\begin{figure}[tb]
-    \centering
-    \includegraphics*{gen-graph-restarts.pdf}
-
-    \caption{The cumulative number of satisfiable instances solved over time, with and without
-    restarts and nogood recording.}
-    \label{figure:restarts}
-\end{figure}
 

### 2018-08-12 paper/paper.tex
Mucking around

@@
 
-\begin{figure}[tb]
+\begin{figure}[p]
     \centering
     \includegraphics*{gen-graph-others.pdf}
+
+    \smallskip
+
+    \centering
+    \includegraphics*{gen-graph-others-zoom.pdf}
 
@@
     \centering
-    \includegraphics*{gen-graph-others-zoom.pdf}
+    \includegraphics*{gen-graph-others-induced.pdf}
 
-    \caption{Above, the cumulative number of instances solved over time, comparing our algorithm
+    \smallskip
+
+    \centering
+    \includegraphics*{gen-graph-others-induced-zoom.pdf}
+
+    \caption{On the top two rows, the cumulative number of instances solved over time, comparing our algorithm
     (both in its basic form, and with the improvements introduced in the remainder of the paper) to
-    other solvers. Below, the same, looking only at satisfiable instances.}
+    other solvers. The second considers only satisfiable instances. On the bottom two rows, the same,
+    for the induced subgraph isomorphism problem.}
     \label{figure:others}

### 2018-09-05 paper/paper.tex
2019
Betterification

@@
 \documentclass[letterpaper]{article} %DO NOT CHANGE THIS
-\usepackage{aaai18}  %Required
+\usepackage{aaai19}  %Required
 \usepackage{times}  %Required
@@
     demonstrate that a random value-ordering heuristic with restarts beats conventional search with
-    the best-known tailored heuristic. This motivates the introduction of a slightly-random
+    the best-known tailored heuristic. This motivates the introduction of a novel slightly-random
     heuristic with similar behaviour to the tailored heuristic: when combined with restarts and
@@
 direct a constraint programming style search
-\cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:conf/ijcai/McCreeshPT16}. This
-heuristic is much better than branching randomly, but is still far from perfect.  We could try to
-introduce tiebreaking mechanisms and similar refinements to strengthen the heuristic. However, given
-the extreme cost of making an incorrect branching choice early in search, we believe it is more
-fruitful to investigate alternatives to simple backtracking (depth-first) search. This paper
-evaluates two such approaches: discrepancy search (which performs poorly), and a new form of search
-(which is extremely beneficial) based around restarts and nogood recording. To explore different
-parts of the search space when restarting, we introduce a novel way of adapting a value-ordering
-heuristic to add a very small amount of randomness. The effects of these changes are shown in the
-final plot of \cref{figure:scatter-random}: we can make a state-of-the-art algorithm perform much
-better on a large number of satisfiable instances, whilst performing worse only rarely on
-satisfiable instances, and never on unsatisfiable instances.  Further, these properties are
-\emph{not} shared by discrepancy searches (\cref{figure:scatter-dds}) or by random value-ordering
-with restarts (the first two plots of \cref{figure:scatter-random}). Finally, we show that this
-strategy is also effective in an optimisation setting, producing benefits in two maximum common
-induced subgraph algorithms.
+\cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:journals/jair/McCreeshPST18}. This
+heuristic is much better than branching randomly, but is still far from perfect. We therefore
+investigate different techniques for offsetting mistakes in value-ordering decisions. We first try
+discrepancy searches, which turn out to be much too expensive to be a practical solution. We then
+look at using random value-ordering, restarts, and nogood recording, which together give better
+aggregate performance at the expense of much worse behaviour on many instances. Next, we introduce a
+new way of adapting the existing value-ordering heuristic to introduce just a small amount of
+randomness.  This lets us make a state-of-the-art subgraph algorithm perform much better
+on a large number of satisfiable instances, whilst performing worse only rarely on satisfiable
+instances, and never on unsatisfiable instances. Finally, we show that this strategy is also
+effective in an optimisation setting, producing benefits in two maximum common induced subgraph
+algorithms. This combination of techniques gives us, for the first time, a practical alternative to
+backtracking search for constraint programming style algorithms where we have a strong
+value-ordering heuristic and where we care both about satisfiable and unsatisfiable instances.
 
@@
 Let $G$ be a graph. We denote its set of vertices by $\vertexset(G)$, and write $v \sim_G w$ to mean
-that vertices $v$ and $w$ are adjacent; we allow a vertex to be adjacent to itself. The
+that vertices $v$ and $w$ are adjacent; we allow a vertex to be adjacent to itself (a \emph{loop}). The
 \emph{neighbourhood} of a vertex, $\neighbourhood_G(v)$, is the set of vertices adjacent to $v$, and
@@
 
-The non-induced subgraph isomorphism problem is to find an injective mapping from the vertices of a
-pattern graph $\mathcal{P}$ to a target graph $\mathcal{T}$, such that adjacent vertices in
-$\mathcal{P}$ are mapped to adjacent vertices in $\mathcal{T}$ (including that vertices with loops
-in $\mathcal{P}$ may only be mapped to vertices with loops in $\mathcal{T}$). We write $v \mapsto
-v'$ to mean that pattern vertex $v$ is mapped to target $v'$ under such a mapping. Finally, we
-define a function $\FuncSty{compatible}(v, w)$ to mean that mapping $v \mapsto w$ is feasible based
-upon both the loop rule, and the neighbourhood degree sequence relation of
-\citet{DBLP:journals/constraints/ZampelliDS10}.
+The \emph{non-induced subgraph isomorphism problem} is to find an injective mapping from the
+vertices of a pattern graph $\mathcal{P}$ to a target graph $\mathcal{T}$, such that adjacent
+vertices in $\mathcal{P}$ are mapped to adjacent vertices in $\mathcal{T}$ (including that vertices
+with loops in $\mathcal{P}$ may only be mapped to vertices with loops in $\mathcal{T}$). The
+\emph{induced} problem additionally requires that non-adjacent vertices are mapped to non-adjacent
+vertices. We write $v \mapsto v'$ to mean that pattern vertex $v$ is mapped to target $v'$ under
+either kind of mapping. Finally, we define a function $\FuncSty{compatible}(v, w)$ to mean that
+mapping $v \mapsto w$ is feasible based upon both the loop rule, and the neighbourhood degree
+sequence relation of \citet{DBLP:journals/constraints/ZampelliDS10}.
 
@@
 for full technical details; that algorithm, in turn, is a simplification of the Glasgow2 algorithm
-\cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}.
+\cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}. We show the non-induced problem; the
+induced problem may be solved by including the loop-complements of $\mathcal{P}$ and $\mathcal{T}$
+in $L$ on \siplineref{supplemental}.
 
@@
 \medskip
-    \nl $\FuncSty{search}$ (Domains $D$, Decisions $B$) $\rightarrow$ Bool
+    \nl $\FuncSty{search}$ (Domains $D$, Trail $B$) $\rightarrow$ Bool
     \tikzmark{restart8s}\KwSty{or} \KwSty{restart}\tikzmark{restart8e} \;
@@
         \nl \tikzmark{restart7s}trigger nogoods with watch $(v \mapsto v')$\tikzmark{restart7e}\siplabel{2wl} \;
-        \nl \ForEach{$D_w \in D - D_v$}{
-            \nl $D_w \gets D_w - v'$ \siplabel{removev} \;
+        \nl \ForEach{$D_w \in D \setminus \{ D_v$ \}}{
+            \nl $D_w \gets D_w \setminus \{ v' \}$ \siplabel{removev} \;
             \nl \ForEach{$(P,\,T) \in L~\KwSty{where}~v \sim_P w$}{\nl$D_w \gets D_w \cap \neighbourhood_T(v')$}
@@
 which occurs on \siplineref{valueordering} of \cref{algorithm:sip}, we will initially select
-vertices from highest degree to lowest degree; we return to value ordering in the second half of
+vertices from highest degree to lowest degree; value ordering is the major theme of the remainder of
 this paper. Note that the variable-ordering heuristic affects both unsatisfiable and satisfiable
@@
 \cref{algorithm:sip} in C++ using the bit-parallel implementation of
-\citet{DBLP:conf/aaai/HoffmannMR17} as a starting point, and compiled it using GCC 6.3.0. We use a
+\citet{DBLP:conf/aaai/HoffmannMR17} as a starting point, and compiled it using GCC 7.2.0. We use a
 deterministic pseudo-random number generator for reproducibility.
@@
 pattern / target pairs. This gives a total of 14,621 instances (rather than the original 5,725). At
-least \input{gen-n-sat.tex}\unskip\ of these instances are known to be satisfiable, and at least
-\input{gen-n-unsat.tex}\unskip\ are unsatisfiable.
+least \input{gen-n-sat.tex}\unskip\ of these instances are known to be satisfiable for the
+non-induced problem, and at least \input{gen-n-unsat.tex}\unskip\ are unsatisfiable.
 
@@
 We begin by verifying that our implementation of the basic \cref{algorithm:sip} is competitive. In
-\cref{figure:others} we show the cumulative number of instances solved over time, using our
-implementation (both with the degree heuristic, and with the modifications described in the
-remainder of this paper), the Glasgow2 and Glasgow3 algorithms from which \cref{algorithm:sip} is derived
+the first and third plots \cref{figure:others} we show the cumulative number of instances solved
+over time for the non-induced and induced problems respectively. We compare our implementation (both
+with the degree heuristic, and with the modifications described in the remainder of this paper), the
+Glasgow2 and Glasgow3 algorithms from which \cref{algorithm:sip} is derived
 \cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}, the PathLAD variation of the LAD
-algorithm \cite{DBLP:journals/ai/Solnon10,DBLP:conf/lion/KotthoffMS16}, and VF2
-\cite{DBLP:journals/pami/CordellaFSV04}, in each case using the original implementation provided
-by the algorithm's authors. For each algorithm, the $y$ value
+algorithm \cite{DBLP:journals/ai/Solnon10,DBLP:conf/lion/KotthoffMS16}, VF2
+\cite{DBLP:journals/pami/CordellaFSV04}, RI \cite{DBLP:journals/bmcbi/BonniciGPSF13}, and VF3
+\cite{DBLP:conf/gbrpr/CarlettiFSV17} (which only supports the induced problem), in each case using
+the original implementation provided by the algorithm's authors. For each algorithm, the $y$ value
 gives the cumulative number of instances which (individually) can be solved in no more than $x$
@@
 
-As expected, the performance of \cref{algorithm:sip} with the degree heuristic is very close to that
-of Glasgow2, although we perform better at lower runtimes because we use simpler preprocessing at
-the top of search. Our implementation also clearly beats PathLAD and VF2, except for very low
-choices of timeout.
+As expected, in both cases the performance of \cref{algorithm:sip} with the degree heuristic is very
+close to that of Glasgow2, although we perform better at lower runtimes because we use simpler
+preprocessing at the top of search. Our implementation also clearly beats PathLAD, VF2, VF3 and RI,
+except for very low choices of timeout.
 
@@
 hard to see the differences between the stronger solvers at higher runtimes. This paper focusses
-upon improving the performance on the remaining hard satisfiable instances, and so in the second
-plot in \cref{figure:others} (and in subsequent cumulative plots) we show only satisfiable
+upon improving the performance on the remaining hard satisfiable instances, and so in the second and
+fourth plots in \cref{figure:others} (and in subsequent cumulative plots) we show only satisfiable
 instances, and use a reduced range on both axes.
@@
 \Cref{figure:value-ordering-heuristics} shows the cumulative number of satisfiable instances
-solved over time using different value-ordering heuristics:
+solved over time using different value-ordering heuristics. For the remainder of this section, we
+show only the non-induced problem; results with the induced variant are similar.
 
@@
 \item[Biased] We branch by selecting a vertex $v'$ from the chosen domain $D_v$ with
-probability \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \text{.} \] This is a new
-heuristic, which continues to prefer vertices of high degree, but will give the same chance of
-being selected next to two vertices of equal degree.  It also introduces an element of randomness,
-which will be needed to make restarts have an effect. \end{description}
+probability \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \text{.} \]\end{description}
 
 \noindent
-The performance of the degree and biased heuristics are very similar: the
-biased heuristic has slightly worse performance below a 400s timeout, and effectively equal
-performance above it. This shows that we can introduce an element of randomness into the degree
-value-ordering heuristic without adversely affecting its performance \emph{in aggregate}.
-\Cref{figure:scatter-heuristics} gives a detailed look at these results. This scatter plot
-shows that despite the aggregate performance being similar, on a case by case basis, the two
-heuristics can make a large difference to the performance of satisfiable instances.
+The biased heuristic is new. It continues to prefer vertices of high degree, but will give the same
+chance of being selected next to two vertices of equal degree.  It also introduces an element of
+randomness, which will be needed to make restarts have an effect.  The performance of the degree and
+biased heuristics are very similar: the biased heuristic has slightly worse performance below a 400s
+timeout, and effectively equal performance above it. This shows that we can introduce an element of
+randomness into the degree value-ordering heuristic without adversely affecting its performance
+\emph{in aggregate} (and more generally, this scheme could be used for any value-ordering
+heuristic which puts choices in some kind of rank with ties.)
+
+\Cref{figure:scatter-heuristics} gives a detailed look at these results. This scatter plot shows
+that despite the aggregate performance being similar, on a case by case basis, the two heuristics
+can make a large difference to the performance of satisfiable instances.
 
@@
 A \emph{discrepancy} is where search goes against a value-ordering heuristic.  Discrepancy searches
-\cite{DBLP:conf/ijcai/HarveyG95,DBLP:conf/aaai/Korf96,DBLP:conf/ijcai/Walsh97} are alternatives to
-backtracking search that initially search disallowing all discrepancies, and then retry search
-allowing an increasing number of discrepancies at each iteration until either a solution is found or
-unsatisfiability is proven. These schemes assume that value-ordering heuristics are usually
-reliable, and that most solutions can be found with only a small number of discrepancies; in such
-cases, the heavy commitment to early branching choices made by conventional backtracking search can
-be extremely costly.
+\cite{DBLP:conf/ijcai/HarveyG95,DBLP:conf/aaai/Korf96,DBLP:conf/ijcai/Walsh97,DBLP:conf/cpaior/KarouiHLN07}
+are alternatives to backtracking search that initially search disallowing all discrepancies, and
+then retry search allowing an increasing number of discrepancies at each iteration until either a
+solution is found or unsatisfiability is proven. These schemes assume that value-ordering heuristics
+are usually reliable, and that most solutions can be found with only a small number of
+discrepancies; in such cases, the heavy commitment to early branching choices made by conventional
+backtracking search can be extremely costly.
 
 \Cref{figure:scatter-dds} shows the effects of adding Walsh's \cite{DBLP:conf/ijcai/Walsh97}
-depth-bounded discrepancy search (DDS) to \cref{algorithm:sip}. Each point represents the solving
+depth-bounded discrepancy search (DDS) to \cref{algorithm:sip} (results with other discrepancy
+search variants are similar). Each point represents the solving
 time for one instance---to avoid noise for easier instances, we measure only time spent during
@@
 extremely beneficial on satisfiable instances.  However, on both unsatisfiable and most satisfiable
-instances, the overheads can be extremely large, and DDS is much worse in aggregate (even when only
-considering satisfiable instances).
-
-These large overheads are to be expected: discrepancy searches are aimed primarily at getting better
-feasible solutions in optimisation problems which are too large for a proof of optimality to be a
-realistic prospect, and they are not well-suited for unsatisfiable decision problems. However, the
-extremely large gains on some satisfiable instances confirm our suspicions that we should find an
-alternative to heavy-commitment backtracking search.
+instances, the overheads can be extremely large, and DDS is much worse in aggregate and is not a
+viable approach (even when only considering satisfiable instances).  These large overheads are to be
+expected: discrepancy searches are aimed primarily at getting better feasible solutions in
+optimisation problems which are too large for a proof of optimality to be a realistic prospect, and
+they are not well-suited for unsatisfiable decision problems. Despite this, the extremely large gains on
+some satisfiable instances confirm our suspicions that we should find an alternative to
+heavy-commitment backtracking search.
 
@@
 \cite{DBLP:journals/jsat/LecoutreSTV07,DBLP:conf/cp/GayHLS15,DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17}.
-In this paper, we instead rely upon randomness in the \emph{value}-ordering heuristic, and continue
+In this paper, we instead rely upon randomness in a special new \emph{value}-ordering heuristic, and continue
 to use smallest domain first with static tiebreaking for variable-ordering.\footnote{It may be
@@
 still---that is, if we are introducing restarts, then it is better to add a small amount of
-randomness to a tailored deterministic heuristic than it is to throw away the heuristic altogether.
+randomness to a tailored heuristic than it is to throw away the heuristic altogether.
 Indeed, the original algorithm can solve 1,983 satisfiable instances by 993.0 seconds, whilst the
@@
 instances that are more than a factor of ten times worse. In other words, as well as improving
-performance, we have made up most of the consistency we lost by introducing randomness.
+performance, we have made up most of the consistency we lost by introducing randomness. This is a
+promising result: for the first time, we have introduced a viable alternative to backtracking search
+in a setting where we both have a strong \emph{value}-ordering heuristic, \emph{and} care about both
+satisfiable and unsatisfiable instances.
 
@@
 value-ordering heuristic? Are there further benefits to be had from also modifying variable-ordering
-heuristics? Are there theoretical rather than just empirical reasons for using an exponential bias
-scheme? And how can such a search process be parallelised?
+heuristics? And how can such a search process be parallelised?
 

### 2018-10-23 paper/paper.tex
Tweaks

@@
 The subgraph isomorphism problem is to decide whether a copy of a small ``pattern'' graph occurs
-inside a larger ``target'' graph. Although \NP-complete, by combining design techniques from
+inside a larger ``target'' graph. Although the problem is \NP-complete, by combining design techniques from
 artificial intelligence with careful algorithm engineering, modern subgraph isomorphism solvers can
@@
 
-\Cref{figure:scatter-heuristics} gives a detailed look at these results. This scatter plot shows
-that despite the aggregate performance being similar, on a case by case basis, the two heuristics
-can make a large difference to the performance of satisfiable instances.
+\Cref{figure:scatter-heuristics} gives a detailed look at these results.  Each point represents the
+solving time for one instance---to avoid noise for easier instances, we measure only time spent
+during search, and exclude time spent in preprocessing and initialisation.  Points below the $x-y$
+diagonal are speedups, whilst points on the top and right axes represent instances which timed out
+after one thousand seconds with one algorithm, but not the other. For satisfiable instances, the
+different point styles show the different families, whilst all unsatisfiable instances are shown as
+dark dots.  This plot shows that despite the aggregate performance being similar, on a case by case
+basis, the two heuristics can make a large difference to the performance of individual satisfiable
+instances. This suggests that although degree is a good heuristic, we should perhaps not commit
+heavily to a vertex of highest degree, but also consider vertices of the same or similar degree.
 
@@
 depth-bounded discrepancy search (DDS) to \cref{algorithm:sip} (results with other discrepancy
-search variants are similar). Each point represents the solving
-time for one instance---to avoid noise for easier instances, we measure only time spent during
-search, and exclude time spent in preprocessing and initialisation.  Points below the $x-y$ diagonal
-are speedups, whilst points on the top and right axes represent instances which timed out after one
-thousand seconds with one algorithm, but not the other. For satisfiable instances, the different
-point styles show the different families, whilst all unsatisfiable instances are shown as dark dots.
-
-The points well below the diagonal line and along the right-hand axis show that DDS can sometimes be
-extremely beneficial on satisfiable instances.  However, on both unsatisfiable and most satisfiable
-instances, the overheads can be extremely large, and DDS is much worse in aggregate and is not a
-viable approach (even when only considering satisfiable instances).  These large overheads are to be
-expected: discrepancy searches are aimed primarily at getting better feasible solutions in
-optimisation problems which are too large for a proof of optimality to be a realistic prospect, and
-they are not well-suited for unsatisfiable decision problems. Despite this, the extremely large gains on
-some satisfiable instances confirm our suspicions that we should find an alternative to
-heavy-commitment backtracking search.
+search variants are similar).  The points well below the diagonal line and along the right-hand axis
+show that DDS can sometimes be extremely beneficial on satisfiable instances.  However, on both
+unsatisfiable and most satisfiable instances, the overheads can be extremely large, and DDS is much
+worse in aggregate and is not a viable approach (even when only considering satisfiable instances).
+These large overheads are to be expected: discrepancy searches are aimed primarily at getting better
+feasible solutions in optimisation problems which are too large for a proof of optimality to be a
+realistic prospect, and they are not well-suited for unsatisfiable decision problems. Despite this,
+the extremely large gains on some satisfiable instances confirm our suspicions that we should find
+an alternative to heavy-commitment backtracking search.
 
@@
 to perform a certain amount of search, and then if no solution has been found (and unsatisfiability
-has not been proven), to abandon search and restart from the beginning. Obviously, such an approach
+has not been proven), to abandon search and restart from the beginning. Such an approach
 can only be beneficial if something changes after restarting---in a constraint programming setting,
@@
 would considerably complicate the empirical analysis.} Using restarts on value-ordering heuristics
-appears to be uncommon---although \citet{DBLP:conf/flairs/RazgonOP07} look at learning
+is uncommon---although \citet{DBLP:conf/flairs/RazgonOP07} look at learning
 value-ordering heuristics from restarts, \citet{DBLP:conf/cp/ChuSS09} use a similar
@@
 \siplineref{countrestarts}) to decide when to restart, rather than counting recursive calls.
-Following established wisdom, we multiply each item in the Luby sequence by a magic constant---we
+Following convention, we multiply each item in the Luby sequence by a constant---we
 used the SMAC automatic parameter tuner \cite{DBLP:conf/lion/HutterHL11} to select the value 660.
-(We also tried a geometric restart scheme, which gave less favourable results.)
 
@@
 
-\subsection{The Effects of Restarts and Nogood Recording}
-
 \begin{figure}[tb]
@@
 
-    \caption{The cumulative number of satisfiable instances solved over time, with and without
+    \caption{The number of satisfiable instances solved over time, with and without
     restarts and nogood recording.}
@@
 \end{figure*}
+
+\subsection{The Effects of Restarts and Nogood Recording}
 
@@
 
-Having looked at subgraph isomorphism in detail, we now briefly discuss the more general maximum
-common induced subgraph problem.  Two recent algorithms for this problem also make use of
-backtracking search with degree as a value-ordering heuristic. The k${\downarrow}$ algorithm
-\cite{DBLP:conf/aaai/HoffmannMR17} attempts to solve the problem by first trying to solve the
-induced subgraph isomorphism problem, and then if that fails, retries allowing a single unmatched
-vertex (and thus using weaker invariants), and so on. Due to its similarity to \cref{algorithm:sip},
-we can introduce the same bias and restart strategy, requiring only small changes
-to handle wildcards in nogoods.
+Having looked at subgraph isomorphism in detail, we now briefly discuss the maximum common induced
+subgraph problem, to see whether our new approach to search has more general applicability.  Two
+recent algorithms for this problem also make use of backtracking search with degree as a
+value-ordering heuristic. The k${\downarrow}$ algorithm \cite{DBLP:conf/aaai/HoffmannMR17} attempts
+to solve the problem by first trying to solve the induced subgraph isomorphism problem, and then if
+that fails, retries allowing a single unmatched vertex (and thus using weaker invariants), and so
+on. Due to its similarity to \cref{algorithm:sip}, we can introduce the same bias and restart
+strategy, requiring only small changes to handle wildcards in nogoods.
 
@@
 instances. Our results show that the key to success is a combination of aggressive restarts, nogood
-recording, and biasing an existing value-ordering heuristic with just a small amount of randomness.
+recording, and biasing an existing value-ordering heuristic with just a pinch of randomness.
 

### 2018-11-09 paper/paper.tex
LNCSify

@@
 
-\def\year{2019}\relax
-%File: formatting-instruction.tex
-\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
-\usepackage{aaai19}  %Required
-\usepackage{times}  %Required
-\usepackage{helvet}  %Required
-\usepackage{courier}  %Required
-\usepackage{url}  %Required
-\usepackage{graphicx}  %Required
-\frenchspacing  %Required
-\setlength{\pdfpagewidth}{8.5in}  %Required
-\setlength{\pdfpageheight}{11in}  %Required
-%PDF Info Is Required:
-  \pdfinfo{
-/Title (Value Ordering, Discrepancies, and Restarts for Subgraph Algorithms)
-/Author (Anonymous and Anonymous and Anonymous)}
-\setcounter{secnumdepth}{0}
+\documentclass[runningheads]{llncs}
 
@@
 
-% \usepackage{showframe}
+\usepackage{showframe}
 
@@
 
-\newcommand{\citet}[1]{\citeauthor{#1} \shortcite{#1}}
+\newcommand{\citet}[2]{#1\cite{#2}}
 
@@
 
-\title{Value Ordering, Discrepancies, and Restarts for Subgraph Algorithms}
-\author{Anonymous \and Anonymous \and Anonymous \\ email address \\ address}
+\begin{document}
 
-\begin{document}
+\title{Beating Simple Backtracking Search in Subgraph Algorithms\thanks{This work was
+supported by the Engineering and Physical Sciences Research Council [grant numbers EP/026842/1 and
+EP/M508056/1].}}
+
+\author{Ciaran McCreesh\orcidID{0000-0002-6106-4871}\and
+Patrick Prosser\orcidID{0000-0003-4460-6912} \and
+James Trimble\orcidID{0000-0001-7282-8745}}
+
+\authorrunning{C. McCreesh et al.}
+
+\institute{University of Glasgow, Glasgow, Scotland\\
+\email{ciaran.mccreesh@glasgow.ac.uk}}
 
@@
 \begin{abstract}
-    Modern subgraph isomorphism solvers use degree-based value-ordering heuristics to direct
-    backtracking search. This kind of search makes a heavy commitment to the first branching choice,
+    Modern subgraph isomorphism solvers use degree as a value-ordering heuristics to direct
+    backtracking search. This makes a heavy commitment to the first branching choice,
     which is often incorrect. We evaluate discrepancy-based searches as an alternative, but find
@@
 the \emph{degree} of $v$, $\deg_G(v)$, is the cardinality of its neighbourhood. When the graph is clear from the context, we omit
-the $G$ subscripts. As per \citet{DBLP:conf/aaai/HoffmannMR17}, $G^{n,\ell}$ denotes the graph with
+the $G$ subscripts. As per \citet{Hoffmann et al.\ }{DBLP:conf/aaai/HoffmannMR17}, $G^{n,\ell}$ denotes the graph with
 vertex set $\vertexset(G)$, and with edges between vertices $v$ and $w$ if there are at least $n$
@@
 mapping $v \mapsto w$ is feasible based upon both the loop rule, and the neighbourhood degree
-sequence relation of \citet{DBLP:journals/constraints/ZampelliDS10}.
+sequence relation of \citet{Zampelli et al.\ }{DBLP:journals/constraints/ZampelliDS10}.
 
@@
 subgraph isomorphism problem. This algorithm is very closely based upon the $k{\downarrow}$
-algorithm of \citet{DBLP:conf/aaai/HoffmannMR17} with $k = 0$, and we refer the reader to that paper
+algorithm of \citet{Hoffmann et al.\ }{DBLP:conf/aaai/HoffmannMR17} with $k = 0$, and we refer the reader to that paper
 for full technical details; that algorithm, in turn, is a simplification of the Glasgow2 algorithm
@@
 vertices. Each domain is initialised at the top of search (\siplinerangeref{initstart}{initend}),
-using the invariants described by \citet{DBLP:journals/constraints/ZampelliDS10} and
-\citet{DBLP:conf/aaai/HoffmannMR17} to restrict its initial values.
+using the invariants described by \citet{Zampelli et al.\ }{DBLP:journals/constraints/ZampelliDS10} and
+\citet{Hoffmann et al.\ }{DBLP:conf/aaai/HoffmannMR17} to restrict its initial values.
 
@@
 
-Our $\FuncSty{propagate}$ routine is also close to that of \citet{DBLP:conf/aaai/HoffmannMR17},
+Our $\FuncSty{propagate}$ routine is also close to that of \citet{Hoffmann et al.\ }{DBLP:conf/aaai/HoffmannMR17},
 differing only in that we carry out the filtering for injectivity
@@
         \node [fill=highlight, rounded corners, fit=(restart2sc) (restart2ec)] { };
-        \coordinate (restart3sc) at ($(pic cs:restart3s) + (0, 0.15)$);
-        \coordinate (restart3ec) at ($(pic cs:restart3e) + (0, 0.00)$);
-        \node [fill=highlight, rounded corners, fit=(restart3sc) (restart3ec)] { };
-        \coordinate (restart4sc) at ($(pic cs:restart4s) + (0, 0.15)$);
-        \coordinate (restart4ec) at ($(pic cs:restart4e) + (0, 0.00)$);
-        \node [fill=highlight, rounded corners, fit=(restart4sc) (restart4ec)] { };
-        \coordinate (restart5sc) at ($(pic cs:restart5s) + (0, 0.15)$);
-        \coordinate (restart5ec) at ($(pic cs:restart5e) - (0, 0.10)$);
-        \coordinate (restart5rc) at ($(pic cs:restart5r) + (0, 0.00)$);
-        \node [fill=highlight, rounded corners, fit=(restart5sc) (restart5ec) (restart5rc)] { };
-        \coordinate (restart6sc) at ($(pic cs:restart6s) + (0, 0.15)$);
-        \coordinate (restart6ec) at ($(pic cs:restart6e) + (0, 0.00)$);
-        \coordinate (restart6rc) at ($(pic cs:restart6r) + (0.8, 0.00)$);
-        \node [fill=highlight, rounded corners, fit=(restart6sc) (restart6ec) (restart6rc)] { };
-        \coordinate (restart7sc) at ($(pic cs:restart7s) + (0, 0.15)$);
-        \coordinate (restart7ec) at ($(pic cs:restart7e) + (0, 0.00)$);
-        \node [fill=highlight, rounded corners, fit=(restart7sc) (restart7ec)] { };
-        \coordinate (restart8sc) at ($(pic cs:restart8s) + (0.05, 0.15)$);
-        \coordinate (restart8ec) at ($(pic cs:restart8e) + (0, 0.00)$);
-        \node [fill=highlight, rounded corners, fit=(restart8sc) (restart8ec)] { };
         \coordinate (restart9sc) at ($(pic cs:restart9s) + (0, 0.15)$);
@@
 }
-\medskip
+\caption{The main subgraph isomorphism algorithm. The shaded code is added for restarting. See
+    \cref{algorithm:sip:search,algorithm:sip:propagate} for the search and propagation routines.}
+\label{algorithm:sip}
+\end{algorithm}
+
+\begin{algorithm}[p]\DontPrintSemicolon\small\SetInd{0.1em}{1em}
+    \begin{tikzpicture}[remember picture,overlay]
+        \coordinate (restart3sc) at ($(pic cs:restart3s) + (0, 0.15)$);
+        \coordinate (restart3ec) at ($(pic cs:restart3e) + (0, 0.00)$);
+        \node [fill=highlight, rounded corners, fit=(restart3sc) (restart3ec)] { };
+        \coordinate (restart4sc) at ($(pic cs:restart4s) + (0, 0.15)$);
+        \coordinate (restart4ec) at ($(pic cs:restart4e) + (0, 0.00)$);
+        \node [fill=highlight, rounded corners, fit=(restart4sc) (restart4ec)] { };
+        \coordinate (restart5sc) at ($(pic cs:restart5s) + (0, 0.15)$);
+        \coordinate (restart5ec) at ($(pic cs:restart5e) - (0, 0.10)$);
+        \coordinate (restart5rc) at ($(pic cs:restart5r) + (0, 0.00)$);
+        \node [fill=highlight, rounded corners, fit=(restart5sc) (restart5ec) (restart5rc)] { };
+        \coordinate (restart6sc) at ($(pic cs:restart6s) + (0, 0.15)$);
+        \coordinate (restart6ec) at ($(pic cs:restart6e) + (0, 0.00)$);
+        \coordinate (restart6rc) at ($(pic cs:restart6r) + (0.8, 0.00)$);
+        \node [fill=highlight, rounded corners, fit=(restart6sc) (restart6ec) (restart6rc)] { };
+        \coordinate (restart8sc) at ($(pic cs:restart8s) + (0.05, 0.15)$);
+        \coordinate (restart8ec) at ($(pic cs:restart8e) + (0, 0.00)$);
+        \node [fill=highlight, rounded corners, fit=(restart8sc) (restart8ec)] { };
+    \end{tikzpicture}
     \nl $\FuncSty{search}$ (Domains $D$, Trail $B$) $\rightarrow$ Bool
@@
 }
-\medskip
+\caption{Search routine for \cref{algorithm:sip}.}
+\label{algorithm:sip:search}
+\end{algorithm}
+
+\begin{algorithm}[tb]\DontPrintSemicolon\small\SetInd{0.1em}{1em}
+    \begin{tikzpicture}[remember picture,overlay]
+        \coordinate (restart7sc) at ($(pic cs:restart7s) + (0, 0.15)$);
+        \coordinate (restart7ec) at ($(pic cs:restart7e) + (0, 0.00)$);
+        \node [fill=highlight, rounded corners, fit=(restart7sc) (restart7ec)] { };
+    \end{tikzpicture}
 \nl $\FuncSty{propagate}$ (Domains $D$) $\rightarrow$ Bool \;
@@
 }
-\caption{Subgraph isomorphism. The shaded code is added for restarting.}
-\label{algorithm:sip}
+\caption{Propagation routine for \cref{algorithm:sip}.}
+\label{algorithm:sip:propagate}
 \end{algorithm}
@@
 \cref{algorithm:sip} in C++ using the bit-parallel implementation of
-\citet{DBLP:conf/aaai/HoffmannMR17} as a starting point, and compiled it using GCC 7.2.0. We use a
+\citet{Hoffmann et al.\ }{DBLP:conf/aaai/HoffmannMR17} as a starting point, and compiled it using GCC 7.2.0. We use a
 deterministic pseudo-random number generator for reproducibility.
 
-We use the dataset introduced by \citet{DBLP:conf/lion/KotthoffMS16} for evaluation. This dataset
+We use the dataset introduced by \citet{Kotthoff et al.\ }{DBLP:conf/lion/KotthoffMS16} for evaluation. This dataset
 brings together a range of randomly-generated and application instance families from earlier papers:
@@
 
-\begin{figure}[p]
-    \centering
+\begin{figure}[tb]
     \includegraphics*{gen-graph-others.pdf}
-
-    \smallskip
-
-    \centering
-    \includegraphics*{gen-graph-others-zoom.pdf}
+    \hfill
+    \includegraphics*{gen-graph-others-induced.pdf}
 
@@
 
-    \centering
-    \includegraphics*{gen-graph-others-induced.pdf}
-
-    \smallskip
-
-    \centering
+    \includegraphics*{gen-graph-others-zoom.pdf}
+    \hfill
     \includegraphics*{gen-graph-others-induced-zoom.pdf}
 
-    \caption{On the top two rows, the cumulative number of instances solved over time, comparing our algorithm
+    \caption{On the top row, the cumulative number of instances solved over time, comparing our algorithm
     (both in its basic form, and with the improvements introduced in the remainder of the paper) to
-    other solvers. The second considers only satisfiable instances. On the bottom two rows, the same,
-    for the induced subgraph isomorphism problem.}
+    other solvers, for the non-induced and induced problems. On the bottom row, the same,
+    considering only satisfiable instances.}
     \label{figure:others}
@@
 
-\Cref{figure:value-ordering-heuristics} shows the cumulative number of satisfiable instances
-solved over time using different value-ordering heuristics. For the remainder of this section, we
-show only the non-induced problem; results with the induced variant are similar.
+The left plot of \cref{figure:value-ordering-heuristics} shows the cumulative number of satisfiable
+instances solved over time using different value-ordering heuristics. For the remainder of this
+section, we show only the non-induced problem; results with the induced variant are similar.
 
@@
 
-\Cref{figure:scatter-heuristics} gives a detailed look at these results.  Each point represents the
-solving time for one instance---to avoid noise for easier instances, we measure only time spent
-during search, and exclude time spent in preprocessing and initialisation.  Points below the $x-y$
-diagonal are speedups, whilst points on the top and right axes represent instances which timed out
-after one thousand seconds with one algorithm, but not the other. For satisfiable instances, the
-different point styles show the different families, whilst all unsatisfiable instances are shown as
-dark dots.  This plot shows that despite the aggregate performance being similar, on a case by case
-basis, the two heuristics can make a large difference to the performance of individual satisfiable
-instances. This suggests that although degree is a good heuristic, we should perhaps not commit
-heavily to a vertex of highest degree, but also consider vertices of the same or similar degree.
+\begin{figure}[tb]
+    \includegraphics*{gen-graph-value-ordering-heuristics.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-heuristics.pdf}
 
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-value-ordering-heuristics.pdf}
-
-    \caption{The cumulative number of satisfiable instances solved over time, using four
-    different value-ordering heuristics.}
+    \caption{On the left, the cumulative number of satisfiable instances solved over time, using
+    four different value-ordering heuristics. On the right, an instance by instance comparison of
+    the degree and biased heuristics on all instances. Points on the outer axes are timeouts, and
+    point style shows instance family.}
     \label{figure:value-ordering-heuristics}
@@
 
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-scatter-heuristics.pdf}
-
-    \caption{An instance by instance comparison of the degree and biased heuristics on all
-    instances. Points on the outer axes are timeouts, and point style shows instance family.}
-    \label{figure:scatter-heuristics}
-\end{figure}
-
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-scatter-dds.pdf}
-
-    \caption{Comparing depth-bounded discrepancy search to backtracking search, both with the degree
-    heuristic.}\label{figure:scatter-dds}
-\end{figure}
+The right-hand plot of \cref{figure:value-ordering-heuristics} gives a detailed look at these
+results.  Each point represents the solving time for one instance---to avoid noise for easier
+instances, we measure only time spent during search, and exclude time spent in preprocessing and
+initialisation.  Points below the $x-y$ diagonal are speedups, whilst points on the top and right
+axes represent instances which timed out after one thousand seconds with one algorithm, but not the
+other. For satisfiable instances, the different point styles show the different families, whilst all
+unsatisfiable instances are shown as dark dots.  This plot shows that despite the aggregate
+performance being similar, on a case by case basis, the two heuristics can make a large difference
+to the performance of individual satisfiable instances. This suggests that although degree is a good
+heuristic, we should perhaps not commit heavily to a vertex of highest degree, but also consider
+vertices of the same or similar degree.
 
@@
 
+\begin{figure}[tb]
+    \includegraphics*{gen-graph-dds.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-dds.pdf}
+
+    \caption{Comparing depth-bounded discrepancy search to backtracking search, both with the degree
+    heuristic.}\label{figure:scatter-dds}
+\end{figure}
+
 \Cref{figure:scatter-dds} shows the effects of adding Walsh's \cite{DBLP:conf/ijcai/Walsh97}
@@
 search variants are similar).  The points well below the diagonal line and along the right-hand axis
-show that DDS can sometimes be extremely beneficial on satisfiable instances.  However, on both
-unsatisfiable and most satisfiable instances, the overheads can be extremely large, and DDS is much
-worse in aggregate and is not a viable approach (even when only considering satisfiable instances).
-These large overheads are to be expected: discrepancy searches are aimed primarily at getting better
-feasible solutions in optimisation problems which are too large for a proof of optimality to be a
-realistic prospect, and they are not well-suited for unsatisfiable decision problems. Despite this,
-the extremely large gains on some satisfiable instances confirm our suspicions that we should find
-an alternative to heavy-commitment backtracking search.
+on the scatter plot show that DDS can sometimes be extremely beneficial on satisfiable instances.
+However, on both unsatisfiable and most satisfiable instances, the overheads can be extremely large,
+and DDS is much worse in aggregate and is not a viable approach (even when only considering
+satisfiable instances).  These large overheads are to be expected: discrepancy searches are aimed
+primarily at getting better feasible solutions in optimisation problems which are too large for a
+proof of optimality to be a realistic prospect, and they are not well-suited for unsatisfiable
+decision problems. Despite this, the extremely large gains on some satisfiable instances confirm our
+suspicions that we should find an alternative to heavy-commitment backtracking search.
 
@@
 would considerably complicate the empirical analysis.} Using restarts on value-ordering heuristics
-is uncommon---although \citet{DBLP:conf/flairs/RazgonOP07} look at learning
-value-ordering heuristics from restarts, \citet{DBLP:conf/cp/ChuSS09} use a similar
-scheme in the context of parallel search, and an early approach by \citet{DBLP:conf/aaai/GomesSK98} does so in an optimisation context.
+is uncommon---although \citet{Razgon et al.\ }{DBLP:conf/flairs/RazgonOP07} look at learning
+value-ordering heuristics from restarts, \citet{Chu et al.\ }{DBLP:conf/cp/ChuSS09} use a similar
+scheme in the context of parallel search, and an early approach by \citet{Gomes et al.\ }{DBLP:conf/aaai/GomesSK98} does so in an optimisation context.
 
@@
 \begin{figure}[tb]
-    \centering
     \includegraphics*{gen-graph-restarts.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-by-family.pdf}
 
-    \caption{The number of satisfiable instances solved over time, with and without
-    restarts and nogood recording.}
-    \label{figure:restarts}
+    \caption{On the left, the number of satisfiable instances solved over time, with and without
+    restarts and nogood recording. To the right, a comparison between the original algorithm and the
+    new algorithm with all features enabled.}
+    \label{figure:old-vs-new}
 \end{figure}
 
-\begin{figure*}[tb]
+\begin{figure}[tb]
     \includegraphics*{gen-graph-scatter-random-goods.pdf}
@@
     \includegraphics*{gen-graph-scatter-random.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-by-family.pdf}
-    \caption{Comparing basic backtracking with the degree heuristic, versus the random and biased
-    heuristics with restarts. The first plot does not use nogood recording, whilst the second and
-    third do.}
-    \label{figure:scatter-random}
-\end{figure*}
+    \caption{On the left, not using nogood recording introduces slowdowns, particularly on
+    unsatisfiable instances. On the right, using a random value-ordering gives much worse
+    performance on many satisfiable instances.}
+    \label{figure:features}
+\end{figure}
 
@@
 
-In \cref{figure:restarts} we show the effects of adding restarts and nogood recording to the
-algorithm. With restarts and nogood recording,
-the random value-ordering heuristic comfortably beats the degree strategy with simple backtracking
-search. In other
-words, although having a good value-ordering heuristic is beneficial, introducing randomness into
-the search is better, if it is done alongside a mechanism to avoid heavy commitment to any
-particular random choice. \Cref{figure:scatter-random} looks at
-this in more detail. On the first plot, without nogood recording, unsatisfiable instances
-experience a slowdown of more than an order of magnitude, whilst on the second plot the
-performance remains unchanged (which shows that the overheads of nogoods and restarting are
-negligible). We therefore use nogood recording in all subsequent experiments with restarts. However,
-although better in aggregate on satisfiable instances, on a case by case basis many instances
-perform substantially worse with randomisation (either with or without restarts).
+In \cref{figure:old-vs-new} we show the effects of adding restarts and nogood recording to the
+algorithm. With restarts and nogood recording, the random value-ordering heuristic comfortably beats
+the degree strategy with simple backtracking search. In other words, although having a good
+value-ordering heuristic is beneficial, introducing randomness into the search is better, if it is
+done alongside a mechanism to avoid heavy commitment to any particular random choice.
+\Cref{figure:old-vs-new} also shows that the biased heuristic together with restarts is better
+still---that is, if we are introducing restarts, then it is better to add a small amount of
+randomness to a tailored heuristic than it is to throw away the heuristic altogether.  Indeed, the
+original algorithm can solve 1,983 satisfiable instances by 993.0 seconds, whilst the biased and
+random restarting algorithms require only 9.2 seconds and 33.8 seconds respectively to solve the
+same number.
 
-\Cref{figure:restarts} also shows that the biased heuristic together with restarts is better
-still---that is, if we are introducing restarts, then it is better to add a small amount of
-randomness to a tailored heuristic than it is to throw away the heuristic altogether.
-Indeed, the original algorithm can solve 1,983 satisfiable instances by 993.0 seconds, whilst the
-biased and random restarting algorithms require only 9.2 seconds and 33.8 seconds respectively to
-solve the same number.
-
-In the more detailed view in the final plot of \cref{figure:scatter-random}, comparing the
-basic algorithm with the degree heuristic to the degree-biased algorithm with restarts, all of the
+In the more detailed view in the right-hand plot of \cref{figure:old-vs-new}, comparing the basic
+algorithm with the degree heuristic to the degree-biased algorithm with restarts, all of the
 unsatisfiable instances are very close to the $x-y$ diagonal, showing that their performance is
@@
 instances that are more than a factor of ten times worse.  In other words, as well as improving
-performance, we have made up most of the consistency we lost by introducing randomness. This is a
-promising result: for the first time, we have introduced a viable alternative to backtracking search
-in a setting where we both have a strong \emph{value}-ordering heuristic, \emph{and} care about both
-satisfiable and unsatisfiable instances.
+performance, we have made up most of the consistency we lost by introducing randomness.
+
+As we might expect, these properties do not hold if any of the combination of changes are disabled.
+In the left-hand plot of \cref{figure:features}, we see large slowdowns on unsatisfiable instances
+when using a random value ordering and restarts but without nogood recording, and on the right-hand
+plot we see many more satisfiable instances above the $x-y$ diagonal when using the random
+value-ordering heuristic as opposed to the degree-biased heuristic.
+
+This is a promising result: for the first time, we have introduced a viable alternative to
+backtracking search in a setting where we both have a strong \emph{value}-ordering heuristic,
+\emph{and} care about both satisfiable and unsatisfiable instances.
 
 \section{Maximum Common Subgraph Algorithms}
+
+\begin{figure}[tb]
+    \centering
+    \includegraphics*{gen-graph-mcs.pdf}
+
+    \medskip
+
+    \includegraphics*{gen-graph-scatter-kdown.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-mcsplit.pdf}
+
+    \caption{Above, the cumulative number of maximum common induced subgraph instances solved by
+    k${\downarrow}$ and McSplit${\downarrow}$ over time, with and without restarts and nogood
+    recording. Below, comparing k${\downarrow}$ (left) and McSplit${\downarrow}$ (right) on an instance by instance
+    basis.}\label{figure:mcs}
+\end{figure}
 
@@
 Performance results from these two modified algorithms, using the same families of instances as in
-the previous section, are shown in \cref{figure:mcs,figure:kdown,figure:mcsplit}. Although we have
+the previous section, are shown in \cref{figure:mcs}. Although we have
 moved from a decision problem to an optimisation problem, the same changes remain clearly
@@
 on only seven.
-
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-mcs.pdf}
-    \caption{The cumulative number of maximum common induced subgraph instances solved by
-    k${\downarrow}$ and McSplit${\downarrow}$ over time, with and without restarts and nogood
-    recording.}\label{figure:mcs}
-\end{figure}
-
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-scatter-kdown.pdf}
-    \caption{Comparing k${\downarrow}$ with and without restarts and nogood recording on an instance
-    by instance basis.}\label{figure:kdown}
-\end{figure}
-
-\begin{figure}[p]
-    \centering
-    \includegraphics*{gen-graph-scatter-mcsplit.pdf}
-    \caption{Comparing McSplit${\downarrow}$ with and without restarts and nogood recording on an instance
-    by instance basis.}\label{figure:mcsplit}
-\end{figure}
 
@@
 
-% The authors would like to thank Ruth Hoffmann, Christophe Lecoutre, Christine Solnon and Craig
-% Reilly for their comments.  This work was supported by the Engineering and Physical Sciences
-% Research Council [grant numbers EP/026842/1 and EP/M508056/1].
+The authors would like to thank Ruth Hoffmann, Christophe Lecoutre, Christine Solnon and Craig
+Reilly for their comments.
 
-The authors would like to thank Anonymous, Anonymous, Anonymous and Anonymous for their comments.
-This work was supported by an anonymous funding agency.
-
-\bibliographystyle{aaai}
+\bibliographystyle{splncs04}
 \bibliography{dblp}

### 2018-11-10 paper/paper.tex
Start bringing parallel in

@@
 \usepackage{amsmath}
-\usepackage[ruled,vlined]{algorithm2e}
+% \usepackage[ruled,vlined]{algorithm2e}
 \usepackage{tikz}
@@
 
-\crefname{algocf}{Algorithm}{Algorithms}
-\Crefname{algocf}{Algorithm}{Algorithms}
+% \crefname{algocf}{Algorithm}{Algorithms}
+% \Crefname{algocf}{Algorithm}{Algorithms}
 
@@
 \title{Beating Simple Backtracking Search in Subgraph Algorithms\thanks{This work was
-supported by the Engineering and Physical Sciences Research Council [grant numbers EP/026842/1 and
-EP/M508056/1].}}
+supported by the Engineering and Physical Sciences Research Council (grant numbers EP/026842/1 and
+EP/M508056/1). This work used the Cirrus UK National Tier-2 HPC Service at EPCC
+(http://www.cirrus.ac.uk) funded by the University of Edinburgh and EPSRC (EP/P020267/1).}}
 
@@
 \begin{abstract}
-    Modern subgraph isomorphism solvers use degree as a value-ordering heuristics to direct
-    backtracking search. This makes a heavy commitment to the first branching choice,
-    which is often incorrect. We evaluate discrepancy-based searches as an alternative, but find
-    them too expensive to be effective in this situation, even on satisfiable instances. We then
-    demonstrate that a random value-ordering heuristic with restarts beats conventional search with
-    the best-known tailored heuristic. This motivates the introduction of a novel slightly-random
-    heuristic with similar behaviour to the tailored heuristic: when combined with restarts and
-    nogood recording, this new search strategy is over a hundred times more effective on satisfiable
-    instances, and does not affect performance on unsatisfiable instances.  Finally, we show that
-    this strategy also improves two maximum common induced subgraph algorithms.
+    The current state of the art in subgraph isomorphism solving involves using degree as a
+    value-ordering heuristics to direct backtracking search. Such a search makes a heavy commitment
+    to the first branching choice, which is often incorrect. We evaluate discrepancy-based searches
+    as an alternative, but find them too expensive to be effective in this situation, even on
+    satisfiable instances. We then demonstrate that a random value-ordering heuristic with restarts
+    beats conventional search with the best-known tailored heuristic. This motivates the
+    introduction of a novel slightly-random heuristic with similar behaviour to the tailored
+    heuristic: when combined with restarts and nogood recording, this new search strategy is over ??
+    times more effective on satisfiable instances, and does not affect performance on
+    unsatisfiable instances. Such a search can also be parallelised in a way which is much
+    simpler than conventional work-stealing: we obtain further speedups of ??  on ??  cores, and
+    ?? over ?? distributed-memory hosts. Finally, we show that this search strategy also improves
+    two maximum common induced subgraph algorithms.
 \end{abstract}
@@
 on a large number of satisfiable instances, whilst performing worse only rarely on satisfiable
-instances, and never on unsatisfiable instances. Finally, we show that this strategy is also
-effective in an optimisation setting, producing benefits in two maximum common induced subgraph
-algorithms. This combination of techniques gives us, for the first time, a practical alternative to
-backtracking search for constraint programming style algorithms where we have a strong
-value-ordering heuristic and where we care both about satisfiable and unsatisfiable instances.
+instances, and never on unsatisfiable instances.
 
-\section{Definitions and Algorithms}
+?? Parallel
 
-Let $G$ be a graph. We denote its set of vertices by $\vertexset(G)$, and write $v \sim_G w$ to mean
-that vertices $v$ and $w$ are adjacent; we allow a vertex to be adjacent to itself (a \emph{loop}). The
-\emph{neighbourhood} of a vertex, $\neighbourhood_G(v)$, is the set of vertices adjacent to $v$, and
-the \emph{degree} of $v$, $\deg_G(v)$, is the cardinality of its neighbourhood. When the graph is clear from the context, we omit
-the $G$ subscripts. As per \citet{Hoffmann et al.\ }{DBLP:conf/aaai/HoffmannMR17}, $G^{n,\ell}$ denotes the graph with
-vertex set $\vertexset(G)$, and with edges between vertices $v$ and $w$ if there are at least $n$
-simple paths of length $\ell$ between vertices $v$ and $w$ in $G$.
+Finally, we show that this strategy is also effective in an optimisation setting, producing benefits
+in two maximum common induced subgraph algorithms. This combination of techniques gives us, for the
+first time, a practical alternative to backtracking search for constraint programming style
+algorithms where we have a strong value-ordering heuristic and where we care both about satisfiable
+and unsatisfiable instances. ?? Parallel
+
+\subsection{Definitions}
+
+% Let $G$ be a graph. We denote its set of vertices by $\vertexset(G)$, and write $v \sim_G w$ to mean
+% that vertices $v$ and $w$ are adjacent; we allow a vertex to be adjacent to itself (a \emph{loop}). The
+% \emph{neighbourhood} of a vertex, $\neighbourhood_G(v)$, is the set of vertices adjacent to $v$, and
+% the \emph{degree} of $v$, $\deg_G(v)$, is the cardinality of its neighbourhood. When the graph is clear from the context, we omit
+% the $G$ subscripts. As per \citet{Hoffmann et al.\ }{DBLP:conf/aaai/HoffmannMR17}, $G^{n,\ell}$ denotes the graph with
+% vertex set $\vertexset(G)$, and with edges between vertices $v$ and $w$ if there are at least $n$
+% simple paths of length $\ell$ between vertices $v$ and $w$ in $G$.
 
 The \emph{non-induced subgraph isomorphism problem} is to find an injective mapping from the
-vertices of a pattern graph $\mathcal{P}$ to a target graph $\mathcal{T}$, such that adjacent
-vertices in $\mathcal{P}$ are mapped to adjacent vertices in $\mathcal{T}$ (including that vertices
-with loops in $\mathcal{P}$ may only be mapped to vertices with loops in $\mathcal{T}$). The
-\emph{induced} problem additionally requires that non-adjacent vertices are mapped to non-adjacent
-vertices. We write $v \mapsto v'$ to mean that pattern vertex $v$ is mapped to target $v'$ under
-either kind of mapping. Finally, we define a function $\FuncSty{compatible}(v, w)$ to mean that
-mapping $v \mapsto w$ is feasible based upon both the loop rule, and the neighbourhood degree
-sequence relation of \citet{Zampelli et al.\ }{DBLP:journals/constraints/ZampelliDS10}.
+vertices of a \emph{pattern} graph $\mathcal{P}$ to a \emph{target} graph $\mathcal{T}$, such that
+adjacent vertices in $\mathcal{P}$ are mapped to adjacent vertices in $\mathcal{T}$ (including that
+vertices with loops in $\mathcal{P}$ may only be mapped to vertices with loops in $\mathcal{T}$).
+The \emph{induced} problem additionally requires that non-adjacent vertices are mapped to
+non-adjacent vertices.
 
-\subsection{A Subgraph Isomorphism Algorithm}
+?? Applications
 
-We now briefly outline the non-shaded parts of \cref{algorithm:sip}, which is how we will solve the
-subgraph isomorphism problem. This algorithm is very closely based upon the $k{\downarrow}$
-algorithm of \citet{Hoffmann et al.\ }{DBLP:conf/aaai/HoffmannMR17} with $k = 0$, and we refer the reader to that paper
-for full technical details; that algorithm, in turn, is a simplification of the Glasgow2 algorithm
-\cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}. We show the non-induced problem; the
-induced problem may be solved by including the loop-complements of $\mathcal{P}$ and $\mathcal{T}$
-in $L$ on \siplineref{supplemental}.
+% We write $v \mapsto v'$ to mean that pattern vertex $v$ is mapped to target $v'$ under
+% either kind of mapping. Finally, we define a function $\FuncSty{compatible}(v, w)$ to mean that
+% mapping $v \mapsto w$ is feasible based upon both the loop rule, and the neighbourhood degree
+% sequence relation of \citet{Zampelli et al.\ }{DBLP:journals/constraints/ZampelliDS10}.
 
-\Cref{algorithm:sip} works by recursively building up a valid mapping of pattern vertices to target
-vertices. For each pattern vertex $v$, we have a domain $D_v$ containing a set of feasible target
-vertices. Each domain is initialised at the top of search (\siplinerangeref{initstart}{initend}),
-using the invariants described by \citet{Zampelli et al.\ }{DBLP:journals/constraints/ZampelliDS10} and
-\citet{Hoffmann et al.\ }{DBLP:conf/aaai/HoffmannMR17} to restrict its initial values.
+\subsection{The Glasgow Subgraph Solver}
 
-The $\FuncSty{search}$ procedure recursively selects an unassigned domain $D$
-(\siplineref{variableordering}), and tries assigning
-it each of its remaining values in turn (\siplineref{valueordering}). The effects of this assignment
-are propagated (\siplineref{dopropagate}). If an
-inconsistency is detected, we continue and try another value; otherwise, we recurse
-(\siplineref{recurse}) until a solution is found. Finally, if we run out of values, we backtrack
-(\siplineref{backtrack}).
+This paper looks at improving the Glasgow Subgraph
+Solver\footnote{\url{https://github.com/ciaranm/glasgow-subgraph-solver/}}, which can solve both the
+non-induced and the induced subgraph isomorphism problems.  The solver is very closely based upon
+the $k{\downarrow}$ algorithm of \citet{Hoffmann et al.\ }{DBLP:conf/aaai/HoffmannMR17} with $k =
+0$, and we refer the reader to that paper for full technical details; that algorithm, in turn, is a
+simplification and re-engineering of an older Glasgow algorithm
+\cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}.  Essentially, the solver is essentially
+a dedicated forward-checking constraint programming implementation specifically for subgraph
+problems. It works with a model having a variable per pattern graph vertex, with domains ranging
+over the target graph vertices, and performs a backtracking search to map each pattern vertex to a
+target vertex whilst propagating adjacency and injectivity constraints (together with further
+implied constraints based upon degrees and paths).  However, it uses specialised bit-parallel data
+structures and algorithms, and propagates constraints in a fixed order rather than using a queue.
 
-Our $\FuncSty{propagate}$ routine is also close to that of \citet{Hoffmann et al.\ }{DBLP:conf/aaai/HoffmannMR17},
-differing only in that we carry out the filtering for injectivity
-(\siplinerangeref{alldiffstart}{alldiffend}) after every assignment rather than once at the end of
-the unit propagation loop. (We make this change because the all-different propagator used is not
-idempotent, and we wish to avoid an odd interaction when we introduce nogood filtering.)
+% We now briefly outline the non-shaded parts of \cref{algorithm:sip}, which is how we will solve the
+% subgraph isomorphism problem. This algorithm is very closely based upon the $k{\downarrow}$
+% algorithm of \citet{Hoffmann et al.\ }{DBLP:conf/aaai/HoffmannMR17} with $k = 0$, and we refer the reader to that paper
+% for full technical details; that algorithm, in turn, is a simplification of the Glasgow2 algorithm
+% \cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}. We show the non-induced problem; the
+% induced problem may be solved by including the loop-complements of $\mathcal{P}$ and $\mathcal{T}$
+% in $L$ on \siplineref{supplemental}.
+% 
+% \Cref{algorithm:sip} works by recursively building up a valid mapping of pattern vertices to target
+% vertices. For each pattern vertex $v$, we have a domain $D_v$ containing a set of feasible target
+% vertices. Each domain is initialised at the top of search (\siplinerangeref{initstart}{initend}),
+% using the invariants described by \citet{Zampelli et al.\ }{DBLP:journals/constraints/ZampelliDS10} and
+% \citet{Hoffmann et al.\ }{DBLP:conf/aaai/HoffmannMR17} to restrict its initial values.
+% 
+% The $\FuncSty{search}$ procedure recursively selects an unassigned domain $D$
+% (\siplineref{variableordering}), and tries assigning
+% it each of its remaining values in turn (\siplineref{valueordering}). The effects of this assignment
+% are propagated (\siplineref{dopropagate}). If an
+% inconsistency is detected, we continue and try another value; otherwise, we recurse
+% (\siplineref{recurse}) until a solution is found. Finally, if we run out of values, we backtrack
+% (\siplineref{backtrack}).
+% 
+% Our $\FuncSty{propagate}$ routine is also close to that of \citet{Hoffmann et al.\ }{DBLP:conf/aaai/HoffmannMR17},
+% differing only in that we carry out the filtering for injectivity
+% (\siplinerangeref{alldiffstart}{alldiffend}) after every assignment rather than once at the end of
+% the unit propagation loop. (We make this change because the all-different propagator used is not
+% idempotent, and we wish to avoid an odd interaction when we introduce nogood filtering.)
 
-\begin{algorithm}[p]\DontPrintSemicolon\small\SetInd{0.1em}{1em}
-    \begin{tikzpicture}[remember picture,overlay]
-        \coordinate (restart1sc) at ($(pic cs:restart1s) + (0, 0.15)$);
-        \coordinate (restart1ec) at ($(pic cs:restart1e) + (0, 0.00)$);
-        \node [fill=highlight, rounded corners, fit=(restart1sc) (restart1ec)] { };
-        \coordinate (restart2sc) at ($(pic cs:restart2s) + (0, 0.15)$);
-        \coordinate (restart2ec) at ($(pic cs:restart2e) + (0, 0.00)$);
-        \node [fill=highlight, rounded corners, fit=(restart2sc) (restart2ec)] { };
-        \coordinate (restart9sc) at ($(pic cs:restart9s) + (0, 0.15)$);
-        \coordinate (restart9ec) at ($(pic cs:restart9e) + (-0.15, 0.00)$);
-        \node [fill=highlight, rounded corners, fit=(restart9sc) (restart9ec)] { };
-    \end{tikzpicture}
+% \begin{algorithm}[p]\DontPrintSemicolon\small\SetInd{0.1em}{1em}
+%     \begin{tikzpicture}[remember picture,overlay]
+%         \coordinate (restart1sc) at ($(pic cs:restart1s) + (0, 0.15)$);
+%         \coordinate (restart1ec) at ($(pic cs:restart1e) + (0, 0.00)$);
+%         \node [fill=highlight, rounded corners, fit=(restart1sc) (restart1ec)] { };
+%         \coordinate (restart2sc) at ($(pic cs:restart2s) + (0, 0.15)$);
+%         \coordinate (restart2ec) at ($(pic cs:restart2e) + (0, 0.00)$);
+%         \node [fill=highlight, rounded corners, fit=(restart2sc) (restart2ec)] { };
+%         \coordinate (restart9sc) at ($(pic cs:restart9s) + (0, 0.15)$);
+%         \coordinate (restart9ec) at ($(pic cs:restart9e) + (-0.15, 0.00)$);
+%         \node [fill=highlight, rounded corners, fit=(restart9sc) (restart9ec)] { };
+%     \end{tikzpicture}
+% 
+%     \nl $\FuncSty{subgraphIsomorphism}$ (Graph $\mathcal{P}$, Graph $\mathcal{T}$) $\rightarrow$ Bool \;
+% \nl \Begin{
+%     \nl \lIf{$\left|\vertexset(\mathcal{P})\right| >
+%     \left|\vertexset(\mathcal{T})\right|$}{$\KwSty{return}~\KwSty{false}$\siplabel{enough}}
+%     \nl Discard isolated vertices in $\mathcal{P}$\siplabel{isolated} \;
+%     \nl \KwSty{global} $L \gets \big[ (\mathcal{P},\ \mathcal{T}), \siplabel{supplemental}
+%         (\mathcal{P}^{1,2}, \mathcal{T}^{1,2}),
+%         (\mathcal{P}^{2,2},\ \mathcal{T}^{2,2}),$ \\
+%         $\hspace*{3em}(\mathcal{P}^{3,2},\ \mathcal{T}^{3,2}),
+%         (\mathcal{P}^{4,2},\ \mathcal{T}^{4,2}) \big]$ \;
+%     \nl \ForEach{\siplabel{initstart}$v \in \vertexset(\mathcal{P})$}{
+%         \nl $D_v \gets \vertexset(\mathcal{T})$\;
+%         \nl \ForEach{$(P,\,T) \in L$}{\nl$D_v \gets \{ w \in D_v : \FuncSty{compatible}(v, w)\}$\siplabel{initend}
+%         }
+%     }
+%     \nl \While{$\KwSty{true}$\siplabel{restartsloop}}{
+%         \nl \lIf{$\KwSty{not}~\FuncSty{propagate}(D)$}{$\KwSty{return}~\KwSty{false}$}
+%         \nl \tikzmark{restart1s}$\KwSty{global}~(R, R') \gets (0, \textnormal{the next Luby sequence value})$\tikzmark{restart1e} \;
+%         \nl $S \gets \FuncSty{search}(\{ E \in D : \left|E\right| > 1 \}, \top)$ \;
+%         \nl \tikzmark{restart9s}\lIf{$S \ne \KwSty{restart}$}{\tikzmark{restart9e}$\KwSty{return}~S$}
+%         \nl \tikzmark{restart2s}trigger nogoods containing only one clause\tikzmark{restart2e} \;
+%     }
+% }
+% \caption{The main subgraph isomorphism algorithm. The shaded code is added for restarting. See
+%     \cref{algorithm:sip:search,algorithm:sip:propagate} for the search and propagation routines.}
+% \label{algorithm:sip}
+% \end{algorithm}
+% 
+% \begin{algorithm}[p]\DontPrintSemicolon\small\SetInd{0.1em}{1em}
+%     \begin{tikzpicture}[remember picture,overlay]
+%         \coordinate (restart3sc) at ($(pic cs:restart3s) + (0, 0.15)$);
+%         \coordinate (restart3ec) at ($(pic cs:restart3e) + (0, 0.00)$);
+%         \node [fill=highlight, rounded corners, fit=(restart3sc) (restart3ec)] { };
+%         \coordinate (restart4sc) at ($(pic cs:restart4s) + (0, 0.15)$);
+%         \coordinate (restart4ec) at ($(pic cs:restart4e) + (0, 0.00)$);
+%         \node [fill=highlight, rounded corners, fit=(restart4sc) (restart4ec)] { };
+%         \coordinate (restart5sc) at ($(pic cs:restart5s) + (0, 0.15)$);
+%         \coordinate (restart5ec) at ($(pic cs:restart5e) - (0, 0.10)$);
+%         \coordinate (restart5rc) at ($(pic cs:restart5r) + (0, 0.00)$);
+%         \node [fill=highlight, rounded corners, fit=(restart5sc) (restart5ec) (restart5rc)] { };
+%         \coordinate (restart6sc) at ($(pic cs:restart6s) + (0, 0.15)$);
+%         \coordinate (restart6ec) at ($(pic cs:restart6e) + (0, 0.00)$);
+%         \coordinate (restart6rc) at ($(pic cs:restart6r) + (0.8, 0.00)$);
+%         \node [fill=highlight, rounded corners, fit=(restart6sc) (restart6ec) (restart6rc)] { };
+%         \coordinate (restart8sc) at ($(pic cs:restart8s) + (0.05, 0.15)$);
+%         \coordinate (restart8ec) at ($(pic cs:restart8e) + (0, 0.00)$);
+%         \node [fill=highlight, rounded corners, fit=(restart8sc) (restart8ec)] { };
+%     \end{tikzpicture}
+%     \nl $\FuncSty{search}$ (Domains $D$, Trail $B$) $\rightarrow$ Bool
+%     \tikzmark{restart8s}\KwSty{or} \KwSty{restart}\tikzmark{restart8e} \;
+% \nl \Begin{
+%     \nl \lIf{$D = \emptyset$}{$\KwSty{return}~\KwSty{true}$}
+%     \nl $D_v \gets \textnormal{a domain from}~D~\textnormal{chosen by variable heuristic}$\siplabel{variableordering} \;
+%     \nl \tikzmark{restart3s}$C \gets \emptyset$\tikzmark{restart3e} \;
+%     \nl \ForEach {$v' \in D_v~\textnormal{ordered by value heuristic}$\siplabel{valueordering}}{
+%         \nl \tikzmark{restart4s}$C \gets C \cup \{ v' \}$\tikzmark{restart4e} \;
+%         \nl $D' \gets \FuncSty{clone}(D)$ \;
+%         \nl $D'_v \gets \{ v' \} $ \;
+%         \nl \If{\FuncSty{propagate}(D')\siplabel{dopropagate}}{
+%             \nl $S \gets \FuncSty{search}(\{ E \in D'\,{:}\,\left|E\right|{>}\,1 \}, B \wedge (v\,{\mapsto}\,v'))$\siplabel{recurse}\;
+%             \nl \lIf{$S = \KwSty{true}$}{$\KwSty{return}~\KwSty{true}$}
+%             \nl \tikzmark{restart5s}\ElseIf{$S = \KwSty{restart}$}{
+%                 \nl \ForEach{$c \in C$}{\nl$\textnormal{post nogood}~B \wedge (v \mapsto c) \Rightarrow \bot$\tikzmark{restart5r}}
+%                 \nl $\KwSty{return}~\KwSty{restart}$\tikzmark{restart5e} \;
+%             }
+%         }
+%     }
+%     \nl \tikzmark{restart6s}\If{$(R \gets R + 1) = R'$\tikzmark{restart6r}}{
+%         \nl post nogood $B$ \;
+%         \nl $\KwSty{return}~\KwSty{restart}\tikzmark{restart6e}$\siplabel{countrestarts}}
+%     \nl $\KwSty{return}~\KwSty{false}$\siplabel{backtrack}
+% }
+% \caption{Search routine for \cref{algorithm:sip}.}
+% \label{algorithm:sip:search}
+% \end{algorithm}
+% 
+% \begin{algorithm}[tb]\DontPrintSemicolon\small\SetInd{0.1em}{1em}
+%     \begin{tikzpicture}[remember picture,overlay]
+%         \coordinate (restart7sc) at ($(pic cs:restart7s) + (0, 0.15)$);
+%         \coordinate (restart7ec) at ($(pic cs:restart7e) + (0, 0.00)$);
+%         \node [fill=highlight, rounded corners, fit=(restart7sc) (restart7ec)] { };
+%     \end{tikzpicture}
+% \nl $\FuncSty{propagate}$ (Domains $D$) $\rightarrow$ Bool \;
+% \nl \Begin{
+%     \nl \While{$D_v \gets $\textnormal{a unit domain from}~$D$}{
+%         \nl $v' \gets \textnormal{the single value in}~D_v$ \;
+%         \nl \tikzmark{restart7s}trigger nogoods with watch $(v \mapsto v')$\tikzmark{restart7e}\siplabel{2wl} \;
+%         \nl \ForEach{$D_w \in D \setminus \{ D_v$ \}}{
+%             \nl $D_w \gets D_w \setminus \{ v' \}$ \siplabel{removev} \;
+%             \nl \ForEach{$(P,\,T) \in L~\KwSty{where}~v \sim_P w$}{\nl$D_w \gets D_w \cap \neighbourhood_T(v')$}
+%             \nl \lIf{$D_w = \emptyset$}{$\KwSty{return}~\KwSty{false}$}
+%         }
+% \nl $(H,\,A,\,n) \gets (\emptyset,\,\emptyset,\,0)$\siplabel{alldiffstart} \;
+% \nl \ForEach{$D_v \in D$ \textnormal{in ascending cardinality order\siplabel{eachdomain}}}{
+%     \nl $(D_v, A,\,n) \gets (D_v \gets D_v \setminus H\siplabel{elimhall}, A \cup D_v,\,n + 1)$ \siplabel{acc} \;
+%     \nl \lIf{$D_v = \emptyset~\vee~|A| < n$}{$\KwSty{return}~\KwSty{false}$\siplabel{failhall}}
+%     \nl \lIf{$|A| = n$}{$(H,\,A,\,n) \gets (H \cup A,\,\emptyset,\,0)$\siplabel{hall}}\siplabel{alldiffend}
+% }
+%     }
+% \nl $\KwSty{return}~\KwSty{true}$ \;
+% }
+% \caption{Propagation routine for \cref{algorithm:sip}.}
+% \label{algorithm:sip:propagate}
+% \end{algorithm}
 
-    \nl $\FuncSty{subgraphIsomorphism}$ (Graph $\mathcal{P}$, Graph $\mathcal{T}$) $\rightarrow$ Bool \;
-\nl \Begin{
-    \nl \lIf{$\left|\vertexset(\mathcal{P})\right| >
-    \left|\vertexset(\mathcal{T})\right|$}{$\KwSty{return}~\KwSty{false}$\siplabel{enough}}
-    \nl Discard isolated vertices in $\mathcal{P}$\siplabel{isolated} \;
-    \nl \KwSty{global} $L \gets \big[ (\mathcal{P},\ \mathcal{T}), \siplabel{supplemental}
-        (\mathcal{P}^{1,2}, \mathcal{T}^{1,2}),
-        (\mathcal{P}^{2,2},\ \mathcal{T}^{2,2}),$ \\
-        $\hspace*{3em}(\mathcal{P}^{3,2},\ \mathcal{T}^{3,2}),
-        (\mathcal{P}^{4,2},\ \mathcal{T}^{4,2}) \big]$ \;
-    \nl \ForEach{\siplabel{initstart}$v \in \vertexset(\mathcal{P})$}{
-        \nl $D_v \gets \vertexset(\mathcal{T})$\;
-        \nl \ForEach{$(P,\,T) \in L$}{\nl$D_v \gets \{ w \in D_v : \FuncSty{compatible}(v, w)\}$\siplabel{initend}
-        }
-    }
-    \nl \While{$\KwSty{true}$\siplabel{restartsloop}}{
-        \nl \lIf{$\KwSty{not}~\FuncSty{propagate}(D)$}{$\KwSty{return}~\KwSty{false}$}
-        \nl \tikzmark{restart1s}$\KwSty{global}~(R, R') \gets (0, \textnormal{the next Luby sequence value})$\tikzmark{restart1e} \;
-        \nl $S \gets \FuncSty{search}(\{ E \in D : \left|E\right| > 1 \}, \top)$ \;
-        \nl \tikzmark{restart9s}\lIf{$S \ne \KwSty{restart}$}{\tikzmark{restart9e}$\KwSty{return}~S$}
-        \nl \tikzmark{restart2s}trigger nogoods containing only one clause\tikzmark{restart2e} \;
-    }
-}
-\caption{The main subgraph isomorphism algorithm. The shaded code is added for restarting. See
-    \cref{algorithm:sip:search,algorithm:sip:propagate} for the search and propagation routines.}
-\label{algorithm:sip}
-\end{algorithm}
+% For the variable-ordering heuristic on \siplineref{variableordering} of
+% \cref{algorithm:sip}, we use smallest domain first, tiebreaking on highest degree
+% \cite{DBLP:journals/ai/HaralickE80,DBLP:journals/jair/McCreeshPST18}. Variable ordering is not the
+% focus of this paper---as far as possible, we use this same dynamic ordering throughout the paper, so
+% as not to affect the performance of the algorithm on unsatisfiable instances.  For value ordering,
+% which occurs on \siplineref{valueordering} of \cref{algorithm:sip}, we will initially select
+% vertices from highest degree to lowest degree; value ordering is the major theme of the remainder of
+% this paper. Note that the variable-ordering heuristic affects both unsatisfiable and satisfiable
+% instances, whilst the value-ordering heuristic is relevant only for satisfiable instances.
 
-\begin{algorithm}[p]\DontPrintSemicolon\small\SetInd{0.1em}{1em}
-    \begin{tikzpicture}[remember picture,overlay]
-        \coordinate (restart3sc) at ($(pic cs:restart3s) + (0, 0.15)$);
-        \coordinate (restart3ec) at ($(pic cs:restart3e) + (0, 0.00)$);
-        \node [fill=highlight, rounded corners, fit=(restart3sc) (restart3ec)] { };
-        \coordinate (restart4sc) at ($(pic cs:restart4s) + (0, 0.15)$);
-        \coordinate (restart4ec) at ($(pic cs:restart4e) + (0, 0.00)$);
-        \node [fill=highlight, rounded corners, fit=(restart4sc) (restart4ec)] { };
-        \coordinate (restart5sc) at ($(pic cs:restart5s) + (0, 0.15)$);
-        \coordinate (restart5ec) at ($(pic cs:restart5e) - (0, 0.10)$);
-        \coordinate (restart5rc) at ($(pic cs:restart5r) + (0, 0.00)$);
-        \node [fill=highlight, rounded corners, fit=(restart5sc) (restart5ec) (restart5rc)] { };
-        \coordinate (restart6sc) at ($(pic cs:restart6s) + (0, 0.15)$);
-        \coordinate (restart6ec) at ($(pic cs:restart6e) + (0, 0.00)$);
-        \coordinate (restart6rc) at ($(pic cs:restart6r) + (0.8, 0.00)$);
-        \node [fill=highlight, rounded corners, fit=(restart6sc) (restart6ec) (restart6rc)] { };
-        \coordinate (restart8sc) at ($(pic cs:restart8s) + (0.05, 0.15)$);
-        \coordinate (restart8ec) at ($(pic cs:restart8e) + (0, 0.00)$);
-        \node [fill=highlight, rounded corners, fit=(restart8sc) (restart8ec)] { };
-    \end{tikzpicture}
-    \nl $\FuncSty{search}$ (Domains $D$, Trail $B$) $\rightarrow$ Bool
-    \tikzmark{restart8s}\KwSty{or} \KwSty{restart}\tikzmark{restart8e} \;
-\nl \Begin{
-    \nl \lIf{$D = \emptyset$}{$\KwSty{return}~\KwSty{true}$}
-    \nl $D_v \gets \textnormal{a domain from}~D~\textnormal{chosen by variable heuristic}$\siplabel{variableordering} \;
-    \nl \tikzmark{restart3s}$C \gets \emptyset$\tikzmark{restart3e} \;
-    \nl \ForEach {$v' \in D_v~\textnormal{ordered by value heuristic}$\siplabel{valueordering}}{
-        \nl \tikzmark{restart4s}$C \gets C \cup \{ v' \}$\tikzmark{restart4e} \;
-        \nl $D' \gets \FuncSty{clone}(D)$ \;
-        \nl $D'_v \gets \{ v' \} $ \;
-        \nl \If{\FuncSty{propagate}(D')\siplabel{dopropagate}}{
-            \nl $S \gets \FuncSty{search}(\{ E \in D'\,{:}\,\left|E\right|{>}\,1 \}, B \wedge (v\,{\mapsto}\,v'))$\siplabel{recurse}\;
-            \nl \lIf{$S = \KwSty{true}$}{$\KwSty{return}~\KwSty{true}$}
-            \nl \tikzmark{restart5s}\ElseIf{$S = \KwSty{restart}$}{
-                \nl \ForEach{$c \in C$}{\nl$\textnormal{post nogood}~B \wedge (v \mapsto c) \Rightarrow \bot$\tikzmark{restart5r}}
-                \nl $\KwSty{return}~\KwSty{restart}$\tikzmark{restart5e} \;
-            }
-        }
-    }
-    \nl \tikzmark{restart6s}\If{$(R \gets R + 1) = R'$\tikzmark{restart6r}}{
-        \nl post nogood $B$ \;
-        \nl $\KwSty{return}~\KwSty{restart}\tikzmark{restart6e}$\siplabel{countrestarts}}
-    \nl $\KwSty{return}~\KwSty{false}$\siplabel{backtrack}
-}
-\caption{Search routine for \cref{algorithm:sip}.}
-\label{algorithm:sip:search}
-\end{algorithm}
-
-\begin{algorithm}[tb]\DontPrintSemicolon\small\SetInd{0.1em}{1em}
-    \begin{tikzpicture}[remember picture,overlay]
-        \coordinate (restart7sc) at ($(pic cs:restart7s) + (0, 0.15)$);
-        \coordinate (restart7ec) at ($(pic cs:restart7e) + (0, 0.00)$);
-        \node [fill=highlight, rounded corners, fit=(restart7sc) (restart7ec)] { };
-    \end{tikzpicture}
-\nl $\FuncSty{propagate}$ (Domains $D$) $\rightarrow$ Bool \;
-\nl \Begin{
-    \nl \While{$D_v \gets $\textnormal{a unit domain from}~$D$}{
-        \nl $v' \gets \textnormal{the single value in}~D_v$ \;
-        \nl \tikzmark{restart7s}trigger nogoods with watch $(v \mapsto v')$\tikzmark{restart7e}\siplabel{2wl} \;
-        \nl \ForEach{$D_w \in D \setminus \{ D_v$ \}}{
-            \nl $D_w \gets D_w \setminus \{ v' \}$ \siplabel{removev} \;
-            \nl \ForEach{$(P,\,T) \in L~\KwSty{where}~v \sim_P w$}{\nl$D_w \gets D_w \cap \neighbourhood_T(v')$}
-            \nl \lIf{$D_w = \emptyset$}{$\KwSty{return}~\KwSty{false}$}
-        }
-\nl $(H,\,A,\,n) \gets (\emptyset,\,\emptyset,\,0)$\siplabel{alldiffstart} \;
-\nl \ForEach{$D_v \in D$ \textnormal{in ascending cardinality order\siplabel{eachdomain}}}{
-    \nl $(D_v, A,\,n) \gets (D_v \gets D_v \setminus H\siplabel{elimhall}, A \cup D_v,\,n + 1)$ \siplabel{acc} \;
-    \nl \lIf{$D_v = \emptyset~\vee~|A| < n$}{$\KwSty{return}~\KwSty{false}$\siplabel{failhall}}
-    \nl \lIf{$|A| = n$}{$(H,\,A,\,n) \gets (H \cup A,\,\emptyset,\,0)$\siplabel{hall}}\siplabel{alldiffend}
-}
-    }
-\nl $\KwSty{return}~\KwSty{true}$ \;
-}
-\caption{Propagation routine for \cref{algorithm:sip}.}
-\label{algorithm:sip:propagate}
-\end{algorithm}
-
-For the variable-ordering heuristic on \siplineref{variableordering} of
-\cref{algorithm:sip}, we use smallest domain first, tiebreaking on highest degree
-\cite{DBLP:journals/ai/HaralickE80,DBLP:journals/jair/McCreeshPST18}. Variable ordering is not the
-focus of this paper---as far as possible, we use this same dynamic ordering throughout the paper, so
-as not to affect the performance of the algorithm on unsatisfiable instances.  For value ordering,
-which occurs on \siplineref{valueordering} of \cref{algorithm:sip}, we will initially select
-vertices from highest degree to lowest degree; value ordering is the major theme of the remainder of
-this paper. Note that the variable-ordering heuristic affects both unsatisfiable and satisfiable
-instances, whilst the value-ordering heuristic is relevant only for satisfiable instances.
-
-\section{Empirical Evaluation}
+\subsection{Experimental Setup}
 
 Our experiments are performed on systems with dual Intel Xeon E5-2697A v4 CPUs and 512GBytes RAM,
-running Ubuntu 17.04. We
-implemented\footnote{DOI removed for anonymous review}
-\cref{algorithm:sip} in C++ using the bit-parallel implementation of
-\citet{Hoffmann et al.\ }{DBLP:conf/aaai/HoffmannMR17} as a starting point, and compiled it using GCC 7.2.0. We use a
-deterministic pseudo-random number generator for reproducibility.
-
-We use the dataset introduced by \citet{Kotthoff et al.\ }{DBLP:conf/lion/KotthoffMS16} for evaluation. This dataset
-brings together a range of randomly-generated and application instance families from earlier papers:
+running Ubuntu 17.04, using GCC 7.2.0 as the compiler. We use a deterministic pseudo-random number
+generator for reproducibility.  We use the dataset introduced by \citet{Kotthoff et al.\
+}{DBLP:conf/lion/KotthoffMS16} for evaluation. This dataset brings together a range of
+randomly-generated and application instance families from earlier papers:
 
@@
 
-\subsection{Baseline Performance}
+\section{Improving Sequential Search}
 
@@
 
-We begin by verifying that our implementation of the basic \cref{algorithm:sip} is competitive. In
-the first and third plots \cref{figure:others} we show the cumulative number of instances solved
-over time for the non-induced and induced problems respectively. We compare our implementation (both
-with the degree heuristic, and with the modifications described in the remainder of this paper), the
-Glasgow2 and Glasgow3 algorithms from which \cref{algorithm:sip} is derived
-\cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}, the PathLAD variation of the LAD
-algorithm \cite{DBLP:journals/ai/Solnon10,DBLP:conf/lion/KotthoffMS16}, VF2
+We begin by verifying that the Glasgow Subgraph Solver is the state of the art. In the first and
+third plots \cref{figure:others} we show the cumulative number of instances solved over time for the
+non-induced and induced problems respectively. We compare the Glasgow Subgraph Solver (both
+unmodified, and with the modifications described in the remainder of this paper), the PathLAD
+variation of the LAD algorithm \cite{DBLP:journals/ai/Solnon10,DBLP:conf/lion/KotthoffMS16}, VF2
 \cite{DBLP:journals/pami/CordellaFSV04}, RI \cite{DBLP:journals/bmcbi/BonniciGPSF13}, and VF3
@@
 be to allow the rightmost algorithm to succeed on $y$ out of the 14,621 instances (bearing in mind
-that the two sets of $y$ instances could be different).
-
-As expected, in both cases the performance of \cref{algorithm:sip} with the degree heuristic is very
-close to that of Glasgow2, although we perform better at lower runtimes because we use simpler
-preprocessing at the top of search. Our implementation also clearly beats PathLAD, VF2, VF3 and RI,
-except for very low choices of timeout.
+that the two sets of $y$ instances could be different).  The plots show that our starting point
+comfortably beats PathLAD, VF2, VF3 and RI, except for very low choices of timeout.
 
@@
 fourth plots in \cref{figure:others} (and in subsequent cumulative plots) we show only satisfiable
-instances, and use a reduced range on both axes.
+instances, and use a reduced range on both axes.  For the remainder of this paper, we show only
+the non-induced problem, which tends to be harder; results with the induced variant are similar.
+
+\subsection{Discrepancy Searches}
+
+A \emph{discrepancy} is where search goes against a value-ordering heuristic.  Discrepancy searches
+\cite{DBLP:conf/ijcai/HarveyG95,DBLP:conf/aaai/Korf96,DBLP:conf/ijcai/Walsh97,DBLP:conf/cpaior/KarouiHLN07}
+are alternatives to backtracking search that initially search disallowing all discrepancies, and
+then retry search allowing an increasing number of discrepancies at each iteration until either a
+solution is found or unsatisfiability is proven. These schemes assume that value-ordering heuristics
+are usually reliable, and that most solutions can be found with only a small number of
+discrepancies; in such cases, the heavy commitment to early branching choices made by conventional
+backtracking search can be extremely costly.
+
+\begin{figure}[tb]
+    \includegraphics*{gen-graph-dds.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-dds.pdf}
+
+    \caption{Comparing depth-bounded discrepancy search to backtracking search, both with the degree
+    heuristic.}\label{figure:scatter-dds}
+\end{figure}
+
+?? Also do LDS?
+
+\Cref{figure:scatter-dds} shows the effects of adding Walsh's \cite{DBLP:conf/ijcai/Walsh97}
+depth-bounded discrepancy search (DDS) to the solver (results with other discrepancy search variants
+are similar).  Each point represents the solving time for one instance---to avoid noise for easier
+instances, we measure only time spent during search, and exclude time spent in preprocessing and
+initialisation.  Points below the $x-y$ diagonal are speedups, whilst points on the top and right
+axes represent instances which timed out after one thousand seconds with one algorithm, but not the
+other. For satisfiable instances, the different point styles show the different families, whilst all
+unsatisfiable instances are shown as dark dots.  The points well below the diagonal line and along
+the right-hand axis on the scatter plot show that DDS can sometimes be extremely beneficial on
+satisfiable instances.  However, on both unsatisfiable and most satisfiable instances, the overheads
+can be extremely large, and DDS is much worse in aggregate and is not a viable approach (even when
+only considering satisfiable instances).  These large overheads are to be expected: discrepancy
+searches are aimed primarily at getting better feasible solutions in optimisation problems which are
+too large for a proof of optimality to be a realistic prospect, and they are not well-suited for
+unsatisfiable decision problems. Despite this, the extremely large gains on some satisfiable
+instances confirm our suspicions that we should find an alternative to heavy-commitment backtracking
+search.
 
@@
 
+?? Lead in
+
 The left plot of \cref{figure:value-ordering-heuristics} shows the cumulative number of satisfiable
-instances solved over time using different value-ordering heuristics. For the remainder of this
-section, we show only the non-induced problem; results with the induced variant are similar.
+instances solved over time using different value-ordering heuristics.
 
@@
 
-The right-hand plot of \cref{figure:value-ordering-heuristics} gives a detailed look at these
-results.  Each point represents the solving time for one instance---to avoid noise for easier
-instances, we measure only time spent during search, and exclude time spent in preprocessing and
-initialisation.  Points below the $x-y$ diagonal are speedups, whilst points on the top and right
-axes represent instances which timed out after one thousand seconds with one algorithm, but not the
-other. For satisfiable instances, the different point styles show the different families, whilst all
-unsatisfiable instances are shown as dark dots.  This plot shows that despite the aggregate
+The right-hand plot of \cref{figure:value-ordering-heuristics} gives a detailed comparison of the
+degree and biased heuristics. It shows that despite the aggregate
 performance being similar, on a case by case basis, the two heuristics can make a large difference
@@
 
-\subsection{Discrepancy Searches}
-
-A \emph{discrepancy} is where search goes against a value-ordering heuristic.  Discrepancy searches
-\cite{DBLP:conf/ijcai/HarveyG95,DBLP:conf/aaai/Korf96,DBLP:conf/ijcai/Walsh97,DBLP:conf/cpaior/KarouiHLN07}
-are alternatives to backtracking search that initially search disallowing all discrepancies, and
-then retry search allowing an increasing number of discrepancies at each iteration until either a
-solution is found or unsatisfiability is proven. These schemes assume that value-ordering heuristics
-are usually reliable, and that most solutions can be found with only a small number of
-discrepancies; in such cases, the heavy commitment to early branching choices made by conventional
-backtracking search can be extremely costly.
-
-\begin{figure}[tb]
-    \includegraphics*{gen-graph-dds.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-dds.pdf}
-
-    \caption{Comparing depth-bounded discrepancy search to backtracking search, both with the degree
-    heuristic.}\label{figure:scatter-dds}
-\end{figure}
-
-\Cref{figure:scatter-dds} shows the effects of adding Walsh's \cite{DBLP:conf/ijcai/Walsh97}
-depth-bounded discrepancy search (DDS) to \cref{algorithm:sip} (results with other discrepancy
-search variants are similar).  The points well below the diagonal line and along the right-hand axis
-on the scatter plot show that DDS can sometimes be extremely beneficial on satisfiable instances.
-However, on both unsatisfiable and most satisfiable instances, the overheads can be extremely large,
-and DDS is much worse in aggregate and is not a viable approach (even when only considering
-satisfiable instances).  These large overheads are to be expected: discrepancy searches are aimed
-primarily at getting better feasible solutions in optimisation problems which are too large for a
-proof of optimality to be a realistic prospect, and they are not well-suited for unsatisfiable
-decision problems. Despite this, the extremely large gains on some satisfiable instances confirm our
-suspicions that we should find an alternative to heavy-commitment backtracking search.
+?? Intuition
 
@@
 to use smallest domain first with static tiebreaking for variable-ordering.\footnote{It may be
-possible to further improve \cref{algorithm:sip} by also introducing randomness or some form of
+possible to further improve the solver by also introducing randomness or some form of
 learning into its variable-ordering heuristic. However, simultaneously introducing a second change
-would considerably complicate the empirical analysis.} Using restarts on value-ordering heuristics
-is uncommon---although \citet{Razgon et al.\ }{DBLP:conf/flairs/RazgonOP07} look at learning
-value-ordering heuristics from restarts, \citet{Chu et al.\ }{DBLP:conf/cp/ChuSS09} use a similar
-scheme in the context of parallel search, and an early approach by \citet{Gomes et al.\ }{DBLP:conf/aaai/GomesSK98} does so in an optimisation context.
+would considerably complicate the empirical analysis. Additionally, the solver's current
+hand-crafted variable-ordering heuristics already beat adaptive heuristics like impact or
+activity-based search.} Using restarts on value-ordering heuristics is uncommon (although
+\citet{Razgon et al.\ }{DBLP:conf/flairs/RazgonOP07} look at learning value-ordering heuristics from
+restarts, \citet{Chu et al.\ }{DBLP:conf/cp/ChuSS09} use a similar scheme in the context of parallel
+search, and an early approach by \citet{Gomes et al.\ }{DBLP:conf/aaai/GomesSK98} does so in an
+optimisation context).
 
-The shaded code shows how to introduce restarts to \cref{algorithm:sip}. The recursive
-$\FuncSty{search}$ procedure is modified to return either true (if the instance is satisfiable),
-false (unsatisfiable), or a third value if some threshold was reached without the solution being
-known.  The main function then calls $\FuncSty{search}$ in a loop (\siplineref{restartsloop}), until
-a solution is found.
+% The shaded code shows how to introduce restarts to \cref{algorithm:sip}. The recursive
+% $\FuncSty{search}$ procedure is modified to return either true (if the instance is satisfiable),
+% false (unsatisfiable), or a third value if some threshold was reached without the solution being
+% known.  The main function then calls $\FuncSty{search}$ in a loop (\siplineref{restartsloop}), until
+% a solution is found.
 
-We use the Luby scheme \cite{DBLP:journals/ipl/LubySZ93} to determine when to restart. We count the
-number of backtracks (that is, when we reach the end of the main $\KwSty{foreach}$ loop on
-\siplineref{countrestarts}) to decide when to restart, rather than counting recursive calls.
-Following convention, we multiply each item in the Luby sequence by a constant---we
-used the SMAC automatic parameter tuner \cite{DBLP:conf/lion/HutterHL11} to select the value 660.
-
-The remainder of the changes are to introduce nogoods. To avoid exploring portions of the search
-space that we have already visitied, every time we restart, we add new constraints to the problem
-which eliminate already-explored subtrees---such a constraint is called a nogood. We generate simple
-decision nogoods, as follows. The $B$ argument to $\FuncSty{search}$ tracks the branching choices
-made so far, in the form of a CNF clause with each literal being a guessed $(v \mapsto v')$
-assignment. Upon backtracking due to a decision to restart, we post a nogood of the form $(v \mapsto
-v') \wedge (w \mapsto w') \wedge (x \mapsto x') \Rightarrow \bot$ for every branch to the left of
-the current (incomplete) branch in the search tree, and when we first make a decision to restart
-before backtracking, we post a similar nogood eliminating the entire subtree explored.
+Preliminary experiments directed us to use the Luby scheme \cite{DBLP:journals/ipl/LubySZ93} to
+determine when to restart.
+%We count the number of backtracks
+%(that is, when we reach the end of the main $\KwSty{foreach}$ loop on
+%\siplineref{countrestarts})
+%to decide when to restart, rather than counting recursive calls.
+Following convention, we multiply each item in the Luby sequence by a constant---we used the SMAC
+automatic parameter tuner \cite{DBLP:conf/lion/HutterHL11} to select the value 660.
+% The remainder of the changes are to introduce nogoods.  The $B$ argument to $\FuncSty{search}$ tracks the branching choices
+% made so far, in the form of a CNF clause with each literal being a guessed $(v \mapsto v')$
+% assignment.
+To avoid exploring portions of the search
+space that we have already visited, every time we restart, we add new constraints to the problem
+which eliminate already-explored subtrees---such a constraint is called a nogood. We generate
+simple decision nogoods. That is, upon backtracking due to a decision to restart, we post a nogood
+of the form $(v \mapsto v') \wedge (w \mapsto w') \wedge (x \mapsto x') \Rightarrow \bot$ for every
+branch to the left of the current (incomplete) branch in the search tree, and when we first make a
+decision to restart before backtracking, we post a similar nogood eliminating the entire subtree
+explored.
 
 Finally, we use the two watched literals technique \cite{DBLP:conf/dac/MoskewiczMZZM01} to
-propagate stored nogoods (\siplineref{2wl}). This has two benefits: the propagation complexity does
-not particularly depend upon the number of stored nogoods, and it does not require any work upon
-backtracking.  Other more sophisticated nogood generation and propagation schemes exist
-\cite{DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17}, but it is not clear whether these will be
-beneficial in a setting where we are not maintaining arc consistency.
+propagate stored nogoods.
+%(\siplineref{2wl}).
+This has two benefits: the propagation complexity does not particularly depend upon the number of
+stored nogoods, and it does not require any work upon backtracking.  Other more sophisticated nogood
+generation and propagation schemes exist \cite{DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17},
+but these would not be helpful here since our solver does not maintain arc consistency or use a
+propagation queue.
+
+\subsection{The Effects of Restarts and Nogood Recording}
 
@@
 \end{figure}
-
-\begin{figure}[tb]
-    \includegraphics*{gen-graph-scatter-random-goods.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-random.pdf}
-    \caption{On the left, not using nogood recording introduces slowdowns, particularly on
-    unsatisfiable instances. On the right, using a random value-ordering gives much worse
-    performance on many satisfiable instances.}
-    \label{figure:features}
-\end{figure}
-
-\subsection{The Effects of Restarts and Nogood Recording}
 
@@
 performance, we have made up most of the consistency we lost by introducing randomness.
+This is a promising result: for the first time, we have introduced a viable alternative to
+backtracking search in a setting where we both have a strong \emph{value}-ordering heuristic,
+\emph{and} care about both satisfiable and unsatisfiable instances.
+
+\begin{figure}[tb]
+    \includegraphics*{gen-graph-scatter-random-goods.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-random.pdf}
+    \caption{On the left, not using nogood recording introduces slowdowns, particularly on
+    unsatisfiable instances. On the right, using a random value-ordering gives much worse
+    performance on many satisfiable instances.}
+    \label{figure:features}
+\end{figure}
 
@@
 
-This is a promising result: for the first time, we have introduced a viable alternative to
-backtracking search in a setting where we both have a strong \emph{value}-ordering heuristic,
-\emph{and} care about both satisfiable and unsatisfiable instances.
+\section{Parallel Search}
+
+?? No sharing, sharing with Luby restarts, sharing with constant restarts, sharing with triggered
+restarts
+
+?? Reproducibility, constant and triggered
+
+?? How many more nodes on unsat instances
+
+?? Scalability to MPI
 
@@
 that fails, retries allowing a single unmatched vertex (and thus using weaker invariants), and so
-on. Due to its similarity to \cref{algorithm:sip}, we can introduce the same bias and restart
-strategy, requiring only small changes to handle wildcards in nogoods.
+on. Due to its similarity to the Glasgow Subgraph Solver, we can introduce the same bias and restart
+strategy.
 
@@
 
-The effectiveness of this approach raises a number of interesting questions and avenues for possible
-future research. For example, will this same technique work for other problems where we have a good
-value-ordering heuristic? Are there further benefits to be had from also modifying variable-ordering
-heuristics? And how can such a search process be parallelised?
+?? Parallel
+
+?? Apply to CP in general
 
@@
 
-The authors would like to thank Ruth Hoffmann, Christophe Lecoutre, Christine Solnon and Craig
+The authors would like to thank Ruth Hoffmann, Christophe Lecoutre, Christine Solnon, and Craig
 Reilly for their comments.

### 2018-11-10 vf3/lad2arg.cc
VF3

@@ -1,0 +1,103 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include <algorithm>
+#include <cstdlib>
+#include <exception>
+#include <iostream>
+#include <numeric>
+#include <string>
+#include <vector>
+
+using std::accumulate;
+using std::cerr;
+using std::cin;
+using std::cout;
+using std::endl;
+using std::exception;
+using std::istream;
+using std::string;
+using std::vector;
+
+class GraphFileError :
+    public exception
+{
+    private:
+        string _what;
+
+    public:
+        GraphFileError(const string & filename, const string & message) throw ();
+
+        auto what() const throw () -> const char *;
+};
+
+GraphFileError::GraphFileError(const string & filename, const string & message) throw () :
+    _what("Error reading graph file '" + filename + "': " + message)
+{
+}
+
+auto GraphFileError::what() const throw () -> const char *
+{
+    return _what.c_str();
+}
+
+namespace
+{
+    auto read_word(istream & infile) -> int
+    {
+        int x;
+        infile >> x;
+        return x;
+    }
+}
+
+auto main(int, char * argv[]) -> int
+{
+    try {
+        uint16_t size = read_word(cin);
+        if (! cin)
+            throw GraphFileError{ "stdin", "error reading size" };
+
+        vector<vector<int> > adj;
+        adj.resize(size, vector<int>(size));
+
+        for (int r = 0 ; r < size ; ++r) {
+            int c_end = read_word(cin);
+            if (! cin)
+                throw GraphFileError{ "stdin", "error reading edges count" };
+
+            for (int c = 0 ; c < c_end ; ++c) {
+                int e = read_word(cin);
+
+                if (e < 0 || e >= size)
+                    throw GraphFileError{ "stdin", "edge index out of bounds" };
+
+                adj[r][e] = adj[e][r] = 1;
+            }
+        }
+
+        string rest;
+        if (cin >> rest)
+            throw GraphFileError{ "stdin", "EOF not reached, next text is \"" + rest + "\"" };
+        if (! cin.eof())
+            throw GraphFileError{ "stdin", "EOF not reached" };
+
+        cout << size << endl;
+        for (int r = 0 ; r < size ; ++r)
+            cout << r << " " << (adj[r][r] ? 2 : 1) << endl;
+        for (int r = 0 ; r < size ; ++r) {
[…]

### 2018-11-10 vf3/solve_vf3_induced.cc

@@ -1,0 +1,146 @@
+#include <stdio.h>
+#include <iostream>
+#include <fstream>
+#include <stdlib.h>
+//#include <unistd.h>
+#include <chrono>
+#include <atomic>
+#include <thread>
+#include <mutex>
+#include <condition_variable>
+#include "match.hpp"
+#include "argloader.hpp"
+#include "argraph.hpp"
+#include "argedit.hpp"
+#include "nodesorter.hpp"
+#include "probability_strategy.hpp"
+#include "vf3_sub_state.hpp"
+#include "nodesorter.hpp"
+#include "nodeclassifier.hpp"
+
+template<> long long VF3SubState<int,int,Empty,Empty>::instance_count = 0;
+static long long state_counter = 0;
+
+template<typename State>
+bool match_with_timeout(State &s, int *pn, node_id c1[], node_id c2[], unsigned long long & nodes,
+        std::atomic<bool> & abort) {
+    if (abort)
+        return false;
+
+    ++nodes;
+
+    if (s.IsGoal())
+    {
+        *pn = s.CoreLen();
+        s.GetCoreSet(c1, c2);
+        return true;
+    }
+
+    if (s.IsDead())
+        return false;
+
+    node_id n1 = NULL_NODE, n2 = NULL_NODE;
+    bool found = false;
+
+    while (!found && s.NextPair(&n1, &n2, n1, n2))
+    {
+        if (s.IsFeasiblePair(n1, n2))
+        {
+            State s1(s);
+            s1.AddPair(n1, n2);
+            found = match_with_timeout(s1, pn, c1, c2, nodes, abort);
+        }
+    }
+
+    return found;
+}
+
+int main(int argc, char** argv)
+{
+
+  char *pattern, *target;
+
+  state_counter = 0;
+  int n = 0;
+
+  if (argc != 4)
+  {
+	  std::cout << "Usage: " << argv[0] << " pattern target timeout \n";
+	  return -1;
+  }
+  
+  pattern = argv[1];
+  target = argv[2];
+  
+  std::ifstream graphInPat(pattern);
+  std::ifstream graphInTarg(target);
+
+  StreamARGLoader<int, Empty> pattloader(graphInPat);
+  StreamARGLoader<int, Empty> targloader(graphInTarg);
+
+  ARGraph<int, Empty> patt_graph(&pattloader);
+  ARGraph<int, Empty> targ_graph(&targloader);
+  
[…]

### 2018-11-10 code/sequential.cc
Up to date Glasgow solver

@@ -1,1058 +1,0 @@
-/* vim: set sw=4 sts=4 et foldmethod=syntax : */
-
-#include "sequential.hh"
-#include "fixed_bit_set.hh"
-#include "template_voodoo.hh"
-
-#include <algorithm>
-#include <array>
-#include <functional>
-#include <limits>
-#include <list>
-#include <map>
-#include <numeric>
-#include <random>
-#include <tuple>
-#include <utility>
-
-using std::array;
-using std::iota;
-using std::fill;
-using std::find_if;
-using std::get;
-using std::greater;
-using std::list;
-using std::max;
-using std::map;
-using std::move;
-using std::mt19937;
-using std::next;
-using std::numeric_limits;
-using std::pair;
-using std::sort;
-using std::string;
-using std::swap;
-using std::to_string;
-using std::tuple;
-using std::uniform_int_distribution;
-using std::uniform_real_distribution;
-using std::vector;
-
-using std::chrono::duration_cast;
-using std::chrono::milliseconds;
-using std::chrono::steady_clock;
-
-namespace
-{
-    enum class Search
-    {
-        Aborted,
-        Unsatisfiable,
-        Satisfiable
-    };
-
-    enum class RestartingSearch
-    {
-        Aborted,
-        Unsatisfiable,
-        Satisfiable,
-        Restart
-    };
-
-    auto degree_sort(const Graph & graph, vector<int> & p, bool reverse) -> void
-    {
-        // pre-calculate degrees
-        vector<int> degrees;
-
-        for (int v = 0 ; v < graph.size() ; ++v)
-            degrees.push_back(graph.degree(v));
-
-        // sort on degree
-        sort(p.begin(), p.end(),
-                [&] (int a, int b) { return (! reverse) ^ (degrees[a] < degrees[b] || (degrees[a] == degrees[b] && a > b)); });
-    }
-
-    using Assignment = pair<unsigned, unsigned>;
-
-    struct Assignments
-    {
-        vector<tuple<Assignment, bool, int> > values;
-
-        bool contains(const Assignment & assignment) const
-        {
-            // this should not be a linear scan...
-            return values.end() != find_if(values.begin(), values.end(), [&] (const auto & a) {
-                    return get<0>(a) == assignment;
-                    });
-        }
-    };
-
-    // A nogood, aways of the form (list of assignments) -> false, where the
-    // last part is implicit. If there are at least two assignments, then the
[…]

### 2018-11-10 glasgow/formats/csv.cc

@@ -1,0 +1,48 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "formats/csv.hh"
+#include "formats/input_graph.hh"
+
+#include <fstream>
+#include <unordered_map>
+#include <vector>
+
+using std::ifstream;
+using std::pair;
+using std::string;
+using std::unordered_map;
+using std::vector;
+
+auto read_csv(ifstream && infile, const string & filename) -> InputGraph
+{
+    InputGraph result{ 0, false, false };
+
+    if (! infile)
+        throw GraphFileError{ filename, "error opening file" };
+
+    unordered_map<string, int> vertices;
+    string line;
+
+    vector<pair<int, int> > edges;
+
+    while (getline(infile, line)) {
+        auto pos = line.find(',');
+        if (string::npos == pos)
+            throw GraphFileError{ filename, "expected a comma but didn't get one" };
+        string left = line.substr(0, pos), right = line.substr(pos + 1);
+        int left_idx = vertices.emplace(left, vertices.size()).first->second;
+        int right_idx = vertices.emplace(right, vertices.size()).first->second;
+        edges.emplace_back(left_idx, right_idx);
+    }
+
+    result.resize(vertices.size());
+
+    for (auto & e : edges)
+        result.add_edge(e.first, e.second);
+
+    for (auto & [v, l] : vertices)
+        result.set_vertex_label(l, v);
+
+    return result;
+}
+

### 2018-11-10 glasgow/formats/dimacs.cc

@@ -1,27 +1,23 @@
 /* vim: set sw=4 sts=4 et foldmethod=syntax : */
 
-#include "dimacs.hh"
-#include "graph.hh"
-#include "graph_file_error.hh"
+#include "formats/dimacs.hh"
+#include "formats/input_graph.hh"
+#include "formats/graph_file_error.hh"
 
-#include <boost/regex.hpp>
 #include <fstream>
+#include <regex>
 
 using std::getline;
 using std::ifstream;
+using std::regex;
+using std::smatch;
 using std::stoi;
 using std::string;
+using std::to_string;
 
-using boost::regex;
-using boost::smatch;
-
-auto read_dimacs(const string & filename) -> Graph
+auto read_dimacs(ifstream && infile, const string & filename) -> InputGraph
 {
-    Graph result{ 0 };
-
-    ifstream infile{ filename };
-    if (! infile)
-        throw GraphFileError{ filename, "unable to open file" };
+    InputGraph result{ 0, false, false };
 
     string line;
     while (getline(infile, line)) {
@@ -60,6 +56,9 @@ auto read_dimacs(const string & filename) -> Graph
 
     if (! infile.eof())
         throw GraphFileError{ filename, "error reading file" };
+
+    for (int v = 0 ; v < result.size() ; ++v)
+        result.set_vertex_label(v, to_string(v + 1));
 
     return result;
 }

### 2018-11-10 glasgow/formats/graph_file_error.cc

@@ -1,11 +1,10 @@
 /* vim: set sw=4 sts=4 et foldmethod=syntax : */
 
-#include "graph_file_error.hh"
-#include "graph.hh"
+#include "formats/graph_file_error.hh"
+#include "formats/input_graph.hh"
 
 #include <fstream>
 
-using std::ifstream;
 using std::string;
 
 GraphFileError::GraphFileError(const string & filename, const string & message) throw () :
@@ -17,5 +16,4 @@ auto GraphFileError::what() const throw () -> const char *
 {
     return _what.c_str();
 }
-
 

### 2018-11-10 glasgow/formats/input_graph.cc

@@ -1,0 +1,92 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "formats/input_graph.hh"
+
+#include <algorithm>
+#include <limits>
+#include <string_view>
+
+using std::distance;
+using std::numeric_limits;
+using std::make_pair;
+using std::max;
+using std::string_view;
+
+InputGraph::InputGraph(int size, bool v, bool e) :
+    _has_vertex_labels(v),
+    _has_edge_labels(e)
+{
+    if (0 != size)
+        resize(size);
+}
+
+auto InputGraph::resize(int size) -> void
+{
+    _size = size;
+    _vertex_labels.resize(size);
+}
+
+auto InputGraph::add_edge(int a, int b) -> void
+{
+    _edges.emplace(make_pair(a, b), "");
+    _edges.emplace(make_pair(b, a), "");
+}
+
+auto InputGraph::add_directed_edge(int a, int b, string_view label) -> void
+{
+    _edges.emplace(make_pair(a, b), label).first->second = label;
+    _edges.emplace(make_pair(b, a), "unlabelled");
+}
+
+auto InputGraph::adjacent(int a, int b) const -> bool
+{
+    return _edges.count({ a, b });
+}
+
+auto InputGraph::size() const -> int
+{
+    return _size;
+}
+
+auto InputGraph::degree(int a) const -> int
+{
+    auto lower = _edges.lower_bound({ a, 0 });
+    auto upper = _edges.upper_bound({ a, numeric_limits<int>::max() });
+    return distance(lower, upper);
+}
+
+auto InputGraph::set_vertex_label(int v, string_view l) -> void
+{
+    _vertex_labels[v] = l;
+}
+
+auto InputGraph::vertex_label(int v) const -> string_view
+{
+    return _vertex_labels[v];
+}
+
+auto InputGraph::edge_label(int a, int b) const -> string_view
+{
+    return _edges.find({a, b})->second;
+}
+
+auto InputGraph::begin_edges() const -> InputGraph::EdgesIterator
+{
+    return _edges.begin();
+}
+
+auto InputGraph::end_edges() const -> InputGraph::EdgesIterator
+{
+    return _edges.end();
+}
+
+auto InputGraph::has_vertex_labels() const -> bool
+{
+    return _has_vertex_labels;
+}
+
+auto InputGraph::has_edge_labels() const -> bool
+{
+    return _has_edge_labels;
+}
+

### 2018-11-10 glasgow/formats/lad.cc

@@ -1,0 +1,98 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "formats/lad.hh"
+#include "formats/input_graph.hh"
+
+#include <fstream>
+#include <map>
+
+using std::ifstream;
+using std::map;
+using std::pair;
+using std::string;
+using std::to_string;
+
+namespace
+{
+    auto read_word(ifstream & infile) -> int
+    {
+        int x;
+        infile >> x;
+        return x;
+    }
+}
+
+auto read_lad(ifstream && infile, const string & filename) -> InputGraph
+{
+    InputGraph result{ 0, false, false };
+
+    result.resize(read_word(infile));
+    if (! infile)
+        throw GraphFileError{ filename, "error reading size" };
+
+    for (int r = 0 ; r < result.size() ; ++r) {
+        int c_end = read_word(infile);
+        if (! infile)
+            throw GraphFileError{ filename, "error reading edges count" };
+
+        for (int c = 0 ; c < c_end ; ++c) {
+            int e = read_word(infile);
+
+            if (e < 0 || e >= result.size())
+                throw GraphFileError{ filename, "edge index out of bounds" };
+
+            result.add_edge(r, e);
+        }
+    }
+
+    string rest;
+    if (infile >> rest)
+        throw GraphFileError{ filename, "EOF not reached, next text is \"" + rest + "\"" };
+    if (! infile.eof())
+        throw GraphFileError{ filename, "EOF not reached" };
+
+    return result;
+}
+
+auto read_labelled_lad(ifstream && infile, const string & filename) -> InputGraph
+{
+    InputGraph result{ 0, true, true };
+
+    result.resize(read_word(infile));
+    if (! infile)
+        throw GraphFileError{ filename, "error reading size" };
+
+    for (int r = 0 ; r < result.size() ; ++r) {
+        int l = read_word(infile);
+        if (! infile)
+            throw GraphFileError{ filename, "error reading label" };
+
+        int c_end = read_word(infile);
+        if (! infile)
+            throw GraphFileError{ filename, "error reading edges count" };
+
+        result.set_vertex_label(r, to_string(l));
+
+        for (int c = 0 ; c < c_end ; ++c) {
+            int e = read_word(infile);
+
+            if (e < 0 || e >= result.size())
+                throw GraphFileError{ filename, "edge index out of bounds" };
+
+            int l = read_word(infile);
+            if (l < 0)
[…]

### 2018-11-10 glasgow/formats/read_file_format.cc

@@ -1,0 +1,119 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "formats/read_file_format.hh"
+#include "formats/dimacs.hh"
+#include "formats/lad.hh"
+#include "formats/csv.hh"
+
+#include <fstream>
+#include <regex>
+#include <sstream>
+#include <vector>
+
+using std::ifstream;
+using std::ios;
+using std::move;
+using std::regex;
+using std::smatch;
+using std::stoi;
+using std::string;
+using std::stringstream;
+using std::to_string;
+using std::vector;
+
+auto detect_format(ifstream & infile, const string & filename) -> string
+{
+    string line;
+    if (! getline(infile, line) || line.empty())
+        throw GraphFileError{ filename, "unable to read file to detect file format" };
+
+    static const regex
+        dimacs_comment{ R"(c(\s.*)?)" },
+        dimacs_problem{ R"(p\s+(edge|col)\s+(\d+)\s+(\d+)?\s*)" },
+        lad_header{ R"(\d+)" },
+        lad_zero_labelled_line{ R"(0 \d+)" },
+        lad_zero_unlabelled_line{ R"(0)" },
+        lad_line{ R"(\d+\s+(\d+\s+)*\d+\s*)" },
+        csv_problem{ R"(\S+,\S+)" };
+
+    smatch match;
+    if (regex_match(line, match, dimacs_comment)) {
+        while (regex_match(line, match, dimacs_comment)) {
+            // looks like a DIMACS comment, ignore
+            if (! getline(infile, line) || line.empty())
+                throw GraphFileError{ filename, "unable to auto-detect file format (entirely c lines?)" };
+        }
+        if (! regex_match(line, match, dimacs_problem))
+            throw GraphFileError{ filename, "unable to auto-detect file format (c line not followed by a p line?)" };
+        return "dimacs";
+    }
+    else if (regex_match(line, match, dimacs_problem))
+        return "dimacs";
+    else if (regex_match(line, match, lad_header)) {
+        if ("0" == line)
+            return "lad";
+
+        // got to figure out whether we're labelled or not
+        if (! getline(infile, line) || line.empty())
+            throw GraphFileError{ filename, "unable to auto-detect file format (number followed by nothing)" };
+        if (regex_match(line, match, lad_zero_labelled_line))
+            return "labelledlad";
+        else if (regex_match(line, match, lad_zero_unlabelled_line))
+            return "lad";
[…]

### 2018-11-10 glasgow/solve_subgraph_isomorphism.cc

@@ -1,8 +1,7 @@
 /* vim: set sw=4 sts=4 et foldmethod=syntax : */
 
-#include "lad.hh"
-#include "dimacs.hh"
-#include "sequential.hh"
+#include "formats/read_file_format.hh"
+#include "solver.hh"
 
 #include <boost/program_options.hpp>
 
@@ -105,49 +104,28 @@ auto run_this(Result_ func(const Data_ &, const Params_ &)) -> function<Result_ 
 
 auto main(int argc, char * argv[]) -> int
 {
-    auto subgraph_isomorphism_algorithms = {
-        make_pair( string{ "simple" },                       sequential_subgraph_isomorphism ),
-        make_pair( string{ "restarting" },                   sequential_subgraph_isomorphism ),
-        make_pair( string{ "customisable-sequential" },      sequential_subgraph_isomorphism )
-    };
-
     try {
         po::options_description display_options{ "Program options" };
         display_options.add_options()
             ("help",                                         "Display help information")
             ("timeout",            po::value<int>(),         "Abort after this many seconds")
-            ("dimacs",                                       "Read DIMACS format instead of LAD")
+            ("format",             po::value<std::string>(), "Specify input file format (auto, lad, labelledlad, dimacs)")
+            ("pattern-format",     po::value<std::string>(), "Specify input file format just for the pattern graph")
+            ("target-format",      po::value<std::string>(), "Specify input file format just for the target graph")
             ("induced",                                      "Solve the induced version")
-            ("enumerate",                                    "Count the number of solutions");
-
-        po::options_description custom_options{ "Options for customisable algorithms (not all combinations make sense)" };
-        custom_options.add_options()
-            ("restarts",                                     "Use restarts")
-            ("dds",                                          "Use dds")
-            ("shuffle",                                      "Use shuffling")
-            ("softmax-shuffle",                              "Use softmax shuffling")
-            ("antiheuristic",                                "Use antiheuristic")
-            ("input-order",                                  "Use input order")
-            ("luby-multiplier",      po::value<unsigned>(),  "Specify a Luby multiplier")
-            ("geometric-multiplier", po::value<double>(),    "Specify a Geometric multiplier")
-            ("geometric-start",      po::value<unsigned>(),  "Specify geometric start value")
[…]

### 2018-11-10 glasgow/solver.cc

@@ -1,0 +1,1178 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "solver.hh"
+#include "fixed_bit_set.hh"
+#include "template_voodoo.hh"
+
+#include <algorithm>
+#include <array>
+#include <functional>
+#include <limits>
+#include <list>
+#include <map>
+#include <numeric>
+#include <optional>
+#include <random>
+#include <type_traits>
+#include <utility>
+
+#include <boost/dynamic_bitset.hpp>
+
+using std::array;
+using std::iota;
+using std::fill;
+using std::find_if;
+using std::greater;
+using std::is_same;
+using std::list;
+using std::max;
+using std::map;
+using std::move;
+using std::mt19937;
+using std::next;
+using std::numeric_limits;
+using std::optional;
+using std::pair;
+using std::sort;
+using std::string;
+using std::swap;
+using std::to_string;
+using std::uniform_int_distribution;
+using std::vector;
+
+using std::chrono::duration_cast;
+using std::chrono::milliseconds;
+using std::chrono::steady_clock;
+
+using boost::dynamic_bitset;
+
+namespace
+{
+    enum class SearchResult
+    {
+        Aborted,
+        Unsatisfiable,
+        Satisfiable,
+        SatisfiableButKeepGoing,
+        Restart
+    };
+
+    struct Assignment
+    {
+        unsigned pattern_vertex;
+        unsigned target_vertex;
+
+        auto operator== (const Assignment & other) const -> bool
+        {
+            return pattern_vertex == other.pattern_vertex && target_vertex == other.target_vertex;
+        }
+
+        auto operator!= (const Assignment & other) const -> bool
+        {
+            return ! (*this == other);
+        }
+    };
+
+    struct AssignmentInformation
+    {
+        Assignment assignment;
+        bool is_decision;
+        int discrepancy_count;
+        int choice_count;
+    };
+
+    struct Assignments
+    {
+        vector<AssignmentInformation> values;
+
+        bool contains(const Assignment & assignment) const
+        {
+            // this should not be a linear scan...
+            return values.end() != find_if(values.begin(), values.end(), [&] (const auto & a) {
+                    return a.assignment == assignment;
+                    });
+        }
+    };
+
+    // A nogood, aways of the form (list of assignments) -> false, where the
+    // last part is implicit. If there are at least two assignments, then the
+    // first two assignments are the watches (and the literals are permuted
[…]

### 2018-11-10 glasgow/solve_subgraph_isomorphism.cc
Support specifying value-ordering

@@ -113,8 +113,14 @@ auto main(int argc, char * argv[]) -> int
             ("pattern-format",     po::value<std::string>(), "Specify input file format just for the pattern graph")
             ("target-format",      po::value<std::string>(), "Specify input file format just for the target graph")
             ("induced",                                      "Solve the induced version")
-            ("enumerate",                                    "Count the number of solutions")
-            ("presolve",                                     "Try presolving (hacky, experimental, possibly useful for easy instances");
+            ("enumerate",                                    "Count the number of solutions");
+
+        po::options_description configuration_options{ "Advanced configuration options" };
+        configuration_options.add_options()
+            ("presolve",                                     "Try presolving (hacky, experimental, possibly useful for easy instances")
+            ("value-ordering",     po::value<std::string>(), "Specify value-ordering heuristic (biased / degree / antidegree / random)");
+
+        display_options.add(configuration_options);
 
         po::options_description all_options{ "All options" };
         all_options.add_options()
@@ -159,6 +165,22 @@ auto main(int argc, char * argv[]) -> int
         params.induced = options_vars.count("induced");
         params.enumerate = options_vars.count("enumerate");
         params.presolve = options_vars.count("presolve");
+
+        if (options_vars.count("value-ordering")) {
+            std::string value_ordering_heuristic = options_vars["value-ordering"].as<std::string>();
+            if (value_ordering_heuristic == "biased")
+                params.value_ordering_heuristic = ValueOrdering::Biased;
+            else if (value_ordering_heuristic == "degree")
+                params.value_ordering_heuristic = ValueOrdering::Degree;
+            else if (value_ordering_heuristic == "antidegree")
+                params.value_ordering_heuristic = ValueOrdering::AntiDegree;
+            else if (value_ordering_heuristic == "random")
+                params.value_ordering_heuristic = ValueOrdering::Random;
+            else {
+                cerr << "Unknown value-ordering heuristic '" << value_ordering_heuristic << "'" << endl;
+                return EXIT_FAILURE;
+            }
+        }
 
         char hostname_buf[255];
         if (0 == gethostname(hostname_buf, 255))

### 2018-11-10 glasgow/solver.cc

@@ -34,6 +34,7 @@ using std::numeric_limits;
 using std::optional;
 using std::pair;
 using std::sort;
+using std::stable_sort;
 using std::string;
 using std::swap;
 using std::to_string;
@@ -568,6 +569,17 @@ namespace
             }
         }
 
+        auto degree_sort(
+                ArrayType_ & branch_v,
+                unsigned branch_v_end,
+                bool reverse
+                ) -> void
+        {
+            stable_sort(branch_v.begin(), branch_v.begin() + branch_v_end, [&] (int a, int b) -> bool {
+                    return (targets_degrees[0][a] >= targets_degrees[0][b]) ^ reverse;
+                    });
+        }
+
         auto restarting_search(
                 Assignments & assignments,
                 const Domains & domains,
@@ -606,7 +618,23 @@ namespace
                 branch_v[branch_v_end++] = f_v;
             }
 
-            softmax_shuffle(branch_v, branch_v_end);
+            switch (params.value_ordering_heuristic) {
+                case ValueOrdering::Degree:
+                    degree_sort(branch_v, branch_v_end, false);
+                    break;
+
+                case ValueOrdering::AntiDegree:
+                    degree_sort(branch_v, branch_v_end, true);
+                    break;
+
+                case ValueOrdering::Biased:
+                    softmax_shuffle(branch_v, branch_v_end);
+                    break;
+
+                case ValueOrdering::Random:
+                    shuffle(branch_v.begin(), branch_v.begin() + branch_v_end, global_rand);
+                    break;
+            }
 
             int discrepancy_count = 0;
             bool actually_hit_a_success = false, actually_hit_a_failure = true;

### 2018-11-10 glasgow/solve_subgraph_isomorphism.cc
Support turning off restarts

@@ -118,6 +118,7 @@ auto main(int argc, char * argv[]) -> int
         po::options_description configuration_options{ "Advanced configuration options" };
         configuration_options.add_options()
             ("presolve",                                     "Try presolving (hacky, experimental, possibly useful for easy instances")
+            ("restarts-constant",  po::value<int>(),         "How often to perform restarts (0 disables restarts)")
             ("value-ordering",     po::value<std::string>(), "Specify value-ordering heuristic (biased / degree / antidegree / random)");
 
         display_options.add(configuration_options);
@@ -165,6 +166,9 @@ auto main(int argc, char * argv[]) -> int
         params.induced = options_vars.count("induced");
         params.enumerate = options_vars.count("enumerate");
         params.presolve = options_vars.count("presolve");
+
+        if (options_vars.count("restarts-constant"))
+            params.restarts_constant = options_vars["restarts-constant"].as<int>();
 
         if (options_vars.count("value-ordering")) {
             std::string value_ordering_heuristic = options_vars["value-ordering"].as<std::string>();

### 2018-11-10 glasgow/solver.cc

@@ -960,10 +960,10 @@ namespace
             while (! done) {
                 long long backtracks_until_restart;
 
-                if (params.enumerate)
+                if (params.enumerate || 0 == params.restarts_constant)
                     backtracks_until_restart = -1;
                 else {
-                    backtracks_until_restart = *current_luby * params.luby_multiplier;
+                    backtracks_until_restart = *current_luby * params.restarts_constant;
                     if (next(current_luby) == luby.end()) {
                         luby.insert(luby.end(), luby.begin(), luby.end());
                         luby.push_back(*luby.rbegin() * 2);

### 2018-11-10 glasgow/solve_subgraph_isomorphism.cc
Support turning off nogoods

@@ -118,6 +118,7 @@ auto main(int argc, char * argv[]) -> int
         po::options_description configuration_options{ "Advanced configuration options" };
         configuration_options.add_options()
             ("presolve",                                     "Try presolving (hacky, experimental, possibly useful for easy instances")
+            ("nogood-size-limit",  po::value<int>(),         "Maximum size of nogood to generate (0 disables nogoods")
             ("restarts-constant",  po::value<int>(),         "How often to perform restarts (0 disables restarts)")
             ("value-ordering",     po::value<std::string>(), "Specify value-ordering heuristic (biased / degree / antidegree / random)");
 
@@ -167,6 +168,8 @@ auto main(int argc, char * argv[]) -> int
         params.enumerate = options_vars.count("enumerate");
         params.presolve = options_vars.count("presolve");
 
+        if (options_vars.count("nogood-size-limit"))
+            params.nogood_size_limit = options_vars["nogood-size-limit"].as<int>();
         if (options_vars.count("restarts-constant"))
             params.restarts_constant = options_vars["restarts-constant"].as<int>();
 

### 2018-11-10 glasgow/solver.cc

@@ -516,7 +516,7 @@ namespace
         auto post_nogood(
                 const Assignments & assignments)
         {
-            if (params.enumerate)
+            if (params.enumerate || params.nogood_size_limit == 0)
                 return;
 
             Nogood nogood;
@@ -524,6 +524,9 @@ namespace
             for (auto & a : assignments.values)
                 if (a.is_decision)
                     nogood.literals.emplace_back(a.assignment);
+
+            if (nogood.literals.size() > params.nogood_size_limit)
+                return;
 
             nogoods.emplace_back(move(nogood));
             need_to_watch.emplace_back(prev(nogoods.end()));

### 2018-11-10 glasgow/solve_subgraph_isomorphism.cc
DDS support

@@ -117,6 +117,7 @@ auto main(int argc, char * argv[]) -> int
 
         po::options_description configuration_options{ "Advanced configuration options" };
         configuration_options.add_options()
+            ("dds",                                          "Use DDS (forces degree value-ordering, no restarts or nogoods)")
             ("presolve",                                     "Try presolving (hacky, experimental, possibly useful for easy instances")
             ("nogood-size-limit",  po::value<int>(),         "Maximum size of nogood to generate (0 disables nogoods")
             ("restarts-constant",  po::value<int>(),         "How often to perform restarts (0 disables restarts)")
@@ -167,6 +168,7 @@ auto main(int argc, char * argv[]) -> int
         params.induced = options_vars.count("induced");
         params.enumerate = options_vars.count("enumerate");
         params.presolve = options_vars.count("presolve");
+        params.dds = options_vars.count("dds");
 
         if (options_vars.count("nogood-size-limit"))
             params.nogood_size_limit = options_vars["nogood-size-limit"].as<int>();

### 2018-11-10 glasgow/solver.cc

@@ -58,6 +58,14 @@ namespace
         Restart
     };
 
+    enum class DiscrepancySearchResult
+    {
+        Aborted,
+        Unsatisfiable,
+        Satisfiable,
+        SatisfiableButKeepGoing
+    };
+
     struct Assignment
     {
         unsigned pattern_vertex;
@@ -640,7 +648,7 @@ namespace
             }
 
             int discrepancy_count = 0;
-            bool actually_hit_a_success = false, actually_hit_a_failure = true;
+            bool actually_hit_a_success = false, actually_hit_a_failure = false;
 
             // for each value remaining...
             for (auto f_v = branch_v.begin(), f_end = branch_v.begin() + branch_v_end ; f_v != f_end ; ++f_v) {
@@ -710,6 +718,113 @@ namespace
                 return SearchResult::SatisfiableButKeepGoing;
             else
                 return SearchResult::Unsatisfiable;
+        }
+
+        auto dds_search(
+                Assignments & assignments,
+                const Domains & domains,
+                unsigned long long & nodes,
+                unsigned long long & propagations,
+                unsigned long long & solution_count,
+                int depth,
+                int discrepancy_k,
+                bool & used_all_discrepancies) -> DiscrepancySearchResult
+        {
+            if (params.abort->load())
+                return DiscrepancySearchResult::Aborted;
+
+            ++nodes;
+
+            // find ourselves a domain, or succeed if we're all assigned
+            const Domain * branch_domain = find_branch_domain(domains);
+            if (! branch_domain) {
+                if (params.enumerate) {
+                    ++solution_count;
+                    return DiscrepancySearchResult::SatisfiableButKeepGoing;
+                }
+                else
+                    return DiscrepancySearchResult::Satisfiable;
+            }
+
+            // pull out the remaining values in this domain for branching
+            auto remaining = branch_domain->values;
+
+            ArrayType_ branch_v;
+            if constexpr (is_same<ArrayType_, vector<int> >::value)
+                branch_v.resize(target_size);
+
+            unsigned branch_v_end = 0;
+            for (auto f_v = remaining.find_first() ; f_v != decltype(remaining)::npos ; f_v = remaining.find_first()) {
+                remaining.reset(f_v);
+                branch_v[branch_v_end++] = f_v;
+            }
+
+            degree_sort(branch_v, branch_v_end, false);
+
+            int discrepancy_count = 0;
+            bool actually_hit_a_success = false;
+
[…]

### 2018-11-13 glasgow/parallel_solver.cc
Don't sync on the first pass

@@ -1042,6 +1042,7 @@ namespace
                     auto current_luby = luby.begin();
                     double current_geometric = params.restarts_constant;
                     unsigned number_of_restarts = 0;
+                    bool first_pass = true;
 
                     // keep going until someone posts an empty nogood, signifying done
                     while (true) {
@@ -1065,42 +1066,46 @@ namespace
 
                         ++number_of_restarts;
 
-                        nogoods_ready_barrier.count_down_and_wait();
+                        if (! first_pass) {
+                            nogoods_ready_barrier.count_down_and_wait();
 
-                        // start watching new nogoods from all threads
-                        for (auto & ss : all_search_data) {
-                            for (auto & new_nogood : ss.need_to_watch) {
-                                // have to copy nogoods not owned by us, they are permuted
-                                Nogoods::iterator n = new_nogood;
-                                if (&ss != &s)
-                                    n = s.nogoods.insert(s.nogoods.end(), *new_nogood);
+                            // start watching new nogoods from all threads
+                            for (auto & ss : all_search_data) {
+                                for (auto & new_nogood : ss.need_to_watch) {
+                                    // have to copy nogoods not owned by us, they are permuted
+                                    Nogoods::iterator n = new_nogood;
+                                    if (&ss != &s)
+                                        n = s.nogoods.insert(s.nogoods.end(), *new_nogood);
 
-                                if (n->literals.empty()) {
-                                    done = true;
-                                    break;
-                                }
-                                else if (1 == n->literals.size()) {
-                                    for (auto & d : domains)
-                                        if (d.v == n->literals[0].pattern_vertex) {
-                                            d.values.reset(n->literals[0].target_vertex);
-                                            d.count = d.values.count();
-                                            break;
-                                        }
-                                }
-                                else {
-                                    s.watches[n->literals[0]].push_back(n);
-                                    s.watches[n->literals[1]].push_back(n);
+                                    if (n->literals.empty()) {
+                                        done = true;
+                                        break;
+                                    }
+                                    else if (1 == n->literals.size()) {
+                                        for (auto & d : domains)
+                                            if (d.v == n->literals[0].pattern_vertex) {
+                                                d.values.reset(n->literals[0].target_vertex);
+                                                d.count = d.values.count();
+                                                break;
+                                            }
+                                    }
+                                    else {
+                                        s.watches[n->literals[0]].push_back(n);
+                                        s.watches[n->literals[1]].push_back(n);
+                                    }
[…]

### 2018-11-13 paper/paper.tex
Work on parallel a bit

@@
 
-\author{Ciaran McCreesh\orcidID{0000-0002-6106-4871}\and
+\author{Blair Archibald\orcidID{????-????-????-????}\and
+Ciaran McCreesh\orcidID{0000-0002-6106-4871}\and
 Patrick Prosser\orcidID{0000-0003-4460-6912} \and
@@
 
-\authorrunning{C. McCreesh et al.}
+\authorrunning{B. Archibald et al.}
 
@@
 
-Our experiments are performed on systems with dual Intel Xeon E5-2697A v4 CPUs and 512GBytes RAM,
-running Ubuntu 17.04, using GCC 7.2.0 as the compiler. We use a deterministic pseudo-random number
-generator for reproducibility.  We use the dataset introduced by \citet{Kotthoff et al.\
-}{DBLP:conf/lion/KotthoffMS16} for evaluation. This dataset brings together a range of
-randomly-generated and application instance families from earlier papers:
+Our experiments are performed on the EPCC Cirrus HPC facility, on systems with dual Intel Xeon
+E5-2695 v4 CPUs and 256GBytes RAM, running Centos 7.3.1611. ?? C++ native threads, MPI. We use GCC
+7.2.0 as the compiler. ?? Steady clock times. We use a deterministic pseudo-random number generator
+for reproducibility.
+
+We use the dataset introduced by \citet{Kotthoff et al.\ }{DBLP:conf/lion/KotthoffMS16} for
+evaluation. This dataset brings together a range of randomly-generated and application instance
+families from earlier papers:
 
@@
 randomness to a tailored heuristic than it is to throw away the heuristic altogether.  Indeed, the
-original algorithm can solve 1,983 satisfiable instances by 993.0 seconds, whilst the biased and
-random restarting algorithms require only 9.2 seconds and 33.8 seconds respectively to solve the
+original algorithm can solve ?? satisfiable instances by ?? seconds, whilst the biased and
+random restarting algorithms require only ?? seconds and ?? seconds respectively to solve the
 same number.
@@
 \begin{figure}[tb]
-    \includegraphics*{gen-graph-scatter-random-goods.pdf}
+    \includegraphics*{gen-graph-scatter-biased-goods.pdf}
     \hfill
@@
 In the left-hand plot of \cref{figure:features}, we see large slowdowns on unsatisfiable instances
-when using a random value ordering and restarts but without nogood recording, and on the right-hand
-plot we see many more satisfiable instances above the $x-y$ diagonal when using the random
-value-ordering heuristic as opposed to the degree-biased heuristic.
+when disabling nogood recording, and on the right-hand plot we see many more satisfiable instances
+above the $x-y$ diagonal when using the random value-ordering heuristic as opposed to the
+degree-biased heuristic.
 
 \section{Parallel Search}
+
+\begin{figure}[tb]
+    \centering
+    \includegraphics*{gen-graph-parallel.pdf}
+
+    \caption{Parallel}\label{figure:parallel}
+\end{figure}
 

### 2018-11-16 glasgow/parallel_solver.cc
Catch restarts a bit faster

@@ -708,6 +708,9 @@ namespace
 
             // for each value remaining...
             for (auto f_v = branch_v.begin(), f_end = branch_v.begin() + branch_v_end ; f_v != f_end ; ++f_v) {
+                if (do_a_restart)
+                    break;
+
                 // modified in-place by appending, we can restore by shrinking
                 auto assignments_size = assignments.values.size();
 

### 2018-11-16 paper/paper.tex
Parallel is go

@@
 
-\title{Beating Simple Backtracking Search in Subgraph Algorithms\thanks{This work was
-supported by the Engineering and Physical Sciences Research Council (grant numbers EP/026842/1 and
-EP/M508056/1). This work used the Cirrus UK National Tier-2 HPC Service at EPCC
+\title{Sequential and Parallel Solution-Biased Search for Subgraph Algorithms\thanks{This work was
+supported by the Engineering and Physical Sciences Research Council (grant numbers EP/P026842/1,
+EP/M508056/1, and EP/N007565). This work used the Cirrus UK National Tier-2 HPC Service at EPCC
 (http://www.cirrus.ac.uk) funded by the University of Edinburgh and EPSRC (EP/P020267/1).}}
 
-\author{Blair Archibald\orcidID{????-????-????-????}\and
+\author{Blair Archibald\orcidID{0000-0003-3699-6658}\and
 Ciaran McCreesh\orcidID{0000-0002-6106-4871}\and
@@
 
-\authorrunning{B. Archibald et al.}
+%\authorrunning{B. Archibald et al.}
 
@@
 
+    \medskip
+
+    \includegraphics*{gen-graph-scatter-par.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-parconsttick.pdf}
+
     \caption{Parallel}\label{figure:parallel}

### 2018-11-16 glasgow/parallel_solver.cc
Unbugger it

@@ -704,12 +704,14 @@ namespace
             }
 
             int discrepancy_count = 0;
-            bool actually_hit_a_success = false, actually_hit_a_failure = false;
+            bool actually_hit_a_success = false, actually_hit_a_failure = false, completed_the_loop = true;
 
             // for each value remaining...
             for (auto f_v = branch_v.begin(), f_end = branch_v.begin() + branch_v_end ; f_v != f_end ; ++f_v) {
-                if (do_a_restart)
+                if (do_a_restart) {
+                    completed_the_loop = false;
                     break;
+                }
 
                 // modified in-place by appending, we can restore by shrinking
                 auto assignments_size = assignments.values.size();
@@ -772,7 +774,8 @@ namespace
             // no values remaining, backtrack, or possibly kick off a restart
             if ((do_a_restart) || (actually_hit_a_failure && backtracks_until_restart > 0 && 0 == --backtracks_until_restart)) {
                 do_a_restart = true;
-                post_nogood(s, assignments);
+                if (completed_the_loop)
+                    post_nogood(s, assignments);
                 return SearchResult::Restart;
             }
             else if (actually_hit_a_success)

### 2018-11-16 glasgow/parallel_solver.cc
Simplify

@@ -704,15 +704,10 @@ namespace
             }
 
             int discrepancy_count = 0;
-            bool actually_hit_a_success = false, actually_hit_a_failure = false, completed_the_loop = true;
+            bool actually_hit_a_success = false, actually_hit_a_failure = false;
 
             // for each value remaining...
             for (auto f_v = branch_v.begin(), f_end = branch_v.begin() + branch_v_end ; f_v != f_end ; ++f_v) {
-                if (do_a_restart) {
-                    completed_the_loop = false;
-                    break;
-                }
-
                 // modified in-place by appending, we can restore by shrinking
                 auto assignments_size = assignments.values.size();
 
@@ -734,7 +729,7 @@ namespace
                 // recursive search
                 auto search_result = restarting_search(s, assignments, new_domains, nodes, propagations,
                         solution_count, depth + 1, backtracks_until_restart, do_a_restart);
-
+                    
                 switch (search_result) {
                     case SearchResult::Satisfiable:
                         return SearchResult::Satisfiable;
@@ -774,8 +769,7 @@ namespace
             // no values remaining, backtrack, or possibly kick off a restart
             if ((do_a_restart) || (actually_hit_a_failure && backtracks_until_restart > 0 && 0 == --backtracks_until_restart)) {
                 do_a_restart = true;
-                if (completed_the_loop)
-                    post_nogood(s, assignments);
+                post_nogood(s, assignments);
                 return SearchResult::Restart;
             }
             else if (actually_hit_a_success)

### 2018-11-18 glasgow/mpi_solver.cc
MPI solver, attempt one

@@ -1,0 +1,1257 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "solver.hh"
+#include "fixed_bit_set.hh"
+#include "template_voodoo.hh"
+
+#include <algorithm>
+#include <array>
+#include <atomic>
+#include <condition_variable>
+#include <functional>
+#include <limits>
+#include <list>
+#include <map>
+#include <mutex>
+#include <numeric>
+#include <optional>
+#include <random>
+#include <thread>
+#include <type_traits>
+#include <utility>
+
+#include <boost/dynamic_bitset.hpp>
+
+#include <boost/mpi/environment.hpp>
+#include <boost/mpi/communicator.hpp>
+#include <boost/mpi/collectives.hpp>
+
+#include <boost/serialization/vector.hpp>
+#include <boost/serialization/list.hpp>
+#include <boost/serialization/utility.hpp>
+
+using std::array;
+using std::atomic;
+using std::condition_variable;
+using std::iota;
+using std::fill;
+using std::find_if;
+using std::greater;
+using std::is_same;
+using std::list;
+using std::max;
+using std::map;
+using std::move;
+using std::mt19937;
+using std::mutex;
+using std::next;
+using std::numeric_limits;
+using std::optional;
+using std::pair;
+using std::scoped_lock;
+using std::sort;
+using std::stable_sort;
+using std::string;
+using std::swap;
+using std::thread;
+using std::to_string;
+using std::uniform_int_distribution;
+using std::unique_lock;
+using std::vector;
+
+using std::chrono::duration_cast;
+using std::chrono::milliseconds;
+using std::chrono::steady_clock;
+
+using boost::dynamic_bitset;
+
+using boost::mpi::all_gather;
+using boost::mpi::communicator;
+
+namespace
+{
+    struct barrier
+    {
+        const unsigned n;
+        int w;
+        long long g;
+        mutex m;
+        condition_variable cv;
+
+        barrier(unsigned nn) :
+            n(nn),
+            w(int(nn)),
+            g(0)
+        {
+        }
+
+        void count_down_and_wait()
+        {
+            unique_lock<mutex> l(m);
+            if (w == -1)
+                return;
+
+            if (0 == --w) {
+                ++g;
+                w = n;
+                cv.notify_all();
+            }
+            else {
+                int old_g = g;
+                while (true) {
+                    cv.wait(l);
+                    if (old_g != g || -1 == w)
+                        break;
[…]

### 2018-11-18 glasgow/solve_subgraph_isomorphism.cc

@@ -2,7 +2,16 @@
 
 #include "formats/read_file_format.hh"
 #include "solver.hh"
-#include "parallel_solver.hh"
+
+#ifdef WITH_MPI
+#    include "mpi_solver.hh"
+#    include <boost/mpi/environment.hpp>
+#    include <boost/mpi/communicator.hpp>
+#    include <boost/mpi/collectives.hpp>
+#else
+#    include "solver.hh"
+#    include "parallel_solver.hh"
+#endif
 
 #include <boost/program_options.hpp>
 
@@ -15,6 +24,7 @@
 #include <iostream>
 #include <mutex>
 #include <thread>
+#include <vector>
 
 #include <unistd.h>
 
@@ -36,6 +46,7 @@ using std::put_time;
 using std::string;
 using std::thread;
 using std::unique_lock;
+using std::vector;
 
 using std::chrono::seconds;
 using std::chrono::steady_clock;
@@ -43,68 +54,17 @@ using std::chrono::system_clock;
 using std::chrono::duration_cast;
 using std::chrono::milliseconds;
 
-/* Helper: return a function that runs the specified algorithm, dealing
- * with timing information and timeouts. */
-template <typename Result_, typename Params_, typename Data_>
-auto run_this_wrapped(const function<Result_ (const Data_ &, const Params_ &)> & func)
-    -> function<Result_ (const Data_ &, Params_ &, bool &, int)>
-{
-    return [func] (const Data_ & data, Params_ & params, bool & aborted, int timeout) -> Result_ {
-        /* For a timeout, we use a thread and a timed CV. We also wake the
-         * CV up if we're done, so the timeout thread can terminate. */
-        thread timeout_thread;
-        mutex timeout_mutex;
-        condition_variable timeout_cv;
-        atomic<bool> abort;
-        abort.store(false);
-        params.abort = &abort;
-        if (0 != timeout) {
-            timeout_thread = thread([&] {
-                    auto abort_time = steady_clock::now() + seconds(timeout);
-                    {
-                        /* Sleep until either we've reached the time limit,
-                         * or we've finished all the work. */
-                        unique_lock<mutex> guard(timeout_mutex);
-                        while (! abort.load()) {
-                            if (cv_status::timeout == timeout_cv.wait_until(guard, abort_time)) {
-                                /* We've woken up, and it's due to a timeout. */
-                                aborted = true;
-                                break;
-                            }
-                        }
-                    }
-                    abort.store(true);
-                    });
-        }
-
-        /* Start the clock */
[…]

### 2018-11-18 paper/paper.tex
More plottery

@@
     The current state of the art in subgraph isomorphism solving involves using degree as a
-    value-ordering heuristics to direct backtracking search. Such a search makes a heavy commitment
-    to the first branching choice, which is often incorrect. We evaluate discrepancy-based searches
-    as an alternative, but find them too expensive to be effective in this situation, even on
-    satisfiable instances. We then demonstrate that a random value-ordering heuristic with restarts
-    beats conventional search with the best-known tailored heuristic. This motivates the
-    introduction of a novel slightly-random heuristic with similar behaviour to the tailored
-    heuristic: when combined with restarts and nogood recording, this new search strategy is over ??
-    times more effective on satisfiable instances, and does not affect performance on
-    unsatisfiable instances. Such a search can also be parallelised in a way which is much
-    simpler than conventional work-stealing: we obtain further speedups of ??  on ??  cores, and
-    ?? over ?? distributed-memory hosts. Finally, we show that this search strategy also improves
-    two maximum common induced subgraph algorithms.
+    value-ordering heuristic to direct backtracking search. Such a search makes a heavy commitment
+    to the first branching choice, which is often incorrect. To mitigate this, we introduce and
+    evaluate a new approach, which we call ``solution-biased search''. By combining a
+    slightly-random value-ordering heuristic, rapid restarts, and nogood recording, we achieve two
+    orders of magnitude faster performance on satisfiable instances, whilst not affecting
+    performance on unsatisfiable instances. This search can also be parallelised in a very simple
+    but effective way, giving us a further speedup of over thirty from thirty-six cores, across both
+    satisfiable and unsatisfiable instances.  Finally, we show that solution-biased search is also
+    suitable for optimisation problems, by using it to improve two maximum common induced subgraph
+    algorithms.
 \end{abstract}
@@
 
+    \medskip
+
+    \includegraphics*{gen-graph-scatter-par-repeat.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-parconsttick-repeat.pdf}
+
     \caption{Parallel}\label{figure:parallel}
+\end{figure}
+
+\begin{figure}[tb]
+    \includegraphics*{gen-graph-scatter-parconst-nodes.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-parconsttick-nodes.pdf}
+
+    \caption{Parallel nodes}\label{figure:parallelnodes}
 \end{figure}

### 2018-11-18 glasgow/parallel_solver.cc
Whitespace

@@ -729,7 +729,7 @@ namespace
                 // recursive search
                 auto search_result = restarting_search(s, assignments, new_domains, nodes, propagations,
                         solution_count, depth + 1, backtracks_until_restart, do_a_restart);
-                    
+
                 switch (search_result) {
                     case SearchResult::Satisfiable:
                         return SearchResult::Satisfiable;

### 2018-11-18 glasgow/mpi_solver.cc
Also unbugger

@@ -726,15 +726,10 @@ namespace
             }
 
             int discrepancy_count = 0;
-            bool actually_hit_a_success = false, actually_hit_a_failure = false, completed_the_loop = true;
+            bool actually_hit_a_success = false, actually_hit_a_failure = false;
 
             // for each value remaining...
             for (auto f_v = branch_v.begin(), f_end = branch_v.begin() + branch_v_end ; f_v != f_end ; ++f_v) {
-                if (do_a_restart) {
-                    completed_the_loop = false;
-                    break;
-                }
-
                 // modified in-place by appending, we can restore by shrinking
                 auto assignments_size = assignments.values.size();
 
@@ -796,8 +791,7 @@ namespace
             // no values remaining, backtrack, or possibly kick off a restart
             if ((do_a_restart) || (actually_hit_a_failure && backtracks_until_restart > 0 && 0 == --backtracks_until_restart)) {
                 do_a_restart = true;
-                if (completed_the_loop)
-                    post_nogood(s, assignments);
+                post_nogood(s, assignments);
                 return SearchResult::Restart;
             }
             else if (actually_hit_a_success)

### 2018-11-18 glasgow/solve_subgraph_isomorphism.cc
Better MPI result sharing

@@ -41,6 +41,7 @@ using std::exception;
 using std::function;
 using std::localtime;
 using std::make_pair;
+using std::move;
 using std::mutex;
 using std::put_time;
 using std::string;
@@ -252,16 +253,13 @@ auto main(int argc, char * argv[]) -> int
 
 #ifdef WITH_MPI
         /* Merge resulty things */
+        Result combined_result;
         vector<Result> all_results;
-        Result empty_result;
-        gather(mpi_comm, 0 == mpi_comm.rank() ? empty_result : result, all_results, 0);
-        if (do_output_here) {
-            for (auto & r : all_results) {
-                if (! r.isomorphism.empty())
-                    result.isomorphism = r.isomorphism;
-                result.extra_stats.insert(result.extra_stats.end(), r.extra_stats.begin(), r.extra_stats.end());
-            }
-        }
+        gather(mpi_comm, result, all_results, 0);
+        if (do_output_here)
+            for (auto & r : all_results)
+                combined_result.merge("", r);
+        result = move(combined_result);
 #endif
 
         if (do_output_here) {

### 2018-11-22 glasgow/solve_subgraph_isomorphism.cc
Restart timer

@@ -83,6 +83,7 @@ auto main(int argc, char * argv[]) -> int
             ("presolve",                                     "Try presolving (hacky, experimental, possibly useful for easy instances")
             ("nogood-size-limit",  po::value<int>(),         "Maximum size of nogood to generate (0 disables nogoods")
             ("restarts-constant",  po::value<int>(),         "How often to perform restarts (0 disables restarts)")
+            ("restart-timer",      po::value<int>(),         "Also restart after this many milliseconds (0 disables)")
             ("geometric-restarts", po::value<double>(),      "Use geometric restarts with the specified multiplier (default is Luby)")
             ("value-ordering",     po::value<std::string>(), "Specify value-ordering heuristic (biased / degree / antidegree / random)");
         display_options.add(configuration_options);
@@ -151,6 +152,8 @@ auto main(int argc, char * argv[]) -> int
             params.restarts_constant = options_vars["restarts-constant"].as<int>();
         if (options_vars.count("geometric-restarts"))
             params.geometric_multiplier = options_vars["geometric-restarts"].as<double>();
+        if (options_vars.count("restart-timer"))
+            params.restart_timer = milliseconds{ options_vars["restart-timer"].as<int>() };
 
         if (options_vars.count("value-ordering")) {
             std::string value_ordering_heuristic = options_vars["value-ordering"].as<std::string>();

### 2018-11-22 glasgow/solver.cc

@@ -46,6 +46,7 @@ using std::vector;
 using std::chrono::duration_cast;
 using std::chrono::milliseconds;
 using std::chrono::steady_clock;
+using std::chrono::operator""ms;
 
 using boost::dynamic_bitset;
 
@@ -600,7 +601,8 @@ namespace
                 unsigned long long & propagations,
                 unsigned long long & solution_count,
                 int depth,
-                long long & backtracks_until_restart) -> SearchResult
+                long long & backtracks_until_restart,
+                steady_clock::time_point * restart_at) -> SearchResult
         {
             if (params.abort->load())
                 return SearchResult::Aborted;
@@ -674,7 +676,7 @@ namespace
 
                 // recursive search
                 auto search_result = restarting_search(assignments, new_domains, nodes, propagations,
-                        solution_count, depth + 1, backtracks_until_restart);
+                        solution_count, depth + 1, backtracks_until_restart, restart_at);
 
                 switch (search_result) {
                     case SearchResult::Satisfiable:
@@ -713,7 +715,8 @@ namespace
             }
 
             // no values remaining, backtrack, or possibly kick off a restart
-            if (actually_hit_a_failure && backtracks_until_restart > 0 && 0 == --backtracks_until_restart) {
+            if (actually_hit_a_failure && ((backtracks_until_restart > 0 && 0 == --backtracks_until_restart)
+                       || (restart_at && steady_clock::now() > *restart_at))) {
                 post_nogood(assignments);
                 return SearchResult::Restart;
             }
@@ -1166,8 +1169,13 @@ namespace
                     if (propagate(domains, assignments)) {
                         auto assignments_copy = assignments;
 
+                        steady_clock::time_point restart_at;
+                        if (0ms != params.restart_timer)
+                            restart_at = steady_clock::now() + params.restart_timer;
+
                         switch (restarting_search(assignments_copy, domains, result.nodes, result.propagations,
-                                    result.solution_count, 0, backtracks_until_restart)) {
+                                    result.solution_count, 0, backtracks_until_restart,
+                                    (params.restart_timer != 0ms) ? &restart_at : nullptr)) {
                             case SearchResult::Satisfiable:
                                 save_result(assignments_copy, result);
                                 result.complete = true;
@@ -1278,7 +1286,7 @@ namespace
                 ++result.propagations;
                 if (propagate(domains_copy, assignments_copy)) {
                     switch (restarting_search(assignments_copy, domains_copy, result.nodes, result.propagations,
[…]

### 2018-11-22 paper/paper.tex
More experiments

@@
     performance on unsatisfiable instances. This search can also be parallelised in a very simple
-    but effective way, giving us a further speedup of over thirty from thirty-six cores, across both
+    but effective way, giving us a further speedup of over ?? from thirty-six cores, across both
     satisfiable and unsatisfiable instances.  Finally, we show that solution-biased search is also
@@
 The subgraph isomorphism problem is to decide whether a copy of a small ``pattern'' graph occurs
-inside a larger ``target'' graph. Although the problem is \NP-complete, by combining design techniques from
-artificial intelligence with careful algorithm engineering, modern subgraph isomorphism solvers can
-often produce exact solutions quickly even on graphs with thousands of vertices. The current single
-strongest subgraph isomorphism solver uses ``highest degree first'' as a value-ordering heuristic to
-direct a constraint programming style search
+inside a larger ``target'' graph. Although the problem is \NP-complete, by combining design
+techniques from artificial intelligence with careful algorithm engineering, modern subgraph
+isomorphism solvers can often produce exact solutions quickly even on graphs with thousands of
+vertices. The current single strongest subgraph isomorphism solver uses ``highest degree first'' as
+a value-ordering heuristic to direct a constraint programming style search
 \cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:journals/jair/McCreeshPST18}. This
-heuristic is much better than branching randomly, but is still far from perfect. We therefore
-investigate different techniques for offsetting mistakes in value-ordering decisions. We first try
-discrepancy searches, which turn out to be much too expensive to be a practical solution. We then
-look at using random value-ordering, restarts, and nogood recording, which together give better
-aggregate performance at the expense of much worse behaviour on many instances. Next, we introduce a
-new way of adapting the existing value-ordering heuristic to introduce just a small amount of
-randomness.  This lets us make a state-of-the-art subgraph algorithm perform much better
-on a large number of satisfiable instances, whilst performing worse only rarely on satisfiable
-instances, and never on unsatisfiable instances.
+heuristic is much better than branching randomly, but is still far from perfect. To offset mistakes
+made by this heuristic, this paper proposes a new perspective on value-ordering: rather than
+defining a search order, we use degree to direct what \emph{proportion} of the search effort should
+be spent in each subproblem. By combining rapid restarts and nogood recording, and introducing a
+small amount of randomness into the value-ordering heuristic, we make a state-of-the-art subgraph
+algorithm perform two orders of magnitude better on a large number of satisfiable instances, whilst
+performing worse only rarely on satisfiable instances, and never on unsatisfiable instances.
 
-?? Parallel
-
-Finally, we show that this strategy is also effective in an optimisation setting, producing benefits
-in two maximum common induced subgraph algorithms. This combination of techniques gives us, for the
-first time, a practical alternative to backtracking search for constraint programming style
-algorithms where we have a strong value-ordering heuristic and where we care both about satisfiable
-and unsatisfiable instances. ?? Parallel
+This new form of search can also be parallelised. By running many threads with different random
+seeds but the same restart schedule, and sharing nogoods only following restarts, we can achieve
+speedups of ??, with a \emph{much} simpler implementation than conventional work-stealing.  Finally,
+we show that this strategy is also effective in an optimisation setting, producing benefits in two
+maximum common induced subgraph algorithms. This combination of techniques gives us, for the first
+time, a practical alternative to backtracking search for constraint programming style algorithms
+where we have a strong value-ordering heuristic and where we care both about satisfiable and
+unsatisfiable instances.
 
@@
 Our experiments are performed on the EPCC Cirrus HPC facility, on systems with dual Intel Xeon
-E5-2695 v4 CPUs and 256GBytes RAM, running Centos 7.3.1611. ?? C++ native threads, MPI. We use GCC
-7.2.0 as the compiler. ?? Steady clock times. We use a deterministic pseudo-random number generator
-for reproducibility.
+E5-2695 v4 CPUs and 256GBytes RAM, running Centos 7.3.1611. We use GCC 7.2.0 as the compiler. For
+paralellism, we use C++ native threads, and for distributed parallelism we also use the SGI MPT
+implementation of MPI. All timing measurements are steady-clock, and we use a deterministic
+pseudo-random number generator for reproducibility.
 
@@
 
-We begin by verifying that the Glasgow Subgraph Solver is the state of the art. In the first and
+We begin with a set of baseline performance measurements. In the first and
 third plots \cref{figure:others} we show the cumulative number of instances solved over time for the
-non-induced and induced problems respectively. We compare the Glasgow Subgraph Solver (both
-unmodified, and with the modifications described in the remainder of this paper), the PathLAD
-variation of the LAD algorithm \cite{DBLP:journals/ai/Solnon10,DBLP:conf/lion/KotthoffMS16}, VF2
+non-induced and induced problems respectively. We compare the Glasgow Subgraph Solver using
+depth-first search (DFS) and with the modifications described in the remainder of this paper
+(solution-biased search, SBS), the PathLAD variation of the LAD algorithm
+\cite{DBLP:journals/ai/Solnon10,DBLP:conf/lion/KotthoffMS16}, VF2
 \cite{DBLP:journals/pami/CordellaFSV04}, RI \cite{DBLP:journals/bmcbi/BonniciGPSF13}, and VF3
@@
 \end{figure}
-
-?? Also do LDS?
 
@@
 
-    \medskip
-
-    \includegraphics*{gen-graph-scatter-par-repeat.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-parconsttick-repeat.pdf}
-
     \caption{Parallel}\label{figure:parallel}
@@
 
-\begin{figure}[tb]
-    \includegraphics*{gen-graph-scatter-parconst-nodes.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-parconsttick-nodes.pdf}
-
-    \caption{Parallel nodes}\label{figure:parallelnodes}
-\end{figure}
+% \begin{figure}[tb]
+%     \includegraphics*{gen-graph-scatter-par-repeat.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-scatter-parconsttick-repeat.pdf}
+%
+%     \includegraphics*{gen-graph-scatter-parconst-nodes.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-scatter-parconsttick-nodes.pdf}
+%
+%     \caption{Parallel nodes}\label{figure:parallelnodes}
+% \end{figure}
 

### 2018-11-22 glasgow/parallel_solver.cc
Parallel restart timer

@@ -54,6 +54,7 @@ using std::vector;
 using std::chrono::duration_cast;
 using std::chrono::milliseconds;
 using std::chrono::steady_clock;
+using std::chrono::operator""ms;
 
 using boost::dynamic_bitset;
 
@@ -654,6 +655,7 @@ namespace
                 unsigned long long & solution_count,
                 int depth,
                 long long & backtracks_until_restart,
+                steady_clock::time_point * restart_at,
                 atomic<bool> & do_a_restart) -> SearchResult
         {
             if (params.abort->load())
@@ -728,7 +730,7 @@ namespace
 
                 // recursive search
                 auto search_result = restarting_search(s, assignments, new_domains, nodes, propagations,
-                        solution_count, depth + 1, backtracks_until_restart, do_a_restart);
+                        solution_count, depth + 1, backtracks_until_restart, restart_at, do_a_restart);
 
                 switch (search_result) {
                     case SearchResult::Satisfiable:
@@ -767,7 +769,8 @@ namespace
             }
 
             // no values remaining, backtrack, or possibly kick off a restart
-            if ((do_a_restart) || (actually_hit_a_failure && backtracks_until_restart > 0 && 0 == --backtracks_until_restart)) {
+            if ((do_a_restart) || (actually_hit_a_failure && ((backtracks_until_restart > 0 && 0 == --backtracks_until_restart)
+                            || (restart_at && steady_clock::now() > *restart_at)))) {
                 do_a_restart = true;
                 post_nogood(s, assignments);
                 return SearchResult::Restart;
@@ -1115,8 +1118,15 @@ namespace
                         if (propagate(s, domains, assignments)) {
                             auto assignments_copy = assignments;
 
+                            steady_clock::time_point restart_at, * restart_at_ptr = nullptr;
+                            if (0ms != params.restart_timer)
+                                if ((! params.triggered_restarts) || (0 == t)) {
+                                    restart_at = steady_clock::now() + params.restart_timer;
+                                    restart_at_ptr = &restart_at;
+                                }
+
                             switch (restarting_search(s, assignments_copy, domains, thread_result.nodes, thread_result.propagations,
-                                        thread_result.solution_count, 0, backtracks_until_restart, do_a_restart)) {
+                                        thread_result.solution_count, 0, backtracks_until_restart, restart_at_ptr, do_a_restart)) {
                                 case SearchResult::Satisfiable:
                                     save_result(assignments_copy, thread_result);
                                     thread_result.extra_stats.emplace_back("stop_reason = satisfiable");

### 2018-11-22 glasgow/mpi_solver.cc
MPI triggers

@@ -62,6 +62,7 @@ using std::vector;
 using std::chrono::duration_cast;
 using std::chrono::milliseconds;
 using std::chrono::steady_clock;
+using std::chrono::operator""ms;
 
 using boost::dynamic_bitset;
 
@@ -676,6 +677,7 @@ namespace
                 unsigned long long & solution_count,
                 int depth,
                 long long & backtracks_until_restart,
+                steady_clock::time_point * restart_at,
                 atomic<bool> & do_a_restart) -> SearchResult
         {
             if (params.abort->load())
@@ -750,7 +752,7 @@ namespace
 
                 // recursive search
                 auto search_result = restarting_search(s, assignments, new_domains, nodes, propagations,
-                        solution_count, depth + 1, backtracks_until_restart, do_a_restart);
+                        solution_count, depth + 1, backtracks_until_restart, restart_at, do_a_restart);
 
                 switch (search_result) {
                     case SearchResult::Satisfiable:
@@ -789,7 +791,8 @@ namespace
             }
 
             // no values remaining, backtrack, or possibly kick off a restart
-            if ((do_a_restart) || (actually_hit_a_failure && backtracks_until_restart > 0 && 0 == --backtracks_until_restart)) {
+            if ((do_a_restart) || (actually_hit_a_failure && ((backtracks_until_restart > 0 && 0 == --backtracks_until_restart)
+                            || (restart_at && steady_clock::now() > *restart_at)))) {
                 do_a_restart = true;
                 post_nogood(s, assignments);
                 return SearchResult::Restart;
@@ -1152,8 +1155,15 @@ namespace
                         if (propagate(s, domains, assignments)) {
                             auto assignments_copy = assignments;
 
+                            steady_clock::time_point restart_at, * restart_at_ptr = nullptr;
+                            if (0ms != params.restart_timer)
+                                if ((! params.triggered_restarts) || (0 == t)) {
+                                    restart_at = steady_clock::now() + params.restart_timer;
+                                    restart_at_ptr = &restart_at;
+                                }
+
                             switch (restarting_search(s, assignments_copy, domains, thread_result.nodes, thread_result.propagations,
-                                        thread_result.solution_count, 0, backtracks_until_restart, do_a_restart)) {
+                                        thread_result.solution_count, 0, backtracks_until_restart, restart_at_ptr, do_a_restart)) {
                                 case SearchResult::Satisfiable:
                                     save_result(assignments_copy, thread_result);
                                     thread_result.extra_stats.emplace_back("stop_reason = satisfiable");

### 2018-11-23 glasgow/solve_subgraph_isomorphism.cc
Measure the Fraser distribution

@@ -79,6 +79,7 @@ auto main(int argc, char * argv[]) -> int
 
         po::options_description configuration_options{ "Advanced configuration options" };
         configuration_options.add_options()
+            ("fraserify",                                    "Output Fraser distribution statistics")
             ("dds",                                          "Use DDS (forces degree value-ordering, no restarts or nogoods)")
             ("presolve",                                     "Try presolving (hacky, experimental, possibly useful for easy instances")
             ("nogood-size-limit",  po::value<int>(),         "Maximum size of nogood to generate (0 disables nogoods")
@@ -145,6 +146,7 @@ auto main(int argc, char * argv[]) -> int
         params.presolve = options_vars.count("presolve");
         params.dds = options_vars.count("dds");
         params.triggered_restarts = options_vars.count("triggered-restarts");
+        params.fraserify = options_vars.count("fraserify");
 
         if (options_vars.count("nogood-size-limit"))
             params.nogood_size_limit = options_vars["nogood-size-limit"].as<int>();

### 2018-11-23 glasgow/solver.cc

@@ -601,8 +601,10 @@ namespace
                 unsigned long long & propagations,
                 unsigned long long & solution_count,
                 int depth,
+                int position_at_previous_depth,
                 long long & backtracks_until_restart,
-                steady_clock::time_point * restart_at) -> SearchResult
+                steady_clock::time_point * restart_at,
+                std::list<std::string> & extra_stats) -> SearchResult
         {
             if (params.abort->load())
                 return SearchResult::Aborted;
@@ -654,6 +656,12 @@ namespace
             int discrepancy_count = 0;
             bool actually_hit_a_success = false, actually_hit_a_failure = false;
 
+            unsigned long long nodes_before = nodes, solutions_before = solution_count;
+            if (params.fraserify && 0 == depth) {
+                extra_stats.emplace_back("fraser_pattern_degree = " + to_string(patterns_degrees[0][branch_domain->v]));
+                extra_stats.emplace_back("fraser_branch_v_end = " + to_string(branch_v_end));
+            }
+
             // for each value remaining...
             for (auto f_v = branch_v.begin(), f_end = branch_v.begin() + branch_v_end ; f_v != f_end ; ++f_v) {
                 // modified in-place by appending, we can restore by shrinking
@@ -674,9 +682,12 @@ namespace
                     continue;
                 }
 
+                if (params.fraserify && 0 == depth)
+                    extra_stats.emplace_back("fraser_target_degree_" + to_string(f_v - branch_v.begin()) + " = " + to_string(targets_degrees[0][*f_v]));
+
                 // recursive search
                 auto search_result = restarting_search(assignments, new_domains, nodes, propagations,
-                        solution_count, depth + 1, backtracks_until_restart, restart_at);
+                        solution_count, depth + 1, f_v - branch_v.begin(), backtracks_until_restart, restart_at, extra_stats);
 
                 switch (search_result) {
                     case SearchResult::Satisfiable:
@@ -712,6 +723,11 @@ namespace
                 }
 
                 ++discrepancy_count;
+            }
+
+            if (params.fraserify && depth == 1) {
+                extra_stats.emplace_back("fraser_target_nodes_" + to_string(position_at_previous_depth) + " = " + to_string(nodes - nodes_before));
+                extra_stats.emplace_back("fraser_target_solution_counts_" + to_string(position_at_previous_depth) + " = " + to_string(solution_count - solutions_before));
             }
 
[…]

### 2018-11-23 paper/paper.tex
Radical rewrite time

@@
 
-\author{Blair Archibald\orcidID{0000-0003-3699-6658}\and
-Ciaran McCreesh\orcidID{0000-0002-6106-4871}\and
-Patrick Prosser\orcidID{0000-0003-4460-6912} \and
-James Trimble\orcidID{0000-0001-7282-8745}}
+\author{Blair Archibald\inst{1}\orcidID{0000-0003-3699-6658}\and
+Fraser Dunlop\inst{2}\orcidID{0000-0002-4485-4871}\and
+Ruth Hoffmann\inst{2}\orcidID{0000-0002-1011-5894}\and
+Ciaran McCreesh\inst{1}\orcidID{0000-0002-6106-4871}\and
+Patrick Prosser\inst{1}\orcidID{0000-0003-4460-6912} \and
+James Trimble\inst{1}\orcidID{0000-0001-7282-8745}}
 
-%\authorrunning{B. Archibald et al.}
+\authorrunning{B. Archibald et al.}
 
-\institute{University of Glasgow, Glasgow, Scotland\\
-\email{ciaran.mccreesh@glasgow.ac.uk}}
+\institute{University of Glasgow, Glasgow, Scotland \and University of St Andrews, St Andrews,
+Scotland \\ \email{ciaran.mccreesh@glasgow.ac.uk}}
 
@@
     orders of magnitude faster performance on satisfiable instances, whilst not affecting
-    performance on unsatisfiable instances. This search can also be parallelised in a very simple
+    performance on unsatisfiable instances. This search can be parallelised in a very simple
     but effective way, giving us a further speedup of over ?? from thirty-six cores, across both
-    satisfiable and unsatisfiable instances.  Finally, we show that solution-biased search is also
+    satisfiable and unsatisfiable instances.  We also show that solution-biased search is also
     suitable for optimisation problems, by using it to improve two maximum common induced subgraph
@@
 
-    \caption{Comparing depth-bounded discrepancy search to backtracking search, both with the degree
-    heuristic.}\label{figure:scatter-dds}
+    \caption{Comparing depth-bounded discrepancy search (DDS) to depth-first backtracking search
+    (DFS), both using degree as the value-ordering heuristic.}\label{figure:scatter-dds}
 \end{figure}
@@
 depth-bounded discrepancy search (DDS) to the solver (results with other discrepancy search variants
-are similar).  Each point represents the solving time for one instance---to avoid noise for easier
-instances, we measure only time spent during search, and exclude time spent in preprocessing and
-initialisation.  Points below the $x-y$ diagonal are speedups, whilst points on the top and right
-axes represent instances which timed out after one thousand seconds with one algorithm, but not the
-other. For satisfiable instances, the different point styles show the different families, whilst all
-unsatisfiable instances are shown as dark dots.  The points well below the diagonal line and along
-the right-hand axis on the scatter plot show that DDS can sometimes be extremely beneficial on
-satisfiable instances.  However, on both unsatisfiable and most satisfiable instances, the overheads
-can be extremely large, and DDS is much worse in aggregate and is not a viable approach (even when
-only considering satisfiable instances).  These large overheads are to be expected: discrepancy
-searches are aimed primarily at getting better feasible solutions in optimisation problems which are
-too large for a proof of optimality to be a realistic prospect, and they are not well-suited for
-unsatisfiable decision problems. Despite this, the extremely large gains on some satisfiable
-instances confirm our suspicions that we should find an alternative to heavy-commitment backtracking
-search.
+are similar). On the scatter plot, each point represents the solving time for one instance---to
+avoid noise for easier instances, we measure only time spent during search, and exclude time spent
+in preprocessing and initialisation.  Points below the $x-y$ diagonal are speedups, whilst points on
+the top and right axes represent instances which timed out after one thousand seconds with one
+algorithm, but not the other. For satisfiable instances, the different point styles show the
+different families, whilst all unsatisfiable instances are shown as dark dots.  The points well
+below the diagonal line and along the right-hand axis on the scatter plot show that DDS can
+sometimes be extremely beneficial on satisfiable instances.  However, on both unsatisfiable and most
+satisfiable instances, the overheads can be extremely large, and DDS is much worse in aggregate and
+is not a viable approach (even when only considering satisfiable instances).  These large overheads
+are to be expected: discrepancy searches are aimed primarily at getting better feasible solutions in
+optimisation problems which are too large for a proof of optimality to be a realistic prospect, and
+they are not well-suited for unsatisfiable decision problems. Despite this, the extremely large
+gains on some satisfiable instances confirm our suspicions that we should find an alternative to
+heavy-commitment backtracking search.
 
 \subsection{Value-Ordering Heuristics}
-
-?? Lead in
-
-The left plot of \cref{figure:value-ordering-heuristics} shows the cumulative number of satisfiable
-instances solved over time using different value-ordering heuristics.
-
-\begin{description}
-\item[Degree] We select vertices from highest degree to lowest degree
-\cite{DBLP:conf/ijcai/McCreeshPT16}.
-
-\item[Random] We select vertices uniformly at random.
-
-\item[Anti] We select vertices from lowest degree to highest degree. This heuristic is used as
-a sanity check: as expected, degree beats random, and random beats anti.
-
-\item[Biased] We branch by selecting a vertex $v'$ from the chosen domain $D_v$ with
-probability \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \text{.} \]\end{description}
-
-\noindent
-The biased heuristic is new. It continues to prefer vertices of high degree, but will give the same
-chance of being selected next to two vertices of equal degree.  It also introduces an element of
-randomness, which will be needed to make restarts have an effect.  The performance of the degree and
-biased heuristics are very similar: the biased heuristic has slightly worse performance below a 400s
-timeout, and effectively equal performance above it. This shows that we can introduce an element of
-randomness into the degree value-ordering heuristic without adversely affecting its performance
-\emph{in aggregate} (and more generally, this scheme could be used for any value-ordering
-heuristic which puts choices in some kind of rank with ties.)
 
@@
 
+Traditionally, value-ordering heuristics are designed to drive search towards the most promising
+region of the search space ?? cite, or the most constrained ?? cite, or the region with the highest solution
+density ?? cite. In subgraph isomorphism, this is done by selecting vertices from highest degree to
+lowest \cite{DBLP:journals/jair/McCreeshPST18}. The left-hand plot of
+\cref{figure:value-ordering-heuristics} demonstrates that this is indeed a good choice: the
+\textbf{Degree} heuristic's line shows much better performance on satisfiable instances than the
+\textbf{Random} (branch randomly) or \textbf{Anti} (branch from lowest degree to highest) heuristic
+lines. (On unsatisfiable instances, the value-ordering heuristic has no effect on performance,
+because a complete search must be performed.)
+
+But what happens if our value-ordering heuristic has to choose between mapping a pattern vertex to,
+for example, three vertices of degree ten, two vertices of degree nine, or five of degree two? When
+driving conventional backtracking search, the degree heuristic would pick one of the vertices of
+degree ten, and we would commit all of our search effort to the exponentially large search tree
+underneath it.  We will show that this is not a wise choice, and that instead, we should \emph{commit
+equal search effort} to each of the three subproblems found by mapping to vertices of degree ten.
+And similarly, should we be certain that vertex of degree ten is so much better than a vertex of
+degree nine that we should commit no effort to degree nine vertices until all the degree ten
+subproblems have been explored? Or might it be better to commit, say, twice as much effort to each
+degree ten subproblems as to each degree nine subproblems, and only a very small amount of effort to
+the degree two subproblems?  To test this hypothesis, we will now introduce a new alternative to
+backtracking search, which we call \emph{solution-biased search}.  This search is made up of three
+components: a new slightly-random value-ordering heuristic, rapid restarts, and nogood recording.
+
+\subsection{Biased Value-Ordering}
+
+We define a new \textbf{Biased} value-ordering heuristic, as follows. When branching, we select a
+vertex $v'$ from the chosen domain $D_v$ with probability \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in
+D_v}{2^{\deg(w)}}} \text{.} \] This heuristic is now equally likely to pick between vertices of
+equal degree, is twice as likely to pick a vertex of degree $d$ as one of degree $d - 1$, and is
+over a thousand times more likely to pick a vertex of degree $d$ than degree $d - 10$.
+\Cref{figure:value-ordering-heuristics} shows that the performance of the degree and biased
+heuristics are very similar: the biased heuristic has slightly worse performance below a ??
+timeout, and effectively equal performance above it. This shows that we can introduce an element of
+randomness into the degree value-ordering heuristic without adversely affecting its performance
+\emph{in aggregate}.
+
 The right-hand plot of \cref{figure:value-ordering-heuristics} gives a detailed comparison of the
-degree and biased heuristics. It shows that despite the aggregate
-performance being similar, on a case by case basis, the two heuristics can make a large difference
-to the performance of individual satisfiable instances. This suggests that although degree is a good
-heuristic, we should perhaps not commit heavily to a vertex of highest degree, but also consider
-vertices of the same or similar degree.
+degree and biased heuristics. It shows that despite the aggregate performance being similar, on a
+case by case basis, the two heuristics can make a large difference to the performance of individual
+satisfiable instances. This suggests that although degree is a good heuristic, we should perhaps not
+commit heavily to a vertex of highest degree, but also consider vertices of the same or similar
+degree.
 
-?? Intuition
+\subsection{Restarting Search and Nogood Recording}
 
-\subsection{Adding Restarts and Nogood Recording}
-
-Another alternative to plain backtracking search is provided by \emph{restarts}. The general idea is
-to perform a certain amount of search, and then if no solution has been found (and unsatisfiability
-has not been proven), to abandon search and restart from the beginning. Such an approach
-can only be beneficial if something changes after restarting---in a constraint programming setting,
-this is usually the variable-ordering heuristic
+Having introduced a new value-ordering heuristic, we must now also move away from depth-first
+backtracking search. We do this by using \emph{restarts} and \emph{nogood recording}. The general
+idea is to perform a certain amount of search, and then if no solution has been found (and
+unsatisfiability has not been proven), to abandon search and restart from the beginning. Such an
+approach can only be beneficial if something changes after restarting---in a constraint programming
+setting, this is usually the variable-ordering heuristic
 \cite{DBLP:journals/jsat/LecoutreSTV07,DBLP:conf/cp/GayHLS15,DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17}.
-In this paper, we instead rely upon randomness in a special new \emph{value}-ordering heuristic, and continue
-to use smallest domain first with static tiebreaking for variable-ordering.\footnote{It may be
-possible to further improve the solver by also introducing randomness or some form of
-learning into its variable-ordering heuristic. However, simultaneously introducing a second change
-would considerably complicate the empirical analysis. Additionally, the solver's current
-hand-crafted variable-ordering heuristics already beat adaptive heuristics like impact or
-activity-based search.} Using restarts on value-ordering heuristics is uncommon (although
-\citet{Razgon et al.\ }{DBLP:conf/flairs/RazgonOP07} look at learning value-ordering heuristics from
-restarts, \citet{Chu et al.\ }{DBLP:conf/cp/ChuSS09} use a similar scheme in the context of parallel
-search, and an early approach by \citet{Gomes et al.\ }{DBLP:conf/aaai/GomesSK98} does so in an
-optimisation context).
+In this paper, we instead rely upon randomness in our new \emph{value}-ordering heuristic, and
+continue to use smallest domain first with static tiebreaking for variable-ordering.\footnote{It may
+be possible to further improve the solver by also introducing randomness or some form of learning
+into its variable-ordering heuristic. However, simultaneously introducing a second change would
+considerably complicate the empirical analysis. Additionally, the solver's current hand-crafted
+variable-ordering heuristics already beat adaptive heuristics like impact or activity-based search.}
+Using restarts on value-ordering heuristics is uncommon (although \citet{Razgon et al.\
+}{DBLP:conf/flairs/RazgonOP07} look at learning value-ordering heuristics from restarts, \citet{Chu
+et al.\ }{DBLP:conf/cp/ChuSS09} use a similar scheme in the context of parallel search, and an early
+approach by \citet{Gomes et al.\ }{DBLP:conf/aaai/GomesSK98} does so in an optimisation context).
 
@@
 % assignment.
+
 To avoid exploring portions of the search
@@
 decision to restart before backtracking, we post a similar nogood eliminating the entire subtree
-explored.
-
-Finally, we use the two watched literals technique \cite{DBLP:conf/dac/MoskewiczMZZM01} to
+explored. We use the two watched literals technique \cite{DBLP:conf/dac/MoskewiczMZZM01} to
 propagate stored nogoods.
@@
 generation and propagation schemes exist \cite{DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17},
-but these would not be helpful here since our solver does not maintain arc consistency or use a
-propagation queue.
+but these are not helpful in this setting (our solver does not maintain arc consistency or use a
+propagation queue).
 
-\subsection{The Effects of Restarts and Nogood Recording}
+\subsection{Solution-Biased Search in Practice}
 
@@
 
-    \caption{On the left, the number of satisfiable instances solved over time, with and without
-    restarts and nogood recording. To the right, a comparison between the original algorithm and the
-    new algorithm with all features enabled.}
+    \caption{On the left, the number of satisfiable instances solved over time, comparing
+    solution-biased search (SBS), random search with restarts (RSR), and the three value-ordering
+    heuristics with conventional depth-first search. To the right, a comparison between the original
+    algorithm and solution-biased search.}
     \label{figure:old-vs-new}
@@
 In \cref{figure:old-vs-new} we show the effects of adding restarts and nogood recording to the
-algorithm. With restarts and nogood recording, the random value-ordering heuristic comfortably beats
-the degree strategy with simple backtracking search. In other words, although having a good
-value-ordering heuristic is beneficial, introducing randomness into the search is better, if it is
-done alongside a mechanism to avoid heavy commitment to any particular random choice.
-\Cref{figure:old-vs-new} also shows that the biased heuristic together with restarts is better
-still---that is, if we are introducing restarts, then it is better to add a small amount of
-randomness to a tailored heuristic than it is to throw away the heuristic altogether.  Indeed, the
-original algorithm can solve ?? satisfiable instances by ?? seconds, whilst the biased and
-random restarting algorithms require only ?? seconds and ?? seconds respectively to solve the
-same number.
+algorithm. With restarts and nogood recording (random search with restarts, RSR), the random
+value-ordering heuristic comfortably beats the degree strategy with depth-first search. In other
+words, although having a good value-ordering heuristic is beneficial, introducing randomness into
+the search is better, if it is done alongside a mechanism to avoid heavy commitment to any
+particular random choice. However, the biased heuristic together
+with restarts (solution-biased search, SBS) is better still---that is, if we are introducing
+restarts, then it is better to add a small amount of randomness to a tailored heuristic than it is
+to throw away the heuristic altogether.  Indeed, the original algorithm can solve ?? satisfiable
+instances by ?? seconds, whilst the biased and random restarting algorithms require only ?? seconds
+and ?? seconds respectively to solve the same number.
 
-In the more detailed view in the right-hand plot of \cref{figure:old-vs-new}, comparing the basic
-algorithm with the degree heuristic to the degree-biased algorithm with restarts, all of the
-unsatisfiable instances are very close to the $x-y$ diagonal, showing that their performance is
-nearly unchanged. On the other hand, there are large numbers of satisfiable instances well below the
-diagonal line, indicating large speedups.  Better yet, there are only a handful of satisfiable
-instances that are more than a factor of ten times worse.  In other words, as well as improving
-performance, we have made up most of the consistency we lost by introducing randomness.
-This is a promising result: for the first time, we have introduced a viable alternative to
-backtracking search in a setting where we both have a strong \emph{value}-ordering heuristic,
-\emph{and} care about both satisfiable and unsatisfiable instances.
+In the more detailed view in the right-hand plot of \cref{figure:old-vs-new}, comparing the original
+algorithm to solution-biased search, all of the unsatisfiable instances are very close to the $x-y$
+diagonal, showing that their performance is nearly unchanged. On the other hand, there are large
+numbers of satisfiable instances well below the diagonal line, indicating large speedups.  Better
+yet, there are only a handful of satisfiable instances that are more than a factor of ten times
+worse.  In other words, as well as improving performance, we have made up most of the consistency we
+lost by introducing randomness.  This is a promising result: for the first time, we have introduced
+a viable alternative to backtracking search in a setting where we both have a strong
+\emph{value}-ordering heuristic, \emph{and} care about both satisfiable and unsatisfiable instances.
 
@@
 
+\subsection{Why ``Solution-Biased'', and why Exponential Bias?}
+
+?? Although we have shown that it provides good results, we have yet to justify where the biased
+formula comes from, or indeed why we call this approach ``solution-biased''. The crux of the
+matter is that whilst we do not know where solutions actually lie in the search space, we \emph{do}
+know that subproblems under a target vertex of degree $d$ tend to contain a small constant times
+more search nodes than those under a target vertex of degree $d - 1$.
+
 \section{Parallel Search}
+
+?? No sharing, sharing with Luby restarts, sharing with constant restarts, sharing with triggered
+restarts
+
+?? Reproducibility, constant and triggered
+
+?? How many more nodes on unsat instances
+
+?? Scalability to MPI
 
@@
 % \end{figure}
-
-?? No sharing, sharing with Luby restarts, sharing with constant restarts, sharing with triggered
-restarts
-
-?? Reproducibility, constant and triggered
-
-?? How many more nodes on unsat instances
-
-?? Scalability to MPI
 

### 2018-11-23 paper/paper.tex
Progress

@@
 hard to see the differences between the stronger solvers at higher runtimes. This paper focusses
-upon improving the performance on the remaining hard satisfiable instances, and so in the second and
-fourth plots in \cref{figure:others} (and in subsequent cumulative plots) we show only satisfiable
-instances, and use a reduced range on both axes.  For the remainder of this paper, we show only
-the non-induced problem, which tends to be harder; results with the induced variant are similar.
+upon improving the performance on the remaining hard satisfiable instances, and so in the bottom two
+plots in \cref{figure:others} (and in subsequent cumulative plots) we show only satisfiable
+instances, and use a reduced range on both axes.  For the remainder of this paper, we show only the
+non-induced problem, which tends to be harder; results with the induced variant are similar.
 

### 2018-11-25 paper/paper.tex
Nearly there

@@
     performance on unsatisfiable instances. This search can be parallelised in a very simple
-    but effective way, giving us a further speedup of over ?? from thirty-six cores, across both
-    satisfiable and unsatisfiable instances.  We also show that solution-biased search is also
-    suitable for optimisation problems, by using it to improve two maximum common induced subgraph
-    algorithms.
+    but effective way: across both satisfiable and unsatisfiable instances, we get a further speedup
+    of over thirty from thirty-six cores, or one hundred from ten distributed-memory hosts.  We also
+    show that solution-biased search is also suitable for optimisation problems, by using it to
+    improve two maximum common induced subgraph algorithms.
 \end{abstract}
@@
 \begin{figure}[tb]
-    \includegraphics*{gen-graph-others.pdf}
+    \includegraphics{gen-graph-others.pdf}
     \hfill
-    \includegraphics*{gen-graph-others-induced.pdf}
+    \includegraphics{gen-graph-others-induced.pdf}
 
@@
 
-    \includegraphics*{gen-graph-others-zoom.pdf}
+    \includegraphics{gen-graph-others-zoom.pdf}
     \hfill
-    \includegraphics*{gen-graph-others-induced-zoom.pdf}
+    \includegraphics{gen-graph-others-induced-zoom.pdf}
 
@@
 \begin{figure}[tb]
-    \includegraphics*{gen-graph-dds.pdf}
+    \includegraphics{gen-graph-dds.pdf}
     \hfill
-    \includegraphics*{gen-graph-scatter-dds.pdf}
+    \includegraphics{gen-graph-scatter-dds.pdf}
 
@@
 \begin{figure}[tb]
-    \includegraphics*{gen-graph-value-ordering-heuristics.pdf}
+    \includegraphics{gen-graph-value-ordering-heuristics.pdf}
     \hfill
-    \includegraphics*{gen-graph-scatter-heuristics.pdf}
+    \includegraphics{gen-graph-scatter-heuristics.pdf}
 
@@
 \begin{figure}[tb]
-    \includegraphics*{gen-graph-restarts.pdf}
+    \includegraphics{gen-graph-restarts.pdf}
     \hfill
-    \includegraphics*{gen-graph-scatter-by-family.pdf}
+    \includegraphics{gen-graph-scatter-by-family.pdf}
 
@@
 \begin{figure}[tb]
-    \includegraphics*{gen-graph-scatter-biased-goods.pdf}
+    \includegraphics{gen-graph-scatter-biased-goods.pdf}
     \hfill
-    \includegraphics*{gen-graph-scatter-random.pdf}
+    \includegraphics{gen-graph-scatter-random.pdf}
     \caption{On the left, not using nogood recording introduces slowdowns, particularly on
@@
 \begin{figure}[tb]
-    \centering
-    \includegraphics*{gen-graph-parallel.pdf}
+    \includegraphics{gen-graph-parallel.pdf}
 
@@
 
-    \includegraphics*{gen-graph-scatter-par.pdf}
+    \includegraphics{gen-graph-scatter-par.pdf}
     \hfill
-    \includegraphics*{gen-graph-scatter-parconsttick.pdf}
+    \includegraphics{gen-graph-scatter-dist.pdf}
 
-    \caption{Parallel}\label{figure:parallel}
+    \caption{Above, the cumulative number of instances solved over time, comparing the sequential
+    algorithm to results using 32 threads on a single machine, and using five or ten distributed
+    memory hosts. Below, instance by instance comparisons.}\label{figure:parallel}
 \end{figure}
-
-% \begin{figure}[tb]
-%     \includegraphics*{gen-graph-scatter-par-repeat.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-scatter-parconsttick-repeat.pdf}
-% 
-%     \includegraphics*{gen-graph-scatter-parconst-nodes.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-scatter-parconsttick-nodes.pdf}
-% 
-%     \caption{Parallel nodes}\label{figure:parallelnodes}
-% \end{figure}
 
@@
     \centering
-    \includegraphics*{gen-graph-mcs.pdf}
+    \includegraphics{gen-graph-mcs.pdf}
 
@@
 
-    \includegraphics*{gen-graph-scatter-kdown.pdf}
+    \includegraphics{gen-graph-scatter-kdown.pdf}
     \hfill
-    \includegraphics*{gen-graph-scatter-mcsplit.pdf}
+    \includegraphics{gen-graph-scatter-mcsplit.pdf}
 

### 2018-11-26 paper/paper.tex
More things

@@
     but effective way: across both satisfiable and unsatisfiable instances, we get a further speedup
-    of over thirty from thirty-six cores, or one hundred from ten distributed-memory hosts.  We also
+    of over thirty from thirty six cores, or ninety five from ten distributed-memory hosts.  We also
     show that solution-biased search is also suitable for optimisation problems, by using it to
@@
 The subgraph isomorphism problem is to decide whether a copy of a small ``pattern'' graph occurs
-inside a larger ``target'' graph. Although the problem is \NP-complete, by combining design
-techniques from artificial intelligence with careful algorithm engineering, modern subgraph
-isomorphism solvers can often produce exact solutions quickly even on graphs with thousands of
-vertices. The current single strongest subgraph isomorphism solver uses ``highest degree first'' as
-a value-ordering heuristic to direct a constraint programming style search
+inside a larger ``target'' graph.  ?? Applications
+
+Although the problem is \NP-complete, by combining design techniques from artificial intelligence
+with careful algorithm engineering, modern subgraph isomorphism solvers can often produce exact
+solutions quickly even on graphs with thousands of vertices. The current single strongest subgraph
+isomorphism solver uses ``highest degree first'' as a value-ordering heuristic to direct a
+constraint programming style search
 \cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:journals/jair/McCreeshPST18}. This
@@
 algorithm perform two orders of magnitude better on a large number of satisfiable instances, whilst
-performing worse only rarely on satisfiable instances, and never on unsatisfiable instances.
+performing worse only rarely on satisfiable instances, and never on unsatisfiable instances.  This
+strategy is also effective in an optimisation setting, producing benefits in two maximum common
+induced subgraph algorithms. Through this combination of techniques we have, for the first time, a
+practical alternative to backtracking search for constraint programming style algorithms where we
+have a strong value-ordering heuristic and where we care both about satisfiable and unsatisfiable
+instances.
 
-This new form of search can also be parallelised. By running many threads with different random
-seeds but the same restart schedule, and sharing nogoods only following restarts, we can achieve
-speedups of ??, with a \emph{much} simpler implementation than conventional work-stealing.  Finally,
-we show that this strategy is also effective in an optimisation setting, producing benefits in two
-maximum common induced subgraph algorithms. This combination of techniques gives us, for the first
-time, a practical alternative to backtracking search for constraint programming style algorithms
-where we have a strong value-ordering heuristic and where we care both about satisfiable and
-unsatisfiable instances.
+This new form of search can also be parallelised, with a \emph{much} simpler implementation than
+conventional work-stealing. By running many threads with different random seeds but the same restart
+schedule, and sharing nogoods only following restarts, we can achieve speedups of thirty one from a
+thirty six core machine, or ninety five by using ten such machines with message passing.
 
@@
 non-adjacent vertices.
-
-?? Applications
 
@@
         problem \cite{DBLP:journals/pr/SolnonDHJ15}.
-    \item[Images and Meshes] contain graphs representing 2D segmented images and 3D object models
-        respectively, again representing a computer vision problem
-        \cite{DBLP:journals/cviu/DamiandSHJS11}.
+    \item[Images and Meshes] contain graphs representing 2D segmented images and 3D object models,
+        again representing a computer vision problem \cite{DBLP:journals/cviu/DamiandSHJS11}.
 \end{description}
@@
 
-We begin with a set of baseline performance measurements. In the first and
-third plots \cref{figure:others} we show the cumulative number of instances solved over time for the
+We begin with a set of baseline performance measurements. In the top
+two plots \cref{figure:others} we show the cumulative number of instances solved over time for the
 non-induced and induced problems respectively. We compare the Glasgow Subgraph Solver using
@@
 Traditionally, value-ordering heuristics are designed to drive search towards the most promising
-region of the search space ?? cite, or the most constrained ?? cite, or the region with the highest solution
-density ?? cite. In subgraph isomorphism, this is done by selecting vertices from highest degree to
-lowest \cite{DBLP:journals/jair/McCreeshPST18}. The left-hand plot of
+region of the search space \cite{DBLP:conf/ecai/Geelen92}, or the most constrained
+\cite{DBLP:conf/aaai/GentMPW96}, or the region with the highest solution density
+\cite{DBLP:journals/jair/PesantQZ12}. In subgraph isomorphism, this is done by selecting vertices
+from highest degree to lowest \cite{DBLP:journals/jair/McCreeshPST18}. The left-hand plot of
 \cref{figure:value-ordering-heuristics} demonstrates that this is indeed a good choice: the
@@
 \textbf{Random} (branch randomly) or \textbf{Anti} (branch from lowest degree to highest) heuristic
-lines. (On unsatisfiable instances, the value-ordering heuristic has no effect on performance,
-because a complete search must be performed.)
+lines. Meanwhile, on unsatisfiable instances, the value-ordering heuristic has no effect on
+performance, because a complete search must be performed.
 
 But what happens if our value-ordering heuristic has to choose between mapping a pattern vertex to,
-for example, three vertices of degree ten, two vertices of degree nine, or five of degree two? When
-driving conventional backtracking search, the degree heuristic would pick one of the vertices of
-degree ten, and we would commit all of our search effort to the exponentially large search tree
-underneath it.  We will show that this is not a wise choice, and that instead, we should \emph{commit
-equal search effort} to each of the three subproblems found by mapping to vertices of degree ten.
-And similarly, should we be certain that vertex of degree ten is so much better than a vertex of
-degree nine that we should commit no effort to degree nine vertices until all the degree ten
-subproblems have been explored? Or might it be better to commit, say, twice as much effort to each
-degree ten subproblems as to each degree nine subproblems, and only a very small amount of effort to
-the degree two subproblems?  To test this hypothesis, we will now introduce a new alternative to
-backtracking search, which we call \emph{solution-biased search}.  This search is made up of three
-components: a new slightly-random value-ordering heuristic, rapid restarts, and nogood recording.
+for example, three target vertices of degree ten, two vertices of degree nine, or five of degree
+two? When driving conventional backtracking search, the degree heuristic would pick one of the
+vertices of degree ten, and we would commit all of our search effort to the exponentially large
+search tree underneath it, not considering any other choice until this tree has been fully explored
+and eliminated.  We will show that this is not a wise choice, and that instead, we should
+\emph{commit equal search effort} to each of the three subproblems found by mapping to vertices of
+degree ten.  And similarly, should we be certain that vertex of degree ten is so much better than a
+vertex of degree nine that we should commit no effort to degree nine vertices until all the degree
+ten subproblems have been explored? Or might it be better to commit, say, twice as much effort to
+each degree ten subproblems as to each degree nine subproblems, and only a very small amount of
+effort to the degree two subproblems?  To test this hypothesis, we will now introduce a new
+alternative to backtracking search, which we call \emph{solution-biased search}.  This search is
+made up of three components: a new slightly-random value-ordering heuristic, rapid restarts, and
+nogood recording. The aim is to perform a complete search, but spending proportionally more time in
+parts of the search tree that are preferred by the value-ordering heuristic.
 
@@
 
-We define a new \textbf{Biased} value-ordering heuristic, as follows. When branching, we select a
-vertex $v'$ from the chosen domain $D_v$ with probability \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in
-D_v}{2^{\deg(w)}}} \text{.} \] This heuristic is now equally likely to pick between vertices of
-equal degree, is twice as likely to pick a vertex of degree $d$ as one of degree $d - 1$, and is
-over a thousand times more likely to pick a vertex of degree $d$ than degree $d - 10$.
-\Cref{figure:value-ordering-heuristics} shows that the performance of the degree and biased
-heuristics are very similar: the biased heuristic has slightly worse performance below a ??
-timeout, and effectively equal performance above it. This shows that we can introduce an element of
-randomness into the degree value-ordering heuristic without adversely affecting its performance
-\emph{in aggregate}.
+Firstly, we define a new \textbf{Biased} value-ordering heuristic, as follows. When branching, we
+select a vertex $v'$ from the chosen domain $D_v$ with probability \[ p(v') =
+\frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \text{.} \] This heuristic is now equally likely
+to pick between vertices of equal degree, is twice as likely to pick a vertex of degree $d$ as one
+of degree $d - 1$, and is over a thousand times more likely to pick a vertex of degree $d$ than
+degree $d - 10$.
 
-The right-hand plot of \cref{figure:value-ordering-heuristics} gives a detailed comparison of the
-degree and biased heuristics. It shows that despite the aggregate performance being similar, on a
-case by case basis, the two heuristics can make a large difference to the performance of individual
-satisfiable instances. This suggests that although degree is a good heuristic, we should perhaps not
-commit heavily to a vertex of highest degree, but also consider vertices of the same or similar
-degree.
+\Cref{figure:value-ordering-heuristics} confirms that this heuristic, when used with backtracking
+search, will solve close to the same number of instances as the degree heuristic would for any given
+choice of timeout. In other words, we can introduce an element of randomness into the degree
+value-ordering heuristic without adversely affecting its performance \emph{in aggregate}.  The
+right-hand plot gives a detailed comparison. It shows that despite the aggregate performance being
+similar, on a case by case basis, the two heuristics can make a large difference to the performance
+for individual satisfiable instances. This justifies our belief that although degree is a good
+heuristic, we should perhaps not commit heavily to a vertex of highest degree, but also consider
+vertices of the same or similar degree.
 
@@
 restarts, then it is better to add a small amount of randomness to a tailored heuristic than it is
-to throw away the heuristic altogether.  Indeed, the original algorithm can solve ?? satisfiable
-instances by ?? seconds, whilst the biased and random restarting algorithms require only ?? seconds
-and ?? seconds respectively to solve the same number.
+to throw away the heuristic altogether.  Indeed, the original algorithm can solve 1983 satisfiable
+instances by 909 seconds, whilst the biased and random restarting algorithms require only 12 seconds
+and 35 seconds respectively to solve the same number.
 
@@
 
-?? No sharing, sharing with Luby restarts, sharing with constant restarts, sharing with triggered
-restarts
-
-?? Reproducibility, constant and triggered
-
-?? How many more nodes on unsat instances
-
-?? Scalability to MPI
-
 \begin{figure}[tb]
@@
 \end{figure}
+
+Exploiting multiple cores to speed up constraint programming solvers remains an active area of
+research, with no universally perfect solution being available. Four of the more common approaches
+are based upon decompositions
+\cite{DBLP:journals/corr/abs-1008-4328,DBLP:journals/jair/MalapertRR16}, work stealing
+\cite{DBLP:conf/cp/MichelSH07,DBLP:conf/cp/ChuSS09,DBLP:conf/cp/McCreeshP15,DBLP:conf/cpaior/HoffmannMNPRS018},
+parallel discrepancy searches \cite{DBLP:conf/cp/MoisanGQ13,DBLP:conf/cpaior/MoisanQG14}, and
+algorithm portfolios \cite{DBLP:conf/lion/LindauerHH15}. Decomposition approaches are unsuitable for
+decision problems, or problems where we have good value-ordering heuristics, because the
+decomposition interferes strongly with the shape of the search tree
+\cite{DBLP:journals/jair/MalapertRR16}. Work stealing, traditionally, also interferes with
+value-ordering \cite{DBLP:journals/topc/McCreeshP15}, although specially designed exceptions exist
+\cite{DBLP:conf/cp/ChuSS09,DBLP:conf/cp/McCreeshP15,DBLP:conf/cpaior/HoffmannMNPRS018}. However,
+these have very complicated implementations. Parallel discrepancy searches are aware of
+value-ordering heuristics, but have other limitations: they struggles on search trees with heavy
+filtering, and rely upon inner search tree nodes being orders of magnitude less expensive to process
+than leaf nodes. Portfolios, meanwhile, typically rely upon running multiple models or heuristics
+simultaneously, and selecting whichever finishes first, whereas here we have a known good model and
+set of heuristics.
+
+\subsection{Shared Memory Parallelism}
+
+Solution-biased search allows for a much simpler parallel implementation, which we now evaluate.  We
+create a number of threads, and give each thread its own random seed, but otherwise each thread
+performs the same sequential search. Then, whenever a thread restarts, it uses a barrier to wait for
+every other thread to also restart. Each thread's new nogoods from that restart can then be
+gathered, and then each thread resumes, with a larger set of nogoods than would have been generated
+in a sequential run. Finally, whenever any single thread terminates, either due to having found a
+solution or proved unsatisfiability, then every other thread may immediately terminate.
+
+A number of factors combine to make this technique feasible:
+
+\begin{itemize}
+    \item Each thread will be run with the same restart schedule, and so will spend approximately
+        the same amount of time between restarts. Because the only synchronisation between threads
+        is at restarts, we expect threads to be kept busy doing useful work.  (This is in contrast
+        to an alternative method for paralellising restarting search \cite{DBLP:conf/aaai/CireKS14},
+        which packs together successive sequence values to produce a balanced workload.)
+    \item Sequentially, on non-trivial instances the algorithm will restart very many times (many
+        tens of thousands, for instances that reach the thousand second timeout).
+    \item Because the search trees we explore are exponentially large, the randomness in the
+        value-ordering heuristic is sufficient to ensure that most of the time, threads are
+        exploring different parts of the search tree.
+    \item The gathering of nogoods to describe the work done so far provides an alternative to
+        requiring either a specific mechanism to allocate work, or expensive synchronisation between threads.
+    \item If sometimes threads do happen to explore part of
+        the same subproblems, this is not a problem: if the instance is satisfiable, either thread
+        might find a solution first, and if the instance is unsatisfiable, we merely introduce some
+        redundancy into the proof.\footnote{For solving a counting or enumeration problem,
+        matters become slightly more complicated, but not devastatingly so.} The combination of
+        rapid restarts and nogood recording is enough to ensure that this is only a small overhead.
+\end{itemize}
+
+\Cref{figure:parallel} shows how this scheme performs in practice. Sequentially, we can solve 14,357
+instances within the thousand second timeout, with the last instance being solved at 939.0 seconds.
+Using thirty six threads on machines with two eighteen core processors, we can solve 14,357
+instances with a timeout of only 74.2 seconds, giving an aggregate speedup of 12.7.
+
+Closer inspection of the results reveals that with this many threads, a considerable proportion of
+the overall search time is spent with threads waiting at the barrier for synchronisation. This is
+because the time taken to carry out search until 660 backtracks are encountered is only
+\emph{roughly} a constant (in practice it usually varies by around a factor of two). Furthermore,
+the Luby sequence includes occasional large multipliers, and if unsatisfiability is proved during
+towards the end of one of these runs, each thread will end up duplicating a large amount of work.
+
+Because we are using nogood recording, an alternative approach is possible. Rather than using the
+Luby sequence for restarts, we could restart after either a constant number of backtracks, or simply
+after a certain time interval has passed. \Cref{figure:parallel} also shows the effects of restarting
+every 100ms. Sequentially, this approach is slightly better than using the Luby sequence, being able to
+solve 14,370 instances, with the last at 996.4 seconds. With thirty six threads, solving this many
+instances takes 31.8 seconds, giving an aggregate speedup of 31.4.
+
+It is important to emphasise that there is no expectation of a linear speedup from this approach,
+particularly for satisfiable instances. Due to the random nature of the value-ordering heuristic,
+one of our extra threads may ``get lucky'' and find a solution much quicker than we would
+sequentially, leading to a superlinear speedup. Conversely, our extra threads may contribute nothing to finding a
+solution---or worse, due to new nogoods altering the choice made by the random branching heuristic,
+we could even get an absolute slowdown. We can see both of these effects in the bottom left plot of
+\cref{figure:parallel}: although we achieve roughly a linear speedup on unsatisfiable instances,
+satisfiable instances show much greater variability. However, this approach does at least mirror our intuition of
+allocating search effort in proportion to where the value-ordering heuristic believes it will be
+most fruitful, and so we should not be too surprised that we see roughly a linear speedup on average
+on harder instances.
+
+Additionally, for easy instances, most of the algorithm's execution time is spent in a preprocessing
+phase. We have not parallelised this, which is why our results are poor below the one second mark.
+
+\subsection{Distributed Memory Parallelism}
+
+To further test the scalability of this technique, we also implemented a distributed-parallelism
+layer on top of the threaded layer, using MPI. In contrast to the huge difficulties of implementing
+work-stealing in a distributed memory setting, this required only the addition of a simple ``all
+gather'' operation to communicate nogoods, and a ``gather'' to collect and combine the results of
+each host. Termination, meanwhile, was handled by posting an empty nogood (and so termination can
+only occur on a restart).
+
+\Cref{figure:parallel} also shows the results of these experiments, using five and ten hosts.
+Because each host has two sockets, the five host results use ten MPI ranks, each with eighteen
+threads, and the ten host results use twenty MPI ranks. Because the supercomputing service we use is
+not designed for huge numbers of very short problems, we ran only instances whose sequential runtime
+was at least one second, and for the sake of plotting results, treat skipped instances as taking one
+second. Due to the job launcher used, it is also not possible to accurately measure ``total''
+runtime including startup costs, and so instead we report the runtime of the rank zero
+process---this figure is somewhat optimistic for very easy instances. With these caveats in mind, we
+see that our results continue to improve as the host count increases, and we come very close to the
+magic goal of an overall speedup of one hundred.
 
@@
 satisfiable and easy), but gives large benefits on many more instances than it penalises: it is over
-an order of magnitude better on nearly a hundred instances, whilst being an order of magnitude worse
+an order of magnitude better on over three hundred instances, whilst being an order of magnitude worse
 on only seven.
@@
 
-\section*{Acknowledgements}
-
-The authors would like to thank Ruth Hoffmann, Christophe Lecoutre, Christine Solnon, and Craig
-Reilly for their comments.
+% \section*{Acknowledgements}
+% 
+% The authors would like to thank Ozgur Akgun, Chris Jefferson, Christophe Lecoutre, Christine
+% Solnon, and Craig Reilly for their comments.
 

### 2018-11-26 paper/paper.tex
Gratuitous referencification

@@
     evaluate a new approach, which we call ``solution-biased search''. By combining a
-    slightly-random value-ordering heuristic, rapid restarts, and nogood recording, we achieve two
-    orders of magnitude faster performance on satisfiable instances, whilst not affecting
-    performance on unsatisfiable instances. This search can be parallelised in a very simple
-    but effective way: across both satisfiable and unsatisfiable instances, we get a further speedup
-    of over thirty from thirty six cores, or ninety five from ten distributed-memory hosts.  We also
-    show that solution-biased search is also suitable for optimisation problems, by using it to
-    improve two maximum common induced subgraph algorithms.
+    slightly-random value-ordering heuristic, rapid restarts, and nogood recording, we design an
+    algorithm which instead uses degree to direct the proportion of search effort spent in different
+    subproblems. This increases performance by two orders of magnitude on satisfiable instances,
+    whilst not affecting performance on unsatisfiable instances. Such a such search can also be
+    parallelised in a very simple but effective way: across both satisfiable and unsatisfiable
+    instances, we get a further speedup of over thirty from thirty six cores, or ninety five from
+    ten distributed-memory hosts. Finally, we show that solution-biased search is also suitable for
+    optimisation problems, by using it to improve two maximum common induced subgraph algorithms.
 \end{abstract}
@@
 The subgraph isomorphism problem is to decide whether a copy of a small ``pattern'' graph occurs
-inside a larger ``target'' graph.  ?? Applications
+inside a larger ``target'' graph. The problem is broadly applicable, with applications in areas
+including in bioinformatics \cite{o:BonniciGPSF13}, chemistry \cite{o:Regin95}, computer vision
+\cite{DBLP:journals/cviu/DamiandSHJS11,DBLP:journals/pr/SolnonDHJ15}, law enforcement
+\cite{DBLP:journals/cacm/CoffmanGM04}, model checking \cite{DBLP:journals/tcs/SevegnaniC15},
+malware detection \cite{DBLP:conf/dimva/BruschiMM06}, compilers
+\cite{DBLP:conf/cgo/MurrayF12,DBLP:conf/cp/BlindellLCS15}, pattern recognition
+\cite{DBLP:journals/ijprai/FoggiaPV14}, program similarity
+comparison \cite{o:DallaPredaV17}, the design of mechanical locks
+\cite{DBLP:journals/siamrev/VomelLBF17}, and graph databases
+\cite{DBLP:journals/jair/McCreeshPST18}.
 
@@
 The \emph{induced} problem additionally requires that non-adjacent vertices are mapped to
-non-adjacent vertices.
+non-adjacent vertices. The \emph{degree} of a vertex is the number of other vertices to which it is
+adjacent.
 
@@
 
-    \caption{On the top row, the cumulative number of instances solved over time, comparing our algorithm
-    (both in its basic form, and with the improvements introduced in the remainder of the paper) to
-    other solvers, for the non-induced and induced problems. On the bottom row, the same,
-    considering only satisfiable instances.}
+    \caption{On the top row, the cumulative number of instances solved over time, comparing the
+    Glasgow Subgraph Solver (both in its basic form, and with the improvements introduced in the
+    remainder of the paper) to other solvers, for the non-induced and induced problems. On the
+    bottom row, the same, considering only satisfiable instances.}
     \label{figure:others}
@@
 
-\begin{figure}[tb]
-    \includegraphics{gen-graph-parallel.pdf}
-
-    \medskip
-
-    \includegraphics{gen-graph-scatter-par.pdf}
-    \hfill
-    \includegraphics{gen-graph-scatter-dist.pdf}
-
-    \caption{Above, the cumulative number of instances solved over time, comparing the sequential
-    algorithm to results using 32 threads on a single machine, and using five or ten distributed
-    memory hosts. Below, instance by instance comparisons.}\label{figure:parallel}
-\end{figure}
-
 Exploiting multiple cores to speed up constraint programming solvers remains an active area of
@@
 \subsection{Shared Memory Parallelism}
+
+\begin{figure}[tb]
+    \includegraphics{gen-graph-parallel.pdf}
+
+    \medskip
+
+    \includegraphics{gen-graph-scatter-par.pdf}
+    \hfill
+    \includegraphics{gen-graph-scatter-dist.pdf}
+
+    \caption{Above, the cumulative number of instances solved over time, comparing the sequential
+    algorithm to results using 32 threads on a single machine, and using five or ten distributed
+    memory hosts. Below, instance by instance comparisons.}\label{figure:parallel}
+\end{figure}
 

### 2018-11-26 paper/paper.tex
Dupe

@@
 inside a larger ``target'' graph. The problem is broadly applicable, with applications in areas
-including in bioinformatics \cite{o:BonniciGPSF13}, chemistry \cite{o:Regin95}, computer vision
+including in bioinformatics \cite{DBLP:journals/bmcbi/BonniciGPSF13}, chemistry \cite{o:Regin95}, computer vision
 \cite{DBLP:journals/cviu/DamiandSHJS11,DBLP:journals/pr/SolnonDHJ15}, law enforcement

### 2018-11-27 paper/paper.tex
More words

@@
 
-A \emph{discrepancy} is where search goes against a value-ordering heuristic.  Discrepancy searches
+A \emph{discrepancy} is where search goes against the advice of a value-ordering heuristic.
+Discrepancy searches
 \cite{DBLP:conf/ijcai/HarveyG95,DBLP:conf/aaai/Korf96,DBLP:conf/ijcai/Walsh97,DBLP:conf/cpaior/KarouiHLN07}
@@
 
-But what happens if our value-ordering heuristic has to choose between mapping a pattern vertex to,
-for example, three target vertices of degree ten, two vertices of degree nine, or five of degree
-two? When driving conventional backtracking search, the degree heuristic would pick one of the
-vertices of degree ten, and we would commit all of our search effort to the exponentially large
+But what happens if our value-ordering heuristic has to choose between mapping a pattern vertex to
+one of, for example, three target vertices of degree ten, two vertices of degree nine, or five of
+degree two? When driving conventional backtracking search, the degree heuristic would pick one of
+the vertices of degree ten, and we would commit all of our search effort to the exponentially large
 search tree underneath it, not considering any other choice until this tree has been fully explored
@@
 for individual satisfiable instances. This justifies our belief that although degree is a good
-heuristic, we should perhaps not commit heavily to a vertex of highest degree, but also consider
+heuristic, we should perhaps not commit heavily to a single vertex of highest degree, but also consider
 vertices of the same or similar degree.
@@
 
-\subsection{Why ``Solution-Biased'', and why Exponential Bias?}
+\subsection{Solution-Biased Search in Theory}
 
-?? Although we have shown that it provides good results, we have yet to justify where the biased
-formula comes from, or indeed why we call this approach ``solution-biased''. The crux of the
-matter is that whilst we do not know where solutions actually lie in the search space, we \emph{do}
-know that subproblems under a target vertex of degree $d$ tend to contain a small constant times
-more search nodes than those under a target vertex of degree $d - 1$.
+Although we have shown that it provides good results, we have yet to justify where the biased
+formula comes from, or indeed why we call this approach ``solution-biased''.  Our goal is to use
+biased randomness in a value-ordering heuristic to spend time in subproblems proportional to an
+estimate of their solution density \cite{DBLP:journals/jair/PesantQZ12}. ?? I feel like we still
+need a sentence here saying why this is a good idea.
+
+To estimate solution density, we need an estimate of how big different subproblems are likely to be,
+and of how many solutions each subproblem is likely to contain. Of course, obtaining exact (or even
+approximate) values for these figures is at least as hard as solving the problem in its entirety.
+For pairs of Erd\H{o}s-R\'enyi
+random graphs with large solution counts (i.e.\ chosen from within the ``easy satisfiable'' region
+\cite{DBLP:journals/jair/McCreeshPST18}), we observe a linear relationship between subproblem size
+and number of solutions. Thus, for graphs from this distribution, we need only an estimator of
+subproblem size. Measurements also suggest that, for pairs of Erd\H{o}s-R\'enyi graphs, subproblems
+under a target vertex of degree $d$ tend to contain a small constant times more search nodes than
+those under a target vertex of degree $d - 1$.  This empirical analysis suggests that an estimator
+that is exponential in $d$ will give our method the desired behaviour, at least for
+Erd\H{o}s-R\'enyi graphs. We expect it may be possible to derive better estimators for particular
+input classes, although over the full range of problem instances, we have verified that estimators
+of the form $e^d$ substantially outperform polynomial and factorial weightings.
 
@@
 now it has not been clear how to do so efficiently in practice, particularly when we already have
-strong search ordering heuristics and when are expecting to encounter large numbers of unsatisfiable
+strong search ordering heuristics and when we are expecting to encounter large numbers of unsatisfiable
 instances. Our results show that the key to success is a combination of aggressive restarts, nogood

### 2018-11-27 paper/paper.tex
Are we nearly there yet?

@@
 strategy is also effective in an optimisation setting, producing benefits in two maximum common
-induced subgraph algorithms. Through this combination of techniques we have, for the first time, a
-practical alternative to backtracking search for constraint programming style algorithms where we
-have a strong value-ordering heuristic and where we care both about satisfiable and unsatisfiable
-instances.
+induced subgraph algorithms.
 
@@
 biased randomness in a value-ordering heuristic to spend time in subproblems proportional to an
-estimate of their solution density \cite{DBLP:journals/jair/PesantQZ12}. ?? I feel like we still
-need a sentence here saying why this is a good idea.
+estimate of their solution density \cite{DBLP:journals/jair/PesantQZ12}. Such an approach is better
+than committing entirely to the area of maximum solution density because estimators only give a
+probability---although we may estimate that one subtree has twice the solution density of another,
+in reality the ``better'' subtree may contain no solutions at all.
 
@@
 
-We have shown that it is both possible and beneficial to introduce a small amount of randomness into
-backtracking search algorithms for hard subgraph problems, without compromising existing search
-order heuristics, and without paying a penalty for unsatisfiable instances. In principle it is
-well-known that introducing some form of randomness \cite{DBLP:conf/aaai/GomesSK98} or reduced
-commitment \cite{DBLP:conf/ijcai/HarveyG95} into backtracking search can be beneficial, but until
-now it has not been clear how to do so efficiently in practice, particularly when we already have
-strong search ordering heuristics and when we are expecting to encounter large numbers of unsatisfiable
-instances. Our results show that the key to success is a combination of aggressive restarts, nogood
-recording, and biasing an existing value-ordering heuristic with just a pinch of randomness.
+The conventional view of value-ordering heuristics is that they define a search order. We have
+proposed an alternative perspective, where value-orderings define a weighting specifying how much
+search effort should be put into different subproblems, based upon a rough estimate of solution
+densities. We have also shown how to turn this perspective into an algorithm, by combining a biased
+random value-ordering heuristic with rapid restarts and nogood recording. This combination of
+techniques gives us, for the first time, a practical alternative to backtracking search where we
+have a strong value-ordering heuristic and where we care both about satisfiable and unsatisfiable
+instances.  A further benefit is the ease with which such a search can be parallelised. By having
+each thread carry out the same search with a different random seed, and sharing nogoods only on
+restarts, we remove the need for intrusive changes to the core search algorithm, eliminate the
+irregularity problem, and still respect the advice of the value-ordering heuristic.
 
-?? Parallel
+We believe that these technique are broadly applicable, beyond subgraph algorithms, and we intend to
+implement them in a full constraint programming solver. We are also interested in making better use
+of statistical knowledge (either \emph{a priori} or learned during search) to further refine the
+biased randomisation process. And finally, we are trying hard to work out whether our new
+perspective also has some relevance to \emph{variable} ordering heuristics.
 
-?? Apply to CP in general
+\section*{Acknowledgements}
 
-% \section*{Acknowledgements}
-% 
-% The authors would like to thank Ozgur Akgun, Chris Jefferson, Christophe Lecoutre, Christine
-% Solnon, and Craig Reilly for their comments.
+The authors would like to thank Ozgur Akgun, Chris Jefferson, Christophe Lecoutre, Christine
+Solnon, and Craig Reilly for their comments.
 

### 2018-11-27 paper/paper.tex
Exciting

@@
 Using thirty six threads on machines with two eighteen core processors, we can solve 14,357
-instances with a timeout of only 74.2 seconds, giving an aggregate speedup of 12.7.
+instances with a timeout of only 74.2 seconds, giving an aggregate speedup
+\cite{DBLP:conf/cpaior/HoffmannMNPRS018} of 12.7.
 
@@
 \begin{figure}[tb]
-    \centering
     \includegraphics{gen-graph-mcs.pdf}

### 2018-11-28 paper/paper.tex
minor fixes to paper

@@
 algorithm perform two orders of magnitude better on a large number of satisfiable instances, whilst
-performing worse only rarely on satisfiable instances, and never on unsatisfiable instances.  This
+performing worse rarely on satisfiable instances, and never on unsatisfiable instances.  This
 strategy is also effective in an optimisation setting, producing benefits in two maximum common
@@
 
-This new form of search can also be parallelised, with a \emph{much} simpler implementation than
+This new form of search can also be parallelised, with a much simpler implementation than
 conventional work-stealing. By running many threads with different random seeds but the same restart
@@
 simplification and re-engineering of an older Glasgow algorithm
-\cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}.  Essentially, the solver is essentially
+\cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}.  Essentially, the solver is
 a dedicated forward-checking constraint programming implementation specifically for subgraph
@@
         These benchmark instances are widely used, but have unusual properties and so broad
-        conclusions should not be drawn based solely upon behaviour on these instances
+        conclusions should not be drawn based solely upon behaviour of these instances
         \cite{DBLP:journals/jair/McCreeshPST18}.
@@
        \cite{DBLP:journals/mscs/LarrosaV02} from the Stanford Graph Database. We include both the
-        50 small graphs, and the 50 larger graphs.
-    \item[Phase] contains hard crafted instances that lie near the satisfiable / unsatisfiable phase
+        50 small graphs, and the 50 large graphs.
+    \item[Phase] contains hand crafted instances that lie near the satisfiable / unsatisfiable phase
         transition \cite{DBLP:journals/jair/McCreeshPST18}.
@@
 space that we have already visited, every time we restart, we add new constraints to the problem
-which eliminate already-explored subtrees---such a constraint is called a nogood. We generate
+which eliminate already-explored subtrees---such a constraint is called a \emph{nogood}. We generate
 simple decision nogoods. That is, upon backtracking due to a decision to restart, we post a nogood
@@
 
-Solution-biased search allows for a much simpler parallel implementation, which we now evaluate.  We
-create a number of threads, and give each thread its own random seed, but otherwise each thread
+Solution-biased search allows for a much simpler parallel implementation. We
+create a number of threads, and give each thread its own random seed, otherwise each thread
 performs the same sequential search. Then, whenever a thread restarts, it uses a barrier to wait for
-every other thread to also restart. Each thread's new nogoods from that restart can then be
-gathered, and then each thread resumes, with a larger set of nogoods than would have been generated
-in a sequential run. Finally, whenever any single thread terminates, either due to having found a
-solution or proved unsatisfiability, then every other thread may immediately terminate.
+every other thread to also restart. At each restart each thread's new nogood can be gathered, and
+then each thread resumes, with a larger set of nogood than in a sequential run. Finally, whenever
+any single thread terminates, either due to having found a solution or proved unsatisfiability,
+then every other thread may immediately terminate.
 
@@
         the same amount of time between restarts. Because the only synchronisation between threads
-        is at restarts, we expect threads to be kept busy doing useful work.  (This is in contrast
+        is at a restart, we expect threads to be busy doing search.  (This is in contrast
         to an alternative method for paralellising restarting search \cite{DBLP:conf/aaai/CireKS14},
         which packs together successive sequence values to produce a balanced workload.)
-    \item Sequentially, on non-trivial instances the algorithm will restart very many times (many
+    \item Sequentially, on non-trivial instances the algorithm will restart often (many
         tens of thousands, for instances that reach the thousand second timeout).
@@
 It is important to emphasise that there is no expectation of a linear speedup from this approach,
-particularly for satisfiable instances. Due to the random nature of the value-ordering heuristic,
+particularly for satisfiable instances. Due to the biased random nature of the value-ordering heuristic,
 one of our extra threads may ``get lucky'' and find a solution much quicker than we would
 sequentially, leading to a superlinear speedup. Conversely, our extra threads may contribute nothing to finding a
-solution---or worse, due to new nogoods altering the choice made by the random branching heuristic,
+solution---or worse, due to new nogoods altering the choice made by the biased random branching heuristic,
 we could even get an absolute slowdown. We can see both of these effects in the bottom left plot of
@@
 
-The authors would like to thank Ozgur Akgun, Chris Jefferson, Christophe Lecoutre, Christine
+The authors would like to thank \"{O}zg\"{u}r Akg\"{u}n, Chris Jefferson, Christophe Lecoutre, Christine
 Solnon, and Craig Reilly for their comments.

### 2018-11-28 paper/paper.tex
Too essential
Thanks for all the fish

--- 2018-11-27 paper/paper.tex
@@
 simplification and re-engineering of an older Glasgow algorithm
-\cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}.  Essentially, the solver is essentially
-a dedicated forward-checking constraint programming implementation specifically for subgraph
-problems. It works with a model having a variable per pattern graph vertex, with domains ranging
-over the target graph vertices, and performs a backtracking search to map each pattern vertex to a
-target vertex whilst propagating adjacency and injectivity constraints (together with further
-implied constraints based upon degrees and paths).  However, it uses specialised bit-parallel data
-structures and algorithms, and propagates constraints in a fixed order rather than using a queue.
+\cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}.  Essentially, the solver is a dedicated
+forward-checking constraint programming implementation specifically for subgraph problems. It works
+with a model having a variable per pattern graph vertex, with domains ranging over the target graph
+vertices, and performs a backtracking search to map each pattern vertex to a target vertex whilst
+propagating adjacency and injectivity constraints (together with further implied constraints based
+upon degrees and paths).  However, it uses specialised bit-parallel data structures and algorithms,
+and propagates constraints in a fixed order rather than using a queue.
 
@@
 
-The authors would like to thank Ozgur Akgun, Chris Jefferson, Christophe Lecoutre, Christine
-Solnon, and Craig Reilly for their comments.
+The authors would like to thank Ozgur Akgun, Chris Jefferson, Sonja Kraiczy, Christophe Lecoutre,
+Christine Solnon, and Craig Reilly for their comments.
 

### 2018-11-29 paper/paper.tex
Edit Ruth's edits

--- 2018-11-28 paper/paper.tex
@@
 algorithm perform two orders of magnitude better on a large number of satisfiable instances, whilst
-performing worse rarely on satisfiable instances, and never on unsatisfiable instances.  This
+performing worse only rarely on satisfiable instances, and never on unsatisfiable instances.  This
 strategy is also effective in an optimisation setting, producing benefits in two maximum common
@@
 
-This new form of search can also be parallelised, with a much simpler implementation than
+This new form of search can also be parallelised, with a \emph{much} simpler implementation than
 conventional work-stealing. By running many threads with different random seeds but the same restart
@@
 performs the same sequential search. Then, whenever a thread restarts, it uses a barrier to wait for
-every other thread to also restart. At each restart each thread's new nogood can be gathered, and
-then each thread resumes, with a larger set of nogood than in a sequential run. Finally, whenever
+every other thread to also restart. At each restart each thread's new nogoods can be gathered, and
+then each thread resumes, with a larger set of nogoods than in a sequential run. Finally, whenever
 any single thread terminates, either due to having found a solution or proved unsatisfiability,
@@
 sequentially, leading to a superlinear speedup. Conversely, our extra threads may contribute nothing to finding a
-solution---or worse, due to new nogoods altering the choice made by the biased random branching heuristic,
+solution---or worse, due to new nogoods altering the choice made by the random branching heuristic,
 we could even get an absolute slowdown. We can see both of these effects in the bottom left plot of

### 2018-11-29 paper/paper.tex
James's changes

@@
     subproblems. This increases performance by two orders of magnitude on satisfiable instances,
-    whilst not affecting performance on unsatisfiable instances. Such a such search can also be
+    whilst not affecting performance on unsatisfiable instances. This algorithm can also be
     parallelised in a very simple but effective way: across both satisfiable and unsatisfiable
@@
 inside a larger ``target'' graph. The problem is broadly applicable, with applications in areas
-including in bioinformatics \cite{DBLP:journals/bmcbi/BonniciGPSF13}, chemistry \cite{o:Regin95}, computer vision
+including bioinformatics \cite{DBLP:journals/bmcbi/BonniciGPSF13}, chemistry \cite{o:Regin95}, computer vision
 \cite{DBLP:journals/cviu/DamiandSHJS11,DBLP:journals/pr/SolnonDHJ15}, law enforcement
@@
 The \emph{non-induced subgraph isomorphism problem} is to find an injective mapping from the
-vertices of a \emph{pattern} graph $\mathcal{P}$ to a \emph{target} graph $\mathcal{T}$, such that
-adjacent vertices in $\mathcal{P}$ are mapped to adjacent vertices in $\mathcal{T}$ (including that
-vertices with loops in $\mathcal{P}$ may only be mapped to vertices with loops in $\mathcal{T}$).
-The \emph{induced} problem additionally requires that non-adjacent vertices are mapped to
-non-adjacent vertices. The \emph{degree} of a vertex is the number of other vertices to which it is
-adjacent.
+vertices of a \emph{pattern} graph $\mathcal{P}$ to the vertices of a \emph{target} graph
+$\mathcal{T}$, such that adjacent vertices in $\mathcal{P}$ are mapped to adjacent vertices in
+$\mathcal{T}$ (including that vertices with loops in $\mathcal{P}$ may only be mapped to vertices
+with loops in $\mathcal{T}$).  The \emph{induced} problem additionally requires that non-adjacent
+vertices are mapped to non-adjacent vertices. The \emph{degree} of a vertex is the number of other
+vertices to which it is adjacent.
 
@@
 We begin with a set of baseline performance measurements. In the top
-two plots \cref{figure:others} we show the cumulative number of instances solved over time for the
+two plots of \cref{figure:others} we show the cumulative number of instances solved over time for the
 non-induced and induced problems respectively. We compare the Glasgow Subgraph Solver using
@@
 are usually reliable, and that most solutions can be found with only a small number of
-discrepancies; in such cases, the heavy commitment to early branching choices made by conventional
+discrepancies. In such cases, the heavy commitment to early branching choices made by conventional
 backtracking search can be extremely costly.
@@
 \emph{commit equal search effort} to each of the three subproblems found by mapping to vertices of
-degree ten.  And similarly, should we be certain that vertex of degree ten is so much better than a
+degree ten.  And similarly, should we be certain that a vertex of degree ten is so much better than a
 vertex of degree nine that we should commit no effort to degree nine vertices until all the degree
@@
 Erd\H{o}s-R\'enyi graphs. We expect it may be possible to derive better estimators for particular
-input classes, although over the full range of problem instances, we have verified that estimators
-of the form $e^d$ substantially outperform polynomial and factorial weightings.
+input classes, although over the full range of problem instances, we have verified that exponential
+estimators substantially outperform polynomial and factorial weightings.
 

### 2018-11-29 paper/paper.tex
Blair's changes

@@
 
-Solution-biased search allows for a much simpler parallel implementation. We
-create a number of threads, and give each thread its own random seed, otherwise each thread
-performs the same sequential search. Then, whenever a thread restarts, it uses a barrier to wait for
-every other thread to also restart. At each restart each thread's new nogoods can be gathered, and
-then each thread resumes, with a larger set of nogoods than in a sequential run. Finally, whenever
-any single thread terminates, either due to having found a solution or proved unsatisfiability,
-then every other thread may immediately terminate.
+Solution-biased search allows for a much simpler parallel implementation. We create a
+number of threads, and give each thread its own random seed; otherwise each thread performs the same
+sequential search. Threads synchronise on restarts, a simple barrier causing each thread to wait for
+every other thread to also restart.  Nogoods from all threads are then gathered and combined before
+search resumes, now with a larger set of nogoods than in a sequential run.  Finally, whenever any
+single thread terminates, either due to having found a solution or proved unsatisfiability, then
+every other thread may immediately terminate.
 
-A number of factors combine to make this technique feasible:
+This technique requires only limited changes to the top level search driver, and none whatsoever to
+the main recursive search algorithm. Notably, it does not require any locking or communication
+during the recursive search, aside from a single atomic boolean flag to assist early termination.  A
+number of factors combine to make this approach feasible:
 
@@
         to an alternative method for paralellising restarting search \cite{DBLP:conf/aaai/CireKS14},
-        which packs together successive sequence values to produce a balanced workload.)
+        which packs together successive sequence values to produce a balanced workload.) This
+        approach therefore avoids the irregular task size issues which usually arise in parallel
+        combinatorial search.
     \item Sequentially, on non-trivial instances the algorithm will restart often (many
@@
 Luby sequence for restarts, we could restart after either a constant number of backtracks, or simply
-after a certain time interval has passed. \Cref{figure:parallel} also shows the effects of restarting
-every 100ms. Sequentially, this approach is slightly better than using the Luby sequence, being able to
-solve 14,370 instances, with the last at 996.4 seconds. With thirty six threads, solving this many
+after a certain time interval has passed---this bounds the maximum possible idle time that threads
+could spend at a barrier. \Cref{figure:parallel} also shows the effects of restarting every 100ms.
+Sequentially, this approach is slightly better than using the Luby sequence, being able to solve
+14,370 instances, with the last at 996.4 seconds. With thirty six threads, solving this many
 instances takes 31.8 seconds, giving an aggregate speedup of 31.4.
@@
 
-To further test the scalability of this technique, we also implemented a distributed-parallelism
-layer on top of the threaded layer, using MPI. In contrast to the huge difficulties of implementing
-work-stealing in a distributed memory setting, this required only the addition of a simple ``all
-gather'' operation to communicate nogoods, and a ``gather'' to collect and combine the results of
-each host. Termination, meanwhile, was handled by posting an empty nogood (and so termination can
-only occur on a restart).
+To further test the scalability of this technique, we also implemented a distributed-memory
+parallelism layer on top of the threaded layer, using MPI. In contrast to the huge difficulties of
+implementing work-stealing in a distributed memory setting, this required only the addition of two
+MPI calls: an ``all gather'' operation to communicate nogoods, and a ``gather'' to collect and
+combine the results of each host. Termination, meanwhile, was handled by posting an empty nogood
+(and so termination can only occur on a restart).
 
@@
     \caption{Above, the cumulative number of maximum common induced subgraph instances solved by
-    k${\downarrow}$ and McSplit${\downarrow}$ over time, with and without restarts and nogood
-    recording. Below, comparing k${\downarrow}$ (left) and McSplit${\downarrow}$ (right) on an instance by instance
+    k${\downarrow}$ and McSplit${\downarrow}$ over time, with the two forms of search. Below,
+    comparing k${\downarrow}$ (left) and McSplit${\downarrow}$ (right) on an instance by instance
     basis.}\label{figure:mcs}

### 2018-11-29 paper/paper.tex
More little tweaks

@@
     parallelised in a very simple but effective way: across both satisfiable and unsatisfiable
-    instances, we get a further speedup of over thirty from thirty six cores, or ninety five from
+    instances, we get a further speedup of over thirty from thirty-six cores, or ninety five from
     ten distributed-memory hosts. Finally, we show that solution-biased search is also suitable for
@@
 conventional work-stealing. By running many threads with different random seeds but the same restart
-schedule, and sharing nogoods only following restarts, we can achieve speedups of thirty one from a
-thirty six core machine, or ninety five by using ten such machines with message passing.
+schedule, and sharing nogoods only following restarts, we can achieve speedups of thirty-one from a
+thirty-six core machine, or ninety-five by using ten such machines with message passing.
 
-\subsection{Definitions}
+\subsection{Background}
 
@@
 % sequence relation of \citet{Zampelli et al.\ }{DBLP:journals/constraints/ZampelliDS10}.
-
-\subsection{The Glasgow Subgraph Solver}
 
@@
 instances within the thousand second timeout, with the last instance being solved at 939.0 seconds.
-Using thirty six threads on machines with two eighteen core processors, we can solve 14,357
+Using thirty-six threads on machines with two eighteen core processors, we can solve 14,357
 instances with a timeout of only 74.2 seconds, giving an aggregate speedup
@@
 Sequentially, this approach is slightly better than using the Luby sequence, being able to solve
-14,370 instances, with the last at 996.4 seconds. With thirty six threads, solving this many
+14,370 instances, with the last at 996.4 seconds. With thirty-six threads, solving this many
 instances takes 31.8 seconds, giving an aggregate speedup of 31.4.

### 2018-11-29 paper/paper.tex
Even more things

@@
     parallelised in a very simple but effective way: across both satisfiable and unsatisfiable
-    instances, we get a further speedup of over thirty from thirty-six cores, or ninety five from
+    instances, we get a further speedup of over thirty from thirty-six cores, or ninety-five from
     ten distributed-memory hosts. Finally, we show that solution-biased search is also suitable for
@@
 
-Firstly, we define a new \textbf{Biased} value-ordering heuristic, as follows. When branching, we
+We first define a new \textbf{Biased} value-ordering heuristic, as follows. When branching, we
 select a vertex $v'$ from the chosen domain $D_v$ with probability \[ p(v') =
@@
 
-To further test the scalability of this technique, we also implemented a distributed-memory
-parallelism layer on top of the threaded layer, using MPI. In contrast to the huge difficulties of
+To further test the scalability of this technique, we also used MPI to implement a distributed-memory
+parallelism layer on top of the threaded layer. In contrast to the huge difficulties of
 implementing work-stealing in a distributed memory setting, this required only the addition of two
@@
 
-\Cref{figure:parallel} also shows the results of these experiments, using five and ten hosts.
-Because each host has two sockets, the five host results use ten MPI ranks, each with eighteen
-threads, and the ten host results use twenty MPI ranks. Because the supercomputing service we use is
-not designed for huge numbers of very short problems, we ran only instances whose sequential runtime
-was at least one second, and for the sake of plotting results, treat skipped instances as taking one
-second. Due to the job launcher used, it is also not possible to accurately measure ``total''
-runtime including startup costs, and so instead we report the runtime of the rank zero
-process---this figure is somewhat optimistic for very easy instances. With these caveats in mind, we
-see that our results continue to improve as the host count increases, and we come very close to the
-magic goal of an overall speedup of one hundred.
+\Cref{figure:parallel} also shows the results of these experiments, using five, ten, and twenty
+hosts.  Because each host has two sockets, the five host results use ten MPI ranks, each with
+eighteen threads, and the ten and twenty host results use twenty and forty MPI ranks respectively.
+The supercomputing service we use is not designed for huge numbers of very short problems, and so we
+ran only instances whose sequential runtime was at least one second; for the sake of plotting
+results, we treat skipped instances as taking one second. Due to the job launcher used, it is also
+not possible to accurately measure ``total'' runtime including startup costs, and so instead we
+report the runtime of the rank zero process---this figure is somewhat optimistic for very easy
+instances. With these caveats in mind, when seeing how long a timeout is needed to solve any 14,370
+instances, we get aggregate speedups of 57.1 from five hosts and 95.5 from ten hosts over a
+sequential baseline; using twenty hosts is slightly slower, due to increased overheads. However,
+when seeing how long a timeout is needed to solve 14,424 instances, distributed parallelism gives us
+further aggregate speedups of 2.2, 3.5, and 7.9 from five, ten, and twenty hosts, over the single
+machine parallel baseline.
 

### 2018-11-29 paper/paper.tex
Fix Ruth's crime against humanity

@@
 
-The authors would like to thank \"{O}zg\"{u}r Akg\"{u}n, Chris Jefferson, Christophe Lecoutre, Christine
-Solnon, and Craig Reilly for their comments.
+The authors would like to thank \"{O}zg\"{u}r Akg\"{u}n, Chris Jefferson, Sonja Kraiczy, Christophe
+Lecoutre, Christine Solnon, and Craig Reilly for their comments.
 

### 2018-11-29 paper/paper.tex
Last tweaks. Definitely this time. Honest

@@
 The subgraph isomorphism problem is to decide whether a copy of a small ``pattern'' graph occurs
-inside a larger ``target'' graph. The problem is broadly applicable, with applications in areas
+inside a larger ``target'' graph. The problem is broadly applicable, arising in areas
 including bioinformatics \cite{DBLP:journals/bmcbi/BonniciGPSF13}, chemistry \cite{o:Regin95}, computer vision
@@
 E5-2695 v4 CPUs and 256GBytes RAM, running Centos 7.3.1611. We use GCC 7.2.0 as the compiler. For
-paralellism, we use C++ native threads, and for distributed parallelism we also use the SGI MPT
+parallelism, we use C++ native threads, and for distributed parallelism we also use the SGI MPT
 implementation of MPI. All timing measurements are steady-clock, and we use a deterministic
@@
 \cite{DBLP:conf/gbrpr/CarlettiFSV17} (which only supports the induced problem), in each case using
-the original implementation provided by the algorithm's authors. For each algorithm, the $y$ value
-gives the cumulative number of instances which (individually) can be solved in no more than $x$
-milliseconds.  The vertical distance between two lines therefore shows how many more instances can
-be solved by one solver than another, if every instance is run separately with the chosen $x$
-timeout. The horizontal distance shows how many times longer the per-instance timeout would need to
-be to allow the rightmost algorithm to succeed on $y$ out of the 14,621 instances (bearing in mind
-that the two sets of $y$ instances could be different).  The plots show that our starting point
-comfortably beats PathLAD, VF2, VF3 and RI, except for very low choices of timeout.
+the original implementation provided by the algorithm's authors.  The plots show that our starting
+point comfortably beats PathLAD, VF2, VF3 and RI, except for very low choices of timeout. For each
+algorithm, the $y$ value gives the cumulative number of instances which (individually) can be solved
+in no more than $x$ milliseconds.  The vertical distance between two lines therefore shows how many
+more instances can be solved by one solver than another, if every instance is run separately with
+the chosen $x$ timeout. The horizontal distance shows how many times longer the per-instance timeout
+would need to be to allow the rightmost algorithm to succeed on $y$ out of the 14,621 instances
+(bearing in mind that the two sets of $y$ instances could be different), and gives a measure called
+\emph{aggregate speedup} \cite{DBLP:conf/cpaior/HoffmannMNPRS018}.
 
@@
 upon improving the performance on the remaining hard satisfiable instances, and so in the bottom two
-plots in \cref{figure:others} (and in subsequent cumulative plots) we show only satisfiable
-instances, and use a reduced range on both axes.  For the remainder of this paper, we show only the
-non-induced problem, which tends to be harder; results with the induced variant are similar.
+plots in \cref{figure:others} (and in subsequent cumulative plots for sequential algorithms) we show
+only satisfiable instances, and use a reduced range on both axes.  For the remainder of this paper,
+we show only the non-induced problem, which tends to be harder; results with the induced variant are
+similar.
 
@@
 are usually reliable, and that most solutions can be found with only a small number of
-discrepancies. In such cases, the heavy commitment to early branching choices made by conventional
+discrepancies. In such cases, the heavy commitment to early branching choices made by
 backtracking search can be extremely costly.
@@
 worse.  In other words, as well as improving performance, we have made up most of the consistency we
-lost by introducing randomness.  This is a promising result: for the first time, we have introduced
-a viable alternative to backtracking search in a setting where we both have a strong
-\emph{value}-ordering heuristic, \emph{and} care about both satisfiable and unsatisfiable instances.
+lost by introducing randomness.
 
@@
 and of how many solutions each subproblem is likely to contain. Of course, obtaining exact (or even
-approximate) values for these figures is at least as hard as solving the problem in its entirety.
-For pairs of Erd\H{o}s-R\'enyi
+approximate) values for these figures is at least as hard as solving the problem in its entirety,
+but we may obtain usable approximations.  For pairs of Erd\H{o}s-R\'enyi
 random graphs with large solution counts (i.e.\ chosen from within the ``easy satisfiable'' region
-\cite{DBLP:journals/jair/McCreeshPST18}), we observe a linear relationship between subproblem size
+\cite{DBLP:journals/jair/McCreeshPST18}), we can observe a linear relationship between subproblem size
 and number of solutions. Thus, for graphs from this distribution, we need only an estimator of
@@
     \caption{Above, the cumulative number of instances solved over time, comparing the sequential
-    algorithm to results using 32 threads on a single machine, and using five or ten distributed
-    memory hosts. Below, instance by instance comparisons.}\label{figure:parallel}
+    algorithm to results using 32 threads on a single machine, and using five, ten or twenty
+    distributed memory hosts. Below, instance by instance comparisons.}\label{figure:parallel}
 \end{figure}
@@
 
-It is important to emphasise that there is no expectation of a linear speedup from this approach,
+It is important to emphasise that there is no expectation of a linear speedup from this approach
+\cite{DBLP:conf/aaai/LiW84,DBLP:journals/tc/LiW86,DBLP:conf/irregular/BruinKT95},
 particularly for satisfiable instances. Due to the biased random nature of the value-ordering heuristic,
@@
 \Cref{figure:parallel} also shows the results of these experiments, using five, ten, and twenty
-hosts.  Because each host has two sockets, the five host results use ten MPI ranks, each with
+hosts.  Because each host has two CPU sockets, the five host results use ten MPI ranks, each with
 eighteen threads, and the ten and twenty host results use twenty and forty MPI ranks respectively.
@@
 techniques gives us, for the first time, a practical alternative to backtracking search where we
-have a strong value-ordering heuristic and where we care both about satisfiable and unsatisfiable
-instances.  A further benefit is the ease with which such a search can be parallelised. By having
-each thread carry out the same search with a different random seed, and sharing nogoods only on
-restarts, we remove the need for intrusive changes to the core search algorithm, eliminate the
-irregularity problem, and still respect the advice of the value-ordering heuristic.
+have a strong \emph{value}-ordering heuristic, and where we care both about satisfiable \emph{and}
+unsatisfiable instances.
+
+A further benefit is the ease with which such a search can be parallelised. By having each thread
+carry out the same search with a different random seed, and sharing nogoods only on restarts, we
+remove the need for intrusive changes to the core search algorithm, eliminate the irregularity
+problem, and still respect the advice of the value-ordering heuristic.
 

### 2019-02-27 paper/paper.tex
Fixes

@@
 are based upon decompositions
-\cite{DBLP:journals/corr/abs-1008-4328,DBLP:journals/jair/MalapertRR16}, work stealing
+\cite{DBLP:journals/corr/abs-1008-4328,DBLP:journals/jair/MalapertRR16}, work-stealing
 \cite{DBLP:conf/cp/MichelSH07,DBLP:conf/cp/ChuSS09,DBLP:conf/cp/McCreeshP15,DBLP:conf/cpaior/HoffmannMNPRS018},
@@
 decomposition interferes strongly with the shape of the search tree
-\cite{DBLP:journals/jair/MalapertRR16}. Work stealing, traditionally, also interferes with
+\cite{DBLP:journals/jair/MalapertRR16}. Work-stealing, traditionally, also interferes with
 value-ordering \cite{DBLP:journals/topc/McCreeshP15}, although specially designed exceptions exist
@@
 instances, we get aggregate speedups of 57.1 from five hosts and 95.5 from ten hosts over a
-sequential baseline; using twenty hosts is slightly slower, due to increased overheads. However,
-when seeing how long a timeout is needed to solve 14,424 instances, distributed parallelism gives us
-further aggregate speedups of 2.2, 3.5, and 7.9 from five, ten, and twenty hosts, over the single
-machine parallel baseline.
+sequential baseline; using twenty hosts is slightly slower, due to increased overheads. Finally, if
+we look at harder instances, by allow a longer sequential timeout, we can solve 14,415 instances
+sequentially with the last at 8,549 seconds; at this difficulty level, we achieve aggregate speedups
+of 82.6, 103.2, and 144.9 from five, ten and twenty hosts.
 

### 2019-02-27 paper/paper.tex
Last tweaks

@@
     parallelised in a very simple but effective way: across both satisfiable and unsatisfiable
-    instances, we get a further speedup of over thirty from thirty-six cores, or ninety-five from
+    instances, we get a further speedup of over thirty from thirty-six cores, and over one hundred from
     ten distributed-memory hosts. Finally, we show that solution-biased search is also suitable for
@@
 conventional work-stealing. By running many threads with different random seeds but the same restart
-schedule, and sharing nogoods only following restarts, we can achieve speedups of thirty-one from a
-thirty-six core machine, or ninety-five by using ten such machines with message passing.
+schedule, and sharing nogoods only following restarts, we can achieve aggregate speedups
+\cite{DBLP:conf/cpaior/HoffmannMNPRS018} of thirty-one from a thirty-six core machine, or over one
+hundred by using ten such machines.
 
@@
 
-To avoid exploring portions of the search
-space that we have already visited, every time we restart, we add new constraints to the problem
-which eliminate already-explored subtrees---such a constraint is called a \emph{nogood}. We generate
-simple decision nogoods. That is, upon backtracking due to a decision to restart, we post a nogood
-of the form $(v \mapsto v') \wedge (w \mapsto w') \wedge (x \mapsto x') \Rightarrow \bot$ for every
-branch to the left of the current (incomplete) branch in the search tree, and when we first make a
-decision to restart before backtracking, we post a similar nogood eliminating the entire subtree
-explored. We use the two watched literals technique \cite{DBLP:conf/dac/MoskewiczMZZM01} to
-propagate stored nogoods.
+To avoid exploring portions of the search space that we have already visited, every time we restart,
+we add new constraints to the problem which eliminate already-explored subtrees---such a constraint
+is called a \emph{nogood}. We generate simple decision nogoods. That is, upon backtracking due to a
+decision to restart, we post a nogood of the form $(v \mapsto v') \wedge (w \mapsto w') \wedge (x
+\mapsto x') \Rightarrow \bot$ for every branch to the left of the current (incomplete) branch at
+every level of the search tree, and when we first make a decision to restart before backtracking, we
+post a similar nogood eliminating the entire subtree explored. We use the two watched literals
+technique \cite{DBLP:conf/dac/MoskewiczMZZM01} to propagate stored nogoods.
 %(\siplineref{2wl}).
@@
     \item The gathering of nogoods to describe the work done so far provides an alternative to
-        requiring either a specific mechanism to allocate work, or expensive synchronisation between threads.
+        requiring either a specific mechanism to allocate work, or expensive synchronisation between
+        threads. Notably, this completely bypasses the typical difficulties of sharing all learned
+        nogoods in learning solvers \cite{DBLP:books/daglib/p/HamadiJS12}.
     \item If sometimes threads do happen to explore part of

### 2019-06-01 talk/talk.tex
A talk

--- 2018-01-16 long-talk/talk.tex
@@
 \usepackage{amssymb}                   % \nexists
-\usepackage{gnuplot-lua-tikz}          % graphs
 \usepackage[vlined]{algorithm2e} % algorithms
 \usepackage{centernot}
-\usepackage{graphicx}
+\usepackage{listings}
+\usepackage{csquotes}
+
+\RequirePackage[tt=false, type1=true]{libertine}
+\RequirePackage[varqu]{zi4}
+\RequirePackage[libertine]{newtxmath}
+\RequirePackage[T1]{fontenc}
 
@@
 
-        \node (logo) [anchor=north east, shift={(-0.6cm,-0.6cm)}] at (current page.north east) {
-            \includegraphics[keepaspectratio=true,scale=0.7]{UoG_keyline.pdf}
+        \node (logo) [anchor=north east, shift={(-0.6cm,-0.4cm)}] at (current page.north east) {
+            \begin{minipage}{0.20\paperwidth}
+                \begin{center}
+                \includegraphics[keepaspectratio=true,scale=0.35]{UoG_keyline.pdf} \\[0.15cm]
+                \includegraphics[keepaspectratio=true,scale=0.8]{UStA_Logo_SP.eps}
+                \end{center}
+            \end{minipage}
         };
 
-        \node [anchor=west, xshift=0.2cm] at (current page.west |- logo.west) {
+        \node [anchor=west, xshift=0.2cm, yshift=0.2cm] at (current page.west |- logo.west) {
             \begin{minipage}{0.65\paperwidth}\raggedright
@@
 
+\makeatletter
+\newenvironment{justborderframe}[2][]{
+    \def\beamer@entrycode{\vspace*{-\headheight}}
+    \setbeamertemplate{headline}
+    {%
+        \begin{beamercolorbox}[colsep=1.5pt]{upper separation line head}
+        \end{beamercolorbox}
+        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
+            leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
+        \end{beamercolorbox}%
+        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
+            leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
+        \end{beamercolorbox}%
+        \begin{beamercolorbox}[colsep=1.5pt]{lower separation line head}
+        \end{beamercolorbox}
+        \vspace*{\headheight}
+    }
+
+    \setbeamertemplate{footline}
+    {%
+        \begin{beamercolorbox}[colsep=1.5pt]{upper separation line foot}
+        \end{beamercolorbox}
+        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
+            leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
+        \end{beamercolorbox}%
+        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
+            leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
+        \end{beamercolorbox}%
+        \begin{beamercolorbox}[colsep=1.5pt]{lower separation line foot}
+        \end{beamercolorbox}
+    }
+
+    \begin{frame}[#1]{}
+    }{
+    \end{frame}
+}
+\makeatother
+
 % }}}
 
-\title[Value Ordering, Discrepancies, and Restarts for Subgraph Algorithms]{Value Ordering,
-Discrepancies, and Restarts for Subgraph Algorithms}
-\author[Ciaran McCreesh, Patrick Prosser and James Trimble]{\textbf{Ciaran McCreesh}, Patrick
-Prosser and \\ James Trimble}
+\author{Blair Archibald \and Fraser Dunlop \and Ruth Hoffmann \and Ciaran McCreesh \and Patrick
+Prosser \and James Trimble}
+\title{Sequential and Parallel Solution-Biased Search for Subgraph Algorithms}
 
@@
         \tikz[overlay, remember picture]
-        \node[at=(current page.south), anchor=south, inner sep=0pt]{\includegraphics[keepaspectratio=true, width=\paperwidth]{background.jpg}};
+        \node[at=(current page.south), anchor=south, inner sep=0pt]{\includegraphics[keepaspectratio=true, height=\paperheight]{background.jpg}};
     }
@@
 
-\section{Subgraph Isomorphism}
+\begin{frame}{Subgraph Isomorphism}
+    \centering
+    \begin{tikzpicture}
+        \node [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {\vphantom{1}};
+        \node [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {\vphantom{1}};
+        \node [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {\vphantom{1}};
+        \node [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {\vphantom{1}};
 
-\begin{frame}{Non-Induced Subgraph Isomorphism}
+        \draw [ultra thick] (Na) -- (Nb);
+        \draw [ultra thick] (Nb) -- (Nc);
+        \draw [ultra thick] (Nc) -- (Nd);
+        \draw [ultra thick] (Nb) -- (Nd);
 
-    \begin{center}\begin{tikzpicture}
-        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {1};
-        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {2};
-        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {3};
-        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {4};
+        \node [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {\vphantom{1}};
+        \node [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {\vphantom{1}};
+        \node [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {\vphantom{1}};
+        \node [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {\vphantom{1}};
+        \node [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {\vphantom{1}};
+        \node [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {\vphantom{1}};
 
-        \node <2> [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {1};
-        \node <2> [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {2};
-        \node <2> [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {3};
-        \node <2> [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {4};
-
-        \draw <1> [thick, color=uofgsandstone!50] (Na) -- (Nb);
-        \draw <1> [thick, color=uofgsandstone!50] (Nb) -- (Nc);
-        \draw <1> [thick, color=uofgsandstone!50] (Nc) -- (Nd);
-        \draw <1> [thick, color=uofgsandstone!50] (Nb) -- (Nd);
-
-        \draw <2> [ultra thick] (Na) -- (Nb);
-        \draw <2> [ultra thick] (Nb) -- (Nc);
-        \draw <2> [ultra thick] (Nc) -- (Nd);
-        \draw <2> [ultra thick] (Nb) -- (Nd);
-
-        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {1};
-        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {2};
-        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {3};
-        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {4};
-        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {5};
-        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {6};
-
-        \node <2> [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {1};
-        \node <2> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {2};
-        \node <2> [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {3};
-        \node <2> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {4};
-        \node <2> [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {5};
-        \node <2> [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {6};
-
-        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N2);
-        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N3);
-        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N4);
-        \draw <1> [thick, color=uofgsandstone!50] (N2) -- (N4);
-        \draw <1> [thick, color=uofgsandstone!50] (N3) -- (N5);
-        \draw <1> [thick, color=uofgsandstone!50] (N3) -- (N6);
-        \draw <1> [thick, color=uofgsandstone!50] (N4) -- (N6);
-        \draw <1> [thick, color=uofgsandstone!50] (N5) -- (N6);
-        \draw <1> [thick, color=uofgsandstone!50] (N2) to [in=45, out=315] (N6);
-        \draw <1> [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
-
-        \draw <2> [thick, color=uofgsandstone!50] (N1) -- (N2);
-        \draw <2> [ultra thick] (N1) -- (N3);
-        \draw <2> [thick, color=uofgsandstone!50] (N1) -- (N4);
-        \draw <2> [thick, color=uofgsandstone!50] (N2) -- (N4);
-        \draw <2> [ultra thick] (N3) -- (N5);
-        \draw <2> [ultra thick] (N3) -- (N6);
-        \draw <2> [thick, color=uofgsandstone!50] (N4) -- (N6);
-        \draw <2> [ultra thick] (N5) -- (N6);
-        \draw <2> [thick, color=uofgsandstone!50] (N2) to [in=45, out=315] (N6);
-        \draw <2> [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
-    \end{tikzpicture}\end{center}
-
+        \draw [thick, color=uofgsandstone!50] (N1) -- (N2);
+        \draw [ultra thick] (N1) -- (N3);
+        \draw [thick, color=uofgsandstone!50] (N1) -- (N4);
+        \draw [thick, color=uofgsandstone!50] (N2) -- (N4);
+        \draw [ultra thick] (N3) -- (N5);
+        \draw [ultra thick] (N3) -- (N6);
+        \draw [thick, color=uofgsandstone!50] (N4) -- (N6);
+        \draw [ultra thick] (N5) -- (N6);
+        \draw [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
+    \end{tikzpicture}
 \end{frame}
 
-\begin{frame}{Benchmark Instances}
+\begin{frame}{Subgraph Finding, as a Constraint Program}
+    \begin{itemize}
+        \item A variable for each pattern vertex. The domains are all of the target vertices.
+        \item At least two sets of constraints:
+            \begin{itemize}
+                \item Adjacent pairs of vertices must be mapped to adjacent pairs of vertices.
+                \item All different.
+            \end{itemize}
+        \item Then we get clever:
+            \begin{itemize}
+                \item Extra constraints about degrees, paths, \ldots
+                \item Very good variable- and value-ordering heuristics
+            \end{itemize}
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{The Glasgow Subgraph Solver}
+    \begin{center}
+        \url{https://github.com/ciaranm/glasgow-subgraph-solver}
+    \end{center}
 
     \begin{itemize}
-        \item Other papers: a subset of 5,725 instances from Christine Solnon's selection:
-            \begin{itemize}
-                \item Randomly generated with different models.
-                \item Real-world graphs.
-                \item Computer vision problems.
-                \item Biochemistry problems.
-                \item Phase transition instances.
+        \item Subgraph isomorphism, and all its variants (induced / non-induced, homomorphism,
+            locally injective, labels, side constraints, directed, \ldots).
+        \item Also special algorithms for clique.
     \end{itemize}
-        \item Here: all 14,621 instances:
-            \begin{itemize}
-                \item $\ge \input{../paper/gen-n-sat.tex}$ satisfiable.
-                \item $\ge \input{../paper/gen-n-unsat.tex}$ unsatisfiable.
-            \end{itemize}
-        \item A lot of them are very easy for good algorithms.
-    \end{itemize}
-
-\end{frame}
-
-\begin{frame}{The Algorithm}
-
-    \begin{itemize}
-        \item A variable for every pattern vertex.
-        \item Domains range over target vertices.
-        \item Clever filtering at the top of search using neighbourhood degree sequences and paths.
-        \item Bit-parallel forward checking search:
-            \begin{itemize}
-                \item Adjacency and pre-computed path count constraints.
-                \item Cheap hacky all-different propagator.
-                \item Smallest domain first, tie-breaking on highest degree.
-            \end{itemize}
-    \end{itemize}
-
-\end{frame}
-
-\begin{frame}{Is It Any Good?}
-
-    \includegraphics<1>{gen-graph-others.pdf}%
-    \includegraphics<2>{gen-graph-others-zoom.pdf}
-
 \end{frame}
@@
     \begin{itemize}
-        \item Highest static degree first.
-        \item But degree spread is usually quite low\ldots
+        \item Largest target degree first.
     \end{itemize}
@@
 \begin{frame}{Sanity Check}
-
-    \includegraphics<1>{gen-graph-value-ordering-heuristics.pdf}
-
+    \begin{center}
+        \includegraphics<1>{gen-graph-value-ordering.pdf}%
+        \includegraphics<2>{gen-graph-value-ordering-unsat.pdf}
+    \end{center}
 \end{frame}
 
-\begin{frame}{Discrepancy Search}
-
-    \includegraphics<1>{gen-graph-scatter-dds.pdf}
-
-\end{frame}
-
-\begin{frame}{Restarts and Nogoods}
-
+\begin{frame}{However\ldots}
     \begin{itemize}
-        \item Restarts to affect \emph{value}-ordering heuristics.
-        \item Two watched literals to avoid repeating ourselves.
-        \item Luby with a magic constant multiplier.
-            \begin{itemize}
-                \item 666, chosen by divine revelation.
-            \end{itemize}
-    \end{itemize}
-
-\end{frame}
-
-\begin{frame}{Random Value-Ordering with Restarts}
-
-    \includegraphics<1>{gen-graph-random-restarts.pdf}%
-    \includegraphics<2>{gen-graph-scatter-random.pdf}
-
-\end{frame}
-
-\begin{frame}{Biased Value-Ordering}
-
-    \only<1>{
-    \begin{itemize}
-        \item Select a vertex $v'$ from the chosen domain $D_v$ with
-            probability \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \text{.} \]
-        \item Looks a lot like $\mathit{softmax}$, which uses base $e$.
-    \end{itemize}
-    }%
-    \includegraphics<2>{gen-graph-value-ordering-heuristics-biased.pdf}
-
-\end{frame}
-
-\begin{frame}{Biased Value-Ordering with Restarts}
-
-    \includegraphics<1>{gen-graph-restarts.pdf}%
-    \includegraphics<2>{gen-graph-scatter-final.pdf}
-
-\end{frame}
-
-\section{Maximum Common Subgraph}
-
-\begin{frame}{Maximum Common Subgraph Algorithms}
-
-    \includegraphics<1>{gen-graph-mcs.pdf}%
-    \includegraphics<2>{gen-graph-scatter-kdown.pdf}%
-    \includegraphics<3>{gen-graph-scatter-mcsplit.pdf}%
-
-\end{frame}
-
-\section{Parallel}
-
-\begin{frame}{Parallel}
-
-    \includegraphics<1>{gen-graph-parallel.pdf}%
-    \includegraphics<2>{gen-graph-as-parallel.pdf}%
-    \includegraphics<3>{gen-graph-scatter-parallel.pdf}%
-
-\end{frame}
-
-\begin{frame}{Properties of This Sort of Parallel Search}
-    \begin{itemize}
-        \item Reproducible (assuming each thread has its own deterministic random number generator)
-        \item Can perform worse than sequential on satisfiable instances:
-            \begin{itemize}
-                \item Nogoods can knock out values from domains, leading to a different random
-                    search order, even if the first thread uses the same seed.
-                \item Challenge for you: design a sane and fast biased random selection algorithm
-                    that is immune to this, and that is statistically the same sequentially. It
-                    feels like this should be possible\ldots
-                \item Also, our all different propagator isn't monotonic or idempotent.
-            \end{itemize}
-
-        \item Similarly, adding extra threads can make it worse.
+        \item What if several vertices have the same degree?
+        \item Is a vertex of degree $10$ really that much better than a vertex of degree $9$?
     \end{itemize}
@@
 
+\begin{frame}{Discrepancy Search?}
+    \begin{center}
+        \includegraphics<1>{gen-graph-value-ordering-dds.pdf}%
+        \includegraphics<2>{gen-graph-value-ordering-dds-unsat.pdf}%
+        \includegraphics<3>{gen-graph-value-ordering-dds-scatter.pdf}
+    \end{center}
+\end{frame}
+
+\begin{frame}{Value-Ordering Heuristics as Distributions}
+    \begin{itemize}
+        \item Traditional view: value-ordering defines a search order.
+        \item New view: value-ordering defines \textcolor{uofgcobalt}{what proportion of the search
+            effort} should be spent on different subproblems.
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{A Slightly Random Value-Ordering Heuristic}
+    \only<1>{
+    \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \]
+    \begin{itemize}
+        \item Equally likely to pick between two vertices of degree $d$.
+        \item Twice as likely to select a vertex of degree $d$ than a vertex of degree $d - 1$.
+        \item In the paper: \textcolor{uofgcobalt}{solution density} and expected distribution of solutions.
+    \end{itemize}}\only<2>{
+    \begin{center}
+        \includegraphics<2>{gen-graph-bias-scatter.pdf}%
+    \end{center}}
+\end{frame}
+
+\begin{frame}{Restarts and Nogood Recording}
+    \begin{itemize}
+        \item Aggressive restarts: every 100ms.
+        \item Nogood recording and 2WL to avoid repeating work.
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Is It Better?}
+    \begin{center}
+        \includegraphics<1>{gen-graph-sbs.pdf}%
+        \includegraphics<2>{gen-graph-sbs-unsat.pdf}%
+        \includegraphics<3>{gen-graph-sbs-scatter.pdf}%
+    \end{center}
+\end{frame}
+
+\begin{frame}{Parallel Search}
+    \begin{itemize}
+        \item Each thread gets its own random seed.
+        \item Barrier synchronise on restarts.
+        \item Share nogoods.
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Is It Even Betterer?}
+    \includegraphics<1>{gen-graph-par-scatter.pdf}%
+    \includegraphics<2>{gen-graph-par.pdf}%
+    \includegraphics<3>{gen-graph-dist.pdf}%
+\end{frame}
+
+\begin{frame}{Broader Perspective}
+    \begin{itemize}
+        \item Also good for maximum common subgraph algorithms.
+        \item Does this work for CP in general?
+        \item Can we finally kill off vanilla sequential backtracking search?
+    \end{itemize}
+\end{frame}
+
+{
+    \usebackgroundtemplate{
+        \tikz[overlay, remember picture]
+        \node[at=(current page.south), anchor=south, inner
+        sep=0pt]{\includegraphics[keepaspectratio=true, width=\paperwidth]{background2.jpg}};
+    }
+
     \begin{frame}[plain,noframenumbering]
-    \begin{center}
-        \vspace*{4em}
-        \url{http://www.dcs.gla.ac.uk/~ciaran} \\
-        \vspace{1em}
-        \href{mailto:ciaran.mccreesh@glasgow.ac.uk}{\nolinkurl{ciaran.mccreesh@glasgow.ac.uk}} \\
-    \end{center}
-
         \begin{tikzpicture}[remember picture, overlay]
@@
             };
+
+            \node [anchor=north west, shift={(0.6cm,-0.8cm)}] at (current page.north west) {
+                \textcolor{white}{\url{http://www.dcs.gla.ac.uk/~ciaran}}
+            };
+
+            \node [anchor=north west, shift={(0.6cm,-1.4cm)}] at (current page.north west) {
+                    \textcolor{white}{\href{mailto:ciaran.mccreesh@glasgow.ac.uk}{\nolinkurl{ciaran.mccreesh@glasgow.ac.uk}}}
+            };
         \end{tikzpicture}
     \end{frame}
+}
 

### 2019-06-07 talk/talk.tex
Tweaking

@@
                 \includegraphics[keepaspectratio=true,scale=0.35]{UoG_keyline.pdf} \\[0.15cm]
-                \includegraphics[keepaspectratio=true,scale=0.8]{UStA_Logo_SP.eps}
+                \includegraphics[keepaspectratio=true,scale=0.07]{UStA_Logo_SP.png}
                 \end{center}
@@
     \begin{tikzpicture}
-        \node [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {\vphantom{1}};
-        \node [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {\vphantom{1}};
-        \node [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {\vphantom{1}};
-        \node [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {\vphantom{1}};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {\vphantom{1}};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {\vphantom{1}};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {\vphantom{1}};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {\vphantom{1}};
+
+        \node <2> [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {\vphantom{1}};
+        \node <2> [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {\vphantom{1}};
+        \node <2> [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {\vphantom{1}};
+        \node <2> [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {\vphantom{1}};
 
@@
 
-        \node [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {\vphantom{1}};
-        \node [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {\vphantom{1}};
-        \node [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {\vphantom{1}};
-        \node [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {\vphantom{1}};
-        \node [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {\vphantom{1}};
-        \node [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {\vphantom{1}};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {\vphantom{1}};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {\vphantom{1}};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {\vphantom{1}};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {\vphantom{1}};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {\vphantom{1}};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {\vphantom{1}};
 
-        \draw [thick, color=uofgsandstone!50] (N1) -- (N2);
-        \draw [ultra thick] (N1) -- (N3);
-        \draw [thick, color=uofgsandstone!50] (N1) -- (N4);
-        \draw [thick, color=uofgsandstone!50] (N2) -- (N4);
-        \draw [ultra thick] (N3) -- (N5);
-        \draw [ultra thick] (N3) -- (N6);
-        \draw [thick, color=uofgsandstone!50] (N4) -- (N6);
-        \draw [ultra thick] (N5) -- (N6);
-        \draw [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
+        \node <2> [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {\vphantom{1}};
+        \node <2> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {\vphantom{1}};
+        \node <2> [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {\vphantom{1}};
+        \node <2> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {\vphantom{1}};
+        \node <2> [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {\vphantom{1}};
+        \node <2> [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {\vphantom{1}};
+
+        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N2);
+        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N3);
+        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N4);
+        \draw <1> [thick, color=uofgsandstone!50] (N2) -- (N4);
+        \draw <1> [thick, color=uofgsandstone!50] (N3) -- (N5);
+        \draw <1> [thick, color=uofgsandstone!50] (N3) -- (N6);
+        \draw <1> [thick, color=uofgsandstone!50] (N4) -- (N6);
+        \draw <1> [thick, color=uofgsandstone!50] (N5) -- (N6);
+        \draw <1> [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
+
+        \draw <2> [thick, color=uofgsandstone!50] (N1) -- (N2);
+        \draw <2> [ultra thick] (N1) -- (N3);
+        \draw <2> [thick, color=uofgsandstone!50] (N1) -- (N4);
+        \draw <2> [thick, color=uofgsandstone!50] (N2) -- (N4);
+        \draw <2> [ultra thick] (N3) -- (N5);
+        \draw <2> [ultra thick] (N3) -- (N6);
+        \draw <2> [thick, color=uofgsandstone!50] (N4) -- (N6);
+        \draw <2> [ultra thick] (N5) -- (N6);
+        \draw <2> [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
     \end{tikzpicture}
@@
                 \item Extra constraints about degrees, paths, \ldots
-                \item Very good variable- and value-ordering heuristics
+                \item Better data structures and propagation queues.
+                \item Very good variable- and value-ordering heuristics.
             \end{itemize}
@@
     \begin{itemize}
+        \item A CP style solver specifically for subgraph algorithms.
         \item Subgraph isomorphism, and all its variants (induced / non-induced, homomorphism,
@@
         \item Also special algorithms for clique.
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Benchmark Instances}
+    \begin{itemize}
+        \item 14,621 instances from Christine Solnon's collection:
+            \begin{itemize}
+                \item Randomly generated with different models.
+                \item Real-world graphs.
+                \item Computer vision problems.
+                \item Biochemistry problems.
+                \item Phase transition instances.
+            \end{itemize}
+        \item At least\ldots
+            \begin{itemize}
+                \item $\ge$ 2,110 satisfiable.
+                \item $\ge$ 12,322 unsatisfiable.
+            \end{itemize}
+        \item A lot of them are very easy for good algorithms.
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Hardware}
+    \begin{itemize}
+        \item HPC, optimised for throughput not reproducibility.\footnote{This work used the Cirrus
+            UK National Tier-2 HPC Service at EPCC (http://www.cirrus.ac.uk) funded by the
+            University of Edinburgh and EPSRC (EP/P020267/1)}
+        \item Dual Intel Xeon E5-2695 v4 CPUs, $2 \times 18$ cores
+        \item 256GBytes RAM
+        \item GCC 7.2.0
+        \item C++ native threads, SGI MPT MPI
     \end{itemize}
@@
 
-\begin{frame}{Sanity Check}
-    \begin{center}
-        \includegraphics<1>{gen-graph-value-ordering.pdf}%
-        \includegraphics<2>{gen-graph-value-ordering-unsat.pdf}
-    \end{center}
+\begin{frame}{Sanity Check}\includegraphics<1>{gen-graph-value-ordering.pdf}\includegraphics<2>{gen-graph-value-ordering-unsat.pdf}
 \end{frame}
@@
     \only<1>{
+    \begin{itemize}
+        \item For a fixed domain $D_v$, pick a vertex $v'$ from a domain $D_v$ with probability
     \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \]
-    \begin{itemize}
         \item Equally likely to pick between two vertices of degree $d$.
@@
     \begin{itemize}
+        \item Can scale to even more hosts with less strict synchronisation.
         \item Also good for maximum common subgraph algorithms.
